### 2021-06-22 README.Rmd
First public version of data package

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# idealista18

<!-- badges: start -->
<!-- badges: end -->

The package idealista18 includes real estate data for three major markets in Spain (Barcelona, Madrid, and Valencia) for the year 2018.

## Installation

<!--
You can install the released version of idealista18 from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("idealista18")
```
-->

You can install the package from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("paezha/idealista18")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(idealista18)
## basic example code
```

### 2021-06-22 R/idealista18.R

@@ -1,0 +1,371 @@
+#' idealista18: A data package with property values in Spain from the Idealista data base.
+#'
+#' This package is an exercise in package creation using
+#' R studio. The package includes a sample function and
+#' a sample dataset with their respective documentation.
+#'
+#' @docType package
+#' @name idealista18
+#' @author Antonio Paez, School of Earth, Environment and Society, McMaster University \email{paezha@@mcmaster.ca}
+#' @references \url{https://github.com/paezha/Reproducible-Research-Workflow}
+NULL
+
+#' Sale prices of residential property in Barcelona in 2018.
+#'
+#' A dataset containing information about 84,280 real estate transactions
+#' in Barcelona during 2018.
+#'
+#' @format An object of class `sf` (inherits from `data.frame`) with 84,280 rows and 57 variables:
+#' \describe{
+#'   \item{ASSETID}{ID of the asset}
+#'   \item{PERIOD}{Date YYYYMM ((it corresponds to Ads in the list in the previous 3 months)}
+#'   \item{PRICE}{Price at idealista (euros??)}
+#'   \item{UNITPRICE}{NEEDS TO BE DEFINED (euros??)}
+#'   \item{ADTYPOLOGYID}{Asset Typology (home, chalet)}
+#'   \item{ADOPERATIONID}{Asset Operation (rent, sales)}
+#'   \item{H3INDEX}{Uber's H3 index with level 10 of resolution}
+#'   \item{LEVEL6NAME}{Municipality Name}
+#'   \item{CONSTRUCTEDAREA}{Home area in sq.m}
+#'   \item{ROOMNUMBER}{Number of rooms (total??)}
+#'   \item{BATHNUMBER}{Number of bathrooms}
+#'   \item{HASTERRACE}{Dummy variable for terrace (takes 1 if there is a terrace, 0 otherwise)}
+#'   \item{HASLIFT}{Dummy variable for lift (takes 1 if there is a lift in the building, 0 otherwise)}
+#'   \item{HASAIRCONDITIONING}{Dummy variable for AA (takes 1 if there is a AA, 0 otherwise)}
[…]

### 2021-06-22 R/perf_reg_metrics.R

@@ -1,0 +1,63 @@
+#' Compute various performance metrics for a regression model
+#'
+#' @param truth A simple features data frame with the data for model estimation.
+#' @param predictions The proportion of the sample for training, a number between 0 and 1.
+#' @param model_label A label for the model
+#' @param split_label A label for the data split used
+#' @return A list with various performance metrics.
+#' @import Metrics
+#' @importFrom dplyr %>%
+#' @importFrom stats median
+#' @export
+#' @examples
+#'
+#' library(idealista18)
+#' data(Valencia_Sale)
+#' mod <- lm(LOGPRICE ~ CONSTRUCTEDAREA, data = Valencia_Sale)
+#' perf_reg_metrics(truth = Valencia_Sale$LOGPRICE, predictions = mod$fitted.values)
+
+# #' @import dplyr Metrics
+
+perf_reg_metrics <- function(truth, predictions, model_label = "", split_label = ""){
+  #COMPUTES PERFORMANCE METRICS FOR REGRESSION MODELS
+  ae <- Metrics::ae(truth, predictions)
+  ape <- Metrics::ape(truth, predictions)
+
+
+  hit_ratio <- data.frame(ape) %>%
+    dplyr::summarise(n = dplyr::n(),
+                     hit_ratio_5 = sum(. < 0.05)/n,
+                     hit_ratio_10 = sum(. < 0.1)/n,
+                     hit_ratio_20 = sum(. < 0.2)/n,
+                     hit_ratio_50 = sum(. < 0.5)/n)
+
+  metrics <- data.frame(estimator = model_label,
+                        split = split_label,
+                        mae = Metrics::mae(
+                          truth,
+                          predictions),
+                        mdae = median(ae),
+                        rmse = Metrics::rmse(
+                          truth,
+                          predictions),
+                        mape = Metrics::mape(
+                          truth,
+                          predictions),
+                        medape = median(ape),
+                        bias = Metrics::bias(
+                          truth,
+                          predictions),
+                        pc_bias = Metrics::percent_bias(
+                          truth,
+                          predictions))
+
+  all.metrics <- cbind(hit_ratio, metrics)  %>%
+    dplyr::select(estimator, split, n,
+                  mae, mdae, rmse,
+                  mape, medape, bias, pc_bias,
+                  hit_ratio_5, hit_ratio_10, hit_ratio_20, hit_ratio_50
+    )
+
+
+  return(all.metrics)
+}

### 2021-06-22 data-in-brief/dib-idealista18.tex
Started Data in Brief paper

%% This is file `dib-template.tex',
%%
%% Copyright 2020 Elsevier Ltd
%%
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for Elsevier's document class `elsarticle'
%% with harvard style bibliographic references
%%
%% $Id: dib-template.tex 185 2020-08-07 09:06:08Z rishi $
%%
%% Use the option review to obtain double line spacing
%\documentclass[times,review,preprint]{elsarticle}

%% Use the options `final' to obtain the final layout
%% Use longtitle option to break abstract to multiple pages if overfull.
%% For Review pdf (With double line spacing)
%\documentclass[times,review]{elsarticle}
%% For abstracts longer than one page.
%\documentclass[times,review,longtitle]{elsarticle}
%% For Review pdf without preprint line
%\documentclass[times,review,nopreprintline]{elsarticle}
%% Final pdf
\documentclass[times,final]{elsarticle}
%%
%\documentclass[times,final,longtitle]{elsarticle}
%%

%%
%% Stylefile to load DIB template
\usepackage{dib}
\usepackage{framed,multirow}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{latexsym}

%% For line numbers
%\usepackage[switch]{lineno}

% Following three lines are needed for this document.
% If you are not loading colors or url, then these are
% not required.
\usepackage{url}
\usepackage{xcolor}
\definecolor{newcolor}{rgb}{.8,.349,.1}

%%
\usepackage{longtable}
\usepackage[colorlinks]{hyperref}

\journal{Data in Brief}

\begin{document}

\verso{Given-name Surname \textit{etal}}

\begin{frontmatter}

\dochead{Data Article}
%The article title must include the word 'data' or 'dataset'.
%Please avoid the use of acronyms and abbreviations where possible.
%For co-submission authors, the title should be unique,
%i.e. not the same as your research paper.
%A maximum of 250 characters is allowed.
\title{idealista18: A data package with real estate information in three major Spanish markets from the Idealista database\tnoteref{tnote1}}%
\tnotetext[tnote1]{This is an example for title footnote coding.}
%Tip: here are a few examples of recent suitable article titles - these are short and clear:
%%Adolescent Rat Social Play: Amygdalar Proteomic and Transcriptomic Data
%%Execution Data Logs of a Supercomputer Workload Over its Extended Lifetime
%%Calgary Preschool Magnetic Resonance Imaging (MRI) Dataset]

%%Authors
\author[1]{David \snm{Rey Blanco}}
\author[1]{Pelayo \snm{González Arbues}\fnref{fn1}}
\fntext[fn1]{This is author footnote for second author.}
\author[2]{Fernando \snm{López Hernández}}
\author[3]{Antonio \snm{Páez}\corref{cor1}}
%% Fourth author's email
\ead{paezha@mcmaster.ca}
\cortext[cor1]{Corresponding author:
  Tel.: +1-905-525-9140 ext 26099}

%%Affiliations
\address[1]{Affiliation 1, Address, City and Postal Code, Country}
\address[2]{Affiliation 2, Address, City and Postal Code, Country}
\address[3]{School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada}

%\received{1 May 2013}
%\finalform{10 May 2013}
%\accepted{13 May 2013}
%\availableonline{15 May 2013}
%\communicated{S. Sarkar}

\begin{abstract}
%%%%
Please Type your abstract here.

\noindent[The Abstract should describe the data collection process, the analysis
performed, the data, and their reuse potential. It should not provide
conclusions or interpretive insights. If your article is being
submitted via another Elsevier journal as a co-submission, please cite
this research article in the abstract.

\noindent\textbf{Tip:} do not use words such as
`study', `results', and `conclusions' because a data article should be
describing your data only.  Min 100 words - Max 500 words.]
%%%%
\end{abstract}

\begin{keyword}
%% Keywords
%[Include 4-8 keywords (or phrases) to facilitate others finding your
%article online.
%\noindent\textbf{Tip:} Try Google Scholar to find which terms are most common in your
%field. In biomedical fields, MeSH terms are a good 'common vocabulary'
%to draw from]
\KWD Property values\sep Spain\sep Spatial analysis\sep Machine learning\sep Hedonic price analysis
\end{keyword}

\end{frontmatter}

%% For linenumbers
%\linenumbers

{\fontsize{7.5pt}{9pt}\selectfont
%%%
\noindent\textbf{Specifications Table}

Every section of this table is mandatory.
Please enter information in the right-hand column and remove all the instructions
\begin{longtable}{|p{33mm}|p{94mm}|}
\hline
\endhead
\hline
\endfoot
Subject                & Geography, Economics\\
\hline
Specific subject area  & Spatial analysis, machine learning, hedonic price analysis\\
\hline
Type of data           & Table\newline\\
%\clearpage
How data were acquired & [State how the data were acquired: E.g. Microscope,
                         SEM, NMR, mass spectrometry, survey* etc.\newline
                         Instruments: E.g. hardware, software, program\newline
                         Make and model and of the instruments used:\newline

                         {\fontsize{7pt}{8pt}\selectfont
                         *\,if you conducted a survey you must submit a copy of the
                         survey(s) used (either provide these as supplementary material
                         file or provide a URL link to the survey
                         in this section of the table).
                         If the survey is not written in English,
                         please provide an English-language translation.}]\\
\hline
Data format            & Spatially masked\\
\hline
Parameters for
data\newline
collection             & [Provide a brief description of which conditions were considered
                         for data collection. Max 400 characters]\\

\hline
Description of
data\newline
collection             & [Provide a brief description of how these data were collected.
                         Max 600 characters]\\
\hline
Data source location   & Institution: Idealista\newline
                         City/Town/Region: Madrid\newline
                         Country: Spain\newline
                         Latitude and longitude (and GPS coordinates, if possible) for collected samples/data:\newline

                         If you are describing secondary data, you are required to provide a list of
                         the primary data sources used in the section.\newline

                         Primary data sources:  ]\\
\hline
\hypertarget{target1}
{Data accessibility}   & Repository name: GitHub\newline
                         Direct URL to data: https://github.com/paezha/idealista18\newline
                         \\
\hline
Related
research\newline
article                & D. Rey Blanco, P. González Arbues, F. López Hernández, A. Páez, Using machine learning to identify spatial market segments: A reproducible study of major Spanish markets, Comput Environ Urban Syst. In Press.\newline
\end{longtable}
}
%%%

\section*{Value of the Data}

[Provide 3-6 bullet points explaining why these data are of value to the scientific community.
Bullet points 1-3 must specifically answer the questions next to the bullet point,
but do not include the question itself in your answer. You may
provide up to three additional bullet points to outline the value of these data.
Please keep points brief, with ideally no more than 400 characters for each point.]

\begin{itemize}
\itemsep=0pt
\parsep=0pt
\item Your first bullet point must explain why these data are useful or important?
\item Your second bullet point must explain who can benefit from these data?
\item Your third point bullet must explain how these data might be used/reused for
further insights and/or development of experiments.
\item In the next three points you may like to explain how these data could
potentially make an impact on society and highlight any other additional value of these data.
\item ....
\end{itemize}

\section*{Data Description}

\noindent [Individually describe each data file (i.e. figure 1, figure 2, table
1, dataset, raw data, supplementary data, etc.) that are included in
this article. Please make sure you refer to every data file and provide
a clear description for each - do not simply list them. No insight,
interpretation, background or conclusions should be included in this
section. Please include legends with any tables, figures or graphs.

\noindent\textbf{Tip:} do not forget to describe any supplementary data files.]

\section*{Experimental Design, Materials and Methods}

\noindent [Offer a complete description of the experimental design and methods
used to acquire these data. Please provide any programs or code files
used for filtering and analyzing these data. It is very important that
this section is as comprehensive as possible. If you are submitting via
another Elsevier journal (a co-submission) you are encouraged to
provide more detail than in your accompanying research article. There
is no character limit for this section; however, no insight,
interpretation, or background should be included in this section.

\noindent\textbf{Tip:} do not describe your data (figures, tables, etc.) in this section,
do this in the Data Description section above.]

\section*{Ethics Statement}
\noindent [Please refer to the journal's
\href{https://www.elsevier.com/journals/data-in-brief/2352-3409/guide-for-authors}{Guide for Authors}
for more information on
the ethical requirements for publication in Data in Brief. In addition
to these requirements:

\noindent\textbf{If the work involved the use of human subjects:}
please include a statement here confirming that informed consent was
obtained for experimentation with human subjects;

\noindent\textbf{If the work involved animal experiments:} please
include a statement confirming that all experiments comply with
the \href{https://www.nc3rs.org.uk/arrive-guidelines}{ARRIVE\ guidelines} and were be carried out in accordance with the
U.K. Animals (Scientific Procedures) Act, 1986 and associated
guidelines, \href{https://ec.europa.eu/environment/chemicals/lab_animals/legislation_en.htm}{EU Directive 2010/63/EU for animal experiments}, or the
National Institutes of Health guide for the care and use of Laboratory
animals (NIH Publications No. 8023, revised 1978)]

\section*{Acknowledgments}
Acknowledgments should be inserted at the end of the paper, before the
references, not as a footnote to the title. Use the unnumbered
Acknowledgements Head style for the Acknowledgments heading.

\section*{Declaration of Competing Interest}

\noindent [All authors are required to report the following information:
\begin{enumerate}
\item[(1)] All third-party financial support for the work this article;

\item[(2)] All financial relationships with any entity that could be
viewed as relevant to data described in this manuscript;

\item[(3)] All sources of revenue with relevance to this work where
payments have been made to authors, or their institutions on their
behalf, within the 36 months prior to submission;

\item[(4)] Any other interactions with the sponsor, outside of the
submitted work;

\item[(5)] Any relevant patents or copyrights (planned, pending or
issued);

\item[(6)] Any other relationships or affiliations that may be
perceived by readers to have influenced, or give the appearance of
potentially influencing, what has been written in this article.
\end{enumerate}
As a general guideline, it is usually better to
disclose a relationship than not. This information will be acknowledged
at publication in the manuscript. If there is no known competing
financial interests or personal relationships that could have appeared
to influence the work reported in this paper, please include this
statement.]

\vskip12pt\noindent
The authors declare that they have no known competing
financial interests or personal relationships which have, or could be
perceived to have, influenced the work reported in this article.

\vskip12pt\noindent
[If there are financial interests/personal relationships which may be
considered as potential competing interests, please declare them here.]

\subsection*{Note}
\label{sec1}
Any instructions relevant to the \verb+elsarticle.cls+ are applicable
here as well. See the online instruction available on:
\makeatletter
\if@twocolumn
\begin{verbatim}
 http://support.river-valley.com/wiki/
 index.php?title=Elsarticle.cls
\end{verbatim}
\else
\begin{verbatim}
 http://support.river-valley.com/wiki/index.php?title=Elsarticle.cls
\end{verbatim}
\fi

\section*{References}

\noindent [References are limited (approx. 15) and excessive self-citation is not
allowed. \textbf{If your data article is co-submitted via another Elsevier
journal, please cite your associated research article here.}

\noindent\textbf{Reference style:}
Text:?Indicate references by number(s) in square brackets in line with
the text. The actual authors can be referred to, but the reference
number(s) must always be given.?

\noindent Example: '..... as demonstrated [3,6]. Barnaby and Jones [8] obtained a different result ....'?

\noindent [Use \verb+\cite+ command to cite a reference list item in text.

\noindent These are examples for reference citations \cite{1}.
\cite{2}.
\cite{4}.]

%% Numbered
%%If
\bibliographystyle{model1-num-names}
\bibliography{refs.bib}

\end{document}

%%

### 2021-06-22 data-in-brief/dib-template.tex
Started Data in Brief paper

--- 2021-06-22 data-in-brief/dib-idealista18.tex
@@
 %A maximum of 250 characters is allowed.
-\title{idealista18: A data package with real estate information in three major Spanish markets from the Idealista database\tnoteref{tnote1}}%
+\title{Article Title\tnoteref{tnote1}}%
 \tnotetext[tnote1]{This is an example for title footnote coding.}
@@
 %%Authors
-\author[1]{David \snm{Rey Blanco}}
-\author[1]{Pelayo \snm{González Arbues}\fnref{fn1}}
+\author[1]{Given-name1 \snm{Surname1}\corref{cor1}}
+\cortext[cor1]{Corresponding author:
+  Tel.: +0-000-000-0000;
+  fax: +0-000-000-0000;}
+\author[1]{Given-name2 \snm{Surname2}\fnref{fn1}}
 \fntext[fn1]{This is author footnote for second author.}
-\author[2]{Fernando \snm{López Hernández}}
-\author[3]{Antonio \snm{Páez}\corref{cor1}}
-%% Fourth author's email
-\ead{paezha@mcmaster.ca}
-\cortext[cor1]{Corresponding author:
-  Tel.: +1-905-525-9140 ext 26099}
+\author[2]{Given-name3 \snm{Surname3}}
+%% Third author's email
+\ead{author3@author.com}
+\author[2]{Given-name4 \snm{Surname4}}
 
@@
 \address[2]{Affiliation 2, Address, City and Postal Code, Country}
-\address[3]{School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada}
 
@@
 %to draw from]
-\KWD Property values\sep Spain\sep Spatial analysis\sep Machine learning\sep Hedonic price analysis
+\KWD Keyword1\sep Keyword2\sep Keyword3
 \end{keyword}
@@
 %\linenumbers
+\section*{Data in Brief Article Template}
+%% main text
+\noindent \href{https://www.journals.elsevier.com/data-in-brief/about-data-in-brief/data-in-brief-faq}%
+{\textit{Data in Brief}}
+is an open access journal that publishes data articles.
+Please note:
+\begin{itemize}
+\item A data article is different to a research article, so it is
+important to \textbf{use the template} below to prepare your manuscript for Data
+in Brief.
+\item A data article should \textbf{simply describe data} without providing
+conclusions or interpretive insights.
+\item Before you start writing your data article you should read the
+guidance on
+\href{https://www.journals.elsevier.com/data-in-brief/policies-and-guidelines/what-data-are-suitable-for-data-in-brief}%
+{What Data are Suitable for Data in Brief}.
+\item It is mandatory that Data in Brief authors share their research data:
+\begin{itemize}
+\item If you have \textbf{raw data} (also referred to as primary, source or
+unprocessed data) relating to any charts, graphs or figures in the
+manuscript, these data must be publicly available, either with the data
+article (e.g. as a supplementary file) or hosted on a trusted data
+repository.
+
+\item If you are describing \textbf{secondary data} you are required to provide
+a list of the primary data sources used \underline{and} to make the full secondary
+dataset publicly available, either with the data article (e.g. as a
+supplementary file) or hosted on a trusted data repository.
+
+\item Although we allow supplementary files, it is preferred that
+authors deposit their data in a trusted data repository ($>$70\% of
+Data in Brief authors now do this). See our?
+\href{https://www.elsevier.com/authors/author-resources/research-data/data-base-linking#repositories}%
+{list of supported data repositories}.
+
+\item For data that, for ethical reasons, require access controls a
+mechanism must be provided so that our Editors and reviewers may access
+these data without revealing their identities to authors (more
+information is provided in the template \hyperlink{target1}{below}).
+\end{itemize}
+\end{itemize}
+
+Have you any questions? See a list of frequently asked questions
+\href{https://www.journals.elsevier.com/data-in-brief/about-data-in-brief/data-in-brief-faq}{here},
+or email our Managing Editors:
+\href{mailto:dib-me@elsevier.com}{dib-me@elsevier.com}. This
+step-by-step?
+\href{https://www.journals.elsevier.com/data-in-brief/about-data-in-brief/how-to-submit-your-research-data-article-data-in-brief}%
+{video}?guide will also tell you how to complete the
+template correctly to maximise your chances of acceptance.
+
+\vskip6pt
+\noindent Authors can submit to Data in Brief in two ways:
+
+\begin{enumerate}
+\item[\bf(1)] \textbf{If you are submitting your data article directly to Data in
+Brief, you can now skip the next section and complete the}
+\hyperlink{target2}{\textbf{Data Article template}}.
+
+\item[\bf(2)] \textbf{If you are submitting your data article to Data in Brief via
+another Elsevier journal as a co-submission (i.e. with a Research
+Article), please read the} \hyperlink{target3}{\textbf{Co-submission Instructions}}
+\textbf{on the next page
+before completing the} \hyperlink{target2}{\textbf{Data Article template}}.
+
+\end{enumerate}
+
+\hypertarget{target3}{}
+\section*{Co-submission Instructions}
+A co-submission to~\textit{Data in Brief}~is done at the same time that you
+submit (or resubmit, after revision) a research article to another
+Elsevier journal. For co-submissions you therefore submit your
+\textit{Data in Brief} data article manuscript via the other journal's
+submission system and \underline{not} directly to
+\textit{Data in Brief} itself.
+
+\vskip6pt\noindent
+The other Elsevier journal's Guide for Authors will state if a
+co-submission is offered by that journal, and any revision letter/email
+you receive from a participating journal will contain an offer to
+submit a data article to \textit{Data in Brief}.
+
+\vskip6pt\noindent
+\textbf{To complete a co-submission you will need to zip your
+~\textit{Data in Brief}~
+manuscript file and all other files relevant to the
+~\textit{Data in Brief}~
+submission (including any supplementary data files) into a single .zip
+file, and upload this as a "Data in Brief"-labelled item in the other
+journal's submission system when you submit manuscript to that journal.
+The .zip file will then be automatically transferred to
+~\textit{Data in Brief}~
+when your research article is accepted for publication in the other
+journal, and when published your original research article and data
+article will link to each other on ScienceDirect.}
+
+\vskip6pt\noindent
+\textbf{As~\textit{Data in Brief}~is open access, a moderate article publication charge
+(APC) fee is payable on publication. For more information about the
+APC, please see
+\href{https://www.elsevier.com/journals/data-in-brief/2352-3409/open-access-journal}%
+{here}.}
+
+\vskip6pt\noindent
+\textit{Data in Brief} \underline{requires} that authors share their research data. This can
+be done by submitting it with the data article (e.g. as a supplementary
+file) or by hosting on a trusted data repository (the latter is
+preferred). Failure to do this will delay publication of your
+co-submission.
+
+\vskip6pt\noindent
+\textbf{If you have any questions, please contact:
+\href{mailto:DIB@Elsevier.com}{DIB@Elsevier.com}}
+
+\vskip6pt\noindent
+Please note, authors should not republish the same data presented in
+their original research article in a \textit{Data in Brief} co-submission, as
+this could constitute duplicate publication; however, \textit{Data in Brief}
+welcomes the publication of any data article that fulfils one or more
+of the following criteria:
+
+\checkmark A description of the supplementary data that would
+previously have been hosted as supplementary electronic files alongside
+your original research article.*
+
+\checkmark A description of the full dataset or additional information
+that will aid reuse of the data.
+
+\checkmark A detailed description of the raw data relating to the
+charts, graphs or figures in your companion research article, if making
+these data available will substantially enhance reproducibility and/or
+reanalysis of the data.
+
+\checkmark Any negative datasets or data from intermediate experiments
+related to your research.
+
+\textsf{X} Review articles or supplemental files from a review article
+are not considered original data and are typically unsuitable for Data
+in Brief.
+
+\vskip12pt\noindent
+* If describing supplementary data that you previously planned to
+publish as supplem
+entary electronic files hosted alongside the original
+research article, it is requested that you either\break
+deposit these in a
+repository (preferred) or submit these to \textit{Data in Brief} alongside the
+data article. \textbf{They should not be published as supplementary files with
+your research article in the other journal}.
+
+\clearpage
+\hypertarget{target2}{}
+\section*{Data Article template}
+\noindent
+Please fill in the template below. All sections are mandatory unless
+otherwise indicated. Please read all instructions in [square brackets]
+carefully and ensure that you delete all instruction text (including
+the questions) from the template before submitting your article.
+
+\vskip6pt\noindent
+Reminder: A data article simply describes data and should not provide
+conclusions or interpretive insights, so \textbf{avoid} using words such as
+`study', `results' and `conclusions'.
+
+\vskip6pt\noindent
+We would welcome feedback on this template and how it might be
+improved. To provide anonymous feedback via a very short survey, please
+click \href{https://forms.office.com/Pages/ResponsePage.aspx?id=P-50kiWUCUGif5-xXBBnXTeXkbO343VFrbpYVBvxdZtUM05UVjIwM0U4WlRKUldCOTNMRUQwOVRHTy4u}%
+{here}.
+
+\vskip6pt\noindent
+{\small\textbf{\textit{Please delete this line and everything above it before submitting your
+article, in addition to anything in [square brackets] below, including
+in the Specifications Table}}\vskip6pt\hrule\vskip12pt}
 
@@
 \endfoot
-Subject                & Geography, Economics\\
+Subject                & [Please select one CATEGORY for your manuscript from the list
+					               available at:\break
+                         \href{https://www.elsevier.com/__data/assets/excel_doc/0012/736977/%
+                               DIB-categories.xlsx}{DIB categories}.]\\
 \hline
-Specific subject area  & Spatial analysis, machine learning, hedonic price analysis\\
+Specific subject area  & [Briefly describe the narrower subject area. Max 150 characters]\\
 \hline
-Type of data           & Table\newline\\
+Type of data           & [List the type(s) of data this article describes.
+                         Simply delete from this list as appropriate:]
+                         Table\newline
+                         Image\newline
+                         Chart\newline
+                         Graph\newline
+                         Figure\newline
+                         [Any other type not listed- please specify]\\
 %\clearpage
@@
 \hline
-Data format            & Spatially masked\\
+Data format            & [List your data format(s). Note, unless you are describing secondary data,
+                         all raw data must be provided (either with this data article or linked to a repository).
+                         Simply delete from this list as appropriate:]\newline
+                         Raw\newline
+                         Analyzed\newline
+                         Filtered\newline
+                         [Any other format not listed- please specify]\\
 \hline
@@
 \hline
-Data source location   & Institution: Idealista\newline
-                         City/Town/Region: Madrid\newline
-                         Country: Spain\newline
+Data source location   & [Fill in the information available, and delete from this list as appropriate:\newline
+
+                         Institution:\newline
+                         City/Town/Region:\newline
+                         Country:\newline
                          Latitude and longitude (and GPS coordinates, if possible) for collected samples/data:\newline
@@
 \hypertarget{target1}
-{Data accessibility}   & Repository name: GitHub\newline
-                         Direct URL to data: https://github.com/paezha/idealista18\newline
-                         \\
+{Data accessibility}   & [State here if the data are either hosted `With the article' or on a public repository.
+                         In the interests of openly sharing data we recommend hosting your data in a
+                         trusted repository ($>$70\% of Data in Brief authors now use a data repository).
+                         See our \href{https://www.elsevier.com/authors/author-resources/research-data/data-base-linking#repositories}{list of supported data repositories}.
+                         We suggest \href{https://data.mendeley.com/}{Mendeley Data} if you do not have a trusted repository.\newline
+
+                         Please delete or complete as appropriate, either:]\newline
+
+                         With the article\newline
+
+                         [Or, if in a public repository:]\newline
+
+                         Repository name: [Name repository]\newline
+                         Data identification number: [provide number]\newline
+                         Direct URL to data: [e.g. https://www.data.edu.com - please note,\newline
+                         this URL should be working at the time of submission]\newline
+
+                         [\textbf{In addition, for data with access controls only:} For data that,
+                         for ethical reasons (i.e. human patient data),
+                         require access controls please describe
+                         how readers can request access these data and provide a link to any
+                         Data Use Agreement (DUA) or upload a copy as a supplementary file.]\newline
+
+                         Instructions for accessing these data:\newline
+
+                         [\textit{Important: if your data have access controls a mechanism must also be
+                         provided so that our Editors and reviewers may access these data
+                         without revealing their identities to authors, please include
+                         these instructions with your submission. Please contact the Managing
+                         Editors (DIB-ME@DIB.com) if you have any questions.}]\\
 \hline
@@
 research\newline
-article                & D.
-Rey Blanco, P.
-González Arbues, F.
-López Hernández, A.
-Páez, Using machine learning to identify spatial market segments: A reproducible study of major Spanish markets, Comput Environ Urban Syst.
-In Press.\newline
+article                & [If your data article is related to a research article - \textbf{especially
+                         if it is a co-submission} - please cite your associated research
+                         article here. Authors should only list \textbf{one article}.\newline
+
+                         Authors' names\newline
+                         Title\newline
+                         Journal\newline
+                         DOI: \textbf{OR} for co-submission manuscripts `In Press'\newline
+
+                         \textbf{For example, for a direct submission:}\newline
+
+                         J. van der Geer, J.A.J. Hanraads, R.A. Lupton, The art of writing a scientific article,
+                         J. Sci. Commun. 163 (2010) 51-59. https://doi.org/10.1016/j.Sc.2010.00372\newline
+
+                         \textbf{Or, for a co-submission (when your related research article has not yet published):}\newline
+
+                         J. van der Geer, J.A.J. Hanraads, R.A. Lupton, The art of writing a
+                         scientific article, J. Sci. Commun. In Press.\newline
+
+                         \textbf{Or, if your data article is not directly related to a research article,
+                         please delete this last row of the table.}]
 \end{longtable}
@@
 
+\subsection*{Reference list using Bib\TeX database file}
+\noindent [If Bib\TeX database file is used for reference data please use
+\begin{verbatim}
+\bibliographystyle{model1-num-names}
+\bibliography{<BibTeX file name>}
+\end{verbatim}
+and use bibtex command to generate list of references.]
+
 %% Numbered
@@
 \bibliographystyle{model1-num-names}
-\bibliography{refs.bib}
+\bibliography{refs}
+
+\subsection*{Reference list using bibliography envrironment}
+
+\noindent [List:?Number the references (numbers in square brackets) in
+the list in the order in which they appear in the text.?
+
+\noindent Examples:?
+
+{\small
+\begin{verbatim}
+\begin{thebibliography}{0}
+\bibitem{1} J. van der Geer, J.A.J. Hanraads, R.A. Lupton, The art of
+writing a scientific article, J. Sci. Commun. 163 (2010) 51-59.
+https://doi.org/10.1016/j.Sc.2010.00372.
+\bibitem{2} Van der Geer, J., Hanraads, J.A.J., Lupton, R.A., 2018. The
+art of writing a scientific article. Heliyon. 19, e00205.
+https://doi.org/10.1016/j.heliyon.2018.e00205.
+\bibitem{3} W. Strunk Jr., E.B. White, The Elements of Style, fourth ed.,
+Longman, New York, 2000.?
+\bibitem{4} G.R. Mettam, L.B. Adams, How to prepare an electronic
+version of your article, in: B.S. Jones, R.Z. Smith (Eds.),
+Introduction to the Electronic Age, E-Publishing Inc., New York, 2009,
+pp. 281-304.
+\bibitem{5} Cancer Research UK, Cancer statistics reports for the UK.
+http://www.cancerresearchuk.org/aboutcancer/statistics/cancerstatsreport/,
+2003 (accessed 13 March 2003).
+\bibitem{6} [dataset] M. Oguro, S. Imahiro, S. Saito, T. Nakashizuka,
+Mortality data for Japanese oak wilt disease and surrounding forest
+compositions, Mendeley Data, v1, 2015.
+https://doi.org/10.17632/xwj98nb39r.1.
+\end{thebibliography}
+\end{verbatim}
+}
+
+\noindent Example:]
+\vspace*{-12pt}
+
+\begin{thebibliography}{0}
+\item[] \textbf{Reference to a journal publication:}
+\bibitem{1} J. van der Geer, J.A.J. Hanraads, R.A. Lupton, The art of
+writing a scientific article, J. Sci. Commun. 163 (2010) 51-59.
+https://doi.org/10.1016/j.Sc.2010.00372.
+
+\textbf{Reference to a journal publication with an article number:}
+\bibitem{2} Van der Geer, J., Hanraads, J.A.J., Lupton, R.A., 2018. The
+art of writing a scientific article. Heliyon. 19, e00205.
+https://doi.org/10.1016/j.heliyon.2018.e00205.
+
+\textbf{Reference to a book:}
+\bibitem{3} W. Strunk Jr., E.B. White, The Elements of Style, fourth ed., Longman, New York, 2000.?
+
+\textbf{Reference to a chapter in an edited book:}
+\bibitem{4} G.R. Mettam, L.B. Adams, How to prepare an electronic
+version of your article, in: B.S. Jones, R.Z. Smith (Eds.),
+Introduction to the Electronic Age, E-Publishing Inc., New York, 2009,
+pp. 281-304.
+
+\textbf{Reference to a website:}
+\bibitem{5} Cancer Research UK, Cancer statistics reports for the UK.\newline
+http://www.cancerresearchuk.org/aboutcancer/statistics/cancerstatsreport/, 2003 (accessed 13 March 2003).
+
+\textbf{Reference to a dataset:}
+\bibitem{6} [dataset] M. Oguro, S. Imahiro, S. Saito, T. Nakashizuka,
+Mortality data for Japanese oak wilt disease and surrounding forest
+compositions, Mendeley Data, v1, 2015.
+https://doi.org/10.17632/xwj98nb39r.1.]
+\end{thebibliography}
 

### 2021-06-23 R/idealista18.R
Simplified datasets (removed redundant or unused fields) Added field comments

@@ -1,8 +1,12 @@
-#' idealista18: A data package with property values in Spain from the Idealista data base.
+#' idealista18: A data package with real estate litings in Spain from the idealista data base.
 #'
-#' This package is an exercise in package creation using
-#' R studio. The package includes a sample function and
-#' a sample dataset with their respective documentation.
+#' This package contains four quarterly extractions with the set of apartments listed on idealista during 2018,
+#' the package contains three data items for the three spanish cities of Madrid, Barcelona and Valencia:
+#'
+#' * Polygons for each city neighborhood
+#' * Listing ads (referred also as assets)
+#' * Points of Interest for each city
+#'
 #'
 #' @docType package
 #' @name idealista18
@@ -12,21 +16,19 @@ NULL
 
 #' Sale prices of residential property in Barcelona in 2018.
 #'
-#' A dataset containing information about 84,280 real estate transactions
+#' A dataset containing information about 84,280 real estate ads on idealista
 #' in Barcelona during 2018.
 #'
 #' @format An object of class `sf` (inherits from `data.frame`) with 84,280 rows and 57 variables:
 #' \describe{
-#'   \item{ASSETID}{ID of the asset}
-#'   \item{PERIOD}{Date YYYYMM ((it corresponds to Ads in the list in the previous 3 months)}
-#'   \item{PRICE}{Price at idealista (euros??)}
-#'   \item{UNITPRICE}{NEEDS TO BE DEFINED (euros??)}
+#'   \item{ASSETID}{Unique identifier of the asset}
+#'   \item{PERIOD}{Date YYYYMM, indicates the quarter when the ad was extracted we used YYYY03 for the 1st Q, YYYY06 the 2nd, YYYY09 for the 3rd and YYYY12 for the 4th}
+#'   \item{PRICE}{Asking price for the ad at idealista expressed in euros}
+#'   \item{UNITPRICE}{Price in euros per square meter}
 #'   \item{ADTYPOLOGYID}{Asset Typology (home, chalet)}
 #'   \item{ADOPERATIONID}{Asset Operation (rent, sales)}
-#'   \item{H3INDEX}{Uber's H3 index with level 10 of resolution}
-#'   \item{LEVEL6NAME}{Municipality Name}
[…]

### 2021-07-04 data-in-brief/dib-idealista18.tex
A first version of the abstract section Value of data section

@@
 %%Affiliations
-\address[1]{Affiliation 1, Address, City and Postal Code, Country}
-\address[2]{Affiliation 2, Address, City and Postal Code, Country}
+\address[1]{idealista, Plaza de las Cortes 5, 28014 Madrid, Spain}
+\address[2]{Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia (Spain)}
 \address[3]{School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada}
@@
 %%%%
-Please Type your abstract here.
+This dataset contains three items for each of the three major cities in Spain: Madrid, Barcelona and Valencia.
+First real estate listings published on idealista portal in 2018.
+All listings have been enriched with cadastral information (i.e.
+building year of construction, built quality materials grade) plus some geographical features such as
+distance to relevant city areas and the coordinates themselves.
+To comply with european personal protection laws, we have processeed some sensible variables yet preserving their spatial properties.
+The second and third items are a list of points of interest and the administrative boundaries for each municipality.
+This dataset is suitable to house market analysis, hedonic house price models and other spatial research related with real estate markets.
 
-\noindent[The Abstract should describe the data collection process, the analysis
-performed, the data, and their reuse potential. It should not provide
-conclusions or interpretive insights. If your article is being
-submitted via another Elsevier journal as a co-submission, please cite
-this research article in the abstract.
+\textbf{<falta cita/referencia al artículo>}
 
-\noindent\textbf{Tip:} do not use words such as
-`study', `results', and `conclusions' because a data article should be
-describing your data only.  Min 100 words - Max 500 words.]
+%\noindent[The Abstract should describe the data collection process, the analysis
+%performed, the data, and their reuse potential. It should not provide
+%conclusions or interpretive insights. If your article is being
+%submitted via another Elsevier journal as a co-submission, please cite
+%this research article in the abstract.
+
+%\noindent\textbf{Tip:} do not use words such as
+%`study', `results', and `conclusions' because a data article should be
+%describing your data only.  Min 100 words - Max 500 words.]
+
 %%%%
@@
 
-[Provide 3-6 bullet points explaining why these data are of value to the scientific community.
-Bullet points 1-3 must specifically answer the questions next to the bullet point,
-but do not include the question itself in your answer. You may
-provide up to three additional bullet points to outline the value of these data.
-Please keep points brief, with ideally no more than 400 characters for each point.]
-
 \begin{itemize}
@@
 \parsep=0pt
-\item Your first bullet point must explain why these data are useful or important?
-\item Your second bullet point must explain who can benefit from these data?
-\item Your third point bullet must explain how these data might be used/reused for
-further insights and/or development of experiments.
-\item In the next three points you may like to explain how these data could
-potentially make an impact on society and highlight any other additional value of these data.
-\item ....
+  \item A cleaned and enriched dataset consisting of real estate listings for three major cities in Spain, constructed to analyze the impacto of using machine learning models to identify spatial market segments to build house price hedonic models.
+  \item The dataset can be used to extend the topic of market segments automatic or semi-automatic identification.
+    \item Official boundaries combined with spatial patterns can be used to analyse the suitability of these boundaries for real estate value purposes.
+  \item The dataset can be enlarged with complementary information to develop hedonic models.
+  \item The data can be processed by quantitative analysis and statistical modeling to study the different factors that affect house prices in the three areas.
+  \item Identification of spatial patterns in the real estate scope using the geo-referenced data points. For either value or urban patterns discovery.
 \end{itemize}

### 2021-07-04 R/idealista18.R

@@ -47,7 +47,7 @@ NULL
 #'   \item{HASDOORMAN}{Dummy variable for doorman (takes 1 if there is a doorman in the building, 0 otherwise)}
 #'   \item{HASGARDEN}{Dummy variable for garden (takes 1 if there is a garden in the building, 0 otherwise)}
 #'   \item{ISDUPLEX}{Dummy variable for duplex (takes 1 if it is a duplex, 0 otherwise)}
-#'   \item{ISSTUDIO}{Dummy variable for studio (takes 1 if it is a studio, 0 otherwise)}
+#'   \item{ISSTUDIO}{Dummy variable for bachelor apartment (studio in Spanish) (takes 1 if it is a single-person apartment, 0 otherwise)}
 #'   \item{ISINTOPFLOOR}{Dummy variable indicating if the apartment is located in the top floor (takes 1 on the top floor 0 otherwise)}
 #'   \item{CONSTRUCTIONYEAR}{Construction year (source: advertiser)}
 #'   \item{FLOORCLEAN}{Indicates flat floornumber starting from the 0 value for groung floor (source: advertiser)}
@@ -109,7 +109,7 @@ NULL
 #'   \item{HASDOORMAN}{Dummy variable for doorman (takes 1 if there is a doorman in the building, 0 otherwise)}
 #'   \item{HASGARDEN}{Dummy variable for garden (takes 1 if there is a garden in the building, 0 otherwise)}
 #'   \item{ISDUPLEX}{Dummy variable for duplex (takes 1 if it is a duplex, 0 otherwise)}
-#'   \item{ISSTUDIO}{Dummy variable for studio (takes 1 if it is a studio, 0 otherwise)}
+#'   \item{ISSTUDIO}{Dummy variable for bachelor apartment (studio in Spanish) (takes 1 if it is a single-person apartment, 0 otherwise)}
 #'   \item{ISINTOPFLOOR}{Dummy variable indicating if the apartment is located in the top floor (takes 1 on the top floor 0 otherwise)}
 #'   \item{CONSTRUCTIONYEAR}{Construction year (source: advertiser)}
 #'   \item{FLOORCLEAN}{Indicates flat floornumber starting from the 0 value for groung floor (source: advertiser)}
@@ -170,7 +170,7 @@ NULL
 #'   \item{HASDOORMAN}{Dummy variable for doorman (takes 1 if there is a doorman in the building, 0 otherwise)}
[…]

### 2021-07-11 data-in-brief/dib-idealista18.tex
Added figures and changed field descriptions

@@
 \usepackage{framed,multirow}
+\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
+\usepackage{graphicx}
 
@@
 
-\verso{Given-name Surname \textit{etal}}
+\verso{D. Rey Blanco \textit{et al.}}
 
@@
 Data source location   & Institution: Idealista\newline
-                         City/Town/Region: Madrid\newline
+                         City/Town/Region: Madrid, Barcelona, Valencia\newline
                          Country: Spain\newline
@@
 
-\noindent [Individually describe each data file (i.e. figure 1, figure 2, table
-1, dataset, raw data, supplementary data, etc.) that are included in
-this article. Please make sure you refer to every data file and provide
-a clear description for each - do not simply list them. No insight,
-interpretation, background or conclusions should be included in this
-section. Please include legends with any tables, figures or graphs.
+\noindent
+% [Individually describe each data file (i.e. figure 1, figure 2, table
+% 1, dataset, raw data, supplementary data, etc.) that are included in
+% this article. Please make sure you refer to every data file and provide
+% a clear description for each - do not simply list them. No insight,
+% interpretation, background or conclusions should be included in this
+% section. Please include legends with any tables, figures or graphs.
 
-\noindent\textbf{Tip:} do not forget to describe any supplementary data files.]
+The data set is composed by three items for three major spanish cities: quarterly single family listings, neighboorhood polygons and a a set of Points of Interest from Open Street Map \cite{OpenStreetMap}.
+All spatial features, such as polygons and points, are expressed in geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.
+
+The first block integrates property published ads,the files contain the complete offering for each of the four quarters in 2018 on idealista web site \cite{idealista}.
+Idealista is the major real estate listing portal in Spain and also present in other southern european countries as Italy and Portugal.
+Each record contains all attributes the listing ad has plus a number of additional attribues from the Spanish cadastre \cite{Catastro}, the names for all these variables start with the prefix \emph{CAD}.
+Cadastral features asssignment is done by assigning the features of the nearest parcel to the coordinates \emph{LATITUDE} and \emph{LONGITUDE}.
+
+\begin{footnotesize}
+\begin{longtable}{p{40mm} c p{63mm}}
+\hline
+\hline
+% \endhead
+% \hline
+% \endfoot
+& &\\
+\textbf{Variable} & \textbf{Mesurement scale} & \textbf{Description}\\
+& &\\
+\hline
+ASSETID & Identifier & Unique identifier of the advertisement\\
+PERIOD & Nominal (Date) & Expressed as YYYYMM, indicates the quarter when the ad was extracted we used YYYY03 for the 1st Q, YYYY06 the 2nd, YYYY09 for the 3rd and YYYY12 for the 4th\\
+PRICE & Interval & Asking price for the ad at idealista expressed in euros\\
+UNITPRICE & Interval & Asking price in euros per square meter (constructed area)\\
+ADTYPOLOGYID & Nominal & Residential building type: multi-family: \emph{home}, single-familiy: \emph{chalet}\\
+ADOPERATIONID & Nominal & Operation type for the ad: \emph{sale} or \emph{rent}\\
+ROOMNUMBER & Ordinal & Number of bedrooms\\
+BATHNUMBER & Ordinal & Number of bathrooms\\
+HASTERRACE & Nominal & Dummy variable for terrace (takes 1 if there is a terrace, 0 otherwise\\
+HASLIFT & Nominal & Dummy variable for lift (takes 1 if there is a lift in the building, 0 otherwise)\\
+HASAIRCONDITIONING & Nominal & Dummy variable for AA (takes 1 if there is a AA, 0 otherwise)\\
+AMENITYID & Nominal & Indicates the amenities included (1 - no furniture, no kitchen amenities, 2 - kitchen amenities, no furniture, 3 - kitchen amenities, furniture)\\
+HASPARKINGSPACE & Nominal & Dummy variable for parking (takes 1 if parking is included in the Ad, 0 otherwise)\\
+ISPARKINGSPACEINCLUDEDINPRICE & Nominal & Dummy variable for parking (takes 1 if parking is included in the Ad, 0 otherwise)\\
+PARKINGSPACEPRICE & Interval & Price of parking space in euros\\
+HASNORTHORIENTATION & Nominal & Dummy variable for orientation (takes 1 if orientation is North in the Ad, 0 otherwise) - Important note: orientation features are not orthogonal features, a house oriented to the north can be also oriented to the east\\
+HASSOUTHORIENTATION & Nominal & Dummy variable for orientation (takes 1 if orientation is South in the Ad, 0 otherwise) - Important note: orientation features are not orthogonal features, a house oriented to the north can be also oriented to the east\\
+HASEASTORIENTATION & Nominal & Dummy variable for orientation (takes 1 if orientation is East in the Ad, 0 otherwise) - Important note: orientation features are not orthogonal features, a house oriented to the north can be also oriented to the east\\
+HASWESTORIENTATION & Nominal & Dummy variable for orientation (takes 1 if orientation is West in the Ad, 0 otherwise) - Important note: orientation features are not orthogonal features, a house oriented to the north can be also oriented to the east\\
+HASBOXROOM & Nominal & Dummy variable for boxroom (takes 1 if boxroom is included in the Ad, 0 otherwise)\\
+HASWARDROBE & Nominal & Dummy variable for wardrobe (takes 1 whether the property has wardrobes, 0 otherwise)\\
+HASSWIMMINGPOOL & Nominal & Dummy variable for swimming pool (takes 1 if swimming pool is included in the Ad, 0 otherwise)\\
+HASDOORMAN & Nominal & Dummy variable for doorman (takes 1 if there is a doorman in the building, 0 otherwise)\\
+HASGARDEN & Nominal & Dummy variable for garden (takes 1 if there is a garden in the building, 0 otherwise)\\
+ISDUPLEX & Nominal & Dummy variable for bachelor apartment (referred as studio in Spain) (takes 1 if it is a bachelor apartment, 0 otherwise)\\
+ISINTOPFLOOR & Nominal & Dummy variable indicating if the apartment is located in the top floor (takes 1 on the top floor 0 otherwise)\\
+CONSTRUCTIONYEAR & Interval & Construction year (source: advertiser)\\
+FLOORCLEAN & Ordinal & Indicates flat floornumber starting from the 0 value for ground floor (source: advertiser)\\
+FLATLOCATIONID & Nominal & Indicates the kind of views the flat has (1 - external, 2 - internal)\\
+CADCONSTRUCTIONYEAR & Interval & Construction year as of cadastral source (source: cadastre), note this figure can differ from the one given by the advertiser\\
+CADMAXBUILDINGFLOOR & Ordinal & Max building floor (source: cadastre)\\
+CADDWELLINGCOUNT & Interval & Dwelling count in the building (source: cadastre)\\
+CADASTRALQUALITYID & Ordinal & Cadastral quality (source: cadastre). 0 Best - 10 Worst\\
+BUILTTYPEID\_1 & Nominal & Dummy value for flat condition: 1 new development and 0 otherwise\\
+BUILTTYPEID\_2 & Nominal & Dummy value for flat condition: 1 second hand to be restored 0 otherwise (\emph{source: advertiser})\\
+BUILTTYPEID\_3 & Nominal & Dummy value for flat condition: 1 second hand in good condition 0 otherwise (\emph{source: advertiser})\\
+DISTANCE\_TO\_CITY\_CENTER & Interval & Distance to center of city in Km\\
+geometry & Geometry & Geometry for the elements. A point with $X,Y$ coordinates\\
+\hline
+\hline
+\end{longtable}
+\end{footnotesize}
+
+In addition to the common features each city will have a set of spatial additional features:
+
+\begin{footnotesize}
+\begin{longtable}{l p{40mm} c p{63mm}}
+\hline
+\hline
+% \endhead
+% \hline
+% \endfoot
+& & &\\
+\textbf{City} & \textbf{Variable} & \textbf{Mesurement scale} & \textbf{Description}\\
+\hline
+& & &\\
+\emph{Madrid} & DISTANCE\_TO\_CASTELLANA & Interval & Distance in km to the Paseo de la Castellana Street \\
+& & &\\
+% \hline
+\emph{Valencia} & DISTANCE\_TO\_METRO & Interval & Distance in km to the nearest subway station \\
+ & DISTANCE\_TO\_BLASCO & Interval & Distance in km to the Blasco Ibáñez Avenue \\
+& & &\\
+% \hline
+\emph{Barcelona} & DISTANCE\_TO\_DIAGONAL & Interval & Distance in km to the Diagonal Avenue \\
+& & &\\
+\hline
+\hline
+\hline
+\end{longtable}
+\end{footnotesize}
+
+The second part contains the polygons for the different neighborhoods for such cities as we can see in the figure \ref{fig:all-polygons}.
+This boundaries are not the official boundaries but the official boundaries used by idealista portal.
+In practical terms we can assume they are the same, since the website simply collapsed those areas if they are small enough in terms of number of ads.
+In the case of Madrid they just collapse four areas in two new ones.
+
+\begin{figure}[h]
+  \caption{Neighborhood boundaries for Madrid, Barcelona and Valencia. Source: own elaboration}
+  \centering
+  \includegraphics[height=4.5cm]{figures/idealista18-all-polygons}
+  \label{fig:all-polygons}
+\end{figure}
+
+We have a total of 73 neigborhoods for Barcelona, 135 for Madrid and 73 for Valencia. Each neighborhood has also two additional variables:
+
+\begin{footnotesize}
+\begin{longtable}{p{40mm} c p{63mm}}
+\hline
+\hline
+% \endhead
+% \hline
+% \endfoot
+& &\\
+\textbf{Variable} & \textbf{Mesurement scale} & \textbf{Description}\\
+\hline
+& &\\
+LOCATIONID & nominal & Unique identifier for the neighborhood \\
+LOCATIONNAME & nominal & Neighborhood name \\
+& &\\
+\hline
+\end{longtable}
+\end{footnotesize}
+% \noindent\textbf{Tip:} do not forget to describe any supplementary data files.]
 
@@
 
-\noindent [Offer a complete description of the experimental design and methods
-used to acquire these data. Please provide any programs or code files
-used for filtering and analyzing these data. It is very important that
-this section is as comprehensive as possible. If you are submitting via
-another Elsevier journal (a co-submission) you are encouraged to
-provide more detail than in your accompanying research article. There
-is no character limit for this section; however, no insight,
-interpretation, or background should be included in this section.
+The data encompassed three major cities in Spain: Madrid, Barcelona and Valencia, for each municipality we provide listing prices for multi-familiy homes on a quarterly basis.
+The files contain the complete offering for each of the four quarters in 2018 on idealista web site \cite{idealista}.
+Idealista is the major real estate listing portal in Spain and also present in other southern european countries as Italy and Portugal.
 
-\noindent\textbf{Tip:} do not describe your data (figures, tables, etc.) in this section,
-do this in the Data Description section above.]
+Given regulatory restriction idealista listings are slighly anonymized, in order to agree regulatory requirements guaranteeing their attributes and spatial properties.
+This process take two steps, the first consists of the obfuscation of prices adding or substracting a random percentage of their original values ranging from -2.5\% to +2.5\%.
+Since asking prices are not normally a completely continuous magnitude (sale prices are usually multiples of 1000 and rent prices are of 10), after the first price modification we finally align prices to multiples of 1000.
+
+% Year of construction is revised as well, given that original year of construction from listings are entered by users in the web site, therefore subject to errors and incomplete (a 40\% missing rate).
+To remove any issue we assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the field value was missing.
+
+Finally we carry out a spatial masking process that intend to keep spatial properties of the original data set.
+We essentially displace the coordinates of each listing with a minimum distance and a maximum distance with the restriction that every new location must fall in within the original neighborbood of the add.
+The process is described in the algorithm below:
+
+\begin{algorithm}[H]
+ \KwData{all idealista listings}
+ \KwResult{all idealista listings with masked coordinates}
+ initialization\;
+ \For{each listing L}{
+  take geographical location of L as $(X,Y)$
+  \Repeat{this stop condition}{
+    take a random angle $\alpha$ from 0 to 360 degrees
+    take a distance $R$ as a random value from 30 to 60 meters
+    determine a new point $(X',Y')$ calculated as a point located $R$ with the angle $\alpha$
+  }
+  set $(X',Y')$ as the new location for the listing L
+ }
+ \caption{Coordinate displacement process for anonymisation purposes}
+\end{algorithm}
+
+In the figure \ref{fig:coordinates-displacement} we display the histogram with the displacement in meters for all listings in the city of Valencia, as average the displacement is 45 meters
+
+\begin{figure}
+  \caption{Coordinate displacement in meters Valencia. Source: own elaboration}
+  \centering
+  \includegraphics[width=6cm, height=4cm]{figures/coordinates/coordinates-valencia}
+  \label{fig:coordinates-displacement}
+\end{figure}
+
+% Define a new subcommand
+
+\begin{figure}
+\centering
+\begin{tabular}{|c|c|c|}
+\hline
+& Original & Moved
+\\
+\hline
+Madrid &
+\includegraphics[width=3.5cm, height=2.8cm]{figures/points-post-madrid}
+&
+\includegraphics[width=4cm, height=2.8cm]{figures/points-post-madrid}
+\\
+%\hline
+Barcelona &
+\includegraphics[width=3cm, height=2.8cm]{figures/points-pre-barcelona}
+&
+\includegraphics[width=3cm, height=2.8cm]{figures/points-post-barcelona}
+\\
+Valencia &
+%\hline
+\includegraphics[width=3cm, height=2.8cm]{figures/points-pre-valencia}
+&
+\includegraphics[width=3cm, height=2.8cm]{figures/points-post-valencia}
+\\
+\hline
+\end{tabular}
+\end{figure}
+
+% \noindent [Offer a complete description of the experimental design and methods
+% used to acquire these data. Please provide any programs or code files
+% used for filtering and analyzing these data. It is very important that
+% this section is as comprehensive as possible. If you are submitting via
+% another Elsevier journal (a co-submission) you are encouraged to
+% provide more detail than in your accompanying research article. There
+% is no character limit for this section; however, no insight,
+% interpretation, or background should be included in this section.
+%
+% \noindent\textbf{Tip:} do not describe your data (figures, tables, etc.) in this section,
+% do this in the Data Description section above.]
 
@@
 
-\section*{References}
+% \section*{References}
 
-\noindent [References are limited (approx. 15) and excessive self-citation is not
-allowed. \textbf{If your data article is co-submitted via another Elsevier
-journal, please cite your associated research article here.}
-
-\noindent\textbf{Reference style:}
-Text:?Indicate references by number(s) in square brackets in line with
-the text. The actual authors can be referred to, but the reference
-number(s) must always be given.?
-
-\noindent Example: '..... as demonstrated [3,6]. Barnaby and Jones [8] obtained a different result ....'?
-
-\noindent [Use \verb+\cite+ command to cite a reference list item in text.
-
-\noindent These are examples for reference citations \cite{1}.
-\cite{2}.
-\cite{4}.]
+% \noindent [References are limited (approx. 15) and excessive self-citation is not
+% allowed. \textbf{If your data article is co-submitted via another Elsevier
+% journal, please cite your associated research article here.}
+%
+% \noindent\textbf{Reference style:}
+% Text:?Indicate references by number(s) in square brackets in line with
+% the text. The actual authors can be referred to, but the reference
+% number(s) must always be given.?
+%
+% \noindent Example: '..... as demonstrated [3,6]. Barnaby and Jones [8] obtained a different result ....'?
+%
+% \noindent [Use \verb+\cite+ command to cite a reference list item in text.
+%
+% \noindent These are examples for reference citations \cite{1}.
+% \cite{2}.
+% \cite{4}.]
 

### 2021-07-11 R/idealista18.R

@@ -1,4 +1,4 @@
-#' idealista18: A data package with real estate litings in Spain from the idealista data base.
+#' idealista18: A data package with real estate listings in Spain from the idealista data base.
 #'
 #' This package contains four quarterly extractions with the set of apartments listed on idealista during 2018,
 #' the package contains three data items for the three spanish cities of Madrid, Barcelona and Valencia:

### 2021-07-12 data-in-brief/dib-idealista18.tex
Updated before and after spatial point pattern distribution

@@
 \hline
-Type of data           & Table\newline\\
+Type of data           & Tables\newline\\
 %\clearpage
-How data were acquired & [State how the data were acquired: E.g. Microscope,
-                         SEM, NMR, mass spectrometry, survey* etc.\newline
-                         Instruments: E.g. hardware, software, program\newline
-                         Make and model and of the instruments used:\newline
-
-                         {\fontsize{7pt}{8pt}\selectfont
-                         *\,if you conducted a survey you must submit a copy of the
-                         survey(s) used (either provide these as supplementary material
-                         file or provide a URL link to the survey
-                         in this section of the table).
-                         If the survey is not written in English,
-                         please provide an English-language translation.}]\\
+How data were acquired & multi-family listing records given by idealista portal \cite{idealista}\newline
+                         Spanish central cadastral registry \cite{Catastro}\newline
+                         Open street map \cite{OpenStreetMap}\newline
+\\
 \hline
@@
 data\newline
-collection             & [Provide a brief description of which conditions were considered
-                         for data collection. Max 400 characters]\\
+collection             & Data has been directly downloaded from the sources, cadastral and idealista data has been merged based on geographical location for each record
+\\
+% [Provide a brief description of which conditions were considered
+%                          for data collection. Max 400 characters]
 
@@
 data\newline
-collection             & [Provide a brief description of how these data were collected.
-                         Max 600 characters]\\
+collection             & idealista provided the complete record set \newline
+                         cadastral information has been downloaded the open records published quarterly\newline
+                         open street map has been downloaded from its open API
+\\
 \hline
@@
                          Country: Spain\newline
-                         Latitude and longitude (and GPS coordinates, if possible) for collected samples/data:\newline
+                         Latitude and longitude samples/data: EPSG:4326\newline
+\\
 
-                         If you are describing secondary data, you are required to provide a list of
-                         the primary data sources used in the section.\newline
-
-                         Primary data sources:  ]\\
+%
+%                          If you are describing secondary data, you are required to provide a list of
+%                          the primary data sources used in the section.\newline
+%
+%                          Primary data sources:  ]\\
 \hline
@@
 Idealista is the major real estate listing portal in Spain and also present in other southern european countries as Italy and Portugal.
-Each record contains all attributes the listing ad has plus a number of additional attribues from the Spanish cadastre \cite{Catastro}, the names for all these variables start with the prefix \emph{CAD}.
+Each record contains all attributes the listing ad has plus a number of additional attribues from the Spanish cadastre \cite{Catastro}, described in the table \ref{table:data-assets-variables}, the names for all these variables start with the prefix \emph{CAD}.
 Cadastral features asssignment is done by assigning the features of the nearest parcel to the coordinates \emph{LATITUDE} and \emph{LONGITUDE}.
@@
 \begin{longtable}{p{40mm} c p{63mm}}
+\caption{Description of the variables in the listing data set}\\
+\label{table:data-assets-variables} \\
 \hline
@@
 \begin{longtable}{l p{40mm} c p{63mm}}
+\caption{Description of variables of neighborhood polygons data set} \\
 \hline
@@
 
-We have a total of 73 neigborhoods for Barcelona, 135 for Madrid and 73 for Valencia. Each neighborhood has also two additional variables:
+We have a total of 73 neigborhoods for Barcelona, 135 for Madrid and 73 for Valencia. Each neighborhood has also two additional variables described in the table \ref{table:data-additional-variables}.
 
@@
 \begin{longtable}{p{40mm} c p{63mm}}
+\caption{Additional variables for each city} \\
+\label{table:data-additional-variables} \\
 \hline
@@
 
-Finally we carry out a spatial masking process that intend to keep spatial properties of the original data set.
+Finally we carry out a spatial masking process described in the algorithm \ref{algo:coordinates-displacement}.
+It intends to keep spatial properties of the original data set.
 We essentially displace the coordinates of each listing with a minimum distance and a maximum distance with the restriction that every new location must fall in within the original neighborbood of the add.
@@
  \caption{Coordinate displacement process for anonymisation purposes}
+ \label{algo:coordinates-displacement}
 \end{algorithm}
 
-In the figure \ref{fig:coordinates-displacement} we display the histogram with the displacement in meters for all listings in the city of Valencia, as average the displacement is 45 meters
+In the figure \ref{fig:coordinates-displacement} we display the histogram with the displacement in meters for all listings in the city of Valencia, as average the displacement is 45 meters.
 
@@
 % Define a new subcommand
+In the figure \ref{fig:points-pre-post-all} we can see the spatial distribution of the original records compared to spatial distribution after masking:
 
-\begin{figure}
+\begin{figure}[h]
+  \caption{Spatial distribution of ads (before and after masking). Source: own elaboration}
   \centering
-\begin{tabular}{|c|c|c|}
-\hline
-& Original & Moved
-\\
-\hline
-Madrid &
-\includegraphics[width=3.5cm, height=2.8cm]{figures/points-post-madrid}
-&
-\includegraphics[width=4cm, height=2.8cm]{figures/points-post-madrid}
-\\
-%\hline
-Barcelona &
-\includegraphics[width=3cm, height=2.8cm]{figures/points-pre-barcelona}
-&
-\includegraphics[width=3cm, height=2.8cm]{figures/points-post-barcelona}
-\\
-Valencia &
-%\hline
-\includegraphics[width=3cm, height=2.8cm]{figures/points-pre-valencia}
-&
-\includegraphics[width=3cm, height=2.8cm]{figures/points-post-valencia}
-\\
-\hline
-\end{tabular}
+  \includegraphics[width=8cm]{figures/points-pre-post-all}
+  \label{fig:points-pre-post-all}
 \end{figure}
@@
 \section*{Acknowledgments}
-Acknowledgments should be inserted at the end of the paper, before the
-references, not as a footnote to the title. Use the unnumbered
-Acknowledgements Head style for the Acknowledgments heading.
+
+The authors thank Alessandro Galesi for his support in the paper revision and Juan Ramon Selva for collecting and cleaning the spatial data.
+
+% Acknowledgments should be inserted at the end of the paper, before the
+% references, not as a footnote to the title. Use the unnumbered
+% Acknowledgements Head style for the Acknowledgments heading.
 
@@
 
-\noindent [All authors are required to report the following information:
-\begin{enumerate}
-\item[(1)] All third-party financial support for the work this article;
+The authors have no conflicts of interest to declare.
 
-\item[(2)] All financial relationships with any entity that could be
-viewed as relevant to data described in this manuscript;
-
-\item[(3)] All sources of revenue with relevance to this work where
-payments have been made to authors, or their institutions on their
-behalf, within the 36 months prior to submission;
-
-\item[(4)] Any other interactions with the sponsor, outside of the
-submitted work;
-
-\item[(5)] Any relevant patents or copyrights (planned, pending or
-issued);
-
-\item[(6)] Any other relationships or affiliations that may be
-perceived by readers to have influenced, or give the appearance of
-potentially influencing, what has been written in this article.
-\end{enumerate}
-As a general guideline, it is usually better to
-disclose a relationship than not. This information will be acknowledged
-at publication in the manuscript. If there is no known competing
-financial interests or personal relationships that could have appeared
-to influence the work reported in this paper, please include this
-statement.]
-
-\vskip12pt\noindent
-The authors declare that they have no known competing
-financial interests or personal relationships which have, or could be
-perceived to have, influenced the work reported in this article.
-
-\vskip12pt\noindent
-[If there are financial interests/personal relationships which may be
-considered as potential competing interests, please declare them here.]
-
-\subsection*{Note}
-\label{sec1}
-Any instructions relevant to the \verb+elsarticle.cls+ are applicable
-here as well. See the online instruction available on:
-\makeatletter
-\if@twocolumn
-\begin{verbatim}
- http://support.river-valley.com/wiki/
- index.php?title=Elsarticle.cls
-\end{verbatim}
-\else
-\begin{verbatim}
- http://support.river-valley.com/wiki/index.php?title=Elsarticle.cls
-\end{verbatim}
-\fi
+% \noindent [All authors are required to report the following information:
+% \begin{enumerate}
+% \item[(1)] All third-party financial support for the work this article;
+%
+% \item[(2)] All financial relationships with any entity that could be
+% viewed as relevant to data described in this manuscript;
+%
+% \item[(3)] All sources of revenue with relevance to this work where
+% payments have been made to authors, or their institutions on their
+% behalf, within the 36 months prior to submission;
+%
+% \item[(4)] Any other interactions with the sponsor, outside of the
+% submitted work;
+%
+% \item[(5)] Any relevant patents or copyrights (planned, pending or
+% issued);
+%
+% \item[(6)] Any other relationships or affiliations that may be
+% perceived by readers to have influenced, or give the appearance of
+% potentially influencing, what has been written in this article.
+% \end{enumerate}
+% As a general guideline, it is usually better to
+% disclose a relationship than not. This information will be acknowledged
+% at publication in the manuscript. If there is no known competing
+% financial interests or personal relationships that could have appeared
+% to influence the work reported in this paper, please include this
+% statement.]
+%
+% \vskip12pt\noindent
+% The authors declare that they have no known competing
+% financial interests or personal relationships which have, or could be
+% perceived to have, influenced the work reported in this article.
+%
+% \vskip12pt\noindent
+% [If there are financial interests/personal relationships which may be
+% considered as potential competing interests, please declare them here.]
+%
+% \subsection*{Note}
+% \label{sec1}
+% Any instructions relevant to the \verb+elsarticle.cls+ are applicable
+% here as well. See the online instruction available on:
+% \makeatletter
+% \if@twocolumn
+% \begin{verbatim}
+%  http://support.river-valley.com/wiki/
+%  index.php?title=Elsarticle.cls
+% \end{verbatim}
+% \else
+% \begin{verbatim}
+%  http://support.river-valley.com/wiki/index.php?title=Elsarticle.cls
+% \end{verbatim}
+% \fi
 

### 2021-07-12 data-in-brief/dib-idealista18.tex
Added poi information

@@
 We essentially displace the coordinates of each listing with a minimum distance and a maximum distance with the restriction that every new location must fall in within the original neighborbood of the add.
-The process is described in the algorithm below:
+As we can see in the figure \ref{fig:points-moved-image} the area where the new point (masked) would be located will be taken randomly within maximum and minimum displacement distances circles.
+In order to preserve the nature of the neighborhood the house we make sure the new point would fall in the original neighborhood, otherwise we look for a new masked place.
+
+\begin{figure}[ht]
+  \caption{Masking coordinates spatial range. Source: own elaboration}
+  \centering
+  \includegraphics[width=5cm, height=4.3cm]{figures/points-moved-image}
+  \label{fig:points-moved-image}
+\end{figure}
 
@@
 
-\begin{figure}
+\begin{figure}[h]
   \caption{Coordinate displacement in meters Valencia. Source: own elaboration}
@@
 \end{figure}
+
+The last item contains a list of Points of Interest for each city, in the case of Madrid it has the city center (Plaza del Sol), a number of points corresponding to the Paseo de la Castellana that is a major Avenue that cuts across the city from South to North, and a list of subway entraces (Metro).
+For Valencia we have the city center, a list of points for the main street called Basco Ibáñez plus the subway entrances.
+Similarly for Barcelona we have the City Center, points of the main street (Avenida Diagonal) and the subway entrances set.
 

### 2021-07-12 data-in-brief/dib-idealista18.tex
Fixed some mistakes and data package documentation

@@
 %A maximum of 250 characters is allowed.
-\title{idealista18: A data package with real estate information in three major Spanish markets from the Idealista database\tnoteref{tnote1}}%
-\tnotetext[tnote1]{This is an example for title footnote coding.}
+\title{idealista18: A data package with real estate information in three major Spanish markets from the Idealista database }%
+% \title{idealista18: A data package with real estate information in three major Spanish markets from the Idealista database \tnoteref{tnote1} }%
+% \tnotetext[tnote1]{This is an example for title footnote coding.}
 %Tip: here are a few examples of recent suitable article titles - these are short and clear:
@@
 \author[1]{David \snm{Rey Blanco}}
-\author[1]{Pelayo \snm{González Arbues}\fnref{fn1}}
-\fntext[fn1]{This is author footnote for second author.}
+\author[1]{Pelayo \snm{González Arbues}}
+% \author[1]{Pelayo \snm{González Arbues}\fnref{fn1}}
+% \fntext[fn1]{This is author footnote for second author.}
 \author[2]{Fernando \snm{López Hernández}}
@@
 This dataset contains three items for each of the three major cities in Spain: Madrid, Barcelona and Valencia.
-First real estate listings published on idealista portal in 2018.
+The first data set contains real estate listings published on idealista portal in 2018.
 All listings have been enriched with cadastral information (i.e.
-building year of construction, built quality materials grade) plus some geographical features such as
-distance to relevant city areas and the coordinates themselves.
-To comply with european personal protection laws, we have processeed some sensible variables yet preserving their spatial properties.
-The second and third items are a list of points of interest and the administrative boundaries for each municipality.
+building year of construction, built quality materials grade) plus some geographical features such as distance to relevant city areas and the coordinates themselves.
+To comply with european personal protection laws, we have processeed some sensitive variables yet preserving their spatial properties.
+The second data sets contains the neighborhood boundaries for each city, and third sets comprise a set of key points of interest for each municipality.
 This dataset is suitable to house market analysis, hedonic house price models and other spatial research related with real estate markets.
@@
 \parsep=0pt
-  \item A cleaned and enriched dataset consisting of real estate listings for three major cities in Spain, constructed to analyze the impacto of using machine learning models to identify spatial market segments to build house price hedonic models.
-  \item The dataset can be used to extend the topic of market segments automatic or semi-automatic identification.
-    \item Official boundaries combined with spatial patterns can be used to analyse the suitability of these boundaries for real estate value purposes.
-  \item The dataset can be enlarged with complementary information to develop hedonic models.
-  \item The data can be processed by quantitative analysis and statistical modeling to study the different factors that affect house prices in the three areas.
+  \item A cleaned and enriched dataset consisting of real estate listings for three major cities in Spain.
+It has been constructed to analyze the impact of using machine learning models to identify spatial market segments when building house price hedonic models.
+  \item The dataset can be used to extend the topic of automatic or semi-automatic identification of house market segments.
+    \item The neighborhood boundaries combined with spatial patterns can be used to analyse the suitability of these boundaries as spatial dummy variables for real estate analyses purposes.
+  \item The dataset can be enlarged with complementary spatial information to develop hedonic models.
+  \item The data can be processed by quantitative analysis and statistical modeling to study the different factors that affect house prices in the three locations.
   \item Identification of spatial patterns in the real estate scope using the geo-referenced data points. For either value or urban patterns discovery.
@@
 
-The data set is composed by three items for three major spanish cities: quarterly single family listings, neighboorhood polygons and a a set of Points of Interest from Open Street Map \cite{OpenStreetMap}.
+The data sets is composed by three items for three major spanish cities: quarterly single family listings, neighboorhood polygons and a set key of Points of Interest for each city.
 All spatial features, such as polygons and points, are expressed in geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.
-
-The first block integrates property published ads,the files contain the complete offering for each of the four quarters in 2018 on idealista web site \cite{idealista}.
+The first block of data integrates properties published on idealista web site \cite{idealista}; each file contains the complete offering for a city for the four quarters in 2018.
 Idealista is the major real estate listing portal in Spain and also present in other southern european countries as Italy and Portugal.
-Each record contains all attributes the listing ad has plus a number of additional attribues from the Spanish cadastre \cite{Catastro}, described in the table \ref{table:data-assets-variables}, the names for all these variables start with the prefix \emph{CAD}.
+Each record contains the key found in listing ad\footnote{we will use the term \emph{ad} or \emph{listing} interchangeably to refer to a house published on the portal} plus a number of additional attribues from the Spanish cadastre \cite{Catastro}.
+The latter are described in the table \ref{table:data-assets-variables}, and the names for all these variables start with the prefix \emph{CAD}.
 Cadastral features asssignment is done by assigning the features of the nearest parcel to the coordinates \emph{LATITUDE} and \emph{LONGITUDE}.
+The measure scales for each variable has been defined according the theoretical framework proposed by \cite{stevens1946theory} that defines four scales: nominal, ordinal, interval and ratio.
 
@@
 ASSETID & Identifier & Unique identifier of the advertisement\\
-PERIOD & Nominal (Date) & Expressed as YYYYMM, indicates the quarter when the ad was extracted we used YYYY03 for the 1st Q, YYYY06 the 2nd, YYYY09 for the 3rd and YYYY12 for the 4th\\
+PERIOD & Nominal (Date) & Expressed as YYYYMM, indicates the quarter when the ad was extracted. We used YYYY03 for the 1st quarter, YYYY06 the 2nd, YYYY09 for the 3rd and YYYY12 for the 4th\\
 PRICE & Interval & Asking price for the ad at idealista expressed in euros\\
@@
 
-In addition to the common features each city will have a set of spatial additional features:
+In addition to the common features each city will have a set of spatial additional features, in particular refered to the distance to a major street.
 
@@
 
+The record count for each city is: 156,016 listings for Madrid and 84,280 for Barcelona and 79,360 for Valencia.
+It is important to note that the same listing can be found in more than one period, what means that a house was uploaded for sale in one quarter but was sold in a subsequent quarter.
+
 The second part contains the polygons for the different neighborhoods for such cities as we can see in the figure \ref{fig:all-polygons}.
-This boundaries are not the official boundaries but the official boundaries used by idealista portal.
+This boundaries are based on the official boundaries but slighly adapted by idealista \footnote{the criterium used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
 In practical terms we can assume they are the same, since the website simply collapsed those areas if they are small enough in terms of number of ads.
@@
 \section*{Experimental Design, Materials and Methods}
+% Year of construction is revised as well, given that original year of construction from listings are entered by users in the web site, therefore subject to errors and incomplete (a 40\% missing rate).
+To remove any issue we assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the field value was missing.
 
@@
 The files contain the complete offering for each of the four quarters in 2018 on idealista web site \cite{idealista}.
-Idealista is the major real estate listing portal in Spain and also present in other southern european countries as Italy and Portugal.
-
-Given regulatory restriction idealista listings are slighly anonymized, in order to agree regulatory requirements guaranteeing their attributes and spatial properties.
+Idealista is the major real estate listing portal in Spain and also present in other southern european countries as Italy and Portugal.Given the spanish regulatory restrictions idealista listings are slighly anonymized, in order to agree regulatory requirements guaranteeing their attributes and spatial properties.
 This process take two steps, the first consists of the obfuscation of prices adding or substracting a random percentage of their original values ranging from -2.5\% to +2.5\%.
-Since asking prices are not normally a completely continuous magnitude (sale prices are usually multiples of 1000 and rent prices are of 10), after the first price modification we finally align prices to multiples of 1000.
-
-% Year of construction is revised as well, given that original year of construction from listings are entered by users in the web site, therefore subject to errors and incomplete (a 40\% missing rate).
-To remove any issue we assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the field value was missing.
-
-Finally we carry out a spatial masking process described in the algorithm \ref{algo:coordinates-displacement}.
-It intends to keep spatial properties of the original data set.
-We essentially displace the coordinates of each listing with a minimum distance and a maximum distance with the restriction that every new location must fall in within the original neighborbood of the add.
+Since asking prices are not normally a completely continuous magnitude (sale prices are usually multiples of 1000 and rent prices are of 10), after the first price modification we finally align prices to multiples of 1000.Finally we carry out a spatial masking process that intends to keep spatial properties of the original data set.
 As we can see in the figure \ref{fig:points-moved-image} the area where the new point (masked) would be located will be taken randomly within maximum and minimum displacement distances circles.
@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
   \caption{Masking coordinates spatial range. Source: own elaboration}
@@
 
-\begin{algorithm}[H]
+The algorithm \ref{algo:coordinates-displacement} iteratively  displace the coordinates of each listing with a minimum distance and a maximum distance with the restriction that every new location must fall in within the original neighborbood of the listing.
+
+\begin{algorithm}[!ht]
  \KwData{all idealista listings}
@@
 
-\begin{figure}[h]
+\begin{figure}[!ht]
   \caption{Coordinate displacement in meters Valencia. Source: own elaboration}
@@
 % Define a new subcommand
-In the figure \ref{fig:points-pre-post-all} we can see the spatial distribution of the original records compared to spatial distribution after masking:
+In the figure \ref{fig:points-pre-post-all} we can see the spatial distribution of the original records compared to spatial distribution after masking.
 
-\begin{figure}[h]
+\begin{figure}[!ht]
   \caption{Spatial distribution of ads (before and after masking). Source: own elaboration}
@@
 \end{figure}
-
-The last item contains a list of Points of Interest for each city, in the case of Madrid it has the city center (Plaza del Sol), a number of points corresponding to the Paseo de la Castellana that is a major Avenue that cuts across the city from South to North, and a list of subway entraces (Metro).
-For Valencia we have the city center, a list of points for the main street called Basco Ibáñez plus the subway entrances.
-Similarly for Barcelona we have the City Center, points of the main street (Avenida Diagonal) and the subway entrances set.
 
@@
 
-\section*{Ethics Statement}
-\noindent [Please refer to the journal's
-\href{https://www.elsevier.com/journals/data-in-brief/2352-3409/guide-for-authors}{Guide for Authors}
-for more information on
-the ethical requirements for publication in Data in Brief. In addition
-to these requirements:
-
-\noindent\textbf{If the work involved the use of human subjects:}
-please include a statement here confirming that informed consent was
-obtained for experimentation with human subjects;
-
-\noindent\textbf{If the work involved animal experiments:} please
-include a statement confirming that all experiments comply with
-the \href{https://www.nc3rs.org.uk/arrive-guidelines}{ARRIVE\ guidelines} and were be carried out in accordance with the
-U.K. Animals (Scientific Procedures) Act, 1986 and associated
-guidelines, \href{https://ec.europa.eu/environment/chemicals/lab_animals/legislation_en.htm}{EU Directive 2010/63/EU for animal experiments}, or the
-National Institutes of Health guide for the care and use of Laboratory
-animals (NIH Publications No. 8023, revised 1978)]
+% \section*{Ethics Statement}
+% \noindent [Please refer to the journal's
+% \href{https://www.elsevier.com/journals/data-in-brief/2352-3409/guide-for-authors}{Guide for Authors}
+% for more information on
+% the ethical requirements for publication in Data in Brief. In addition
+% to these requirements:
+%
+% \noindent\textbf{If the work involved the use of human subjects:}
+% please include a statement here confirming that informed consent was
+% obtained for experimentation with human subjects;
+%
+% \noindent\textbf{If the work involved animal experiments:} please
+% include a statement confirming that all experiments comply with
+% the \href{https://www.nc3rs.org.uk/arrive-guidelines}{ARRIVE\ guidelines} and were be carried out in accordance with the
+% U.K. Animals (Scientific Procedures) Act, 1986 and associated
+% guidelines, \href{https://ec.europa.eu/environment/chemicals/lab_animals/legislation_en.htm}{EU Directive 2010/63/EU for animal experiments}, or the
+% National Institutes of Health guide for the care and use of Laboratory
+% animals (NIH Publications No. 8023, revised 1978)]
 
@@
 
-The authors have no conflicts of interest to declare.
+The authors declare that they have no known competing financial interests or personal relationships which have, or could be perceived to have, influenced the work reported in this article.
 

### 2021-07-12 R/idealista18.R

@@ -16,7 +16,7 @@ NULL
 
 #' Sale prices of residential property in Barcelona in 2018.
 #'
-#' A dataset containing information about 84,280 real estate ads on idealista
+#' A dataset containing information about 84,280 real estate listings on idealista
 #' in Barcelona during 2018.
 #'
 #' @format An object of class `sf` (inherits from `data.frame`) with 84,280 rows and 57 variables:
@@ -78,7 +78,7 @@ NULL
 
 #' Sale prices of residential property in Madrid in 2018.
 #'
-#' A dataset containing information about 156,016 real estate transactions
+#' A dataset containing information about 156,016 real estate listings on idealista
 #' in Madrid during 2018.
 #'
 #' @format An object of class `sf` (inherits from `data.frame`) with 156,016 rows and 57 variables:
@@ -120,7 +120,8 @@ NULL
 #'   \item{CADASTRALQUALITYID}{Cadastral quality (source: cadastre)}
 #'   \item{BUILTTYPEID_1}{Dummy value for flat condition: 1 new development 0 otherwise (source: advertiser))}
 #'   \item{BUILTTYPEID_2}{Dummy value for flat condition: 1 second hand to be restored 0 otherwise (source: advertiser))}
-#'   \item{BUILTTYPEID_3}{Dummy value for flat condition: 1 second hand in good condition 0 otherwise (source: advertiser))}#'   \item{DISTANCE_TO_CITY_CENTER}{Distance to center of city in km}
+#'   \item{BUILTTYPEID_3}{Dummy value for flat condition: 1 second hand in good condition 0 otherwise (source: advertiser))}#'
+#'   \item{DISTANCE_TO_CITY_CENTER}{Distance to center of city in km}
 #'   \item{DISTANCE_TO_CASTELLANA}{Distance to Paseo de la Castellana in km; la Castellana is a major street that cuts across the city from South to North}
 #'   \item{geometry}{geometry of simple features}
 #'   }
@@ -139,8 +140,8 @@ NULL
 
 #' Sale prices of residential property in Valencia in 2018.
 #'
-#' A dataset containing information about 79,360 real estate transactions
-#' in Madrid during 2018.
+#' A dataset containing information about 79,360 real estate listings on idealista
[…]

### 2021-11-05 data-in-brief/dib-idealista18.tex
Updated edits to dib-idealista.tex
Started editing Data in Brief paper

@@
 
-The data sets is composed by three items for three major spanish cities: quarterly single family listings, neighboorhood polygons and a set key of Points of Interest for each city.
+This data package is composed of data objects corresponding to three major Spanish cities: quarterly single family listings, neighboorhood polygons and a set key of Points of Interest for each city.
 All spatial features, such as polygons and points, are expressed in geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.

### 2021-11-08 data-in-brief/dib-idealista18.tex
Edited Data in Brief

--- 2021-07-12 data-in-brief/dib-idealista18.tex
@@
 %A maximum of 250 characters is allowed.
-\title{idealista18: A data package with real estate information in three major Spanish markets from the Idealista database }%
+\title{A geo-referenced micro-data set of real estate listings for the three largest Spanish cities from the Idealista website}%
 % \title{idealista18: A data package with real estate information in three major Spanish markets from the Idealista database \tnoteref{tnote1} }%
@@
 \author[1]{David \snm{Rey Blanco}}
-\author[1]{Pelayo \snm{González Arbues}}
+\author[1]{Pelayo \snm{González Arbués}}
 % \author[1]{Pelayo \snm{González Arbues}\fnref{fn1}}
@@
 \address[1]{idealista, Plaza de las Cortes 5, 28014 Madrid, Spain}
-\address[2]{Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia (Spain)}
+\address[2]{Facultad de CC de la Empresa, Universidad Politécnica de Cartagena, C/ Real, 3. 30201 Cartagena, Murcia (Spain)}
 \address[3]{School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada}
@@
 %%%%
-This dataset contains three items for each of the three major cities in Spain: Madrid, Barcelona and Valencia.
-The first data set contains real estate listings published on idealista portal in 2018.
-All listings have been enriched with cadastral information (i.e.
-building year of construction, built quality materials grade) plus some geographical features such as distance to relevant city areas and the coordinates themselves.
-To comply with european personal protection laws, we have processeed some sensitive variables yet preserving their spatial properties.
-The second data sets contains the neighborhood boundaries for each city, and third sets comprise a set of key points of interest for each municipality.
-This dataset is suitable to house market analysis, hedonic house price models and other spatial research related with real estate markets.
+This data article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain.
+The observations are for the three largest cities in the country: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations) and Valencia (n = 33,622 observations).
+These data were originally published on \url{idealista.com} real estate website.
+The data sets include the coordinates of properties (latitude and longitude), asking prices of each listd dwelling, and several variables of indoor characteristics.
+The listings were enriched with official information (building year of construction and built quality materials grade) plus other relevant geographical features such as distance to urban points of interest.
+Along with real estate listings, the data product also includes neighborhood boundaries for each city.
+These data sets are companions to a paper that examines the potential of machine learning techniques for spatial market segmentation in hedonic price analysis.
+The data product is offered in the form of a fully documented `R` package.
+This open data product is available for scientific and educational purposes, in particular for geo-spatial studies.
 
-\textbf{<falta cita/referencia al artículo>}
+\textbf{<article reference to be inserted>}
 
@@
 %to draw from]
-\KWD Property values\sep Spain\sep Spatial analysis\sep Machine learning\sep Hedonic price analysis
+\KWD Hedonic price analysis\sep Spain\sep Spatial econometrics\sep Machine learning\sep Geo-spatial analysis
 \end{keyword}
@@
 \hline
-Type of data           & Tables\newline\\
+Type of data           & Tables: dataframes and `sf` objects in `R` format\newline\\
 %\clearpage
-How data were acquired & multi-family listing records given by idealista portal \cite{idealista}\newline
+How data were acquired & Dwelling listing records provided by \url{idealista.com} \cite{idealista}\newline
                          Spanish central cadastral registry \cite{Catastro}\newline
@@
 \hline
-Data format            & Spatially masked\\
+Data format            & `R` package, named \{idealista18\}\\
 \hline
@@
 data\newline
-collection             & Data has been directly downloaded from the sources, cadastral and idealista data has been merged based on geographical location for each record
+collection             & Data has been directly downloaded from the sources, cadastral and idealista website data has been merged based on geographical location for each record.
 \\
@@
 {Data accessibility}   & Repository name: GitHub\newline
-                         Direct URL to data: https://github.com/paezha/idealista18\newline
+                         Direct URL to data: \url{https://github.com/paezha/idealista18}\newline
                          \\
@@
 
-The data sets is composed by three items for three major spanish cities: quarterly single family listings, neighboorhood polygons and a set key of Points of Interest for each city.
-All spatial features, such as polygons and points, are expressed in geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.
-The first block of data integrates properties published on idealista web site \cite{idealista}; each file contains the complete offering for a city for the four quarters in 2018.
-Idealista is the major real estate listing portal in Spain and also present in other southern european countries as Italy and Portugal.
-Each record contains the key found in listing ad\footnote{we will use the term \emph{ad} or \emph{listing} interchangeably to refer to a house published on the portal} plus a number of additional attribues from the Spanish cadastre \cite{Catastro}.
-The latter are described in the table \ref{table:data-assets-variables}, and the names for all these variables start with the prefix \emph{CAD}.
+This open data product \cite{arribas2021open} is composed of several data objects corresponding to three major Spanish cities: dwelling listings, neighboorhood polygons and a set of Points of Interest (POI) for each city\footnote{The data has been provided by Idealista, the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal.}.
+The data set is distributed in the form of an `R` package, named \{idealista18\}.
+All spatial objects such as polygons and points are distributed as simple features objects (class `sf` in `R`).
+Spatial objects include geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.
+
+The first block of data integrates properties published on idealista real state website \cite{idealista}; each file contains the complete set of listings for each of the three cities in the package, corresponding to the four quarters of year 2018.
+Each record contains the key found in the listing ad\footnote{The term \emph{ad} or \emph{listing} is used interchangeably to refer to a property advertised on the website} plus a number of additional attribues from the Spanish cadastre \cite{Catastro}.
+Cadastral information is described in Table \ref{table:data-assets-variables}, including the the prefix \emph{CAD} in the variable name.
 Cadastral features asssignment is done by assigning the features of the nearest parcel to the coordinates \emph{LATITUDE} and \emph{LONGITUDE}.
-The measure scales for each variable has been defined according the theoretical framework proposed by \cite{stevens1946theory} that defines four scales: nominal, ordinal, interval and ratio.
+The measure scales for each variable are defined according to the theoretical framework proposed by \cite{stevens1946theory} that defines four scales: nominal, ordinal, interval and ratio.
 
@@
 \begin{longtable}{p{40mm} c p{63mm}}
-\caption{Description of the variables in the listing data set}\\
+\caption{Description of the variables in the listing data sets}\\
 \label{table:data-assets-variables} \\
@@
 UNITPRICE & Interval & Asking price in euros per square meter (constructed area)\\
-ADTYPOLOGYID & Nominal & Residential building type: multi-family: \emph{home}, single-familiy: \emph{chalet}\\
-ADOPERATIONID & Nominal & Operation type for the ad: \emph{sale} or \emph{rent}\\
 ROOMNUMBER & Ordinal & Number of bedrooms\\
@@
 HASLIFT & Nominal & Dummy variable for lift (takes 1 if there is a lift in the building, 0 otherwise)\\
-HASAIRCONDITIONING & Nominal & Dummy variable for AA (takes 1 if there is a AA, 0 otherwise)\\
+HASAIRCONDITIONING & Nominal & Dummy variable for air conditioner (takes 1 if there is air conditioner, 0 otherwise)\\
 AMENITYID & Nominal & Indicates the amenities included (1 - no furniture, no kitchen amenities, 2 - kitchen amenities, no furniture, 3 - kitchen amenities, furniture)\\
@@
 ISPARKINGSPACEINCLUDEDINPRICE & Nominal & Dummy variable for parking (takes 1 if parking is included in the Ad, 0 otherwise)\\
-PARKINGSPACEPRICE & Interval & Price of parking space in euros\\
+PARKINGSPACEPRICE & Interval & Asking price of parking space in euros\\
 HASNORTHORIENTATION & Nominal & Dummy variable for orientation (takes 1 if orientation is North in the Ad, 0 otherwise) - Important note: orientation features are not orthogonal features, a house oriented to the north can be also oriented to the east\\
@@
 BUILTTYPEID\_3 & Nominal & Dummy value for flat condition: 1 second hand in good condition 0 otherwise (\emph{source: advertiser})\\
-DISTANCE\_TO\_CITY\_CENTER & Interval & Distance to center of city in Km\\
+LONGITUDE & Interval & Longitude, geographical coordinate\\
+LATITUDE & Interval & Latitude geographical coordinate\\
 geometry & Geometry & Geometry for the elements. A point with $X,Y$ coordinates\\
@@
 
-In addition to the common features each city will have a set of spatial additional features, in particular refered to the distance to a major street.
+In addition to the information shown in Table \ref{table:data-assets-variables}, distances from each dwelling to Point of Interest are included.
+Table \ref{table:data-distance-variables} lists these variables for each city.
 
 \begin{footnotesize}
-\begin{longtable}{l p{40mm} c p{63mm}}
+\begin{longtable}{l p{35mm} c p{46mm}}
 \caption{Description of variables of neighborhood polygons data set} \\
+\label{table:data-distance-variables} \\
 \hline
@@
 & & &\\
-\emph{Madrid} & DISTANCE\_TO\_CASTELLANA & Interval & Distance in km to the Paseo de la Castellana Street \\
+\emph{Madrid} & DISTANCE\_TO\_CITY\_CENTER & Interval & Distance in Km to the city center (Puerta del Sol) \\ & DISTANCE\_TO\_METRO & Interval & Distance in Km to the nearest subway station \\ & DISTANCE\_TO\_CASTELLANA & Interval & Distance in Km to the Paseo de la Castellana Street \\
 & & &\\
 % \hline
-\emph{Valencia} & DISTANCE\_TO\_METRO & Interval & Distance in km to the nearest subway station \\
- & DISTANCE\_TO\_BLASCO & Interval & Distance in km to the Blasco Ibáñez Avenue \\
+\emph{Valencia} & DISTANCE\_TO\_CITY\_CENTER & Interval & Distance in Km to the city center (Plaza del Ayuntamiento) \\ & DISTANCE\_TO\_METRO & Interval & Distance in Km to the nearest subway station \\
+ & DISTANCE\_TO\_BLASCO & Interval & Distance in Km to the Blasco Ibáñez Avenue \\
 & & &\\
 % \hline
-\emph{Barcelona} & DISTANCE\_TO\_DIAGONAL & Interval & Distance in km to the Diagonal Avenue \\
+\emph{Barcelona} & DISTANCE\_TO\_CITY\_CENTER & Interval & Distance in Km to the city center (Plaza de España) \\ & DISTANCE\_TO\_METRO & Interval & Distance in Km to the nearest subway station \\ & DISTANCE\_TO\_DIAGONAL & Interval & Distance in Km to the Diagonal Avenue \\
 & & &\\
@@
 
-The record count for each city is: 156,016 listings for Madrid and 84,280 for Barcelona and 79,360 for Valencia.
-It is important to note that the same listing can be found in more than one period, what means that a house was uploaded for sale in one quarter but was sold in a subsequent quarter.
+The record counts for each city in 2018 are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for Valencia.
+Note that the same listing can be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter.
 
-The second part contains the polygons for the different neighborhoods for such cities as we can see in the figure \ref{fig:all-polygons}.
-This boundaries are based on the official boundaries but slighly adapted by idealista \footnote{the criterium used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
-In practical terms we can assume they are the same, since the website simply collapsed those areas if they are small enough in terms of number of ads.
-In the case of Madrid they just collapse four areas in two new ones.
+The second block of data include the spatial features of the three cities divided in neighborhoods.
+Figure \ref{fig:all-polygons} shows the different neighborhoods for the three cities.
+The boundaries are based on the official boundaries but slighly adapted by idealista \footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
+In practical terms we can assume they are the same, since the website simply collapses areas when they are sufficiently small in terms of number of ads.
+In the case of Madrid they just collapse four areas into two new ones.
 
 \begin{figure}[h]
-  \caption{Neighborhood boundaries for Madrid, Barcelona and Valencia. Source: own elaboration}
+  \caption{Neighborhood boundaries for Madrid, Barcelona and Valencia}
   \centering
-  \includegraphics[height=4.5cm]{figures/idealista18-all-polygons}
+  \includegraphics[height=6.5cm]{figures/idealista18-all-polygons}
   \label{fig:all-polygons}
@@
 
-We have a total of 73 neigborhoods for Barcelona, 135 for Madrid and 73 for Valencia. Each neighborhood has also two additional variables described in the table \ref{table:data-additional-variables}.
+There are a total of 73 neigborhoods in Barcelona, 135 in Madrid and 73 in Valencia. Each neighborhood has also two additional variables described in the Table \ref{table:data-additional-variables}.
 
@@
 
+The last block of data included in the data package is a set of Point of Interest of each city in `R` list format.
+These lists include three elements: (i) the coordinates of the city center (see Table \ref{table:data-additional-variables} that identify the central business district); (ii) a set of points that define the main street of each city; and (iii) the coordinates of metro stations.
+
 \section*{Experimental Design, Materials and Methods}
@@
 
-The data encompassed three major cities in Spain: Madrid, Barcelona and Valencia, for each municipality we provide listing prices for multi-familiy homes on a quarterly basis.
-The files contain the complete offering for each of the four quarters in 2018 on idealista web site \cite{idealista}.
-Idealista is the major real estate listing portal in Spain and also present in other southern european countries as Italy and Portugal.Given the spanish regulatory restrictions idealista listings are slighly anonymized, in order to agree regulatory requirements guaranteeing their attributes and spatial properties.
-This process take two steps, the first consists of the obfuscation of prices adding or substracting a random percentage of their original values ranging from -2.5\% to +2.5\%.
-Since asking prices are not normally a completely continuous magnitude (sale prices are usually multiples of 1000 and rent prices are of 10), after the first price modification we finally align prices to multiples of 1000.Finally we carry out a spatial masking process that intends to keep spatial properties of the original data set.
-As we can see in the figure \ref{fig:points-moved-image} the area where the new point (masked) would be located will be taken randomly within maximum and minimum displacement distances circles.
-In order to preserve the nature of the neighborhood the house we make sure the new point would fall in the original neighborhood, otherwise we look for a new masked place.
+As noted above, the data package contains the complete offering in three major Spanish cities for each of the four quarters of 2018, as drawn from idealista web site \cite{idealista}.
+To comply with Spanish regulations, the listings are slightly modified to preserve their anonimity.
+The masking process consists of two steps, as follows: first the prices are obfuscated with the addition or substraction of a random percentage of their original values ranging from -2.5\% to +2.5\%.
+Since asking prices are not normally a completely continuous variable (sale prices are usually multiples of 1000 and rent prices are of 10), after the first price modification we finally align prices to multiples of 1000.
+Secondly, a spatial masking process was also implemented with the intention of keeping spatial properties of the original data set.
+The coordinates of each listing were displaced using a stochastic procedure.
+Effectively, the listings were recoded using coordinates contained in a maximm and minimum displacement circles, as shown in Figure \ref{fig:points-moved-image}.
+To preserve membership in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates are in the original neighborhood of the listing.
 
@@
 
-The algorithm \ref{algo:coordinates-displacement} iteratively  displace the coordinates of each listing with a minimum distance and a maximum distance with the restriction that every new location must fall in within the original neighborbood of the listing.
+The algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall in a different neighborhood.
+This ensures that neighborhood attributes are preserved.
 
@@
 
-In the figure \ref{fig:coordinates-displacement} we display the histogram with the displacement in meters for all listings in the city of Valencia, as average the displacement is 45 meters.
+Figure \ref{fig:coordinates-displacement} shows the histogram of displacements in meters for all listings in the city of Valencia; the average distance between the original and masked coordinates is 45 meters.
 
@@
 % Define a new subcommand
-In the figure \ref{fig:points-pre-post-all} we can see the spatial distribution of the original records compared to spatial distribution after masking.
+Figure \ref{fig:points-pre-post-all} shows the spatial distribution of the original records compared to spatial distribution after masking.
 
@@
   \centering
-  \includegraphics[width=8cm]{figures/points-pre-post-all}
+  \includegraphics[width=12cm]{figures/points-pre-post-all}
   \label{fig:points-pre-post-all}
@@
 
-The authors thank Alessandro Galesi for his support in the paper revision and Juan Ramon Selva for collecting and cleaning the spatial data.
+The authors wish to thank Alessandro Galesi for his support in the paper revision and Juan Ramon Selva for collecting and cleaning the spatial data.
 
@@
 
-The authors declare that they have no known competing financial interests or personal relationships which have, or could be perceived to have, influenced the work reported in this article.
+David Rey Blanco and Pelayo González Arbués are emplyed by idealista.
+They have been granted permission to share the data presented in this articule.
+None of the authors have known competing financial interests or personal relationships which have, or could be perceived to have, influenced the work reported in this article.
 

### 2021-11-10 data-in-brief/dib-idealista18.tex
Anonymized the Data in Brief article
Submission to Computers, Environment and Urban Systems

@@
 
-\verso{D. Rey Blanco \textit{et al.}}
+%\verso{D. Rey Blanco \textit{et al.}}
+\verso{Author One \textit{et al.}}
 
@@
 %%Authors
-\author[1]{David \snm{Rey Blanco}}
-\author[1]{Pelayo \snm{González Arbués}}
+\author[1]{Author \snm{One}}
+\author[2]{Author \snm{Two}}
+%\author[1]{David \snm{Rey Blanco}}
+%\author[2]{Pelayo \snm{Arbués González}}
 % \author[1]{Pelayo \snm{González Arbues}\fnref{fn1}}
 % \fntext[fn1]{This is author footnote for second author.}
-\author[2]{Fernando \snm{López Hernández}}
-\author[3]{Antonio \snm{Páez}\corref{cor1}}
+%\author[3]{Fernando \snm{López Hernández}}
+%\author[4]{Antonio \snm{Páez}\corref{cor1}}
 %% Fourth author's email
-\ead{paezha@mcmaster.ca}
-\cortext[cor1]{Corresponding author:
-  Tel.: +1-905-525-9140 ext 26099}
+%\ead{paezha@mcmaster.ca}
+%\cortext[cor1]{Corresponding author:
+%  Tel.: +1-905-525-9140 ext 26099}
 
 %%Affiliations
-\address[1]{idealista, Plaza de las Cortes 5, 28014 Madrid, Spain}
-\address[2]{Facultad de CC de la Empresa, Universidad Politécnica de Cartagena, C/ Real, 3. 30201 Cartagena, Murcia (Spain)}
-\address[3]{School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada}
+%%\address[1]{idealista, Plaza de las Cortes 5, 28014 Madrid, Spain}
+%%\address[2]{Facultad de CC de la Empresa, Universidad Politécnica de Cartagena, C/ Real, 3. 30201 Cartagena, Murcia (Spain)}
+%%\address[3]{School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada}
 
@@
 
-\textbf{<article reference to be inserted>}
+\textbf{article reference to be inserted}
 
@@
 research\newline
-article                & D.
-Rey Blanco, P.
-González Arbues, F.
-López Hernández, A.
-Páez, Using machine learning to identify spatial market segments: A reproducible study of major Spanish markets, Comput Environ Urban Syst.
-In Press.\newline
+article                & Author One et al., Using machine learning to identify spatial market segments: A reproducible study of major Spanish markets, Comput Environ Urban Syst. In Press.\newline
 \end{longtable}
@@
 
-The authors wish to thank Alessandro Galesi for his support in the paper revision and Juan Ramon Selva for collecting and cleaning the spatial data.
+%The authors wish to thank Alessandro Galesi for his support in the paper revision and Juan Ramon Selva for collecting and cleaning the spatial data.
+The authors wish to thank X for their support in the paper revision and Y for collecting and cleaning the spatial data.
 
@@
 
-David Rey Blanco and Pelayo González Arbués are emplyed by idealista.
+Author One and Author Two are employed by idealista.
+They have been granted permission to share the data presented in this article.
+None of the authors have known competing financial interests or personal relationships which have, or could be perceived to have, influenced the work reported in this article.
+%David Rey Blanco and Pelayo González Arbués are employed by idealista.
 They have been granted permission to share the data presented in this articule.

### 2022-02-05 EPB/EPB.Rmd
Nueva version para EP-B

---
title: A geo-referenced micro-data set of real estate listings for the three largest Spanish cities
runninghead: Rey-Blanco \emph{et al}.
author:
- name: D. Rey-Blanco
  num: 1
- name: P. González-Arbues
  num: 1
- name: F. López
  num: 2
- name: A. Páez*
  num: 4
address:
- num: 1
  org: Idealista, Plaza de las Cortes 5, 28014 Madrid, Spain
- num: 2
  org: Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia, Spain
- num: 3
  org: School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada
corrauth: "Antonio Páez, School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada."
email: paezha@mcmaster.ca
abstract: "This data article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain. These data were originally published on idealista.com real estate website. The observations are obtained for the three largest Spanish cities: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations) and Valencia (n = 33,622 observations). The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics. The listings were enriched with official information (building year of construction and built quality materials grade) plus other relevant geographical features such as distance to urban points of interest. Along with real estate listings, the data product also includes neighborhood boundaries for each city. The data product is offered in the form of a fully documented R package. This open data product is available for scientific and educational purposes, in particular for geo-spatial studies."
keywords: Housing market; idealista.com; geo-referenced data; open data; hedonic price analysis; Spain
classoption:
  - Royal
  - times
bibliography: bibEPB.bib
bibliographystyle: sageh

output:
  rticles::sage_article:
    keep_tex: yes

header-includes:
  - \usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
---

# Introduction

El interés pos el desarrollo de modelos hedónicos de preicción del precio de lvivienda que incluyen la compenente espaciál,  y en general de la importancia de la geografía para realizar un correcto analisis del mercado inmobiliario, ha siod un tópico de creciente interes [@lopez2015; @crespo2013local].

Por otra parte la posibilidad de disponer de datos a nivel micro/urbano de la vivienda ha sido dificil y en algunos casso se ha tenido que recurrir a dudosos procesos de webscraping para poder obtener grandes volúmnees de informacion que permitan. En muchos casos estos procesos de webscrappin pueden contener datos faltantes, errores, datos duplicados etc,

El creciente interés por datos en abierto [@arribas2021]

The main objective of this paper is present a sort description of the an open data set of a geo-referenced dwelling listings.

This data set is excelente para aplicar modelos hedónicos con efectos espaciales, identificar submercados de vivienda, aplicar técnicas de machine learning.

La importancia de disponer de datos en abierto

Pocos datos en abierto de precios de vivienda

Pocos con informacion georeferenciada a nivel de punto

The data set is distributed in the form of an `R` package, named **idealista18** available from https://github.com/paezha/idealista18 . All spatial objects such as polygons and points are distributed as simple features objects (class `sf` in R). The data has been provided by Idealista, the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal.

POR QUE HACEN UNA CONTRIBUCIÓN A LA COMUNIDAD URBAN ANALITICS

# Data description

The open data set 'idealista18' is composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid and Valencia. For each city, dwelling listings, neighborhood polygons and a set of points of interest (POI) has been included in the R package. The next subsections describe each data set.

## Dwelling listings

The dwelling listing of each city, include a set of characteristics of each dwelling that was published on idealista real state website (https://www.idealista.com/) and are included in 'idealista18' package as an sf object [@Pebesma]. The data set corresponding to each city is named with the name of the city followed by '_Sale'. Each of these files contains the complete set of listings, corresponding to the four quarters of year 2018. The record counts for each city in 2018 are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for Valencia. Table \ref{tab:number-ads} show the number of ads included in the data set for city and quarter. Note that is possible that the same dwelling can be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter. The variable ASSETID, included in each sf object is the unique identifier of the dwelling.

```{r number-ads, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
H <- cbind(H,rowSums(H))
text_tbl <- data.frame(
City = c("Barcelona","Madrid","Valencia"),
H,
check.names = FALSE)
colnames(text_tbl) <- c("City", "First", "Second ", "Thirdr", "Fourth","Total ads")
xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "4em", italic = FALSE) %>%
column_spec(2, width = "3em") %>%
column_spec(3, width = "3em", latex_column_spec = "c") %>%
column_spec(4, width = "3em", latex_column_spec = "c") %>%
column_spec(5, width = "3em", latex_column_spec = "c") %>%
column_spec(6, width = "2em", latex_column_spec = "c")
```

Each record of the dwelling listing contains a set of indoor characteristics supplied by advertiser (e.g. price, surface, rooms, basic features, etc) beside with the exact localization of the dwelling (see Section \ref{sec:anonymizing}). Table \ref{tab:variables} list some of the main indoor variables included in the dwelling listing with a short description and the mean value of each variable. This dwelling listing was enriched with a number of additional attributes from the Spanish cadastre [@Catastro]. Cadastral information is described in Table \ref{tab:variables}, including the the prefix CAD in the variable name. Cadastral features assignment is done by assigning the features of the nearest parcel to the coordinates. The year of construction of dwelling was revised, given that original year of construction from listings are entered by users in the web site, therefore subject to errors and incomplete (a 40\% missing rate). To remove any issue we assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the field value was missing. Additionally, the distance of each dwelling to three urban points of interest was included in the data set: distance to city center, distance to the closed metro station and distance to the main street (Diagonal street for Barcelona, Castellana street for Madrid and Blasco Ibañez street for Valencia). The last rows of Table \ref{tab:variables} show the mean values of this variables.

```{r table-variables, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
dummyvariables <- c(3,4,5,6,7,27,30,31,32,33,37,38,39)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("ksjdhfal","Asking price per m^2 (euros)","Surface (m^2)",
                       "Number of bedrooms","Number of bathrooms",
                       "Construction year (advertiser)","Construction year (cadastre)",
                       "Max build floor","Dwelling count in the building",
                       "Cadastral quality. 0 Best-10 Worst",
                       "Distance to city center","Distance to subway station","Distance to main street"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables]), na.rm = TRUE),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables]), na.rm = TRUE),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables]), na.rm = TRUE),
check.names = FALSE)

xtable::xtable(text_tbl, caption = "List of quantitative variables included in the dwelling listing for the three Spanish cities. See the help facility in the \\textbf{idealista18} R package for details and formal definitions. Some variables has been excluded of this table for save space, check the full list in  \\textbf{idealista18} R package. \\label{tab:variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F, font_size=8) %>%
column_spec(1, width = "13em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

To conclude the description of the dwelling listing, Table \ref{tab:Dummy-variables} include the more relevant variables that include information about basic characteristics of the dwellings. All variables listed in this table are dummy variables and the percentaje of dwelling with the characteristis

```{r table-dummy-variables, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
dummyvariables <- c(8,9,10,12,15,16,17,18,19,20,21,22,23,24,25,26,34,35,36)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("=1 if has terrace","=1 if has lift",
                       "=1 if has air conditioning","=1 if has parking",
                       "=1 if has north orientation","=1 if has south orientation",
                       "=1 if has east orientation","=1 if has west orientation",
                       "=1 if has boxroom","=1 if has wardrobe",
                       "=1 if has swinningpool","=1 if has doorman",
                       "=1 if has garden","=1 if is duplex",
                       "=1 if is studio","=1 is in the top floor",
                       "=1 if is new","=1 is second hand to be restored","=1 is second hand good conditions"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables])),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables])),
check.names = FALSE)

xtable::xtable(text_tbl, caption = "List of dummy variables with the percentage of dwelling with a specific characteristic. See the help facility in the \\textbf{idealista18} R package for details and formal definitions. Some dummy variables has been excluded of this table for save space. \\label{tab:Dummy-variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F, font_size = 8) %>%
column_spec(1, width = "12em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

## Neighboorhood polygons

The second block of data included in the 'idealista18' R package are the spatial features of the three cities divided in neighborhoods. Figure \ref{fig:all-polygons} shows the different neighborhoods for the three cities. The boundaries are based on the official boundaries but slightly adapted by idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}. In practical terms we can assume they are the same, since the website simply collapses areas when they are sufficiently small in terms of number of ads. In the case of Madrid they just collapse four areas into two new ones.

A total of

```{r, echo = FALSE, fig.width=5, fig.cap="\\label{fig:all-polygons} (A) Weekly intra-province"}
library(ggplot2)
library(gridExtra)
plot1 <- ggplot(data=Barcelona_Polygons) + geom_sf(color = "black",size=.2) + 
  theme_bw(base_size = 2) + 
  ggtitle("                Barcelona") +
  theme(plot.title = element_text(size = 10),
        panel.border = element_blank(),
    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
plot2 <- ggplot(data=Madrid_Polygons) + geom_sf(color = "black",size=.2) + 
  theme_bw(base_size = 2) +
   ggtitle("                Madrid") +
    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
plot3 <- ggplot(data=Valencia_Polygons) + geom_sf(color = "black",size=.2) + 
  theme_bw(base_size = 2) +
   ggtitle("                Valencia") +
    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
grid.arrange(plot1, plot2, plot3, ncol = 3, nrow = 1)
```

There are a total of 73 neigborhoods in Barcelona, 135 in Madrid and 73 in Valencia. The sf object include an unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).

## Points of Interest

The last block of data included in the data package is a set of Point of Interest of each city in `R` list R objetc. These lists include three elements: (i) the coordinates of the city center (see that identify the central business district); (ii) a set of points that define the main street of each city; and (iii) the coordinates of metro stations.

# Anonymizing the data set {#sec:anonymizing}

To comply with Spanish regulations, two variables are slightly modified to preserve their anonymity. A masking process is apply to asking prices and localization.

Whit respect to the asking prices, the original values are obfuscated with the addition or subtraction of a random percentage of their original values ranging from -2.5\% to +2.5\%. Since asking prices are usually multiples of 1000, after the first price modification, the prices was aligned to multiples of 1000.

With respect to the dwelling localization, a spatial masking process was implemented with the intention of keeping spatial properties of the original data set. The coordinates of each listing were displaced using a stochastic procedure. Effectively, the listings were recorded using coordinates contained in a maximum and minimum displacement circles, as shown in Figure \ref{fig:points-moved-image}. To preserve membership in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates are in the original neighborhood of the listing.

```{=tex}
\begin{figure}[!ht]
  \caption{Masking coordinates. Spatial range}
  \centering
  \includegraphics[width=6cm, height=4.9cm]{EPB_files/points-moved-image.png}
  \label{fig:points-moved-image}
\end{figure}
```

The algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall in a different neighborhood. This ensures that neighborhood attributes are preserved.

```{=tex}
\begin{algorithm}[!ht]
 \KwData{all idealista listings}
 \KwResult{all idealista listings with masked coordinates}
 initialization\;
 \For{each listing L}{
  take geographical location of L as $(X,Y)$
  \Repeat{this stop condition}{
    take a random angle $\alpha$ from 0 to 360 degrees
    take a distance $R$ as a random value from 30 to 60 meters
    determine a new point $(X',Y')$ calculated as a point located $R$ with the angle $\alpha$
  }
  set $(X',Y')$ as the new location for the listing L
 }
 \caption{Coordinate displacement process for anonymisation purposes}
 \label{algo:coordinates-displacement}
\end{algorithm}
```

Figure \ref{fig:coordinates-displacement} shows the histogram of displacements in meters for all listings in the city of Valencia; the average distance between the original and masked coordinates is 45 meters.

```{=tex}
\begin{figure}[!ht]
  \caption{Masking coordinates. Spatial range}
  \centering
  \includegraphics[width=6cm, height=4.9cm]{EPB_files/coordinates-valencia.png}
  \label{fig:points-moved-image}
\end{figure}
```

## The geo-referenced micro-data set

Spatial objects include geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.

# Acknowledgments

The authors wish to thank Alessandro Galesi for their support in the paper revision and Juan Ramon Selva for collecting and cleaning the spatial data. This work has been partially funded by the Spanish Ministry of Economy and Competitiveness Grants PID2019-107800GB- 100

### 2022-02-05 EPB/EPB.tex
Nueva version para EP-B

--- 2022-02-05 EPB/EPB.Rmd
@@
----
-title: A geo-referenced micro-data set of real estate listings for the three largest Spanish cities
-runninghead: Rey-Blanco \emph{et al}.
-author:
-- name: D. Rey-Blanco
-  num: 1
-- name: P. González-Arbues
-  num: 1
-- name: F. López
-  num: 2
-- name: A. Páez*
-  num: 4
-address:
-- num: 1
-  org: Idealista, Plaza de las Cortes 5, 28014 Madrid, Spain
-- num: 2
-  org: Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia, Spain
-- num: 3
-  org: School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada
-corrauth: "Antonio Páez, School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada."
-email: paezha@mcmaster.ca
-abstract: "This data article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain.
-These data were originally published on idealista.com real estate website.
-The observations are obtained for the three largest Spanish cities: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations) and Valencia (n = 33,622 observations).
-The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics.
-The listings were enriched with official information (building year of construction and built quality materials grade) plus other relevant geographical features such as distance to urban points of interest.
-Along with real estate listings, the data product also includes neighborhood boundaries for each city.
-The data product is offered in the form of a fully documented R package.
-This open data product is available for scientific and educational purposes, in particular for geo-spatial studies."
-keywords: Housing market; idealista.com; geo-referenced data; open data; hedonic price analysis; Spain
-classoption:
-  - Royal
-  - times
-bibliography: bibEPB.bib
-bibliographystyle: sageh
+\documentclass[Royal,times,sageh]{sagej}
 
-output:
-  rticles::sage_article:
-    keep_tex: yes
+\usepackage{moreverb,url,natbib, multirow, tabularx}
+\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
 
-header-includes:
-  - \usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
----
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
-# Introduction
+\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
-El interés pos el desarrollo de modelos hedónicos de preicción del precio de lvivienda que incluyen la compenente espaciál,  y en general de la importancia de la geografía para realizar un correcto analisis del mercado inmobiliario, ha siod un tópico de creciente interes [@lopez2015; @crespo2013local].
+\begin{document}
 
-Por otra parte la posibilidad de disponer de datos a nivel micro/urbano de la vivienda ha sido dificil y en algunos casso se ha tenido que recurrir a dudosos procesos de webscraping para poder obtener grandes volúmnees de informacion que permitan.
-En muchos casos estos procesos de webscrappin pueden contener datos faltantes, errores, datos duplicados etc,
+\setcitestyle{aysep={,}}
 
-El creciente interés por datos en abierto [@arribas2021]
+\title{A geo-referenced micro-data set of real estate listings for the
+three largest Spanish cities}
 
-The main objective of this paper is present a sort description of the an open data set of a geo-referenced dwelling listings.
+\runninghead{Rey-Blanco \emph{et al}.}
 
-This data set is excelente para aplicar modelos hedónicos con efectos espaciales, identificar submercados de vivienda, aplicar técnicas de machine learning.
+\author{D. Rey-Blanco\affilnum{1}, P. González-Arbues\affilnum{1}, F.
+López\affilnum{2}, A. Páez*\affilnum{4}}
+
+\affiliation{\affilnum{1}{Idealista, Plaza de las Cortes 5, 28014
+Madrid, Spain}\\\affilnum{2}{Facultad de CC de la Empresa, C/ Real, 3.
+30201 Cartagena, Murcia, Spain}\\\affilnum{3}{School of Earth,
+Environment and Society, McMaster University, 1280 Main St W, Hamilton,
+Ontario L8S 4K1 Canada}}
+
+\corrauth{Antonio Páez, School of Earth, Environment and Society,
+McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada.}
+
+\email{\href{mailto:paezha@mcmaster.ca}{\nolinkurl{paezha@mcmaster.ca}}}
+
+\begin{abstract}
+This data article shares an open data product with big geo-referenced
+micro-data sets of 2018 real estate listings in Spain. These data were
+originally published on idealista.com real estate website. The
+observations are obtained for the three largest Spanish cities: Madrid
+(n = 94,815 observations), Barcelona (n = 61,486 observations) and
+Valencia (n = 33,622 observations). The data sets include the
+coordinates of properties (latitude and longitude), asking prices of
+each listed dwelling, and several variables of indoor characteristics.
+The listings were enriched with official information (building year of
+construction and built quality materials grade) plus other relevant
+geographical features such as distance to urban points of interest.
+Along with real estate listings, the data product also includes
+neighborhood boundaries for each city. The data product is offered in
+the form of a fully documented R package. This open data product is
+available for scientific and educational purposes, in particular for
+geo-spatial studies.
+\end{abstract}
+
+\keywords{Housing market; idealista.com; geo-referenced data; open data;
+hedonic price analysis; Spain}
+
+\maketitle
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+El interés pos el desarrollo de modelos hedónicos de preicción del
+precio de lvivienda que incluyen la compenente espaciál, y en general de
+la importancia de la geografía para realizar un correcto analisis del
+mercado inmobiliario, ha siod un tópico de creciente interes
+\citep{lopez2015, crespo2013local}.
+
+Por otra parte la posibilidad de disponer de datos a nivel micro/urbano
+de la vivienda ha sido dificil y en algunos casso se ha tenido que
+recurrir a dudosos procesos de webscraping para poder obtener grandes
+volúmnees de informacion que permitan. En muchos casos estos procesos de
+webscrappin pueden contener datos faltantes, errores, datos duplicados
+etc,
+
+El creciente interés por datos en abierto \citep{arribas2021}
+
+The main objective of this paper is present a sort description of the an
+open data set of a geo-referenced dwelling listings.
+
+This data set is excelente para aplicar modelos hedónicos con efectos
+espaciales, identificar submercados de vivienda, aplicar técnicas de
+machine learning.
 
@@
 
-The data set is distributed in the form of an `R` package, named **idealista18** available from https://github.com/paezha/idealista18 .
-All spatial objects such as polygons and points are distributed as simple features objects (class `sf` in R).
-The data has been provided by Idealista, the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal.
+The data set is distributed in the form of an \texttt{R} package, named
+\textbf{idealista18} available from
+\url{https://github.com/paezha/idealista18} . All spatial objects such
+as polygons and points are distributed as simple features objects (class
+\texttt{sf} in R). The data has been provided by Idealista, the major
+real estate listing website in Spain, and present in other southern
+european countries as Italy and Portugal.
 
@@
 
-# Data description
+\hypertarget{data-description}{%
+\section{Data description}\label{data-description}}
 
-The open data set 'idealista18' is composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid and Valencia.
-For each city, dwelling listings, neighborhood polygons and a set of points of interest (POI) has been included in the R package.
-The next subsections describe each data set.
+The open data set `idealista18' is composed of nine objects, three
+objects for each of the three main Spanish cities: Barcelona, Madrid and
+Valencia. For each city, dwelling listings, neighborhood polygons and a
+set of points of interest (POI) has been included in the R package. The
+next subsections describe each data set.
 
-## Dwelling listings
+\hypertarget{dwelling-listings}{%
+\subsection{Dwelling listings}\label{dwelling-listings}}
 
-The dwelling listing of each city, include a set of characteristics of each dwelling that was published on idealista real state website (https://www.idealista.com/) and are included in 'idealista18' package as an sf object [@Pebesma].
-The data set corresponding to each city is named with the name of the city followed by '_Sale'.
-Each of these files contains the complete set of listings, corresponding to the four quarters of year 2018.
-The record counts for each city in 2018 are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for Valencia.
-Table \ref{tab:number-ads} show the number of ads included in the data set for city and quarter.
-Note that is possible that the same dwelling can be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter.
-The variable ASSETID, included in each sf object is the unique identifier of the dwelling.
+The dwelling listing of each city, include a set of characteristics of
+each dwelling that was published on idealista real state website
+(\url{https://www.idealista.com/}) and are included in `idealista18'
+package as an sf object \citep{Pebesma}. The data set corresponding to
+each city is named with the name of the city followed by '\_Sale'. Each
+of these files contains the complete set of listings, corresponding to
+the four quarters of year 2018. The record counts for each city in 2018
+are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for
+Valencia. Table \ref{tab:number-ads} show the number of ads included in
+the data set for city and quarter. Note that is possible that the same
+dwelling can be found in more than one period when a property listed for
+sale in one quarter was sold in a subsequent quarter. The variable
+ASSETID, included in each sf object is the unique identifier of the
+dwelling.
 
-```{r number-ads, echo = FALSE, message = FALSE}
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
-H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
-H <- cbind(H,rowSums(H))
-text_tbl <- data.frame(
-City = c("Barcelona","Madrid","Valencia"),
-H,
-check.names = FALSE)
-colnames(text_tbl) <- c("City", "First", "Second ", "Thirdr", "Fourth","Total ads")
-xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") %>%
-xtable2kable(include.rownames = FALSE) %>%
-kable_styling(full_width = F) %>%
-column_spec(1, width = "4em", italic = FALSE) %>%
-column_spec(2, width = "3em") %>%
-column_spec(3, width = "3em", latex_column_spec = "c") %>%
-column_spec(4, width = "3em", latex_column_spec = "c") %>%
-column_spec(5, width = "3em", latex_column_spec = "c") %>%
-column_spec(6, width = "2em", latex_column_spec = "c")
-```
+\begin{table}[ht]
+\centering
+\begin{tabular}{>{\raggedright\arraybackslash}p{4em}>{\raggedleft\arraybackslash}p{3em}cccc}
+  \hline
+City & First & Second  & Thirdr & Fourth & Total ads \\ 
+  \hline
+Barcelona & 17826 & 7951 & 12375 & 23334 & 61486 \\ 
+  Madrid & 21920 & 12652 & 15973 & 44270 & 94815 \\ 
+  Valencia & 9305 & 4655 & 5644 & 14018 & 33622 \\ 
+   \hline
+\end{tabular}
+\caption{Number of dwelling  listing ads for each city and quarter. \label{tab:number-ads}} 
+\end{table}
 
-Each record of the dwelling listing contains a set of indoor characteristics supplied by advertiser (e.g.
-price, surface, rooms, basic features, etc) beside with the exact localization of the dwelling (see Section \ref{sec:anonymizing}).
-Table \ref{tab:variables} list some of the main indoor variables included in the dwelling listing with a short description and the mean value of each variable.
-This dwelling listing was enriched with a number of additional attributes from the Spanish cadastre [@Catastro].
-Cadastral information is described in Table \ref{tab:variables}, including the the prefix CAD in the variable name.
-Cadastral features assignment is done by assigning the features of the nearest parcel to the coordinates.
-The year of construction of dwelling was revised, given that original year of construction from listings are entered by users in the web site, therefore subject to errors and incomplete (a 40\% missing rate).
-To remove any issue we assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the field value was missing.
-Additionally, the distance of each dwelling to three urban points of interest was included in the data set: distance to city center, distance to the closed metro station and distance to the main street (Diagonal street for Barcelona, Castellana street for Madrid and Blasco Ibañez street for Valencia).
-The last rows of Table \ref{tab:variables} show the mean values of this variables.
+Each record of the dwelling listing contains a set of indoor
+characteristics supplied by advertiser (e.g.~price, surface, rooms,
+basic features, etc) beside with the exact localization of the dwelling
+(see Section \ref{sec:anonymizing}). Table \ref{tab:variables} list some
+of the main indoor variables included in the dwelling listing with a
+short description and the mean value of each variable. This dwelling
+listing was enriched with a number of additional attributes from the
+Spanish cadastre \citep{Catastro}. Cadastral information is described in
+Table \ref{tab:variables}, including the the prefix CAD in the variable
+name. Cadastral features assignment is done by assigning the features of
+the nearest parcel to the coordinates. The year of construction of
+dwelling was revised, given that original year of construction from
+listings are entered by users in the web site, therefore subject to
+errors and incomplete (a 40\% missing rate). To remove any issue we
+assign cadastral construction year from the nearest cadastral parcel
+whenever value has an outstanding value (date is after publication date
+or year of construction is before 1500) or when the field value was
+missing. Additionally, the distance of each dwelling to three urban
+points of interest was included in the data set: distance to city
+center, distance to the closed metro station and distance to the main
+street (Diagonal street for Barcelona, Castellana street for Madrid and
+Blasco Ibañez street for Valencia). The last rows of Table
+\ref{tab:variables} show the mean values of this variables.
 
-```{r table-variables, echo = FALSE, message = FALSE}
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
-dummyvariables <- c(3,4,5,6,7,27,30,31,32,33,37,38,39)
-text_tbl <- data.frame(
-Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
-"Sort Description" = c("ksjdhfal","Asking price per m^2 (euros)","Surface (m^2)",
-                       "Number of bedrooms","Number of bathrooms",
-                       "Construction year (advertiser)","Construction year (cadastre)",
-                       "Max build floor","Dwelling count in the building",
-                       "Cadastral quality. 0 Best-10 Worst",
-                       "Distance to city center","Distance to subway station","Distance to main street"),
-# Barcelona
-Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables]), na.rm = TRUE),
-# Madrid
-Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables]), na.rm = TRUE),
-# Valencia
-"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables]), na.rm = TRUE),
-check.names = FALSE)
+\begin{table}[ht]
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}{>{\raggedright\arraybackslash}p{13em}>{\raggedright\arraybackslash}p{14em}ccc}
+  \hline
+Variable & Sort Description & Barcelona & Madrid & Valencia \\ 
+  \hline
+PRICE & ksjdhfal & 395770.58 & 396110.11 & 199678.31 \\ 
+  UNITPRICE & Asking price per m\verb|^|2 (euros) & 4044.86 & 3661.05 & 1714.54 \\ 
+  CONSTRUCTEDAREA & Surface (m\verb|^|2) & 95.46 & 101.40 & 108.95 \\ 
+  ROOMNUMBER & Number of bedrooms & 2.86 & 2.58 & 3.07 \\ 
+  BATHNUMBER & Number of bathrooms & 1.52 & 1.59 & 1.59 \\ 
+  CONSTRUCTIONYEAR & Construction year (advertiser) & 1952.58 & 1964.69 & 1969.43 \\ 
+  CADCONSTRUCTIONYEAR & Construction year (cadastre) & 1952.19 & 1965.70 & 1970.55 \\ 
+  CADMAXBUILDINGFLOOR & Max build floor & 6.85 & 6.38 & 7.04 \\ 
+  CADDWELLINGCOUNT & Dwelling count in the building & 28.56 & 39.19 & 36.83 \\ 
+  CADASTRALQUALITYID & Cadastral quality. 0 Best-10 Worst & 4.31 & 4.85 & 5.34 \\ 
+  DISTANCE\_TO\_CITY\_CENTER & Distance to city center & 2.80 & 4.49 & 2.09 \\ 
+  DISTANCE\_TO\_METRO & Distance to subway station & 0.27 & 0.48 & 0.64 \\ 
+  DISTANCE\_TO\_DIAGONAL & Distance to main street & 1.77 & 2.68 & 2.07 \\ 
+   \hline
+\end{tabular}
+\caption{List of quantitative variables included in the dwelling listing for the three Spanish cities.
+See the help facility in the \textbf{idealista18} R package for details and formal definitions.
+Some variables has been excluded of this table for save space, check the full list in  \textbf{idealista18} R package.
+\label{tab:variables}}
+\end{table}
 
-xtable::xtable(text_tbl, caption = "List of quantitative variables included in the dwelling listing for the three Spanish cities.
-See the help facility in the \\textbf{idealista18} R package for details and formal definitions.
-Some variables has been excluded of this table for save space, check the full list in  \\textbf{idealista18} R package.
-\\label{tab:variables}") %>%
-xtable2kable(include.rownames = FALSE) %>%
-kable_styling(full_width = F, font_size=8) %>%
-column_spec(1, width = "13em", italic = FALSE) %>%
-column_spec(2, width = "14em") %>%
-column_spec(3, width = "2em", latex_column_spec = "c") %>%
-column_spec(4, width = "2em", latex_column_spec = "c") %>%
-column_spec(5, width = "2em", latex_column_spec = "c")
-```
+To conclude the description of the dwelling listing, Table
+\ref{tab:Dummy-variables} include the more relevant variables that
+include information about basic characteristics of the dwellings. All
+variables listed in this table are dummy variables and the percentaje of
+dwelling with the characteristis
 
-To conclude the description of the dwelling listing, Table \ref{tab:Dummy-variables} include the more relevant variables that include information about basic characteristics of the dwellings.
-All variables listed in this table are dummy variables and the percentaje of dwelling with the characteristis
+\begin{table}[ht]
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}{>{\raggedright\arraybackslash}p{12em}>{\raggedright\arraybackslash}p{14em}ccc}
+  \hline
+Variable & Sort Description & Barcelona & Madrid & Valencia \\ 
+  \hline
+HASTERRACE & =1 if has terrace & 0.33 & 0.36 & 0.25 \\ 
+  HASLIFT & =1 if has lift & 0.74 & 0.70 & 0.79 \\ 
+  HASAIRCONDITIONING & =1 if has air conditioning & 0.47 & 0.45 & 0.47 \\ 
+  HASPARKINGSPACE & =1 if has parking & 0.08 & 0.23 & 0.17 \\ 
+  HASNORTHORIENTATION & =1 if has north orientation & 0.13 & 0.11 & 0.13 \\ 
+  HASSOUTHORIENTATION & =1 if has south orientation & 0.31 & 0.24 & 0.19 \\ 
+  HASEASTORIENTATION & =1 if has east orientation & 0.24 & 0.20 & 0.25 \\ 
+  HASWESTORIENTATION & =1 if has west orientation & 0.16 & 0.15 & 0.15 \\ 
+  HASBOXROOM & =1 if has boxroom & 0.12 & 0.26 & 0.13 \\ 
+  HASWARDROBE & =1 if has wardrobe & 0.30 & 0.57 & 0.53 \\ 
+  HASSWIMMINGPOOL & =1 if has swinningpool & 0.03 & 0.15 & 0.07 \\ 
+  HASDOORMAN & =1 if has doorman & 0.08 & 0.25 & 0.05 \\ 
+  HASGARDEN & =1 if has garden & 0.04 & 0.18 & 0.06 \\ 
+  ISDUPLEX & =1 if is duplex & 0.03 & 0.03 & 0.02 \\ 
+  ISSTUDIO & =1 if is studio & 0.02 & 0.03 & 0.01 \\ 
+  ISINTOPFLOOR & =1 is in the top floor & 0.02 & 0.02 & 0.01 \\ 
+  BUILTTYPEID\_1 & =1 if is new & 0.01 & 0.03 & 0.03 \\ 
+  BUILTTYPEID\_2 & =1 is second hand to be restored & 0.17 & 0.19 & 0.13 \\ 
+  BUILTTYPEID\_3 & =1 is second hand good conditions & 0.82 & 0.78 & 0.83 \\ 
+   \hline
+\end{tabular}
+\caption{List of dummy variables with the percentage of dwelling with a specific characteristic.
+See the help facility in the \textbf{idealista18} R package for details and formal definitions.
+Some dummy variables has been excluded of this table for save space.
+\label{tab:Dummy-variables}}
+\end{table}
 
-```{r table-dummy-variables, echo = FALSE, message = FALSE}
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
-dummyvariables <- c(8,9,10,12,15,16,17,18,19,20,21,22,23,24,25,26,34,35,36)
-text_tbl <- data.frame(
-Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
-"Sort Description" = c("=1 if has terrace","=1 if has lift",
-                       "=1 if has air conditioning","=1 if has parking",
-                       "=1 if has north orientation","=1 if has south orientation",
-                       "=1 if has east orientation","=1 if has west orientation",
-                       "=1 if has boxroom","=1 if has wardrobe",
-                       "=1 if has swinningpool","=1 if has doorman",
-                       "=1 if has garden","=1 if is duplex",
-                       "=1 if is studio","=1 is in the top floor",
-                       "=1 if is new","=1 is second hand to be restored","=1 is second hand good conditions"),
-# Barcelona
-Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
-# Madrid
-Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables])),
-# Valencia
-"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables])),
-check.names = FALSE)
+\hypertarget{neighboorhood-polygons}{%
+\subsection{Neighboorhood polygons}\label{neighboorhood-polygons}}
 
-xtable::xtable(text_tbl, caption = "List of dummy variables with the percentage of dwelling with a specific characteristic.
-See the help facility in the \\textbf{idealista18} R package for details and formal definitions.
-Some dummy variables has been excluded of this table for save space.
-\\label{tab:Dummy-variables}") %>%
-xtable2kable(include.rownames = FALSE) %>%
-kable_styling(full_width = F, font_size = 8) %>%
-column_spec(1, width = "12em", italic = FALSE) %>%
-column_spec(2, width = "14em") %>%
-column_spec(3, width = "2em", latex_column_spec = "c") %>%
-column_spec(4, width = "2em", latex_column_spec = "c") %>%
-column_spec(5, width = "2em", latex_column_spec = "c")
-```
-
-## Neighboorhood polygons
-
-The second block of data included in the 'idealista18' R package are the spatial features of the three cities divided in neighborhoods.
-Figure \ref{fig:all-polygons} shows the different neighborhoods for the three cities.
-The boundaries are based on the official boundaries but slightly adapted by idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
-In practical terms we can assume they are the same, since the website simply collapses areas when they are sufficiently small in terms of number of ads.
-In the case of Madrid they just collapse four areas into two new ones.
+The second block of data included in the `idealista18' R package are the
+spatial features of the three cities divided in neighborhoods. Figure
+\ref{fig:all-polygons} shows the different neighborhoods for the three
+cities. The boundaries are based on the official boundaries but slightly
+adapted by
+idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
+In practical terms we can assume they are the same, since the website
+simply collapses areas when they are sufficiently small in terms of
+number of ads. In the case of Madrid they just collapse four areas into
+two new ones.
 
@@
 
-```{r, echo = FALSE, fig.width=5, fig.cap="\\label{fig:all-polygons} (A) Weekly intra-province"}
-library(ggplot2)
-library(gridExtra)
-plot1 <- ggplot(data=Barcelona_Polygons) + geom_sf(color = "black",size=.2) + 
-  theme_bw(base_size = 2) + 
-  ggtitle("                Barcelona") +
-  theme(plot.title = element_text(size = 10),
-        panel.border = element_blank(),
-    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
-plot2 <- ggplot(data=Madrid_Polygons) + geom_sf(color = "black",size=.2) + 
-  theme_bw(base_size = 2) +
-   ggtitle("                Madrid") +
-    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
-    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
-plot3 <- ggplot(data=Valencia_Polygons) + geom_sf(color = "black",size=.2) + 
-  theme_bw(base_size = 2) +
-   ggtitle("                Valencia") +
-    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
-    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
-grid.arrange(plot1, plot2, plot3, ncol = 3, nrow = 1)
-```
+\begin{figure}
+\centering
+\includegraphics{EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
+\caption{\label{fig:all-polygons} (A) Weekly intra-province}
+\end{figure}
 
-There are a total of 73 neigborhoods in Barcelona, 135 in Madrid and 73 in Valencia. The sf object include an unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).
+There are a total of 73 neigborhoods in Barcelona, 135 in Madrid and 73
+in Valencia. The sf object include an unique identifier (LOCATIONID) and
+the neighborhood name (LOCATIONNAME).
 
-## Points of Interest
+\hypertarget{points-of-interest}{%
+\subsection{Points of Interest}\label{points-of-interest}}
 
-The last block of data included in the data package is a set of Point of Interest of each city in `R` list R objetc.
-These lists include three elements: (i) the coordinates of the city center (see that identify the central business district); (ii) a set of points that define the main street of each city; and (iii) the coordinates of metro stations.
+The last block of data included in the data package is a set of Point of
+Interest of each city in \texttt{R} list R objetc. These lists include
+three elements: (i) the coordinates of the city center (see that
+identify the central business district); (ii) a set of points that
+define the main street of each city; and (iii) the coordinates of metro
+stations.
 
-# Anonymizing the data set {#sec:anonymizing}
+\hypertarget{sec:anonymizing}{%
+\section{Anonymizing the data set}\label{sec:anonymizing}}
 
-To comply with Spanish regulations, two variables are slightly modified to preserve their anonymity. A masking process is apply to asking prices and localization.
+To comply with Spanish regulations, two variables are slightly modified
+to preserve their anonymity. A masking process is apply to asking prices
+and localization.
 
-Whit respect to the asking prices, the original values are obfuscated with the addition or subtraction of a random percentage of their original values ranging from -2.5\% to +2.5\%.
-Since asking prices are usually multiples of 1000, after the first price modification, the prices was aligned to multiples of 1000.
+Whit respect to the asking prices, the original values are obfuscated
+with the addition or subtraction of a random percentage of their
+original values ranging from -2.5\% to +2.5\%. Since asking prices are
+usually multiples of 1000, after the first price modification, the
+prices was aligned to multiples of 1000.
 
-With respect to the dwelling localization, a spatial masking process was implemented with the intention of keeping spatial properties of the original data set.
-The coordinates of each listing were displaced using a stochastic procedure.
-Effectively, the listings were recorded using coordinates contained in a maximum and minimum displacement circles, as shown in Figure \ref{fig:points-moved-image}.
-To preserve membership in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates are in the original neighborhood of the listing.
+With respect to the dwelling localization, a spatial masking process was
+implemented with the intention of keeping spatial properties of the
+original data set. The coordinates of each listing were displaced using
+a stochastic procedure. Effectively, the listings were recorded using
+coordinates contained in a maximum and minimum displacement circles, as
+shown in Figure \ref{fig:points-moved-image}. To preserve membership in
+a neighborhood, the spatial masking procedure was constrained to ensure
+that the masked coordinates are in the original neighborhood of the
+listing.
 
-```{=tex}
 \begin{figure}[!ht]
@@
 \end{figure}
-```
 
-The algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall in a different neighborhood.
-This ensures that neighborhood attributes are preserved.
+The algorithm \ref{algo:coordinates-displacement} iteratively displaces
+the coordinates of each listing with a minimum distance and a maximum
+distance with the restriction that the new coordinates do not fall in a
+different neighborhood. This ensures that neighborhood attributes are
+preserved.
 
-```{=tex}
 \begin{algorithm}[!ht]
@@
 \end{algorithm}
-```
 
-Figure \ref{fig:coordinates-displacement} shows the histogram of displacements in meters for all listings in the city of Valencia; the average distance between the original and masked coordinates is 45 meters.
+Figure \ref{fig:coordinates-displacement} shows the histogram of
+displacements in meters for all listings in the city of Valencia; the
+average distance between the original and masked coordinates is 45
+meters.
 
-```{=tex}
 \begin{figure}[!ht]
@@
 \end{figure}
-```
 
-## The geo-referenced micro-data set
+\hypertarget{the-geo-referenced-micro-data-set}{%
+\subsection{The geo-referenced micro-data
+set}\label{the-geo-referenced-micro-data-set}}
 
-Spatial objects include geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.
+Spatial objects include geodetic coordinates using the \emph{EPSG:4326}
+coordinate reference system.
 
-# Acknowledgments
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
 
-The authors wish to thank Alessandro Galesi for their support in the paper revision and Juan Ramon Selva for collecting and cleaning the spatial data.
-This work has been partially funded by the Spanish Ministry of Economy and Competitiveness Grants PID2019-107800GB- 100
+The authors wish to thank Alessandro Galesi for their support in the
+paper revision and Juan Ramon Selva for collecting and cleaning the
+spatial data. This work has been partially funded by the Spanish
+Ministry of Economy and Competitiveness Grants PID2019-107800GB- 100
+
+\bibliographystyle{sageh}
+\bibliography{bibEPB.bib}
+
+\end{document}

### 2022-02-05 data-in-brief/Tables and Figures.Rmd
Nueva version para EP-B

---
title: "Tables and Figures"
author: "Fernando & Compay"
date: "2/4/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Descriptive table

```{r total, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
H <-rbind(H,colSums(H))
text_tbl <- data.frame(
City = c("Barcelona","Madrid","Valencia","Total Ads"),
H,
check.names = FALSE)

xtable::xtable(text_tbl, digits=0, caption = "List of dummy variables with the percentage of dwelling with a specific characteristic. See the help facility in the \\textbf{idealista18} R package for details and formal definitions. Some dummy variables has been excluded of this table for save space. \\label{tab:Dummy-variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "12em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

```{r table-dummy-variables, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
dummyvariables <- c(4,5,6,7,27,30,31,37,38,39)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("Asking price per m^2 (euros)","Surface (m^2)",
                       "Number of bedrooms","Number of bathrooms",
                       "Año construcción","Construction year",
                       "Max build floor",
                       "Distance to city center","Distance to subway station","Distance to main street"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables])),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables])),
check.names = FALSE)

xtable::xtable(text_tbl, caption = "List of dummy variables with the percentage of dwelling with a specific characteristic. See the help facility in the \\textbf{idealista18} R package for details and formal definitions. Some dummy variables has been excluded of this table for save space. \\label{tab:Dummy-variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "12em", italic = FALSE) %>%
column_spec(2, width = "14em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

```{r table-variables, echo = FALSE, message = FALSE}
library(idealista18)
library(kableExtra)
library(xtable)
library(sf)
dummyvariables <- c(8,9,10,12,13,15)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("=1 if has terrace","=1 if has lift","=1 if has AA",
                       "=1 if has parking","=1 if has parking include in price","=1 if has north orientation"),
# Barcelona
mean = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
# Madrid
mean = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
# Valencia
"mean"=colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
check.names = FALSE)

xtable::xtable(text_tbl, caption = "List of dummy varaibles. See the help facility in the \\textbf{idealista18} package for details and formal definitions \\label{tab:Variables}") %>%
xtable2kable(include.rownames = FALSE) %>%
kable_styling(full_width = F) %>%
column_spec(1, width = "18em", italic = FALSE) %>%
column_spec(2, width = "16em") %>%
column_spec(3, width = "2em", latex_column_spec = "c") %>%
column_spec(4, width = "2em", latex_column_spec = "c") %>%
column_spec(5, width = "2em", latex_column_spec = "c")
```

### 2022-02-05 R/idealista18.R

@@ -34,8 +34,6 @@ NULL
 #'   \item{PERIOD}{Date YYYYMM, indicates the quarter when the ad was extracted we used YYYY03 for the 1st Q, YYYY06 the 2nd, YYYY09 for the 3rd and YYYY12 for the 4th}
 #'   \item{PRICE}{Asking price for the ad at idealista expressed in euros}
 #'   \item{UNITPRICE}{Price in euros per square meter}
-#'   \item{ADTYPOLOGYID}{Asset Typology (home, chalet)}
-#'   \item{ADOPERATIONID}{Asset Operation (rent, sales)}
 #'   \item{CONSTRUCTEDAREA}{Home built area in sq.m}
 #'   \item{ROOMNUMBER}{Number of bedrooms}
 #'   \item{BATHNUMBER}{Number of bathrooms}
@@ -95,9 +93,7 @@ NULL
 #'   \item{ASSETID}{Unique identifier of the asset}
 #'   \item{PERIOD}{Date YYYYMM, indicates the quarter when the ad was extracted we used YYYY03 for the 1st Q, YYYY06 the 2nd, YYYY09 for the 3rd and YYYY12 for the 4th}
 #'   \item{PRICE}{Price at idealista expressed in euros}
-#'   \item{UNITPRICE}{Price in euros per square meter}
-#'   \item{ADTYPOLOGYID}{Asset Typology (home, chalet)}
-#'   \item{ADOPERATIONID}{Asset Operation (rent, sales)}
+#'   \item{UNITPRICE}{Price in euros per square meter}}
 #'   \item{CONSTRUCTEDAREA}{Home built area in sq.m}
 #'   \item{ROOMNUMBER}{Number of bedrooms}
 #'   \item{BATHNUMBER}{Number of bathrooms}
@@ -158,8 +154,6 @@ NULL
 #'   \item{PERIOD}{Date YYYYMM, indicates the quarter when the ad was extracted we used YYYY03 for the 1st Q, YYYY06 the 2nd, YYYY09 for the 3rd and YYYY12 for the 4th}
 #'   \item{PRICE}{Price at idealista expressed in euros}
 #'   \item{UNITPRICE}{Price in euros per square meter}
-#'   \item{ADTYPOLOGYID}{Asset Typology (home, chalet)}
-#'   \item{ADOPERATIONID}{Asset Operation (rent, sales)}
[…]

### 2022-02-06 EPB/EPB.Rmd
Segunda version de EP-B

@@
 The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics.
-The listings were enriched with official information (building year of construction and built quality materials grade) plus other relevant geographical features such as distance to urban points of interest.
+The listings were enriched with official information from the Spanish cadastre (e.g.
+built quality materials grade) plus other relevant geographical features such as distance to urban points of interest.
 Along with real estate listings, the data product also includes neighborhood boundaries for each city.
@@
 This open data product is available for scientific and educational purposes, in particular for geo-spatial studies."
-keywords: Housing market; idealista.com; geo-referenced data; open data; hedonic price analysis; Spain
+keywords: Housing market; idealista.com; geo-referenced data; point level data; open data; hedonic price analysis; Spain
 classoption:
@@
 
-El interés pos el desarrollo de modelos hedónicos de preicción del precio de lvivienda que incluyen la compenente espaciál,  y en general de la importancia de la geografía para realizar un correcto analisis del mercado inmobiliario, ha siod un tópico de creciente interes [@lopez2015; @crespo2013local].
+The interest about the determinants of housing market and housing prices has been a growing research area in the last decades, with a vast literature both theoretical as well as empirical.
+Include the spatial component to analyze the real state market, and in general the incorporate geographic variables, has been significant improvements in the understanding of this market.
+But to really understand the determinants of housing market it is essential to have information/data at level-point.
+Therefore, it is increasingly common that the spatial analysis to be development in urban environment with geo-referenced micro-data sets [e.g.
+@lopez2015; @crespo2013local].
+But on the other hand, the availability of this type of open data at point level is limited and not to much data set contain latitude/longitude coordinates of each dwelling.
+In some cases the researchers it has had to resort to web scraping processes in order to obtain large volumes of information that allow get robust analysis [@gupta2022take; @arbia2020spatial; @Li2019; @lopez2015].
+These web scraping processes can include a lot of missing data, download errors, duplicate records, etc.
+Furthermore, in general the authors of this researches do not share the data sets.
 
-Por otra parte la posibilidad de disponer de datos a nivel micro/urbano de la vivienda ha sido dificil y en algunos casso se ha tenido que recurrir a dudosos procesos de webscraping para poder obtener grandes volúmnees de informacion que permitan.
-En muchos casos estos procesos de webscrappin pueden contener datos faltantes, errores, datos duplicados etc,
+In the same vein, in the last years there are a growing interest in open data in geography and data science [@arribasl2021editorial; @arribas2021] and in general by the reproducible or replicable research [@paez2021open].
+But to make open science is necessary have free software and open data.
+While great efforts have been made to make free software available to researchers (e.g.
+R or Python), not much data is out in the open.
+In the particular case of the real estate market, to our knowledge, no to much open micro-data set of housing markets are available.
 
-El creciente interés por datos en abierto [@arribas2021]
+To overcome these limitations, this paper present a sort description of the an open micro-data set of a geo-referenced dwelling listings.
+The data has been provided by Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information of 189,923 dwelling localized in the three major Spanish cities.
+To the date, this data product is the bigger open geo-referenced data set of housing market in Spain.
+Moreover, note that the data set is supply directly by Idealista and therefore is clean and free of download errors.
+The listings has been enriched with official information from the Spanish cadastre plus other relevant geographical features such as distance to urban points of interest.
+The data set is distributed in the form of an `R` package, named **idealista18** that can be accessed at Github repository\footnote{The direct URL to the Github repository is https://github.com/paezha/idealista18}.
 
-The main objective of this paper is present a sort description of the an open data set of a geo-referenced dwelling listings.
-
-This data set is excelente para aplicar modelos hedónicos con efectos espaciales, identificar submercados de vivienda, aplicar técnicas de machine learning.
-
-La importancia de disponer de datos en abierto
-
-Pocos datos en abierto de precios de vivienda
-
-Pocos con informacion georeferenciada a nivel de punto
-
-The data set is distributed in the form of an `R` package, named **idealista18** available from https://github.com/paezha/idealista18 .
-All spatial objects such as polygons and points are distributed as simple features objects (class `sf` in R).
-The data has been provided by Idealista, the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal.
-
-POR QUE HACEN UNA CONTRIBUCIÓN A LA COMUNIDAD URBAN ANALITICS
+<!-- @cohen2008spatial (508 observations) -->
+<!-- @lazrak2014market (19000 observations) -->
 
@@
 
-The open data set 'idealista18' is composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid and Valencia.
+The open data set 'idealista18' is R package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid and Valencia.
 For each city, dwelling listings, neighborhood polygons and a set of points of interest (POI) has been included in the R package.
-The next subsections describe each data set.
+The next subsections describe each object.
+A full description of the data is available in the helps of the package.
 
@@
 The dwelling listing of each city, include a set of characteristics of each dwelling that was published on idealista real state website (https://www.idealista.com/) and are included in 'idealista18' package as an sf object [@Pebesma].
-The data set corresponding to each city is named with the name of the city followed by '_Sale'.
-Each of these files contains the complete set of listings, corresponding to the four quarters of year 2018.
+The data set corresponding to each city is named with the name of the city followed by '_Sale' (e.g.
+Madrid_Sale) and include a total of 42 variables.
+Each sf objects contains the complete set of listings, corresponding to the four quarters of year 2018.
 The record counts for each city in 2018 are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for Valencia.
-Table \ref{tab:number-ads} show the number of ads included in the data set for city and quarter.
+Table \ref{tab:number-ads} show the number of dwelling listing ads included in the data set for city and quarter.
 Note that is possible that the same dwelling can be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter.
-The variable ASSETID, included in each sf object is the unique identifier of the dwelling.
+The variable ASSETID, included in the sf objects is the unique identifier of the dwelling.
 
@@
 
-Each record of the dwelling listing contains a set of indoor characteristics supplied by advertiser (e.g.
-price, surface, rooms, basic features, etc) beside with the exact localization of the dwelling (see Section \ref{sec:anonymizing}).
+Each record of the dwelling listing contains a set of indoor characteristics supplied by advertiser in the Idealista web (e.g.
+price, surface, rooms, basic features, etc) including the exact localization of the dwelling (see Section [Anonymizing the data set](#anonymizing)).
 Table \ref{tab:variables} list some of the main indoor variables included in the dwelling listing with a short description and the mean value of each variable.
@@
 Cadastral features assignment is done by assigning the features of the nearest parcel to the coordinates.
-The year of construction of dwelling was revised, given that original year of construction from listings are entered by users in the web site, therefore subject to errors and incomplete (a 40\% missing rate).
-To remove any issue we assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the field value was missing.
-Additionally, the distance of each dwelling to three urban points of interest was included in the data set: distance to city center, distance to the closed metro station and distance to the main street (Diagonal street for Barcelona, Castellana street for Madrid and Blasco Ibañez street for Valencia).
+The year of construction of dwelling (CONSTRUCTIONYEAR) supply by the advertiser was revised given that the year of construction are entered by users in the web site and therefore subject to errors and incomplete data (a 40\% missing rate).
+To solve this issue we include an alternative variable (CADCONSTRUCTIONYEAR) assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the value supply by the advertiser was missing.
+Additionally, the distance of each dwelling to three urban points of interest was included in the dwelling listing: distance to city center, distance to the closed metro station and distance to the main street (Diagonal street for Barcelona, Castellana street for Madrid and Blasco Ibañez street for Valencia).
 The last rows of Table \ref{tab:variables} show the mean values of this variables.
@@
 Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
-"Sort Description" = c("ksjdhfal","Asking price per m^2 (euros)","Surface (m^2)",
+"Sort Description" = c("Asking price","Asking price per m^2 (euros)","Surface (m^2)",
                        "Number of bedrooms","Number of bathrooms",
@@
 check.names = FALSE)
-
-xtable::xtable(text_tbl, caption = "List of quantitative variables included in the dwelling listing for the three Spanish cities.
+text_tbl[dim(text_tbl)[1],1]<-"DISTANCE_TO_(MAINSTREET)"
+xtable::xtable(text_tbl, caption = "List, sort description and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities.
 See the help facility in the \\textbf{idealista18} R package for details and formal definitions.
@@
 
-To conclude the description of the dwelling listing, Table \ref{tab:Dummy-variables} include the more relevant variables that include information about basic characteristics of the dwellings.
-All variables listed in this table are dummy variables and the percentaje of dwelling with the characteristis
+In addition to the variables listed in Table \ref{tab:variables} the sf object include a set of dummy variables with information abour basic characteristics of the dwelling.
+Table \ref{tab:Dummy-variables} show the more relevant variables included in the sf object.
 
@@
                        "=1 if is studio","=1 is in the top floor",
-                       "=1 if is new","=1 is second hand to be restored","=1 is second hand good conditions"),
+                       "=1 if is new contruction","=1 is second hand to be restored","=1 is second hand good conditions"),
 # Barcelona
@@
 
-xtable::xtable(text_tbl, caption = "List of dummy variables with the percentage of dwelling with a specific characteristic.
+xtable::xtable(text_tbl, caption = "List of dummy variables, sort description and ratio of dwelling with the specific characteristic.
 See the help facility in the \\textbf{idealista18} R package for details and formal definitions.
@@
 The second block of data included in the 'idealista18' R package are the spatial features of the three cities divided in neighborhoods.
-Figure \ref{fig:all-polygons} shows the different neighborhoods for the three cities.
-The boundaries are based on the official boundaries but slightly adapted by idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
+There are an sf object for each city with the name of the city and the suffix '_Polygons' (e.g.
+Valencia_Polygons).
+The Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities.
+The neighborhoods are based on the official boundaries but slightly adapted by Idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
 In practical terms we can assume they are the same, since the website simply collapses areas when they are sufficiently small in terms of number of ads.
@@
 
-A total of
+```{r, echo = FALSE,message=FALSE,fig.width=5, fig.cap="\\label{fig:all-polygons}Quantile maps of the number of dwellings in each neighborhood. Boundary for Madrid, Barcelona and Valencia"}
 
-```{r, echo = FALSE, fig.width=5, fig.cap="\\label{fig:all-polygons} (A) Weekly intra-province"}
-library(ggplot2)
-library(gridExtra)
-plot1 <- ggplot(data=Barcelona_Polygons) + geom_sf(color = "black",size=.2) + 
-  theme_bw(base_size = 2) + 
-  ggtitle("                Barcelona") +
-  theme(plot.title = element_text(size = 10),
-        panel.border = element_blank(),
-    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
-plot2 <- ggplot(data=Madrid_Polygons) + geom_sf(color = "black",size=.2) + 
-  theme_bw(base_size = 2) +
-   ggtitle("                Madrid") +
-    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
-    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
-plot3 <- ggplot(data=Valencia_Polygons) + geom_sf(color = "black",size=.2) + 
-  theme_bw(base_size = 2) +
-   ggtitle("                Valencia") +
-    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
-    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
-grid.arrange(plot1, plot2, plot3, ncol = 3, nrow = 1)
+library("ggplot2")
+library("dplyr")
+library("gridExtra")
+# Barcelona
+Barcelona_counts <- Barcelona_Polygons %>% 
+  mutate(counts = lengths(st_intersects(., Barcelona_Sale)))
+q <- quantile(Barcelona_counts$counts)
+Barcelona_Polygons$Quantile<- as.factor((Barcelona_counts$counts > q[2]) + (Barcelona_counts$counts > q[3]) +(Barcelona_counts$counts >= q[4]) + 1)
+plot4 <- ggplot(data = Barcelona_Polygons) +
+  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
+  theme_bw(base_size = 6) +
+  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
+  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
+        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
+# Madrid
+Madrid_counts <- Madrid_Polygons %>% 
+  mutate(counts = lengths(st_intersects(., Madrid_Sale)))
+q <- quantile(Madrid_counts$counts)
+Madrid_Polygons$Quantile<- as.factor((Madrid_counts$counts > q[2]) + (Madrid_counts$counts > q[3]) +(Madrid_counts$counts >= q[4]) + 1)
+plot5 <- ggplot(data = Madrid_Polygons) +
+  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
+  theme_bw(base_size = 6) +
+  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
+  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
+        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
+# Valencia
+Valencia_counts <- Valencia_Polygons %>% 
+  mutate(counts = lengths(st_intersects(., Valencia_Sale)))
+q <- quantile(Valencia_counts$counts)
+Valencia_Polygons$Quantile<- as.factor((Valencia_counts$counts > q[2]) + (Valencia_counts$counts > q[3]) +(Valencia_counts$counts >= q[4]) + 1)
+plot6 <- ggplot(data = Valencia_Polygons) +
+  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
+  theme_bw(base_size = 6) +
+  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
+  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
+        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
+####################
+# plot1 <- ggplot(data=Barcelona_Polygons) + geom_sf(color = "black",size=.2) + 
+#   theme_bw(base_size = 2) + 
+#   ggtitle("                Barcelona") +
+#   theme(plot.title = element_text(size = 10),
+#         panel.border = element_blank(),
+#     axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
+# plot2 <- ggplot(data=Madrid_Polygons) + geom_sf(color = "black",size=.2) + 
+#   theme_bw(base_size = 2) +
+#    ggtitle("                Madrid") +
+#     theme(plot.title = element_text(size = 10),panel.border = element_blank(),
+#     axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
+# plot3 <- ggplot(data=Valencia_Polygons) + geom_sf(color = "black",size=.2) + 
+#   theme_bw(base_size = 2) +
+#    ggtitle("                Valencia") +
+#     theme(plot.title = element_text(size = 10),panel.border = element_blank(),
+#     axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
+grid.arrange(plot4, plot5, plot6, ncol = 3, nrow = 1)
 ```
 
-There are a total of 73 neigborhoods in Barcelona, 135 in Madrid and 73 in Valencia. The sf object include an unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).
+There are a total of 69 neighborhoods in Barcelona, 135 in Madrid and 73 in Valencia. The sf object include an unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).
 
@@
 
-The last block of data included in the data package is a set of Point of Interest of each city in `R` list R objetc.
-These lists include three elements: (i) the coordinates of the city center (see that identify the central business district); (ii) a set of points that define the main street of each city; and (iii) the coordinates of metro stations.
+The last block of data included in the data package is a set of Point of Interest of each city in `R` list R object.
+The name of the list include the name of the city with the suffix '_POIS' (e.g.
+Barcelona_POIS).
+These lists include three elements: (i) the coordinates of the city center, identifying the central business district; (ii) a set of points that define the main street of each city; and (iii) the coordinates of metro stations.
 
-# Anonymizing the data set {#sec:anonymizing}
+# Anonymizing the data set {#anonymizing}
 
-To comply with Spanish regulations, two variables are slightly modified to preserve their anonymity. A masking process is apply to asking prices and localization.
+To comply with Spanish regulations, two variables are slightly modified to preserve their anonymity. A masking process is apply to asking prices and localization (coordinates).
 
 Whit respect to the asking prices, the original values are obfuscated with the addition or subtraction of a random percentage of their original values ranging from -2.5\% to +2.5\%.
-Since asking prices are usually multiples of 1000, after the first price modification, the prices was aligned to multiples of 1000.
-
-With respect to the dwelling localization, a spatial masking process was implemented with the intention of keeping spatial properties of the original data set.
-The coordinates of each listing were displaced using a stochastic procedure.
-Effectively, the listings were recorded using coordinates contained in a maximum and minimum displacement circles, as shown in Figure \ref{fig:points-moved-image}.
-To preserve membership in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates are in the original neighborhood of the listing.
-
-```{=tex}
-\begin{figure}[!ht]
-  \caption{Masking coordinates. Spatial range}
-  \centering
-  \includegraphics[width=6cm, height=4.9cm]{EPB_files/points-moved-image.png}
-  \label{fig:points-moved-image}
-\end{figure}
-```
-
-The algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall in a different neighborhood.
-This ensures that neighborhood attributes are preserved.
+Since asking prices are usually multiples of 1,000, after the first price modification, the prices was aligned to multiples of 1,000.
 
@@
 
-Figure \ref{fig:coordinates-displacement} shows the histogram of displacements in meters for all listings in the city of Valencia; the average distance between the original and masked coordinates is 45 meters.
+With respect to the dwelling localization, a spatial masking process was implemented with the intention of keeping spatial properties of the original data set.
+The coordinates of each listing were displaced using a stochastic procedure.
+Effectively, the listings were recorded using coordinates contained in a maximum and minimum displacement circles, as shown in Figure \ref{fig:Anonymizing} (left).
+To preserve membership in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates are in the original neighborhood of the listing.
 
-```{=tex}
-\begin{figure}[!ht]
-  \caption{Masking coordinates. Spatial range}
-  \centering
-  \includegraphics[width=6cm, height=4.9cm]{EPB_files/coordinates-valencia.png}
-  \label{fig:points-moved-image}
-\end{figure}
+The Algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall in a different neighborhood.
+This ensures that neighborhood attributes are preserved.
+
+Figure \ref{fig:Anonymizing} (rigth) shows the histogram of displacements in meters for all listings in the city of Valencia; the average distance between the original and masked coordinates is 45 meters.
+
+```{r, echo = FALSE, out.width=c("29%","37%"), out.height="20%",fig.cap="\\label{fig:Anonymizing}(Left) Masking coordinates.
+Spatial range.
+(Rigth) Coordinate displacement in meters Valencia",fig.show='hold',fig.align='center'}
+knitr::include_graphics(c("EPB_files/points-moved-image.png","EPB_files/coordinates-valencia.png"))
 ```
 
-## The geo-referenced micro-data set
+## Conclusion
 
-Spatial objects include geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.
+The data set present in this paper is excellent for applying hedonic models with spatial effects, identifying housing submarkets, applying machine learning techniques, etc..
+
+# Declaration of Competing Interest}
+
+Author One and author Two are employed by idealista.
+They have been granted permission to share the data presented in this article.
+None of the authors have known competing financial interests or personal relationships which have, or could be perceived to have, influenced the work reported in this article.
 

### 2022-02-06 EPB/EPB.tex
Segunda version de EP-B

@@
 each listed dwelling, and several variables of indoor characteristics.
-The listings were enriched with official information (building year of
-construction and built quality materials grade) plus other relevant
+The listings were enriched with official information from the Spanish
+cadastre (e.g.~built quality materials grade) plus other relevant
 geographical features such as distance to urban points of interest.
@@
 
-\keywords{Housing market; idealista.com; geo-referenced data; open data;
-hedonic price analysis; Spain}
+\keywords{Housing market; idealista.com; geo-referenced data; point
+level data; open data; hedonic price analysis; Spain}
 
@@
 
-El interés pos el desarrollo de modelos hedónicos de preicción del
-precio de lvivienda que incluyen la compenente espaciál, y en general de
-la importancia de la geografía para realizar un correcto analisis del
-mercado inmobiliario, ha siod un tópico de creciente interes
-\citep{lopez2015, crespo2013local}.
+The interest about the determinants of housing market and housing prices
+has been a growing research area in the last decades, with a vast
+literature both theoretical as well as empirical. Include the spatial
+component to analyze the real state market, and in general the
+incorporate geographic variables, has been significant improvements in
+the understanding of this market. But to really understand the
+determinants of housing market it is essential to have information/data
+at level-point. Therefore, it is increasingly common that the spatial
+analysis to be development in urban environment with geo-referenced
+micro-data sets \citep[e.g.][]{lopez2015, crespo2013local}. But on the
+other hand, the availability of this type of open data at point level is
+limited and not to much data set contain latitude/longitude coordinates
+of each dwelling. In some cases the researchers it has had to resort to
+web scraping processes in order to obtain large volumes of information
+that allow get robust analysis
+\citep{gupta2022take, arbia2020spatial, Li2019, lopez2015}. These web
+scraping processes can include a lot of missing data, download errors,
+duplicate records, etc. Furthermore, in general the authors of this
+researches do not share the data sets.
 
-Por otra parte la posibilidad de disponer de datos a nivel micro/urbano
-de la vivienda ha sido dificil y en algunos casso se ha tenido que
-recurrir a dudosos procesos de webscraping para poder obtener grandes
-volúmnees de informacion que permitan. En muchos casos estos procesos de
-webscrappin pueden contener datos faltantes, errores, datos duplicados
-etc,
+In the same vein, in the last years there are a growing interest in open
+data in geography and data science
+\citep{arribasl2021editorial, arribas2021} and in general by the
+reproducible or replicable research \citep{paez2021open}. But to make
+open science is necessary have free software and open data. While great
+efforts have been made to make free software available to researchers
+(e.g.~R or Python), not much data is out in the open. In the particular
+case of the real estate market, to our knowledge, no to much open
+micro-data set of housing markets are available.
 
-El creciente interés por datos en abierto \citep{arribas2021}
-
-The main objective of this paper is present a sort description of the an
-open data set of a geo-referenced dwelling listings.
-
-This data set is excelente para aplicar modelos hedónicos con efectos
-espaciales, identificar submercados de vivienda, aplicar técnicas de
-machine learning.
-
-La importancia de disponer de datos en abierto
-
-Pocos datos en abierto de precios de vivienda
-
-Pocos con informacion georeferenciada a nivel de punto
-
-The data set is distributed in the form of an \texttt{R} package, named
-\textbf{idealista18} available from
-\url{https://github.com/paezha/idealista18} . All spatial objects such
-as polygons and points are distributed as simple features objects (class
-\texttt{sf} in R). The data has been provided by Idealista, the major
-real estate listing website in Spain, and present in other southern
-european countries as Italy and Portugal.
-
-POR QUE HACEN UNA CONTRIBUCIÓN A LA COMUNIDAD URBAN ANALITICS
+To overcome these limitations, this paper present a sort description of
+the an open micro-data set of a geo-referenced dwelling listings. The
+data has been provided by Idealista
+company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal}
+and contain information of 189,923 dwelling localized in the three major
+Spanish cities. To the date, this data product is the bigger open
+geo-referenced data set of housing market in Spain. Moreover, note that
+the data set is supply directly by Idealista and therefore is clean and
+free of download errors. The listings has been enriched with official
+information from the Spanish cadastre plus other relevant geographical
+features such as distance to urban points of interest. The data set is
+distributed in the form of an \texttt{R} package, named
+\textbf{idealista18} that can be accessed at Github
+repository\footnote{The direct URL to the Github repository is https://github.com/paezha/idealista18}.
 
@@
 
-The open data set `idealista18' is composed of nine objects, three
-objects for each of the three main Spanish cities: Barcelona, Madrid and
-Valencia. For each city, dwelling listings, neighborhood polygons and a
-set of points of interest (POI) has been included in the R package. The
-next subsections describe each data set.
+The open data set `idealista18' is R package composed of nine objects,
+three objects for each of the three main Spanish cities: Barcelona,
+Madrid and Valencia. For each city, dwelling listings, neighborhood
+polygons and a set of points of interest (POI) has been included in the
+R package. The next subsections describe each object. A full description
+of the data is available in the helps of the package.
 
@@
 package as an sf object \citep{Pebesma}. The data set corresponding to
-each city is named with the name of the city followed by '\_Sale'. Each
-of these files contains the complete set of listings, corresponding to
-the four quarters of year 2018. The record counts for each city in 2018
-are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for
-Valencia. Table \ref{tab:number-ads} show the number of ads included in
-the data set for city and quarter. Note that is possible that the same
-dwelling can be found in more than one period when a property listed for
-sale in one quarter was sold in a subsequent quarter. The variable
-ASSETID, included in each sf object is the unique identifier of the
-dwelling.
+each city is named with the name of the city followed by '\_Sale'
+(e.g.~Madrid\_Sale) and include a total of 42 variables. Each sf objects
+contains the complete set of listings, corresponding to the four
+quarters of year 2018. The record counts for each city in 2018 are:
+94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for
+Valencia. Table \ref{tab:number-ads} show the number of dwelling listing
+ads included in the data set for city and quarter. Note that is possible
+that the same dwelling can be found in more than one period when a
+property listed for sale in one quarter was sold in a subsequent
+quarter. The variable ASSETID, included in the sf objects is the unique
+identifier of the dwelling.
 
@@
 Each record of the dwelling listing contains a set of indoor
-characteristics supplied by advertiser (e.g.~price, surface, rooms,
-basic features, etc) beside with the exact localization of the dwelling
-(see Section \ref{sec:anonymizing}). Table \ref{tab:variables} list some
-of the main indoor variables included in the dwelling listing with a
-short description and the mean value of each variable. This dwelling
-listing was enriched with a number of additional attributes from the
-Spanish cadastre \citep{Catastro}. Cadastral information is described in
-Table \ref{tab:variables}, including the the prefix CAD in the variable
-name. Cadastral features assignment is done by assigning the features of
-the nearest parcel to the coordinates. The year of construction of
-dwelling was revised, given that original year of construction from
-listings are entered by users in the web site, therefore subject to
-errors and incomplete (a 40\% missing rate). To remove any issue we
+characteristics supplied by advertiser in the Idealista web (e.g.~price,
+surface, rooms, basic features, etc) including the exact localization of
+the dwelling (see Section \protect\hyperlink{anonymizing}{Anonymizing
+the data set}). Table \ref{tab:variables} list some of the main indoor
+variables included in the dwelling listing with a short description and
+the mean value of each variable. This dwelling listing was enriched with
+a number of additional attributes from the Spanish cadastre
+\citep{Catastro}. Cadastral information is described in Table
+\ref{tab:variables}, including the the prefix CAD in the variable name.
+Cadastral features assignment is done by assigning the features of the
+nearest parcel to the coordinates. The year of construction of dwelling
+(CONSTRUCTIONYEAR) supply by the advertiser was revised given that the
+year of construction are entered by users in the web site and therefore
+subject to errors and incomplete data (a 40\% missing rate). To solve
+this issue we include an alternative variable (CADCONSTRUCTIONYEAR)
 assign cadastral construction year from the nearest cadastral parcel
 whenever value has an outstanding value (date is after publication date
-or year of construction is before 1500) or when the field value was
-missing. Additionally, the distance of each dwelling to three urban
-points of interest was included in the data set: distance to city
-center, distance to the closed metro station and distance to the main
-street (Diagonal street for Barcelona, Castellana street for Madrid and
-Blasco Ibañez street for Valencia). The last rows of Table
-\ref{tab:variables} show the mean values of this variables.
+or year of construction is before 1500) or when the value supply by the
+advertiser was missing. Additionally, the distance of each dwelling to
+three urban points of interest was included in the dwelling listing:
+distance to city center, distance to the closed metro station and
+distance to the main street (Diagonal street for Barcelona, Castellana
+street for Madrid and Blasco Ibañez street for Valencia). The last rows
+of Table \ref{tab:variables} show the mean values of this variables.
 
@@
   \hline
-PRICE & ksjdhfal & 395770.58 & 396110.11 & 199678.31 \\ 
+PRICE & Asking price & 395770.58 & 396110.11 & 199678.31 \\ 
   UNITPRICE & Asking price per m\verb|^|2 (euros) & 4044.86 & 3661.05 & 1714.54 \\ 
@@
   DISTANCE\_TO\_METRO & Distance to subway station & 0.27 & 0.48 & 0.64 \\ 
-  DISTANCE\_TO\_DIAGONAL & Distance to main street & 1.77 & 2.68 & 2.07 \\ 
+  DISTANCE\_TO\_(MAINSTREET) & Distance to main street & 1.77 & 2.68 & 2.07 \\ 
    \hline
 \end{tabular}
-\caption{List of quantitative variables included in the dwelling listing for the three Spanish cities.
+\caption{List, sort description and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities.
 See the help facility in the \textbf{idealista18} R package for details and formal definitions.
@@
 
-To conclude the description of the dwelling listing, Table
-\ref{tab:Dummy-variables} include the more relevant variables that
-include information about basic characteristics of the dwellings. All
-variables listed in this table are dummy variables and the percentaje of
-dwelling with the characteristis
+In addition to the variables listed in Table \ref{tab:variables} the sf
+object include a set of dummy variables with information abour basic
+characteristics of the dwelling. Table \ref{tab:Dummy-variables} show
+the more relevant variables included in the sf object.
 
@@
   ISINTOPFLOOR & =1 is in the top floor & 0.02 & 0.02 & 0.01 \\ 
-  BUILTTYPEID\_1 & =1 if is new & 0.01 & 0.03 & 0.03 \\ 
+  BUILTTYPEID\_1 & =1 if is new contruction & 0.01 & 0.03 & 0.03 \\ 
   BUILTTYPEID\_2 & =1 is second hand to be restored & 0.17 & 0.19 & 0.13 \\ 
@@
 \end{tabular}
-\caption{List of dummy variables with the percentage of dwelling with a specific characteristic.
+\caption{List of dummy variables, sort description and ratio of dwelling with the specific characteristic.
 See the help facility in the \textbf{idealista18} R package for details and formal definitions.
@@
 The second block of data included in the `idealista18' R package are the
-spatial features of the three cities divided in neighborhoods. Figure
-\ref{fig:all-polygons} shows the different neighborhoods for the three
-cities. The boundaries are based on the official boundaries but slightly
-adapted by
-idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
+spatial features of the three cities divided in neighborhoods. There are
+an sf object for each city with the name of the city and the suffix
+'\_Polygons' (e.g.~Valencia\_Polygons). The Figure
+\ref{fig:all-polygons} shows the quantile maps of the number of
+dwellings in the listing for the different neighborhoods in the three
+cities. The neighborhoods are based on the official boundaries but
+slightly adapted by
+Idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
 In practical terms we can assume they are the same, since the website
@@
 
-A total of
-
 \begin{figure}
@@
 \includegraphics{EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:all-polygons} (A) Weekly intra-province}
+\caption{\label{fig:all-polygons}Quantile maps of the number of
+dwellings in each neighborhood. Boundary for Madrid, Barcelona and
+Valencia}
 \end{figure}
 
-There are a total of 73 neigborhoods in Barcelona, 135 in Madrid and 73
+There are a total of 73 neighborhoods in Barcelona, 135 in Madrid and 73
 in Valencia. The sf object include an unique identifier (LOCATIONID) and
@@
 The last block of data included in the data package is a set of Point of
-Interest of each city in \texttt{R} list R objetc. These lists include
-three elements: (i) the coordinates of the city center (see that
-identify the central business district); (ii) a set of points that
-define the main street of each city; and (iii) the coordinates of metro
-stations.
+Interest of each city in \texttt{R} list R object. The name of the list
+include the name of the city with the suffix '\_POIS'
+(e.g.~Barcelona\_POIS). These lists include three elements: (i) the
+coordinates of the city center, identifying the central business
+district; (ii) a set of points that define the main street of each city;
+and (iii) the coordinates of metro stations.
 
-\hypertarget{sec:anonymizing}{%
-\section{Anonymizing the data set}\label{sec:anonymizing}}
+\hypertarget{anonymizing}{%
+\section{Anonymizing the data set}\label{anonymizing}}
 
@@
 to preserve their anonymity. A masking process is apply to asking prices
-and localization.
+and localization (coordinates).
 
@@
 original values ranging from -2.5\% to +2.5\%. Since asking prices are
-usually multiples of 1000, after the first price modification, the
-prices was aligned to multiples of 1000.
-
-With respect to the dwelling localization, a spatial masking process was
-implemented with the intention of keeping spatial properties of the
-original data set. The coordinates of each listing were displaced using
-a stochastic procedure. Effectively, the listings were recorded using
-coordinates contained in a maximum and minimum displacement circles, as
-shown in Figure \ref{fig:points-moved-image}. To preserve membership in
-a neighborhood, the spatial masking procedure was constrained to ensure
-that the masked coordinates are in the original neighborhood of the
-listing.
-
-\begin{figure}[!ht]
-  \caption{Masking coordinates. Spatial range}
-  \centering
-  \includegraphics[width=6cm, height=4.9cm]{EPB_files/points-moved-image.png}
-  \label{fig:points-moved-image}
-\end{figure}
-
-The algorithm \ref{algo:coordinates-displacement} iteratively displaces
-the coordinates of each listing with a minimum distance and a maximum
-distance with the restriction that the new coordinates do not fall in a
-different neighborhood. This ensures that neighborhood attributes are
-preserved.
+usually multiples of 1,000, after the first price modification, the
+prices was aligned to multiples of 1,000.
 
@@
 
-Figure \ref{fig:coordinates-displacement} shows the histogram of
+With respect to the dwelling localization, a spatial masking process was
+implemented with the intention of keeping spatial properties of the
+original data set. The coordinates of each listing were displaced using
+a stochastic procedure. Effectively, the listings were recorded using
+coordinates contained in a maximum and minimum displacement circles, as
+shown in Figure \ref{fig:Anonymizing} (left). To preserve membership in
+a neighborhood, the spatial masking procedure was constrained to ensure
+that the masked coordinates are in the original neighborhood of the
+listing.
+
+The Algorithm \ref{algo:coordinates-displacement} iteratively displaces
+the coordinates of each listing with a minimum distance and a maximum
+distance with the restriction that the new coordinates do not fall in a
+different neighborhood. This ensures that neighborhood attributes are
+preserved.
+
+Figure \ref{fig:Anonymizing} (rigth) shows the histogram of
 displacements in meters for all listings in the city of Valencia; the
@@
 
-\begin{figure}[!ht]
-  \caption{Masking coordinates. Spatial range}
-  \centering
-  \includegraphics[width=6cm, height=4.9cm]{EPB_files/coordinates-valencia.png}
-  \label{fig:points-moved-image}
+\begin{figure}
+
+{\centering \includegraphics[width=0.29\linewidth,height=0.2\textheight]{EPB_files/points-moved-image} \includegraphics[width=0.37\linewidth,height=0.2\textheight]{EPB_files/coordinates-valencia}
+
+}
+
+\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Rigth) Coordinate displacement in meters Valencia}\label{fig:unnamed-chunk-2}
 \end{figure}
 
-\hypertarget{the-geo-referenced-micro-data-set}{%
-\subsection{The geo-referenced micro-data
-set}\label{the-geo-referenced-micro-data-set}}
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
 
-Spatial objects include geodetic coordinates using the \emph{EPSG:4326}
-coordinate reference system.
+The data set present in this paper is excellent for applying hedonic
+models with spatial effects, identifying housing submarkets, applying
+machine learning techniques, etc..
+
+\hypertarget{declaration-of-competing-interest}{%
+\section{Declaration of Competing
+Interest\}}\label{declaration-of-competing-interest}}
+
+Author One and author Two are employed by idealista. They have been
+granted permission to share the data presented in this article. None of
+the authors have known competing financial interests or personal
+relationships which have, or could be perceived to have, influenced the
+work reported in this article.
 

### 2022-02-06 data-in-brief/dib-idealista18.tex
Segunda version de EP-B

@@
 %%
-%\documentclass[times,final,longtitle]{elsarticle}
+% \documentclass[times,final,longtitle]{elsarticle}
 %%
@@
 %%
+
+\newcommand{\pkg}[1]{{\normalfont\fontseries{b}\selectfont #1}} \let\proglang=\textsf \let\code=\texttt
+
 %% Stylefile to load DIB template
@@
 %%Authors
-\author[1]{Author \snm{One}}
-\author[2]{Author \snm{Two}}
-%\author[1]{David \snm{Rey Blanco}}
-%\author[2]{Pelayo \snm{Arbués González}}
+\author[1]{David \snm{Rey Blanco}}
+\author[1]{Pelayo \snm{González Arbues}}
 % \author[1]{Pelayo \snm{González Arbues}\fnref{fn1}}
 % \fntext[fn1]{This is author footnote for second author.}
-%\author[3]{Fernando \snm{López Hernández}}
-%\author[4]{Antonio \snm{Páez}\corref{cor1}}
+\author[2]{Fernando \snm{López Hernández}}
+\author[3]{Antonio \snm{Páez}\corref{cor1}}
 %% Fourth author's email
-%\ead{paezha@mcmaster.ca}
-%\cortext[cor1]{Corresponding author:
-%  Tel.: +1-905-525-9140 ext 26099}
+\ead{paezha@mcmaster.ca}
+\cortext[cor1]{Corresponding author:
+  Tel.: +1-905-525-9140 ext 26099}
 
 %%Affiliations
-%%\address[1]{idealista, Plaza de las Cortes 5, 28014 Madrid, Spain}
-%%\address[2]{Facultad de CC de la Empresa, Universidad Politécnica de Cartagena, C/ Real, 3. 30201 Cartagena, Murcia (Spain)}
-%%\address[3]{School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada}
-
+\address[1]{idealista, Plaza de las Cortes 5, 28014 Madrid, Spain}
+\address[2]{Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia (Spain)}
+\address[3]{School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada}
 %\received{1 May 2013}
@@
 This data article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain.
-The observations are for the three largest cities in the country: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations) and Valencia (n = 33,622 observations).
 These data were originally published on \url{idealista.com} real estate website.
+The observations are for the three largest Spanish cities: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations) and Valencia (n = 33,622 observations).
 The data sets include the coordinates of properties (latitude and longitude), asking prices of each listd dwelling, and several variables of indoor characteristics.
@@
 Along with real estate listings, the data product also includes neighborhood boundaries for each city.
-These data sets are companions to a paper that examines the potential of machine learning techniques for spatial market segmentation in hedonic price analysis.
-The data product is offered in the form of a fully documented `R` package.
+The data product is offered in the form of a fully documented \proglang{R} package.
 This open data product is available for scientific and educational purposes, in particular for geo-spatial studies.
-
-\textbf{article reference to be inserted}
-
-%\noindent[The Abstract should describe the data collection process, the analysis
-%performed, the data, and their reuse potential. It should not provide
-%conclusions or interpretive insights. If your article is being
-%submitted via another Elsevier journal as a co-submission, please cite
-%this research article in the abstract.
-
-%\noindent\textbf{Tip:} do not use words such as
-%`study', `results', and `conclusions' because a data article should be
-%describing your data only.  Min 100 words - Max 500 words.]
-
-%%%%
 \end{abstract}
@@
 
-{\fontsize{7.5pt}{9pt}\selectfont
+{\fontsize{7.5pt}{9pt}
+
 %%%
-\noindent\textbf{Specifications Table}
+\textbf{Specifications Table}
 
-Every section of this table is mandatory.
-Please enter information in the right-hand column and remove all the instructions
 \begin{longtable}{|p{33mm}|p{94mm}|}
@@
 Subject                & Geography, Economics\\
-\hline
 Specific subject area  & Spatial analysis, machine learning, hedonic price analysis\\
-\hline
-Type of data           & Tables: dataframes and `sf` objects in `R` format\newline\\
+Type of data           & Tables: dataframes and `sf` objects in `R` format\\
 %\clearpage
@@
                          Spanish central cadastral registry \cite{Catastro}\newline
-                         Open street map \cite{OpenStreetMap}\newline
-\\
+                         Open street map \cite{OpenStreetMap}\\
 \hline
-Data format            & `R` package, named \{idealista18\}\\
+Data format            & `R` package, named \pkg{idealista18}\\
 \hline
@@
 collection             & idealista provided the complete record set \newline
-                         cadastral information has been downloaded the open records published quarterly\newline
-                         open street map has been downloaded from its open API
+                         cadastral information has been downloaded the open records published quarterly. Open street map has been downloaded from its open API
 \\
@@
                          Country: Spain\newline
-                         Latitude and longitude samples/data: EPSG:4326\newline
-\\
+                         Latitude and longitude samples/data: EPSG:4326\\
 
@@
 {Data accessibility}   & Repository name: GitHub\newline
-                         Direct URL to data: \url{https://github.com/paezha/idealista18}\newline
-                         \\
+                         Direct URL to data: \url{https://github.com/paezha/idealista18}\\
 \hline
@@
 
-This open data product \cite{arribas2021open} is composed of several data objects corresponding to three major Spanish cities: dwelling listings, neighboorhood polygons and a set of Points of Interest (POI) for each city\footnote{The data has been provided by Idealista, the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal.}.
-The data set is distributed in the form of an `R` package, named \{idealista18\}.
-All spatial objects such as polygons and points are distributed as simple features objects (class `sf` in `R`).
+This open data product \cite{arribas2021open} is composed of several data objects corresponding to three major Spanish cities: dwelling listings, neighboorhood polygons and a set of points of interest (POI) for each city\footnote{The data has been provided by Idealista, the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal.}.
+The data set is distributed in the form of an `R` package, named \pkg{idealista18}.
+All spatial objects such as polygons and points are distributed as simple features objects (class `sf` in proglang{R}).
 Spatial objects include geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.
 
-The first block of data integrates properties published on idealista real state website \cite{idealista}; each file contains the complete set of listings for each of the three cities in the package, corresponding to the four quarters of year 2018.
+The first block of data, integrates properties published on idealista real state website \cite{idealista}.
+Each file contains the complete set of listings for each of the three cities in the package, corresponding to the four quarters of year 2018.
 Each record contains the key found in the listing ad\footnote{The term \emph{ad} or \emph{listing} is used interchangeably to refer to a property advertised on the website} plus a number of additional attribues from the Spanish cadastre \cite{Catastro}.
@@
 \begin{longtable}{p{40mm} c p{63mm}}
+
+\label{table:data-assets-variables} \\
 \caption{Description of the variables in the listing data sets}\\
-\label{table:data-assets-variables} \\
-\hline
 \hline

### 2022-02-07 EPB/EPB.Rmd
Menores cambios en version 2

@@
 Along with real estate listings, the data product also includes neighborhood boundaries for each city.
-The data product is offered in the form of a fully documented R package.
-This open data product is available for scientific and educational purposes, in particular for geo-spatial studies."
-keywords: Housing market; idealista.com; geo-referenced data; point level data; open data; hedonic price analysis; Spain
+The data product is offered in the form of a fully documented R package and is available for scientific and educational purposes, in particular for geo-spatial studies."
+keywords: Housing market; idealista.com; geo-referenced data; point-level data; open data; Spain
 classoption:
@@
 The interest about the determinants of housing market and housing prices has been a growing research area in the last decades, with a vast literature both theoretical as well as empirical.
-Include the spatial component to analyze the real state market, and in general the incorporate geographic variables, has been significant improvements in the understanding of this market.
+Include the spatial component to analyze the real state market, and in general the incorporate geographic variables, has included significant improvements in the understanding of this market.
 But to really understand the determinants of housing market it is essential to have information/data at level-point.
-Therefore, it is increasingly common that the spatial analysis to be development in urban environment with geo-referenced micro-data sets [e.g.
-@lopez2015; @crespo2013local].
+Therefore, it is increasingly common that the spatial analysis to be development in urban environment with geo-referenced micro-data sets [@lopez2015].
 But on the other hand, the availability of this type of open data at point level is limited and not to much data set contain latitude/longitude coordinates of each dwelling.
@@
 These web scraping processes can include a lot of missing data, download errors, duplicate records, etc.
-Furthermore, in general the authors of this researches do not share the data sets.
+Furthermore, in general the authors of those researches do not share the data sets.
 
@@
 While great efforts have been made to make free software available to researchers (e.g.
-R or Python), not much data is out in the open.
-In the particular case of the real estate market, to our knowledge, no to much open micro-data set of housing markets are available.
+R or Python), not much data is out in open.
+In the particular case of the real estate market, to our knowledge, no to much open micro-data set of housing markets are available, with some exceptions [@Song2021].
 
 To overcome these limitations, this paper present a sort description of the an open micro-data set of a geo-referenced dwelling listings.
-The data has been provided by Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information of 189,923 dwelling localized in the three major Spanish cities.
-To the date, this data product is the bigger open geo-referenced data set of housing market in Spain.
+The data has been provided by Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information of 189,923 dwellings localized in the three major Spanish cities.
+To the date, this data product is the bigger open geo-referenced micro-data set of housing market in Spain.
 Moreover, note that the data set is supply directly by Idealista and therefore is clean and free of download errors.
 The listings has been enriched with official information from the Spanish cadastre plus other relevant geographical features such as distance to urban points of interest.
-The data set is distributed in the form of an `R` package, named **idealista18** that can be accessed at Github repository\footnote{The direct URL to the Github repository is https://github.com/paezha/idealista18}.
+The data set is distributed in the form of an `R` package, named 'idealista18' that can be accessed at Github repository\footnote{The direct URL to the Github repository is https://github.com/paezha/idealista18}.
 
@@
 The open data set 'idealista18' is R package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid and Valencia.
-For each city, dwelling listings, neighborhood polygons and a set of points of interest (POI) has been included in the R package.
+For each city, dwelling listings, neighborhood polygons and a set of points of interest has been included in the R package.
 The next subsections describe each object.
@@
 
-The dwelling listing of each city, include a set of characteristics of each dwelling that was published on idealista real state website (https://www.idealista.com/) and are included in 'idealista18' package as an sf object [@Pebesma].
-The data set corresponding to each city is named with the name of the city followed by '_Sale' (e.g.
+The dwelling listing of each city, include a set of characteristics of each dwelling that was published on idealista real state website as an ad.
+The dwelling listing has been included in 'idealista18' package as an sf object [@Pebesma].
+The name of the sf object containing the dwelling listing include the name of the city followed by '_Sale' (e.g.
 Madrid_Sale) and include a total of 42 variables.
-Each sf objects contains the complete set of listings, corresponding to the four quarters of year 2018.
-The record counts for each city in 2018 are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for Valencia.
+Each sf objects include the complete set of listings, corresponding to the four quarters of year 2018.
 Table \ref{tab:number-ads} show the number of dwelling listing ads included in the data set for city and quarter.
+The record counts for each city are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for Valencia.
 Note that is possible that the same dwelling can be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter.
-The variable ASSETID, included in the sf objects is the unique identifier of the dwelling.
+The variable ASSETID, included in the sf objects, is the unique identifier of the dwelling.
 
@@
 check.names = FALSE)
-colnames(text_tbl) <- c("City", "First", "Second ", "Thirdr", "Fourth","Total ads")
+colnames(text_tbl) <- c("City\\Quarter", "First", "Second ", "Thirdr", "Fourth","Total ads")
 xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") %>%
@@
 price, surface, rooms, basic features, etc) including the exact localization of the dwelling (see Section [Anonymizing the data set](#anonymizing)).
-Table \ref{tab:variables} list some of the main indoor variables included in the dwelling listing with a short description and the mean value of each variable.
+Table \ref{tab:variables} list the main indoor variables with a short description and the mean value of each variable.
 This dwelling listing was enriched with a number of additional attributes from the Spanish cadastre [@Catastro].
@@
 Cadastral features assignment is done by assigning the features of the nearest parcel to the coordinates.
-The year of construction of dwelling (CONSTRUCTIONYEAR) supply by the advertiser was revised given that the year of construction are entered by users in the web site and therefore subject to errors and incomplete data (a 40\% missing rate).
-To solve this issue we include an alternative variable (CADCONSTRUCTIONYEAR) assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the value supply by the advertiser was missing.
-Additionally, the distance of each dwelling to three urban points of interest was included in the dwelling listing: distance to city center, distance to the closed metro station and distance to the main street (Diagonal street for Barcelona, Castellana street for Madrid and Blasco Ibañez street for Valencia).
+The year of construction of dwelling (CONSTRUCTIONYEAR) supply by the advertiser was revised given that the year of construction is entered by users in the web site and therefore subject to errors and incomplete data (a 40\% missing rate).
+To solve this issue an alternative variable (CADCONSTRUCTIONYEAR) is included assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the value supply by the advertiser was missing.
+
+Additionally, the distance of each dwelling to three urban points of interest was included in the sf object: distance to city center, distance to the closed metro station and distance to the main street (Diagonal street for Barcelona, Castellana street for Madrid and Blasco Ibañez street for Valencia).
 The last rows of Table \ref{tab:variables} show the mean values of this variables.
@@
 ```{r table-variables, echo = FALSE, message = FALSE}
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
 dummyvariables <- c(3,4,5,6,7,27,30,31,32,33,37,38,39)
@@
 
-In addition to the variables listed in Table \ref{tab:variables} the sf object include a set of dummy variables with information abour basic characteristics of the dwelling.
+In addition to the variables listed in Table \ref{tab:variables} the sf object include a set of dummy variables with information about basic characteristics of the dwelling.
 Table \ref{tab:Dummy-variables} show the more relevant variables included in the sf object.
@@
 ```{r table-dummy-variables, echo = FALSE, message = FALSE}
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
 dummyvariables <- c(8,9,10,12,15,16,17,18,19,20,21,22,23,24,25,26,34,35,36)
@@
 The second block of data included in the 'idealista18' R package are the spatial features of the three cities divided in neighborhoods.
-There are an sf object for each city with the name of the city and the suffix '_Polygons' (e.g.
-Valencia_Polygons).
+There are an sf object for each city with the name of the city and the suffix '_Polygons'.
 The Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities.
@@
 
-```{r, echo = FALSE,message=FALSE,fig.width=5, fig.cap="\\label{fig:all-polygons}Quantile maps of the number of dwellings in each neighborhood. Boundary for Madrid, Barcelona and Valencia"}
+```{r, echo = FALSE,message=FALSE,fig.width=5, fig.cap="\\label{fig:all-polygons}Quantile maps of the number of dwellings in each neighborhood.
+Boundary for Barcelona (Left), Madrid (Center) and Valencia (Rigth)."}
 
@@
   geom_sf(aes(fill = Quantile),color = "black",size=.2) +
+  ylim(39.405,39.52) +
   theme_bw(base_size = 6) +
@@
 
-The last block of data included in the data package is a set of Point of Interest of each city in `R` list R object.
-The name of the list include the name of the city with the suffix '_POIS' (e.g.
-Barcelona_POIS).
-These lists include three elements: (i) the coordinates of the city center, identifying the central business district; (ii) a set of points that define the main street of each city; and (iii) the coordinates of metro stations.
+The last block of data included in the data package is a set of Point of Interest of each city as an object of the class list.
+The name of the list include the name of the city with the suffix '_POIS'.
+These lists include three elements: (i) the coordinates of the city center, identifying the central business district; (ii) a set of coordinates that define the main street of each city; and (iii) the coordinates of metro stations.
 
@@
 
-Figure \ref{fig:Anonymizing} (rigth) shows the histogram of displacements in meters for all listings in the city of Valencia; the average distance between the original and masked coordinates is 45 meters.
+Figure \ref{fig:Anonymizing} (right) shows the histogram of displacements in meters for all listings in the city of Valencia; the average distance between the original and masked coordinates is 45 meters.
 
@@
 
-## Conclusion
+# Conclusion
 
-The data set present in this paper is excellent for applying hedonic models with spatial effects, identifying housing submarkets, applying machine learning techniques, etc..
+This paper describe a data product of geo-referenced micro-data set of the three major Spanish cities.
+This is an excellent data product for understand the complex mechanisms related with the housing market and housing price.
+The researches can applying hedonic models with spatial effects, identifying housing submarkets or applying machine learning techniques.
+Also, the data product can be used for education proposes and teaching activities.
 
-# Declaration of Competing Interest}
+# Declaration of Competing Interest
 
-Author One and author Two are employed by idealista.
+Author One and author Two are employed by Idealista.
 They have been granted permission to share the data presented in this article.
@@
 
-The authors wish to thank Alessandro Galesi for their support in the paper revision and Juan Ramon Selva for collecting and cleaning the spatial data.
-This work has been partially funded by the Spanish Ministry of Economy and Competitiveness Grants PID2019-107800GB- 100
+The authors wish to thank Alessandro Galesi for their support in the paper revision and Juan Ramón Selva for collecting and cleaning the spatial data.
+This work has been partially funded by the Spanish Ministry of Economy and Competitiveness Grants PID2019-107800GB-100

### 2022-02-07 EPB/EPB.tex
Menores cambios en version 2

@@
 neighborhood boundaries for each city. The data product is offered in
-the form of a fully documented R package. This open data product is
-available for scientific and educational purposes, in particular for
-geo-spatial studies.
+the form of a fully documented R package and is available for scientific
+and educational purposes, in particular for geo-spatial studies.
 \end{abstract}
 
-\keywords{Housing market; idealista.com; geo-referenced data; point
-level data; open data; hedonic price analysis; Spain}
+\keywords{Housing market; idealista.com; geo-referenced data;
+point-level data; open data; Spain}
 
@@
 component to analyze the real state market, and in general the
-incorporate geographic variables, has been significant improvements in
-the understanding of this market. But to really understand the
+incorporate geographic variables, has included significant improvements
+in the understanding of this market. But to really understand the
 determinants of housing market it is essential to have information/data
@@
 analysis to be development in urban environment with geo-referenced
-micro-data sets \citep[e.g.][]{lopez2015, crespo2013local}. But on the
-other hand, the availability of this type of open data at point level is
-limited and not to much data set contain latitude/longitude coordinates
-of each dwelling. In some cases the researchers it has had to resort to
-web scraping processes in order to obtain large volumes of information
-that allow get robust analysis
+micro-data sets \citep{lopez2015}. But on the other hand, the
+availability of this type of open data at point level is limited and not
+to much data set contain latitude/longitude coordinates of each
+dwelling. In some cases the researchers it has had to resort to web
+scraping processes in order to obtain large volumes of information that
+allow get robust analysis
 \citep{gupta2022take, arbia2020spatial, Li2019, lopez2015}. These web
 scraping processes can include a lot of missing data, download errors,
-duplicate records, etc. Furthermore, in general the authors of this
+duplicate records, etc. Furthermore, in general the authors of those
 researches do not share the data sets.
@@
 efforts have been made to make free software available to researchers
-(e.g.~R or Python), not much data is out in the open. In the particular
-case of the real estate market, to our knowledge, no to much open
-micro-data set of housing markets are available.
+(e.g.~R or Python), not much data is out in open. In the particular case
+of the real estate market, to our knowledge, no to much open micro-data
+set of housing markets are available, with some exceptions
+\citep{Song2021}.
 
@@
 company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal}
-and contain information of 189,923 dwelling localized in the three major
-Spanish cities. To the date, this data product is the bigger open
-geo-referenced data set of housing market in Spain. Moreover, note that
-the data set is supply directly by Idealista and therefore is clean and
-free of download errors. The listings has been enriched with official
-information from the Spanish cadastre plus other relevant geographical
-features such as distance to urban points of interest. The data set is
-distributed in the form of an \texttt{R} package, named
-\textbf{idealista18} that can be accessed at Github
+and contain information of 189,923 dwellings localized in the three
+major Spanish cities. To the date, this data product is the bigger open
+geo-referenced micro-data set of housing market in Spain. Moreover, note
+that the data set is supply directly by Idealista and therefore is clean
+and free of download errors. The listings has been enriched with
+official information from the Spanish cadastre plus other relevant
+geographical features such as distance to urban points of interest. The
+data set is distributed in the form of an \texttt{R} package, named
+`idealista18' that can be accessed at Github
 repository\footnote{The direct URL to the Github repository is https://github.com/paezha/idealista18}.
@@
 Madrid and Valencia. For each city, dwelling listings, neighborhood
-polygons and a set of points of interest (POI) has been included in the
-R package. The next subsections describe each object. A full description
+polygons and a set of points of interest has been included in the R
+package. The next subsections describe each object. A full description
 of the data is available in the helps of the package.
@@
 The dwelling listing of each city, include a set of characteristics of
-each dwelling that was published on idealista real state website
-(\url{https://www.idealista.com/}) and are included in `idealista18'
-package as an sf object \citep{Pebesma}. The data set corresponding to
-each city is named with the name of the city followed by '\_Sale'
+each dwelling that was published on idealista real state website as an
+ad. The dwelling listing has been included in `idealista18' package as
+an sf object \citep{Pebesma}. The name of the sf object containing the
+dwelling listing include the name of the city followed by '\_Sale'
 (e.g.~Madrid\_Sale) and include a total of 42 variables. Each sf objects
-contains the complete set of listings, corresponding to the four
-quarters of year 2018. The record counts for each city in 2018 are:
-94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for
-Valencia. Table \ref{tab:number-ads} show the number of dwelling listing
-ads included in the data set for city and quarter. Note that is possible
-that the same dwelling can be found in more than one period when a
-property listed for sale in one quarter was sold in a subsequent
-quarter. The variable ASSETID, included in the sf objects is the unique
-identifier of the dwelling.
+include the complete set of listings, corresponding to the four quarters
+of year 2018. Table \ref{tab:number-ads} show the number of dwelling
+listing ads included in the data set for city and quarter. The record
+counts for each city are: 94,815 listings for Madrid, 61,486 for
+Barcelona and 33,622 for Valencia. Note that is possible that the same
+dwelling can be found in more than one period when a property listed for
+sale in one quarter was sold in a subsequent quarter. The variable
+ASSETID, included in the sf objects, is the unique identifier of the
+dwelling.
 
@@
   \hline
-City & First & Second  & Thirdr & Fourth & Total ads \\ 
+City$\backslash$Quarter & First & Second  & Thirdr & Fourth & Total ads \\ 
   \hline
@@
 the dwelling (see Section \protect\hyperlink{anonymizing}{Anonymizing
-the data set}). Table \ref{tab:variables} list some of the main indoor
-variables included in the dwelling listing with a short description and
-the mean value of each variable. This dwelling listing was enriched with
-a number of additional attributes from the Spanish cadastre
-\citep{Catastro}. Cadastral information is described in Table
-\ref{tab:variables}, including the the prefix CAD in the variable name.
-Cadastral features assignment is done by assigning the features of the
-nearest parcel to the coordinates. The year of construction of dwelling
-(CONSTRUCTIONYEAR) supply by the advertiser was revised given that the
-year of construction are entered by users in the web site and therefore
-subject to errors and incomplete data (a 40\% missing rate). To solve
-this issue we include an alternative variable (CADCONSTRUCTIONYEAR)
-assign cadastral construction year from the nearest cadastral parcel
-whenever value has an outstanding value (date is after publication date
-or year of construction is before 1500) or when the value supply by the
-advertiser was missing. Additionally, the distance of each dwelling to
-three urban points of interest was included in the dwelling listing:
-distance to city center, distance to the closed metro station and
-distance to the main street (Diagonal street for Barcelona, Castellana
-street for Madrid and Blasco Ibañez street for Valencia). The last rows
-of Table \ref{tab:variables} show the mean values of this variables.
+the data set}). Table \ref{tab:variables} list the main indoor variables
+with a short description and the mean value of each variable. This
+dwelling listing was enriched with a number of additional attributes
+from the Spanish cadastre \citep{Catastro}. Cadastral information is
+described in Table \ref{tab:variables}, including the the prefix CAD in
+the variable name. Cadastral features assignment is done by assigning
+the features of the nearest parcel to the coordinates. The year of
+construction of dwelling (CONSTRUCTIONYEAR) supply by the advertiser was
+revised given that the year of construction is entered by users in the
+web site and therefore subject to errors and incomplete data (a 40\%
+missing rate). To solve this issue an alternative variable
+(CADCONSTRUCTIONYEAR) is included assign cadastral construction year
+from the nearest cadastral parcel whenever value has an outstanding
+value (date is after publication date or year of construction is before
+1500) or when the value supply by the advertiser was missing.
+
+Additionally, the distance of each dwelling to three urban points of
+interest was included in the sf object: distance to city center,
+distance to the closed metro station and distance to the main street
+(Diagonal street for Barcelona, Castellana street for Madrid and Blasco
+Ibañez street for Valencia). The last rows of Table \ref{tab:variables}
+show the mean values of this variables.
 
@@
 In addition to the variables listed in Table \ref{tab:variables} the sf
-object include a set of dummy variables with information abour basic
+object include a set of dummy variables with information about basic
 characteristics of the dwelling. Table \ref{tab:Dummy-variables} show
@@
 an sf object for each city with the name of the city and the suffix
-'\_Polygons' (e.g.~Valencia\_Polygons). The Figure
-\ref{fig:all-polygons} shows the quantile maps of the number of
-dwellings in the listing for the different neighborhoods in the three
-cities. The neighborhoods are based on the official boundaries but
-slightly adapted by
+'\_Polygons'. The Figure \ref{fig:all-polygons} shows the quantile maps
+of the number of dwellings in the listing for the different
+neighborhoods in the three cities. The neighborhoods are based on the
+official boundaries but slightly adapted by
 Idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
@@
 \caption{\label{fig:all-polygons}Quantile maps of the number of
-dwellings in each neighborhood. Boundary for Madrid, Barcelona and
-Valencia}
+dwellings in each neighborhood. Boundary for Barcelona (Left), Madrid
+(Center) and Valencia (Rigth).}
 \end{figure}
 
-There are a total of 73 neighborhoods in Barcelona, 135 in Madrid and 73
+There are a total of 69 neighborhoods in Barcelona, 135 in Madrid and 73
 in Valencia. The sf object include an unique identifier (LOCATIONID) and
@@
 The last block of data included in the data package is a set of Point of
-Interest of each city in \texttt{R} list R object. The name of the list
-include the name of the city with the suffix '\_POIS'
-(e.g.~Barcelona\_POIS). These lists include three elements: (i) the
-coordinates of the city center, identifying the central business
-district; (ii) a set of points that define the main street of each city;
-and (iii) the coordinates of metro stations.
+Interest of each city as an object of the class list. The name of the
+list include the name of the city with the suffix '\_POIS'. These lists
+include three elements: (i) the coordinates of the city center,
+identifying the central business district; (ii) a set of coordinates
+that define the main street of each city; and (iii) the coordinates of
+metro stations.
 
@@
 
-Figure \ref{fig:Anonymizing} (rigth) shows the histogram of
+Figure \ref{fig:Anonymizing} (right) shows the histogram of
 displacements in meters for all listings in the city of Valencia; the
@@
 \hypertarget{conclusion}{%
-\subsection{Conclusion}\label{conclusion}}
+\section{Conclusion}\label{conclusion}}
 
-The data set present in this paper is excellent for applying hedonic
-models with spatial effects, identifying housing submarkets, applying
-machine learning techniques, etc..
+This paper describe a data product of geo-referenced micro-data set of
+the three major Spanish cities. This is an excellent data product for
+understand the complex mechanisms related with the housing market and
+housing price. The researches can applying hedonic models with spatial
+effects, identifying housing submarkets or applying machine learning
+techniques. Also, the data product can be used for education proposes
+and teaching activities.
 
@@
 \section{Declaration of Competing
-Interest\}}\label{declaration-of-competing-interest}}
+Interest}\label{declaration-of-competing-interest}}
 
-Author One and author Two are employed by idealista. They have been
+Author One and author Two are employed by Idealista. They have been
 granted permission to share the data presented in this article. None of
@@
 The authors wish to thank Alessandro Galesi for their support in the
-paper revision and Juan Ramon Selva for collecting and cleaning the
+paper revision and Juan Ramón Selva for collecting and cleaning the
 spatial data. This work has been partially funded by the Spanish
-Ministry of Economy and Competitiveness Grants PID2019-107800GB- 100
+Ministry of Economy and Competitiveness Grants PID2019-107800GB-100
 

### 2022-02-18 EPB/EPB.Rmd
Version EPB con inglés corregido

--- 2022-02-05 EPB/EPB.Rmd
@@
 ---
-title: A geo-referenced micro-data set of real estate listings for the three largest Spanish cities
+title: A geo-referenced micro-data set of real estate listings for Spain’s three largest cities
 runninghead: Rey-Blanco \emph{et al}.
@@
 - name: A. Páez*
-  num: 4
+  num: 3
 address:
@@
 email: paezha@mcmaster.ca
-abstract: "This data article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain.
-These data were originally published on idealista.com real estate website.
-The observations are obtained for the three largest Spanish cities: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations) and Valencia (n = 33,622 observations).
+abstract: "This article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain.
+These data were originally published on the idealista.com real estate website.
+The observations were obtained for the three largest cities in Spain: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations), and Valencia (n = 33,622 observations).
 The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics.
-The listings were enriched with official information (building year of construction and built quality materials grade) plus other relevant geographical features such as distance to urban points of interest.
-Along with real estate listings, the data product also includes neighborhood boundaries for each city.
-The data product is offered in the form of a fully documented R package.
-This open data product is available for scientific and educational purposes, in particular for geo-spatial studies."
-keywords: Housing market; idealista.com; geo-referenced data; open data; hedonic price analysis; Spain
+The listings were enriched with official information from the Spanish cadastre (e.g., building material quality) plus other relevant geographical features, such as distance to urban points of interest.
+Along with the real estate listings, the data product also includes neighborhood boundaries for each city.
+The data product is offered as a fully documented R package and is available for scientific and educational purposes, particularly for geo-spatial studies"
+keywords: Housing market; idealista.com; geo-referenced data; point-level data; open data; Spain
 classoption:
@@
 
-El interés pos el desarrollo de modelos hedónicos de preicción del precio de lvivienda que incluyen la compenente espaciál,  y en general de la importancia de la geografía para realizar un correcto analisis del mercado inmobiliario, ha siod un tópico de creciente interes [@lopez2015; @crespo2013local].
+Interest in the characteristics of the housing market and housing prices has been a growing area of research in recent decades, generating a vast amount of theoretical and empirical literature.
+Including the spatial component to analyze the real estate market and incorporating geographic variables has significantly improved the understanding of this market.
+But to really understand the characteristics of the housing market, it is essential to have information/data at the point level.
+Therefore, it is becoming common for spatial analysis of urban environments to be developed with geo- referenced micro-data sets [@lopez2015].
+However, the availability of this type of open data at the point level is limited, and not many data sets contain latitude/longitude coordinates for each dwelling.
+In some cases, researchers have had to resort to web scraping processes to obtain the large volumes of information that permit robust analyses [@gupta2022take; @arbia2020spatial; @Li2019; @lopez2015].
+These  web  scraping  processes  can  include  missing  data, download errors, duplicate records, etc.
+Furthermore, the authors of this research do not generally share the data sets.
 
-Por otra parte la posibilidad de disponer de datos a nivel micro/urbano de la vivienda ha sido dificil y en algunos casso se ha tenido que recurrir a dudosos procesos de webscraping para poder obtener grandes volúmnees de informacion que permitan.
-En muchos casos estos procesos de webscrappin pueden contener datos faltantes, errores, datos duplicados etc,
+We are also witnessing a growing interest in open data in geography and data science [@arribasl2021editorial; @arribas2021]  using reproducible or replicable  research  [@paez2021open].
+But  to  work  openly in  science, it  is  necessary to  have  free software and open data.
+While great efforts have been made to make free software available to researchers (e.g., R or Python), not much data is currently out in the open.
+In the particular case of the real estate market, to our knowledge, there are few open micro-data sets of housing markets available [@Song2021].
 
-El creciente interés por datos en abierto [@arribas2021]
+To overcome these limitations, this paper presents a sort description of an open micro-data set of geo-referenced dwelling listings.
+The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain’s three largest cities.
+To date, this data product is the biggest open geo-referenced micro-data set of the housing market in Spain.
+Moreover, the data set has been supplied directly by Idealista, and therefore is clean and free of download errors.
+The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest.
+The data set is distributed as an R package, named ‘idealista18’, which can be accessed from the Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
 
-The main objective of this paper is present a sort description of the an open data set of a geo-referenced dwelling listings.
-
-This data set is excelente para aplicar modelos hedónicos con efectos espaciales, identificar submercados de vivienda, aplicar técnicas de machine learning.
-
-La importancia de disponer de datos en abierto
-
-Pocos datos en abierto de precios de vivienda
-
-Pocos con informacion georeferenciada a nivel de punto
-
-The data set is distributed in the form of an `R` package, named **idealista18** available from https://github.com/paezha/idealista18 .
-All spatial objects such as polygons and points are distributed as simple features objects (class `sf` in R).
-The data has been provided by Idealista, the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal.
-
-POR QUE HACEN UNA CONTRIBUCIÓN A LA COMUNIDAD URBAN ANALITICS
+<!-- @cohen2008spatial (508 observations) -->
+<!-- @lazrak2014market (19000 observations) -->
 
@@
 
-The open data set 'idealista18' is composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid and Valencia.
-For each city, dwelling listings, neighborhood polygons and a set of points of interest (POI) has been included in the R package.
-The next subsections describe each data set.
+The open data set ‘idealista18’ is an R package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia.
+For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the R package.
+The following subsections describe each object.
+A full description of the data is available in the help section of the package.
 
@@
 
-The dwelling listing of each city, include a set of characteristics of each dwelling that was published on idealista real state website (https://www.idealista.com/) and are included in 'idealista18' package as an sf object [@Pebesma].
-The data set corresponding to each city is named with the name of the city followed by '_Sale'.
-Each of these files contains the complete set of listings, corresponding to the four quarters of year 2018.
-The record counts for each city in 2018 are: 94,815 listings for Madrid, 61,486 for Barcelona and 33,622 for Valencia.
-Table \ref{tab:number-ads} show the number of ads included in the data set for city and quarter.
-Note that is possible that the same dwelling can be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter.
-The variable ASSETID, included in each sf object is the unique identifier of the dwelling.
+The dwelling listing of each city includes a set of characteristics for each dwelling published on the idealista real estate website as an ad.
+The dwelling listing has been included in the ‘idealista18’ package as an sf object [@Pebesma].
+The name of the sf object containing the dwelling listing includes the name of the city, followed by '_Sale' (e.g., Madrid_Sale) and includes a total of 42 variables.
+Each sf object includes the complete set of listings corresponding to the four quarters of the year 2018.
+Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter.
+The record counts for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622 for Valencia.
+Note that the same dwelling may be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter.
+The variable ASSETID, included in the sf objects, is the unique identifier of the dwelling.
 
@@
 
-Each record of the dwelling listing contains a set of indoor characteristics supplied by advertiser (e.g.
-price, surface, rooms, basic features, etc) beside with the exact localization of the dwelling (see Section \ref{sec:anonymizing}).
-Table \ref{tab:variables} list some of the main indoor variables included in the dwelling listing with a short description and the mean value of each variable.
-This dwelling listing was enriched with a number of additional attributes from the Spanish cadastre [@Catastro].
-Cadastral information is described in Table \ref{tab:variables}, including the the prefix CAD in the variable name.
-Cadastral features assignment is done by assigning the features of the nearest parcel to the coordinates.
-The year of construction of dwelling was revised, given that original year of construction from listings are entered by users in the web site, therefore subject to errors and incomplete (a 40\% missing rate).
-To remove any issue we assign cadastral construction year from the nearest cadastral parcel whenever value has an outstanding value (date is after publication date or year of construction is before 1500) or when the field value was missing.
-Additionally, the distance of each dwelling to three urban points of interest was included in the data set: distance to city center, distance to the closed metro station and distance to the main street (Diagonal street for Barcelona, Castellana street for Madrid and Blasco Ibañez street for Valencia).
-The last rows of Table \ref{tab:variables} show the mean values of this variables.
+Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including the exact location of the dwelling (see Section [Anonymizing the data set](#anonymizing)).
+Table \ref{tab:variables} lists the main indoor variables with a short description and the mean value of each variable.
+The dwelling listings were enriched with a number of additional attributes from the Spanish cadastre [@Catastro].
+The cadastral information is described in Table \ref{tab:variables}, with the prefix CAD in the variable name.
+The cadastral features were assigned by applying the features of the nearest parcel to the coordinates.
+The year the dwelling was built (CONSTRUCTIONYEAR) given by the advertiser was revised since the year of construction is entered on the website by users, and it is therefore subject to errors and incomplete data (40\% missing data).
+To resolve this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included, assigning the cadastral construction year from the nearest cadastral parcel whenever the value was outstanding (date was after publication date or year of construction was before 1500) or when the value supplied by the advertiser was missing.
+
+Additionally, the distance of each dwelling to three urban points of interest was included in the sf object: distance to the city center, distance to the closest metro station, and distance to a major street (La Diagonal for Barcelona, La Castellana for Madrid, and Blasco Ibañez for Valencia).
+The last rows of Table \ref{tab:variables} show the mean values of these variables.
 
 ```{r table-variables, echo = FALSE, message = FALSE}
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
 dummyvariables <- c(3,4,5,6,7,27,30,31,32,33,37,38,39)
@@
                        "Cadastral quality. 0 Best-10 Worst",
-                       "Distance to city center","Distance to subway station","Distance to main street"),
+                       "Distance to city center","Distance to subway station","Distance to major street"),
 # Barcelona
@@
 check.names = FALSE)
-
-xtable::xtable(text_tbl, caption = "List of quantitative variables included in the dwelling listing for the three Spanish cities.
-See the help facility in the \\textbf{idealista18} R package for details and formal definitions.
-Some variables has been excluded of this table for save space, check the full list in  \\textbf{idealista18} R package.
-\\label{tab:variables}") %>%
+text_tbl[dim(text_tbl)[1],1]<-"DISTANCE_TO_(MAINSTREET)"
+xtable::xtable(text_tbl, caption = "List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities.
+See the help section in the \\textbf{idealista18} R package for details and formal definitions.
+Some variables have been excluded from this table to save space.
+Check the full list in the \\textbf{idealista18} package.\\label{tab:variables}") %>%
 xtable2kable(include.rownames = FALSE) %>%
@@
 
-To conclude the description of the dwelling listing, Table \ref{tab:Dummy-variables} include the more relevant variables that include information about basic characteristics of the dwellings.
-All variables listed in this table are dummy variables and the percentaje of dwelling with the characteristis
+In addition to the variables listed in Table \ref{tab:variables}, the sf object includes a set of dummy variables with information about the basic characteristics of the dwelling.
+Table \ref{tab:Dummy-variables} shows the more relevant variables included in the sf object.
 
 ```{r table-dummy-variables, echo = FALSE, message = FALSE}
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
 dummyvariables <- c(8,9,10,12,15,16,17,18,19,20,21,22,23,24,25,26,34,35,36)
@@
                        "=1 if has boxroom","=1 if has wardrobe",
-                       "=1 if has swinningpool","=1 if has doorman",
+                       "=1 if has swimmingpool","=1 if has doorman",
                        "=1 if has garden","=1 if is duplex",
-                       "=1 if is studio","=1 is in the top floor",
-                       "=1 if is new","=1 is second hand to be restored","=1 is second hand good conditions"),
+                       "=1 if is studio","=1 is on the top floor",
+                       "=1 if is new contruction","=1 is second hand to be restored","=1 is second hand in good condition"),
 # Barcelona
@@
 
-xtable::xtable(text_tbl, caption = "List of dummy variables with the percentage of dwelling with a specific characteristic.
-See the help facility in the \\textbf{idealista18} R package for details and formal definitions.
-Some dummy variables has been excluded of this table for save space.
-\\label{tab:Dummy-variables}") %>%
+xtable::xtable(text_tbl, caption = "List of dummy variables, sort description, and ratios of dwellings with the specific characteristics.
+See the help section in the \\textbf{idealista18} R package for details and formal definitions.
+Some dummy variables have been excluded from this table to save space\\label{tab:Dummy-variables}") %>%
 xtable2kable(include.rownames = FALSE) %>%
@@
 
-The second block of data included in the 'idealista18' R package are the spatial features of the three cities divided in neighborhoods.
-Figure \ref{fig:all-polygons} shows the different neighborhoods for the three cities.
-The boundaries are based on the official boundaries but slightly adapted by idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
-In practical terms we can assume they are the same, since the website simply collapses areas when they are sufficiently small in terms of number of ads.
-In the case of Madrid they just collapse four areas into two new ones.
+The second block of data included in the ‘idealista18’ R package is the spatial features of the three cities divided into neighborhoods.
+There is an sf object for each city with the name of the city and the suffix '_Polygons'.
+Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities.
+The neighborhoods are based on the official boundaries but slightly changed by Idealista\footnote{There are two criteria used to make this division.
+If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}.
+In practical terms, we can assume they are the same since the website combines areas when there are few ads for that area.
+In the case of Madrid, they combined four areas into two.
 
-A total of
-
-```{r, echo = FALSE, fig.width=5, fig.cap="\\label{fig:all-polygons} (A) Weekly intra-province"}
-library(ggplot2)
-library(gridExtra)
-plot1 <- ggplot(data=Barcelona_Polygons) + geom_sf(color = "black",size=.2) + 
-  theme_bw(base_size = 2) + 
-  ggtitle("                Barcelona") +
-  theme(plot.title = element_text(size = 10),
-        panel.border = element_blank(),
-    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
-plot2 <- ggplot(data=Madrid_Polygons) + geom_sf(color = "black",size=.2) + 
-  theme_bw(base_size = 2) +
-   ggtitle("                Madrid") +
-    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
-    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
-plot3 <- ggplot(data=Valencia_Polygons) + geom_sf(color = "black",size=.2) + 
-  theme_bw(base_size = 2) +
-   ggtitle("                Valencia") +
-    theme(plot.title = element_text(size = 10),panel.border = element_blank(),
-    axis.text = element_blank(),panel.grid = element_blank(), axis.ticks = element_blank())
-grid.arrange(plot1, plot2, plot3, ncol = 3, nrow = 1)
+```{r, echo = FALSE,message=FALSE,fig.width=5, fig.cap="\\label{fig:all-polygons}Quantile maps of the number of dwellings in each neighborhood.
+Boundary for Barcelona (Left), Madrid (Center), and Valencia (Right)."}
+library("ggplot2")
+library("dplyr")
+library("gridExtra")
+# Barcelona
+Barcelona_counts <- Barcelona_Polygons %>% 
+  mutate(counts = lengths(st_intersects(., Barcelona_Sale)))
+q <- quantile(Barcelona_counts$counts)
+Barcelona_Polygons$Quantile<- as.factor((Barcelona_counts$counts > q[2]) + (Barcelona_counts$counts > q[3]) +(Barcelona_counts$counts >= q[4]) + 1)
+plot4 <- ggplot(data = Barcelona_Polygons) +
+  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
+  theme_bw(base_size = 6) +
+  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
+  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
+        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
+# Madrid
+Madrid_counts <- Madrid_Polygons %>% 
+  mutate(counts = lengths(st_intersects(., Madrid_Sale)))
+q <- quantile(Madrid_counts$counts)
+Madrid_Polygons$Quantile<- as.factor((Madrid_counts$counts > q[2]) + (Madrid_counts$counts > q[3]) +(Madrid_counts$counts >= q[4]) + 1)
+plot5 <- ggplot(data = Madrid_Polygons) +
+  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
+  theme_bw(base_size = 6) +
+  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
+  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
+        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
+# Valencia
+Valencia_counts <- Valencia_Polygons %>% 
+  mutate(counts = lengths(st_intersects(., Valencia_Sale)))
+q <- quantile(Valencia_counts$counts)
+Valencia_Polygons$Quantile<- as.factor((Valencia_counts$counts > q[2]) + (Valencia_counts$counts > q[3]) +(Valencia_counts$counts >= q[4]) + 1)
+plot6 <- ggplot(data = Valencia_Polygons) +
+  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
+  ylim(39.405,39.52) +
+  theme_bw(base_size = 6) +
+  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
+  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
+        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
+grid.arrange(plot4, plot5, plot6, ncol = 3, nrow = 1)
 ```
 
-There are a total of 73 neigborhoods in Barcelona, 135 in Madrid and 73 in Valencia. The sf object include an unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).
+There are a total of 69 neighborhoods in Barcelona, 135 in Madrid, and 73 in Valencia. The sf object includes a unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).
 
@@
 
-The last block of data included in the data package is a set of Point of Interest of each city in `R` list R objetc.
-These lists include three elements: (i) the coordinates of the city center (see that identify the central business district); (ii) a set of points that define the main street of each city; and (iii) the coordinates of metro stations.
+The last block of data included in the data package is a set of Points of Interest in each city as an object of the class list.
+The name of the list includes the name of the city with the suffix '_POIS'.
+These lists include three elements: (i) the coordinates of the city center, the central business district; (ii) a set of coordinates that define the main street of each city; and (iii) the coordinates of the metro stations.
 
-# Anonymizing the data set {#sec:anonymizing}
+# Anonymizing the data set {#anonymizing}
 
-To comply with Spanish regulations, two variables are slightly modified to preserve their anonymity. A masking process is apply to asking prices and localization.
+To comply with Spanish regulations, two variables were slightly modified to provide anonymity. A masking process was applied to asking prices and location (coordinates).
 
-Whit respect to the asking prices, the original values are obfuscated with the addition or subtraction of a random percentage of their original values ranging from -2.5\% to +2.5\%.
-Since asking prices are usually multiples of 1000, after the first price modification, the prices was aligned to multiples of 1000.
-
-With respect to the dwelling localization, a spatial masking process was implemented with the intention of keeping spatial properties of the original data set.
-The coordinates of each listing were displaced using a stochastic procedure.
-Effectively, the listings were recorded using coordinates contained in a maximum and minimum displacement circles, as shown in Figure \ref{fig:points-moved-image}.
-To preserve membership in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates are in the original neighborhood of the listing.
-
-```{=tex}
-\begin{figure}[!ht]
-  \caption{Masking coordinates. Spatial range}
-  \centering
-  \includegraphics[width=6cm, height=4.9cm]{EPB_files/points-moved-image.png}
-  \label{fig:points-moved-image}
-\end{figure}
-```
-
-The algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall in a different neighborhood.
-This ensures that neighborhood attributes are preserved.
+In terms of the asking prices, the original values were obfuscated with the addition or subtraction of a random percentage of their original values, ranging from -2.5% to +2.5%.
+Since asking prices are usually multiples of 1,000, after the first price modification, the prices were aligned to multiples of 1,000.
 
@@
 
-Figure \ref{fig:coordinates-displacement} shows the histogram of displacements in meters for all listings in the city of Valencia; the average distance between the original and masked coordinates is 45 meters.
+With respect to the location of the dwelling, a spatial masking process was implemented  to maintain the spatial properties of the original data set.
+The coordinates of each listing were displaced using a stochastic procedure.
+The listings were recorded using coordinates contained in maximum and minimum displacement circles, as shown in Figure \ref{fig:Anonymizing} (left).
+To preserve inclusion in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates remained in the original neighborhood of the listing.
 
-```{=tex}
-\begin{figure}[!ht]
-  \caption{Masking coordinates. Spatial range}
-  \centering
-  \includegraphics[width=6cm, height=4.9cm]{EPB_files/coordinates-valencia.png}
-  \label{fig:points-moved-image}
-\end{figure}
+Algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall into a different neighborhood.
+This ensures that neighborhood attributes are preserved.
+
+Figure \ref{fig:Anonymizing} (right) shows the histogram of the displacements in meters for all the listings in the city of Valencia.
+The average distance between the original and masked coordinates is 45 meters.
+
+```{r, echo = FALSE, out.width=c("29%","37%"), out.height="20%",fig.cap="\\label{fig:Anonymizing}(Left) Masking coordinates.
+Spatial range.
+(Right) Coordinate displacement in meters (Valencia)",fig.show='hold',fig.align='center'}
+knitr::include_graphics(c("EPB_files/points-moved-image.png","EPB_files/coordinates-valencia.png"))
 ```
 
-## The geo-referenced micro-data set
+# Conclusion
 
-Spatial objects include geodetic coordinates using the \emph{EPSG:4326} coordinate reference system.
+This paper describes a data product of a geo-referenced micro-data set of Spain’s three largest cities.
+This is an excellent data product to help understand the complex mechanisms related to the housing market and housing prices.
+Researchers can apply hedonic models with spatial effects, identifying housing submarkets or machine learning techniques.
+The data product can also be used for educational proposes and teaching activities.
+
+# Declaration of Competing Interest
+
+The authors wish to thank Alessandro Galesi for his support in the paper revision and  Juan  Ramón  Selva  for  collecting  and  cleaning  the  spatial  data.
+This  work  has been partially funded by the Spanish Ministry of Economy and Competitiveness Grant PID2019-107800GB-100.
 

### 2022-02-18 EPB/EPB.tex
Version EPB con inglés corregido

@@
 
-\title{A geo-referenced micro-data set of real estate listings for the
-three largest Spanish cities}
+\title{A geo-referenced micro-data set of real estate listings for
+Spain's three largest cities}
 
@@
 \author{D. Rey-Blanco\affilnum{1}, P. González-Arbues\affilnum{1}, F.
-López\affilnum{2}, A. Páez*\affilnum{4}}
+López\affilnum{2}, A. Páez*\affilnum{3}}
 
@@
 \begin{abstract}
-This data article shares an open data product with big geo-referenced
+This article shares an open data product with big geo-referenced
 micro-data sets of 2018 real estate listings in Spain. These data were
-originally published on idealista.com real estate website. The
-observations are obtained for the three largest Spanish cities: Madrid
-(n = 94,815 observations), Barcelona (n = 61,486 observations) and
+originally published on the idealista.com real estate website. The
+observations were obtained for the three largest cities in Spain: Madrid
+(n = 94,815 observations), Barcelona (n = 61,486 observations), and
 Valencia (n = 33,622 observations). The data sets include the
@@
 The listings were enriched with official information from the Spanish
-cadastre (e.g.~built quality materials grade) plus other relevant
-geographical features such as distance to urban points of interest.
-Along with real estate listings, the data product also includes
-neighborhood boundaries for each city. The data product is offered in
-the form of a fully documented R package and is available for scientific
-and educational purposes, in particular for geo-spatial studies.
+cadastre (e.g., building material quality) plus other relevant
+geographical features, such as distance to urban points of interest.
+Along with the real estate listings, the data product also includes
+neighborhood boundaries for each city. The data product is offered as a
+fully documented R package and is available for scientific and
+educational purposes, particularly for geo-spatial studies
 \end{abstract}
@@
 
-The interest about the determinants of housing market and housing prices
-has been a growing research area in the last decades, with a vast
-literature both theoretical as well as empirical. Include the spatial
-component to analyze the real state market, and in general the
-incorporate geographic variables, has included significant improvements
-in the understanding of this market. But to really understand the
-determinants of housing market it is essential to have information/data
-at level-point. Therefore, it is increasingly common that the spatial
-analysis to be development in urban environment with geo-referenced
-micro-data sets \citep{lopez2015}. But on the other hand, the
-availability of this type of open data at point level is limited and not
-to much data set contain latitude/longitude coordinates of each
-dwelling. In some cases the researchers it has had to resort to web
-scraping processes in order to obtain large volumes of information that
-allow get robust analysis
+Interest in the characteristics of the housing market and housing prices
+has been a growing area of research in recent decades, generating a vast
+amount of theoretical and empirical literature. Including the spatial
+component to analyze the real estate market and incorporating geographic
+variables has significantly improved the understanding of this market.
+But to really understand the characteristics of the housing market, it
+is essential to have information/data at the point level. Therefore, it
+is becoming common for spatial analysis of urban environments to be
+developed with geo- referenced micro-data sets \citep{lopez2015}.
+However, the availability of this type of open data at the point level
+is limited, and not many data sets contain latitude/longitude
+coordinates for each dwelling. In some cases, researchers have had to
+resort to web scraping processes to obtain the large volumes of
+information that permit robust analyses
 \citep{gupta2022take, arbia2020spatial, Li2019, lopez2015}. These web
-scraping processes can include a lot of missing data, download errors,
-duplicate records, etc. Furthermore, in general the authors of those
-researches do not share the data sets.
+scraping processes can include missing data, download errors, duplicate
+records, etc. Furthermore, the authors of this research do not generally
+share the data sets.
 
-In the same vein, in the last years there are a growing interest in open
-data in geography and data science
-\citep{arribasl2021editorial, arribas2021} and in general by the
-reproducible or replicable research \citep{paez2021open}. But to make
-open science is necessary have free software and open data. While great
-efforts have been made to make free software available to researchers
-(e.g.~R or Python), not much data is out in open. In the particular case
-of the real estate market, to our knowledge, no to much open micro-data
-set of housing markets are available, with some exceptions
-\citep{Song2021}.
+We are also witnessing a growing interest in open data in geography and
+data science \citep{arribasl2021editorial, arribas2021} using
+reproducible or replicable research \citep{paez2021open}. But to work
+openly in science, it is necessary to have free software and open data.
+While great efforts have been made to make free software available to
+researchers (e.g., R or Python), not much data is currently out in the
+open. In the particular case of the real estate market, to our
+knowledge, there are few open micro-data sets of housing markets
+available \citep{Song2021}.
 
-To overcome these limitations, this paper present a sort description of
-the an open micro-data set of a geo-referenced dwelling listings. The
-data has been provided by Idealista
+To overcome these limitations, this paper presents a sort description of
+an open micro-data set of geo-referenced dwelling listings. The data
+have been provided by the Idealista
 company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal}
-and contain information of 189,923 dwellings localized in the three
-major Spanish cities. To the date, this data product is the bigger open
-geo-referenced micro-data set of housing market in Spain. Moreover, note
-that the data set is supply directly by Idealista and therefore is clean
-and free of download errors. The listings has been enriched with
-official information from the Spanish cadastre plus other relevant
-geographical features such as distance to urban points of interest. The
-data set is distributed in the form of an \texttt{R} package, named
-`idealista18' that can be accessed at Github
-repository\footnote{The direct URL to the Github repository is https://github.com/paezha/idealista18}.
+and contain information about 189,923 dwellings located in Spain's three
+largest cities. To date, this data product is the biggest open
+geo-referenced micro-data set of the housing market in Spain. Moreover,
+the data set has been supplied directly by Idealista, and therefore is
+clean and free of download errors. The listings have been enriched with
+official information from the Spanish cadastre along with other relevant
+geographical features, such as distance to urban points of interest. The
+data set is distributed as an R package, named `idealista18', which can
+be accessed from the Github
+repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
 
@@
 
-The open data set `idealista18' is R package composed of nine objects,
-three objects for each of the three main Spanish cities: Barcelona,
-Madrid and Valencia. For each city, dwelling listings, neighborhood
-polygons and a set of points of interest has been included in the R
-package. The next subsections describe each object. A full description
-of the data is available in the helps of the package.
+The open data set `idealista18' is an R package composed of nine
+objects, three objects for each of the three main Spanish cities:
+Barcelona, Madrid, and Valencia. For each city, dwelling listings,
+neighborhood polygons, and a set of points of interest have been
+included in the R package. The following subsections describe each
+object. A full description of the data is available in the help section
+of the package.
 
@@
 
-The dwelling listing of each city, include a set of characteristics of
-each dwelling that was published on idealista real state website as an
-ad. The dwelling listing has been included in `idealista18' package as
+The dwelling listing of each city includes a set of characteristics for
+each dwelling published on the idealista real estate website as an ad.
+The dwelling listing has been included in the `idealista18' package as
 an sf object \citep{Pebesma}. The name of the sf object containing the
-dwelling listing include the name of the city followed by '\_Sale'
-(e.g.~Madrid\_Sale) and include a total of 42 variables. Each sf objects
-include the complete set of listings, corresponding to the four quarters
-of year 2018. Table \ref{tab:number-ads} show the number of dwelling
-listing ads included in the data set for city and quarter. The record
-counts for each city are: 94,815 listings for Madrid, 61,486 for
-Barcelona and 33,622 for Valencia. Note that is possible that the same
-dwelling can be found in more than one period when a property listed for
-sale in one quarter was sold in a subsequent quarter. The variable
-ASSETID, included in the sf objects, is the unique identifier of the
-dwelling.
+dwelling listing includes the name of the city, followed by '\_Sale'
+(e.g., Madrid\_Sale) and includes a total of 42 variables. Each sf
+object includes the complete set of listings corresponding to the four
+quarters of the year 2018. Table \ref{tab:number-ads} shows the number
+of dwelling listing ads included in the data set for each city and
+quarter. The record counts for each city are: 94,815 listings for
+Madrid, 61,486 for Barcelona, and 33,622 for Valencia. Note that the
+same dwelling may be found in more than one period when a property
+listed for sale in one quarter was sold in a subsequent quarter. The
+variable ASSETID, included in the sf objects, is the unique identifier
+of the dwelling.
 
@@
 Each record of the dwelling listing contains a set of indoor
-characteristics supplied by advertiser in the Idealista web (e.g.~price,
-surface, rooms, basic features, etc) including the exact localization of
-the dwelling (see Section \protect\hyperlink{anonymizing}{Anonymizing
-the data set}). Table \ref{tab:variables} list the main indoor variables
-with a short description and the mean value of each variable. This
-dwelling listing was enriched with a number of additional attributes
-from the Spanish cadastre \citep{Catastro}. Cadastral information is
-described in Table \ref{tab:variables}, including the the prefix CAD in
-the variable name. Cadastral features assignment is done by assigning
-the features of the nearest parcel to the coordinates. The year of
-construction of dwelling (CONSTRUCTIONYEAR) supply by the advertiser was
-revised given that the year of construction is entered by users in the
-web site and therefore subject to errors and incomplete data (a 40\%
-missing rate). To solve this issue an alternative variable
-(CADCONSTRUCTIONYEAR) is included assign cadastral construction year
-from the nearest cadastral parcel whenever value has an outstanding
-value (date is after publication date or year of construction is before
-1500) or when the value supply by the advertiser was missing.
+characteristics supplied by the advertisers on the Idealista website
+(e.g., price, surface area, number of bedrooms, basic features, etc.),
+including the exact location of the dwelling (see Section
+\protect\hyperlink{anonymizing}{Anonymizing the data set}). Table
+\ref{tab:variables} lists the main indoor variables with a short
+description and the mean value of each variable. The dwelling listings
+were enriched with a number of additional attributes from the Spanish
+cadastre \citep{Catastro}. The cadastral information is described in
+Table \ref{tab:variables}, with the prefix CAD in the variable name. The
+cadastral features were assigned by applying the features of the nearest
+parcel to the coordinates. The year the dwelling was built
+(CONSTRUCTIONYEAR) given by the advertiser was revised since the year of
+construction is entered on the website by users, and it is therefore
+subject to errors and incomplete data (40\% missing data). To resolve
+this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included,
+assigning the cadastral construction year from the nearest cadastral
+parcel whenever the value was outstanding (date was after publication
+date or year of construction was before 1500) or when the value supplied
+by the advertiser was missing.
 
 Additionally, the distance of each dwelling to three urban points of
-interest was included in the sf object: distance to city center,
-distance to the closed metro station and distance to the main street
-(Diagonal street for Barcelona, Castellana street for Madrid and Blasco
-Ibañez street for Valencia). The last rows of Table \ref{tab:variables}
-show the mean values of this variables.
+interest was included in the sf object: distance to the city center,
+distance to the closest metro station, and distance to a major street
+(La Diagonal for Barcelona, La Castellana for Madrid, and Blasco Ibañez
+for Valencia). The last rows of Table \ref{tab:variables} show the mean
+values of these variables.
 
@@
   DISTANCE\_TO\_METRO & Distance to subway station & 0.27 & 0.48 & 0.64 \\ 
-  DISTANCE\_TO\_(MAINSTREET) & Distance to main street & 1.77 & 2.68 & 2.07 \\ 
+  DISTANCE\_TO\_(MAINSTREET) & Distance to major street & 1.77 & 2.68 & 2.07 \\ 
    \hline
 \end{tabular}
-\caption{List, sort description and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities.
-See the help facility in the \textbf{idealista18} R package for details and formal definitions.
-Some variables has been excluded of this table for save space, check the full list in  \textbf{idealista18} R package.
-\label{tab:variables}}
+\caption{List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities.
+See the help section in the \textbf{idealista18} R package for details and formal definitions.
+Some variables have been excluded from this table to save space.
+Check the full list in the \textbf{idealista18} package.\label{tab:variables}}
 \end{table}
 
-In addition to the variables listed in Table \ref{tab:variables} the sf
-object include a set of dummy variables with information about basic
-characteristics of the dwelling. Table \ref{tab:Dummy-variables} show
-the more relevant variables included in the sf object.
+In addition to the variables listed in Table \ref{tab:variables}, the sf
+object includes a set of dummy variables with information about the
+basic characteristics of the dwelling. Table \ref{tab:Dummy-variables}
+shows the more relevant variables included in the sf object.
 
@@
   HASWARDROBE & =1 if has wardrobe & 0.30 & 0.57 & 0.53 \\ 
-  HASSWIMMINGPOOL & =1 if has swinningpool & 0.03 & 0.15 & 0.07 \\ 
+  HASSWIMMINGPOOL & =1 if has swimmingpool & 0.03 & 0.15 & 0.07 \\ 
   HASDOORMAN & =1 if has doorman & 0.08 & 0.25 & 0.05 \\ 
@@
   ISSTUDIO & =1 if is studio & 0.02 & 0.03 & 0.01 \\ 
-  ISINTOPFLOOR & =1 is in the top floor & 0.02 & 0.02 & 0.01 \\ 
+  ISINTOPFLOOR & =1 is on the top floor & 0.02 & 0.02 & 0.01 \\ 
   BUILTTYPEID\_1 & =1 if is new contruction & 0.01 & 0.03 & 0.03 \\ 
   BUILTTYPEID\_2 & =1 is second hand to be restored & 0.17 & 0.19 & 0.13 \\ 
-  BUILTTYPEID\_3 & =1 is second hand good conditions & 0.82 & 0.78 & 0.83 \\ 
+  BUILTTYPEID\_3 & =1 is second hand in good condition & 0.82 & 0.78 & 0.83 \\ 
    \hline
 \end{tabular}
-\caption{List of dummy variables, sort description and ratio of dwelling with the specific characteristic.
-See the help facility in the \textbf{idealista18} R package for details and formal definitions.
-Some dummy variables has been excluded of this table for save space.
-\label{tab:Dummy-variables}}
+\caption{List of dummy variables, sort description, and ratios of dwellings with the specific characteristics.
+See the help section in the \textbf{idealista18} R package for details and formal definitions.
+Some dummy variables have been excluded from this table to save space\label{tab:Dummy-variables}}
 \end{table}
@@
 
-The second block of data included in the `idealista18' R package are the
-spatial features of the three cities divided in neighborhoods. There are
-an sf object for each city with the name of the city and the suffix
-'\_Polygons'. The Figure \ref{fig:all-polygons} shows the quantile maps
-of the number of dwellings in the listing for the different
-neighborhoods in the three cities. The neighborhoods are based on the
-official boundaries but slightly adapted by
-Idealista\footnote{The criterion used to adapt this division is double, if an area is small enough and similar enough to another they merge both areas, on the other hand if the official area is not homogeneous it is then divided in a series of new polygons}.
-In practical terms we can assume they are the same, since the website
-simply collapses areas when they are sufficiently small in terms of
-number of ads. In the case of Madrid they just collapse four areas into
-two new ones.
+The second block of data included in the `idealista18' R package is the
+spatial features of the three cities divided into neighborhoods. There
+is an sf object for each city with the name of the city and the suffix
+'\_Polygons'. Figure \ref{fig:all-polygons} shows the quantile maps of
+the number of dwellings in the listing for the different neighborhoods
+in the three cities. The neighborhoods are based on the official
+boundaries but slightly changed by
+Idealista\footnote{There are two criteria used to make this division.
+If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}.
+In practical terms, we can assume they are the same since the website
+combines areas when there are few ads for that area. In the case of
+Madrid, they combined four areas into two.
 
@@
 dwellings in each neighborhood. Boundary for Barcelona (Left), Madrid
-(Center) and Valencia (Rigth).}
+(Center), and Valencia (Right).}
 \end{figure}
 
-There are a total of 69 neighborhoods in Barcelona, 135 in Madrid and 73
-in Valencia. The sf object include an unique identifier (LOCATIONID) and
-the neighborhood name (LOCATIONNAME).
+There are a total of 69 neighborhoods in Barcelona, 135 in Madrid, and
+73 in Valencia. The sf object includes a unique identifier (LOCATIONID)
+and the neighborhood name (LOCATIONNAME).
 
@@
 
-The last block of data included in the data package is a set of Point of
-Interest of each city as an object of the class list. The name of the
-list include the name of the city with the suffix '\_POIS'. These lists
-include three elements: (i) the coordinates of the city center,
-identifying the central business district; (ii) a set of coordinates
-that define the main street of each city; and (iii) the coordinates of
-metro stations.
+The last block of data included in the data package is a set of Points
+of Interest in each city as an object of the class list. The name of the
+list includes the name of the city with the suffix '\_POIS'. These lists
+include three elements: (i) the coordinates of the city center, the
+central business district; (ii) a set of coordinates that define the
+main street of each city; and (iii) the coordinates of the metro
+stations.
 
@@
 
-To comply with Spanish regulations, two variables are slightly modified
-to preserve their anonymity. A masking process is apply to asking prices
-and localization (coordinates).
+To comply with Spanish regulations, two variables were slightly modified
+to provide anonymity. A masking process was applied to asking prices and
+location (coordinates).
 
-Whit respect to the asking prices, the original values are obfuscated
-with the addition or subtraction of a random percentage of their
-original values ranging from -2.5\% to +2.5\%. Since asking prices are
-usually multiples of 1,000, after the first price modification, the
-prices was aligned to multiples of 1,000.
+In terms of the asking prices, the original values were obfuscated with
+the addition or subtraction of a random percentage of their original
+values, ranging from -2.5\% to +2.5\%. Since asking prices are usually
+multiples of 1,000, after the first price modification, the prices were
+aligned to multiples of 1,000.
 
@@
 
-With respect to the dwelling localization, a spatial masking process was
-implemented with the intention of keeping spatial properties of the
-original data set. The coordinates of each listing were displaced using
-a stochastic procedure. Effectively, the listings were recorded using
-coordinates contained in a maximum and minimum displacement circles, as
-shown in Figure \ref{fig:Anonymizing} (left). To preserve membership in
-a neighborhood, the spatial masking procedure was constrained to ensure
-that the masked coordinates are in the original neighborhood of the
-listing.
+With respect to the location of the dwelling, a spatial masking process
+was implemented to maintain the spatial properties of the original data
+set. The coordinates of each listing were displaced using a stochastic
+procedure. The listings were recorded using coordinates contained in
+maximum and minimum displacement circles, as shown in Figure
+\ref{fig:Anonymizing} (left). To preserve inclusion in a neighborhood,
+the spatial masking procedure was constrained to ensure that the masked
+coordinates remained in the original neighborhood of the listing.
 
-The Algorithm \ref{algo:coordinates-displacement} iteratively displaces
-the coordinates of each listing with a minimum distance and a maximum
-distance with the restriction that the new coordinates do not fall in a
-different neighborhood. This ensures that neighborhood attributes are
+Algorithm \ref{algo:coordinates-displacement} iteratively displaces the
+coordinates of each listing with a minimum distance and a maximum
+distance with the restriction that the new coordinates do not fall into
+a different neighborhood. This ensures that neighborhood attributes are
 preserved.
 
-Figure \ref{fig:Anonymizing} (right) shows the histogram of
-displacements in meters for all listings in the city of Valencia; the
-average distance between the original and masked coordinates is 45
+Figure \ref{fig:Anonymizing} (right) shows the histogram of the
+displacements in meters for all the listings in the city of Valencia.
+The average distance between the original and masked coordinates is 45
 meters.
@@
 
-\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Rigth) Coordinate displacement in meters Valencia}\label{fig:unnamed-chunk-2}
+\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)}\label{fig:unnamed-chunk-2}
 \end{figure}
@@
 
-This paper describe a data product of geo-referenced micro-data set of
-the three major Spanish cities. This is an excellent data product for
-understand the complex mechanisms related with the housing market and
-housing price. The researches can applying hedonic models with spatial
-effects, identifying housing submarkets or applying machine learning
-techniques. Also, the data product can be used for education proposes
-and teaching activities.
+This paper describes a data product of a geo-referenced micro-data set
+of Spain's three largest cities. This is an excellent data product to
+help understand the complex mechanisms related to the housing market and
+housing prices. Researchers can apply hedonic models with spatial
+effects, identifying housing submarkets or machine learning techniques.
+The data product can also be used for educational proposes and teaching
+activities.
 
@@
 
-Author One and author Two are employed by Idealista. They have been
-granted permission to share the data presented in this article. None of
-the authors have known competing financial interests or personal
-relationships which have, or could be perceived to have, influenced the
-work reported in this article.
+The authors wish to thank Alessandro Galesi for his support in the paper
+revision and Juan Ramón Selva for collecting and cleaning the spatial
+data. This work has been partially funded by the Spanish Ministry of
+Economy and Competitiveness Grant PID2019-107800GB-100.
 

### 2022-02-18 R/idealista18.R

@@ -175,7 +175,7 @@ NULL
 #'   \item{HASGARDEN}{Dummy variable for garden (takes 1 if there is a garden in the building, 0 otherwise)}
 #'   \item{ISDUPLEX}{Dummy variable for duplex (takes 1 if it is a duplex, 0 otherwise)}
 #'   \item{ISSTUDIO}{Dummy variable for bachelor apartment (studio in Spanish) (takes 1 if it is a single-person apartment, 0 otherwise)}
-#'   \item{ISINTOPFLOOR}{Dummy variable indicating if the apartment is located in the top floor (takes 1 on the top floor 0 otherwise)}
+#'   \item{ISINTOPFLOOR}{Dummy variable indicating if the apartment is located on the top floor (takes 1 on the top floor 0 otherwise)}
 #'   \item{CONSTRUCTIONYEAR}{Construction year (source: advertiser)}
 #'   \item{FLOORCLEAN}{Indicates flat floornumber starting from the 0 value for groung floor (source: advertiser)}
 #'   \item{FLATLOCATIONID}{Indicates the kind of views the flat has (1 - external, 2 - internal)}

### 2022-02-18 EPB/EPB.Rmd
Declaración de intereses

@@
 
-The authors wish to thank Alessandro Galesi for his support in the paper revision and  Juan  Ramón  Selva  for  collecting  and  cleaning  the  spatial  data.
-This  work  has been partially funded by the Spanish Ministry of Economy and Competitiveness Grant PID2019-107800GB-100.
+Author One and author Two are employed by Idealista.
+They have been granted permission to share the data presented in this article.
+None of the authors have financial interests or personal relationships which have, or could be perceived to have, influenced the work reported in this article.
 

### 2022-02-18 EPB/EPB.tex
Declaración intereses 2

@@
 
-The authors wish to thank Alessandro Galesi for his support in the paper
-revision and Juan Ramón Selva for collecting and cleaning the spatial
-data. This work has been partially funded by the Spanish Ministry of
-Economy and Competitiveness Grant PID2019-107800GB-100.
+Author One and author Two are employed by Idealista. They have been
+granted permission to share the data presented in this article. None of
+the authors have financial interests or personal relationships which
+have, or could be perceived to have, influenced the work reported in
+this article.
 

### 2022-02-24 EPB/main_EPB.Rmd
Anonimization submission

--- 2022-02-18 EPB/EPB.Rmd
@@
 author:
-- name: D. Rey-Blanco
-  num: 1
-- name: P. González-Arbues
-  num: 1
-- name: F. López
-  num: 2
-- name: A. Páez*
-  num: 3
-address:
-- num: 1
-  org: Idealista, Plaza de las Cortes 5, 28014 Madrid, Spain
-- num: 2
-  org: Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia, Spain
-- num: 3
-  org: School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada
-corrauth: "Antonio Páez, School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada."
-email: paezha@mcmaster.ca
+- name: 
 abstract: "This article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain.
@@
 The data product is offered as a fully documented R package and is available for scientific and educational purposes, particularly for geo-spatial studies"
+
 keywords: Housing market; idealista.com; geo-referenced data; point-level data; open data; Spain
@@
 The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest.
-The data set is distributed as an R package, named ‘idealista18’, which can be accessed from the Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
-
-<!-- @cohen2008spatial (508 observations) -->
-<!-- @lazrak2014market (19000 observations) -->
+The data set is distributed as an R package, named ‘idealista18’, which can be accessed from the Github repository.
 
@@
 The data product can also be used for educational proposes and teaching activities.
-
-# Declaration of Competing Interest
-
-The authors wish to thank Alessandro Galesi for his support in the paper revision and  Juan  Ramón  Selva  for  collecting  and  cleaning  the  spatial  data.
-This  work  has been partially funded by the Spanish Ministry of Economy and Competitiveness Grant PID2019-107800GB-100.
-
-# Acknowledgments
-
-The authors wish to thank Alessandro Galesi for their support in the paper revision and Juan Ramón Selva for collecting and cleaning the spatial data.
-This work has been partially funded by the Spanish Ministry of Economy and Competitiveness Grants PID2019-107800GB-100

### 2022-02-24 EPB/main_EPB.tex
Anonimization submission

--- 2022-02-18 EPB/EPB.tex
@@
 
-\author{D. Rey-Blanco\affilnum{1}, P. González-Arbues\affilnum{1}, F.
-López\affilnum{2}, A. Páez*\affilnum{3}}
+\author{\affilnum{}}
 
-\affiliation{\affilnum{1}{Idealista, Plaza de las Cortes 5, 28014
-Madrid, Spain}\\\affilnum{2}{Facultad de CC de la Empresa, C/ Real, 3.
-30201 Cartagena, Murcia, Spain}\\\affilnum{3}{School of Earth,
-Environment and Society, McMaster University, 1280 Main St W, Hamilton,
-Ontario L8S 4K1 Canada}}
-
-\corrauth{Antonio Páez, School of Earth, Environment and Society,
-McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada.}
-
-\email{\href{mailto:paezha@mcmaster.ca}{\nolinkurl{paezha@mcmaster.ca}}}
+\affiliation{}
 
@@
 data set is distributed as an R package, named `idealista18', which can
-be accessed from the Github
-repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
+be accessed from the Github repository.
 
@@
 \centering
-\includegraphics{EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
+\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
 \caption{\label{fig:all-polygons}Quantile maps of the number of
@@
 
-\hypertarget{declaration-of-competing-interest}{%
-\section{Declaration of Competing
-Interest}\label{declaration-of-competing-interest}}
-
-The authors wish to thank Alessandro Galesi for his support in the paper
-revision and Juan Ramón Selva for collecting and cleaning the spatial
-data. This work has been partially funded by the Spanish Ministry of
-Economy and Competitiveness Grant PID2019-107800GB-100.
-
-\hypertarget{acknowledgments}{%
-\section{Acknowledgments}\label{acknowledgments}}
-
-The authors wish to thank Alessandro Galesi for their support in the
-paper revision and Juan Ramón Selva for collecting and cleaning the
-spatial data. This work has been partially funded by the Spanish
-Ministry of Economy and Competitiveness Grants PID2019-107800GB-100
-
 \bibliographystyle{sageh}

### 2022-02-24 EPB/title_page_EPB.tex
Anonimization submission

--- 2022-02-18 EPB/EPB.tex
@@
-\documentclass[Royal,times,sageh]{sagej}
+\documentclass[Royal,times,]{sagej}
 
@@
 \usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 
@@
 
-\author{D. Rey-Blanco\affilnum{1}, P. González-Arbues\affilnum{1}, F.
+\author{D. Rey-Blanco\affilnum{1}, P. Arbues\affilnum{1}, F.
 López\affilnum{2}, A. Páez*\affilnum{3}}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Interest in the characteristics of the housing market and housing prices
-has been a growing area of research in recent decades, generating a vast
-amount of theoretical and empirical literature. Including the spatial
-component to analyze the real estate market and incorporating geographic
-variables has significantly improved the understanding of this market.
-But to really understand the characteristics of the housing market, it
-is essential to have information/data at the point level. Therefore, it
-is becoming common for spatial analysis of urban environments to be
-developed with geo- referenced micro-data sets \citep{lopez2015}.
-However, the availability of this type of open data at the point level
-is limited, and not many data sets contain latitude/longitude
-coordinates for each dwelling. In some cases, researchers have had to
-resort to web scraping processes to obtain the large volumes of
-information that permit robust analyses
-\citep{gupta2022take, arbia2020spatial, Li2019, lopez2015}. These web
-scraping processes can include missing data, download errors, duplicate
-records, etc. Furthermore, the authors of this research do not generally
-share the data sets.
-
-We are also witnessing a growing interest in open data in geography and
-data science \citep{arribasl2021editorial, arribas2021} using
-reproducible or replicable research \citep{paez2021open}. But to work
-openly in science, it is necessary to have free software and open data.
-While great efforts have been made to make free software available to
-researchers (e.g., R or Python), not much data is currently out in the
-open. In the particular case of the real estate market, to our
-knowledge, there are few open micro-data sets of housing markets
-available \citep{Song2021}.
-
-To overcome these limitations, this paper presents a sort description of
-an open micro-data set of geo-referenced dwelling listings. The data
-have been provided by the Idealista
-company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal}
-and contain information about 189,923 dwellings located in Spain's three
-largest cities. To date, this data product is the biggest open
-geo-referenced micro-data set of the housing market in Spain. Moreover,
-the data set has been supplied directly by Idealista, and therefore is
-clean and free of download errors. The listings have been enriched with
-official information from the Spanish cadastre along with other relevant
-geographical features, such as distance to urban points of interest. The
-data set is distributed as an R package, named `idealista18', which can
-be accessed from the Github
-repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
-
-\hypertarget{data-description}{%
-\section{Data description}\label{data-description}}
-
-The open data set `idealista18' is an R package composed of nine
-objects, three objects for each of the three main Spanish cities:
-Barcelona, Madrid, and Valencia. For each city, dwelling listings,
-neighborhood polygons, and a set of points of interest have been
-included in the R package. The following subsections describe each
-object. A full description of the data is available in the help section
-of the package.
-
-\hypertarget{dwelling-listings}{%
-\subsection{Dwelling listings}\label{dwelling-listings}}
-
-The dwelling listing of each city includes a set of characteristics for
-each dwelling published on the idealista real estate website as an ad.
-The dwelling listing has been included in the `idealista18' package as
-an sf object \citep{Pebesma}. The name of the sf object containing the
-dwelling listing includes the name of the city, followed by '\_Sale'
-(e.g., Madrid\_Sale) and includes a total of 42 variables. Each sf
-object includes the complete set of listings corresponding to the four
-quarters of the year 2018. Table \ref{tab:number-ads} shows the number
-of dwelling listing ads included in the data set for each city and
-quarter. The record counts for each city are: 94,815 listings for
-Madrid, 61,486 for Barcelona, and 33,622 for Valencia. Note that the
-same dwelling may be found in more than one period when a property
-listed for sale in one quarter was sold in a subsequent quarter. The
-variable ASSETID, included in the sf objects, is the unique identifier
-of the dwelling.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{>{\raggedright\arraybackslash}p{4em}>{\raggedleft\arraybackslash}p{3em}cccc}
-  \hline
-City$\backslash$Quarter & First & Second  & Thirdr & Fourth & Total ads \\ 
-  \hline
-Barcelona & 17826 & 7951 & 12375 & 23334 & 61486 \\ 
-  Madrid & 21920 & 12652 & 15973 & 44270 & 94815 \\ 
-  Valencia & 9305 & 4655 & 5644 & 14018 & 33622 \\ 
-   \hline
-\end{tabular}
-\caption{Number of dwelling  listing ads for each city and quarter. \label{tab:number-ads}} 
-\end{table}
-
-Each record of the dwelling listing contains a set of indoor
-characteristics supplied by the advertisers on the Idealista website
-(e.g., price, surface area, number of bedrooms, basic features, etc.),
-including the exact location of the dwelling (see Section
-\protect\hyperlink{anonymizing}{Anonymizing the data set}). Table
-\ref{tab:variables} lists the main indoor variables with a short
-description and the mean value of each variable. The dwelling listings
-were enriched with a number of additional attributes from the Spanish
-cadastre \citep{Catastro}. The cadastral information is described in
-Table \ref{tab:variables}, with the prefix CAD in the variable name. The
-cadastral features were assigned by applying the features of the nearest
-parcel to the coordinates. The year the dwelling was built
-(CONSTRUCTIONYEAR) given by the advertiser was revised since the year of
-construction is entered on the website by users, and it is therefore
-subject to errors and incomplete data (40\% missing data). To resolve
-this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included,
-assigning the cadastral construction year from the nearest cadastral
-parcel whenever the value was outstanding (date was after publication
-date or year of construction was before 1500) or when the value supplied
-by the advertiser was missing.
-
-Additionally, the distance of each dwelling to three urban points of
-interest was included in the sf object: distance to the city center,
-distance to the closest metro station, and distance to a major street
-(La Diagonal for Barcelona, La Castellana for Madrid, and Blasco Ibañez
-for Valencia). The last rows of Table \ref{tab:variables} show the mean
-values of these variables.
-
-\begin{table}[ht]
-\centering
-\fontsize{8}{10}\selectfont
-\begin{tabular}{>{\raggedright\arraybackslash}p{13em}>{\raggedright\arraybackslash}p{14em}ccc}
-  \hline
-Variable & Sort Description & Barcelona & Madrid & Valencia \\ 
-  \hline
-PRICE & Asking price & 395770.58 & 396110.11 & 199678.31 \\ 
-  UNITPRICE & Asking price per m\verb|^|2 (euros) & 4044.86 & 3661.05 & 1714.54 \\ 
-  CONSTRUCTEDAREA & Surface (m\verb|^|2) & 95.46 & 101.40 & 108.95 \\ 
-  ROOMNUMBER & Number of bedrooms & 2.86 & 2.58 & 3.07 \\ 
-  BATHNUMBER & Number of bathrooms & 1.52 & 1.59 & 1.59 \\ 
-  CONSTRUCTIONYEAR & Construction year (advertiser) & 1952.58 & 1964.69 & 1969.43 \\ 
-  CADCONSTRUCTIONYEAR & Construction year (cadastre) & 1952.19 & 1965.70 & 1970.55 \\ 
-  CADMAXBUILDINGFLOOR & Max build floor & 6.85 & 6.38 & 7.04 \\ 
-  CADDWELLINGCOUNT & Dwelling count in the building & 28.56 & 39.19 & 36.83 \\ 
-  CADASTRALQUALITYID & Cadastral quality. 0 Best-10 Worst & 4.31 & 4.85 & 5.34 \\ 
-  DISTANCE\_TO\_CITY\_CENTER & Distance to city center & 2.80 & 4.49 & 2.09 \\ 
-  DISTANCE\_TO\_METRO & Distance to subway station & 0.27 & 0.48 & 0.64 \\ 
-  DISTANCE\_TO\_(MAINSTREET) & Distance to major street & 1.77 & 2.68 & 2.07 \\ 
-   \hline
-\end{tabular}
-\caption{List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities.
-See the help section in the \textbf{idealista18} R package for details and formal definitions.
-Some variables have been excluded from this table to save space.
-Check the full list in the \textbf{idealista18} package.\label{tab:variables}}
-\end{table}
-
-In addition to the variables listed in Table \ref{tab:variables}, the sf
-object includes a set of dummy variables with information about the
-basic characteristics of the dwelling. Table \ref{tab:Dummy-variables}
-shows the more relevant variables included in the sf object.
-
-\begin{table}[ht]
-\centering
-\fontsize{8}{10}\selectfont
-\begin{tabular}{>{\raggedright\arraybackslash}p{12em}>{\raggedright\arraybackslash}p{14em}ccc}
-  \hline
-Variable & Sort Description & Barcelona & Madrid & Valencia \\ 
-  \hline
-HASTERRACE & =1 if has terrace & 0.33 & 0.36 & 0.25 \\ 
-  HASLIFT & =1 if has lift & 0.74 & 0.70 & 0.79 \\ 
-  HASAIRCONDITIONING & =1 if has air conditioning & 0.47 & 0.45 & 0.47 \\ 
-  HASPARKINGSPACE & =1 if has parking & 0.08 & 0.23 & 0.17 \\ 
-  HASNORTHORIENTATION & =1 if has north orientation & 0.13 & 0.11 & 0.13 \\ 
-  HASSOUTHORIENTATION & =1 if has south orientation & 0.31 & 0.24 & 0.19 \\ 
-  HASEASTORIENTATION & =1 if has east orientation & 0.24 & 0.20 & 0.25 \\ 
-  HASWESTORIENTATION & =1 if has west orientation & 0.16 & 0.15 & 0.15 \\ 
-  HASBOXROOM & =1 if has boxroom & 0.12 & 0.26 & 0.13 \\ 
-  HASWARDROBE & =1 if has wardrobe & 0.30 & 0.57 & 0.53 \\ 
-  HASSWIMMINGPOOL & =1 if has swimmingpool & 0.03 & 0.15 & 0.07 \\ 
-  HASDOORMAN & =1 if has doorman & 0.08 & 0.25 & 0.05 \\ 
-  HASGARDEN & =1 if has garden & 0.04 & 0.18 & 0.06 \\ 
-  ISDUPLEX & =1 if is duplex & 0.03 & 0.03 & 0.02 \\ 
-  ISSTUDIO & =1 if is studio & 0.02 & 0.03 & 0.01 \\ 
-  ISINTOPFLOOR & =1 is on the top floor & 0.02 & 0.02 & 0.01 \\ 
-  BUILTTYPEID\_1 & =1 if is new contruction & 0.01 & 0.03 & 0.03 \\ 
-  BUILTTYPEID\_2 & =1 is second hand to be restored & 0.17 & 0.19 & 0.13 \\ 
-  BUILTTYPEID\_3 & =1 is second hand in good condition & 0.82 & 0.78 & 0.83 \\ 
-   \hline
-\end{tabular}
-\caption{List of dummy variables, sort description, and ratios of dwellings with the specific characteristics.
-See the help section in the \textbf{idealista18} R package for details and formal definitions.
-Some dummy variables have been excluded from this table to save space\label{tab:Dummy-variables}}
-\end{table}
-
-\hypertarget{neighboorhood-polygons}{%
-\subsection{Neighboorhood polygons}\label{neighboorhood-polygons}}
-
-The second block of data included in the `idealista18' R package is the
-spatial features of the three cities divided into neighborhoods. There
-is an sf object for each city with the name of the city and the suffix
-'\_Polygons'. Figure \ref{fig:all-polygons} shows the quantile maps of
-the number of dwellings in the listing for the different neighborhoods
-in the three cities. The neighborhoods are based on the official
-boundaries but slightly changed by
-Idealista\footnote{There are two criteria used to make this division.
-If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}.
-In practical terms, we can assume they are the same since the website
-combines areas when there are few ads for that area. In the case of
-Madrid, they combined four areas into two.
-
-\begin{figure}
-\centering
-\includegraphics{EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:all-polygons}Quantile maps of the number of
-dwellings in each neighborhood. Boundary for Barcelona (Left), Madrid
-(Center), and Valencia (Right).}
-\end{figure}
-
-There are a total of 69 neighborhoods in Barcelona, 135 in Madrid, and
-73 in Valencia. The sf object includes a unique identifier (LOCATIONID)
-and the neighborhood name (LOCATIONNAME).
-
-\hypertarget{points-of-interest}{%
-\subsection{Points of Interest}\label{points-of-interest}}
-
-The last block of data included in the data package is a set of Points
-of Interest in each city as an object of the class list. The name of the
-list includes the name of the city with the suffix '\_POIS'. These lists
-include three elements: (i) the coordinates of the city center, the
-central business district; (ii) a set of coordinates that define the
-main street of each city; and (iii) the coordinates of the metro
-stations.
-
-\hypertarget{anonymizing}{%
-\section{Anonymizing the data set}\label{anonymizing}}
-
-To comply with Spanish regulations, two variables were slightly modified
-to provide anonymity. A masking process was applied to asking prices and
-location (coordinates).
-
-In terms of the asking prices, the original values were obfuscated with
-the addition or subtraction of a random percentage of their original
-values, ranging from -2.5\% to +2.5\%. Since asking prices are usually
-multiples of 1,000, after the first price modification, the prices were
-aligned to multiples of 1,000.
-
-\begin{algorithm}[!ht]
- \KwData{all idealista listings}
- \KwResult{all idealista listings with masked coordinates}
- initialization\;
- \For{each listing L}{
-  take geographical location of L as $(X,Y)$
-  \Repeat{this stop condition}{
-    take a random angle $\alpha$ from 0 to 360 degrees
-    take a distance $R$ as a random value from 30 to 60 meters
-    determine a new point $(X',Y')$ calculated as a point located $R$ with the angle $\alpha$
-  }
-  set $(X',Y')$ as the new location for the listing L
- }
- \caption{Coordinate displacement process for anonymisation purposes}
- \label{algo:coordinates-displacement}
-\end{algorithm}
-
-With respect to the location of the dwelling, a spatial masking process
-was implemented to maintain the spatial properties of the original data
-set. The coordinates of each listing were displaced using a stochastic
-procedure. The listings were recorded using coordinates contained in
-maximum and minimum displacement circles, as shown in Figure
-\ref{fig:Anonymizing} (left). To preserve inclusion in a neighborhood,
-the spatial masking procedure was constrained to ensure that the masked
-coordinates remained in the original neighborhood of the listing.
-
-Algorithm \ref{algo:coordinates-displacement} iteratively displaces the
-coordinates of each listing with a minimum distance and a maximum
-distance with the restriction that the new coordinates do not fall into
-a different neighborhood. This ensures that neighborhood attributes are
-preserved.
-
-Figure \ref{fig:Anonymizing} (right) shows the histogram of the
-displacements in meters for all the listings in the city of Valencia.
-The average distance between the original and masked coordinates is 45
-meters.
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.29\linewidth,height=0.2\textheight]{EPB_files/points-moved-image} \includegraphics[width=0.37\linewidth,height=0.2\textheight]{EPB_files/coordinates-valencia}
-
-}
-
-\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)}\label{fig:unnamed-chunk-2}
-\end{figure}
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-This paper describes a data product of a geo-referenced micro-data set
-of Spain's three largest cities. This is an excellent data product to
-help understand the complex mechanisms related to the housing market and
-housing prices. Researchers can apply hedonic models with spatial
-effects, identifying housing submarkets or machine learning techniques.
-The data product can also be used for educational proposes and teaching
-activities.
-
 \hypertarget{declaration-of-competing-interest}{%
@@
 
-\bibliographystyle{sageh}
-\bibliography{bibEPB.bib}
-
 \end{document}

### 2022-07-12 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd
Copied decision to `editorial-materials` in preparation to begin revising the paper

---
title: "Using machine learning to identify spatial market segments: A reproducible study of major Spanish markets - Response to referees"
bibliography: bibliography.bib
output: pdf_document
---

<!--
```{r install-data-package, include = FALSE}
# Run only once if needed to install data package `sobiEquity`
if (!require("idealista18", character.only = TRUE)) {
      devtools::install_github("https://github.com/paezha/idealista18")
  }
```

```{r setup,  cache=FALSE, warning = FALSE, include=FALSE, message = FALSE}
library(idealista18)
library(tidyverse)

set.seed(1234)
```
-->

> Dear Authors, Thank you for the opportunity to review your paper. This manuscript is about the publishing of a one-year relocated and randomised listing price data set for three cities in Spain. Both the asking price and coordinates information are anonymised which limits the usefulness of the data for teaching and academic research. Given the house price data is partly from the Idealista company, it is not clear under which kind of licence the data is being issued and to what extent it may be considered open. The unknown quality of this one year randomised data is the main issue for potential users. Having read the paper many times, I believe that the authors have created the idealista18 package (https://github.com/paezha/idealista18). The Idealista18 Github readme file is too simple to guidance users on how to access the data with the R package. So I am not sure it is the right one. The authors are to be commended for their intention to make open real estate data available in an R package which should render it easily accessible for teaching and research. Below is a list of comments and questions that I had which may need to be addressed in future editions of this work.

> 1.	Please reference the package and dataset and include in the text and in the package the license under which the data is released. The Idealista18 github site is unclear on how to access the data with the package.

> 2.	Please state your anticipated research impacts from the publication of this data set in the introduction.  It would be useful to state how this one-year relocated, and randomised house price data might be best used by academics in research and teaching. The relocated and randomised data may be difficult to further enrich with other variables, the authors need to show how the 44 variables is enough for future users.

>3.	The method of randomisation is described but it would be advantageous to publish the code for this so that researchers can evaluate the impact of this kind of randomisation on this and other data sets. Does the method, as seems likely, have a differential effect on data from high density areas as compared with low?
To what extent are analyses likely to be confined to neighbourhood levels of resolution given the method of spatial randomisation? It would be useful to have some idea of the size of the neighbourhoods to facilitate comparison with other spatial data sets

>4.	More details on the data collection methods of Idealista would be useful in interpreting the data. Does Idealista concentrate on particular market segments or not? Are the list prices good indicators of actual sale prices? Are there sources from which actual paid prices could be identified?

>5.	It would be useful for the authors to evaluate the datasets in the light of the FAIR guiding principles for research data stewardship (findability, accessibility, interoperability, and reusability).

>6.	Given the asking price has resized the original asking price with a random percentage between -2.5% and +2.5%, please elaborate on how this influences quantitative research results and findings with this data.

>7.	Please kindly add in a usage notes section how this data can be useful for the EPB audience. For example, introduce how this anonymised house price data can be integrated with other data and research.
The original data could be a nice data set for a series of housing related papers. If the authors are unable to publish this original asking price, it maybe worthwhile to publish the research based on this data rather than using the randomised data. House price is locationally sensitive, but both the location and price data in this manuscript are altered by anonymisation. The authors need to show for which analyses the data remains valid despite anonymisation.

# References

### 2022-07-12 EPB/title_page_EPB.Rmd

--- 2022-02-18 EPB/EPB.Rmd
@@
   num: 1
-- name: P. González-Arbues
+- name: P. Arbues
   num: 1
@@
   - times
-bibliography: bibEPB.bib
-bibliographystyle: sageh
 
@@
 ---
-
-# Introduction
-
-Interest in the characteristics of the housing market and housing prices has been a growing area of research in recent decades, generating a vast amount of theoretical and empirical literature.
-Including the spatial component to analyze the real estate market and incorporating geographic variables has significantly improved the understanding of this market.
-But to really understand the characteristics of the housing market, it is essential to have information/data at the point level.
-Therefore, it is becoming common for spatial analysis of urban environments to be developed with geo- referenced micro-data sets [@lopez2015].
-However, the availability of this type of open data at the point level is limited, and not many data sets contain latitude/longitude coordinates for each dwelling.
-In some cases, researchers have had to resort to web scraping processes to obtain the large volumes of information that permit robust analyses [@gupta2022take; @arbia2020spatial; @Li2019; @lopez2015].
-These  web  scraping  processes  can  include  missing  data, download errors, duplicate records, etc.
-Furthermore, the authors of this research do not generally share the data sets.
-
-We are also witnessing a growing interest in open data in geography and data science [@arribasl2021editorial; @arribas2021]  using reproducible or replicable  research  [@paez2021open].
-But  to  work  openly in  science, it  is  necessary to  have  free software and open data.
-While great efforts have been made to make free software available to researchers (e.g., R or Python), not much data is currently out in the open.
-In the particular case of the real estate market, to our knowledge, there are few open micro-data sets of housing markets available [@Song2021].
-
-To overcome these limitations, this paper presents a sort description of an open micro-data set of geo-referenced dwelling listings.
-The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain’s three largest cities.
-To date, this data product is the biggest open geo-referenced micro-data set of the housing market in Spain.
-Moreover, the data set has been supplied directly by Idealista, and therefore is clean and free of download errors.
-The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest.
-The data set is distributed as an R package, named ‘idealista18’, which can be accessed from the Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
-
-<!-- @cohen2008spatial (508 observations) -->
-<!-- @lazrak2014market (19000 observations) -->
-
-# Data description
-
-The open data set ‘idealista18’ is an R package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia.
-For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the R package.
-The following subsections describe each object.
-A full description of the data is available in the help section of the package.
-
-## Dwelling listings
-
-The dwelling listing of each city includes a set of characteristics for each dwelling published on the idealista real estate website as an ad.
-The dwelling listing has been included in the ‘idealista18’ package as an sf object [@Pebesma].
-The name of the sf object containing the dwelling listing includes the name of the city, followed by '_Sale' (e.g., Madrid_Sale) and includes a total of 42 variables.
-Each sf object includes the complete set of listings corresponding to the four quarters of the year 2018.
-Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter.
-The record counts for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622 for Valencia.
-Note that the same dwelling may be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter.
-The variable ASSETID, included in the sf objects, is the unique identifier of the dwelling.
-
-```{r number-ads, echo = FALSE, message = FALSE}
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
-H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
-H <- cbind(H,rowSums(H))
-text_tbl <- data.frame(
-City = c("Barcelona","Madrid","Valencia"),
-H,
-check.names = FALSE)
-colnames(text_tbl) <- c("City\\Quarter", "First", "Second ", "Thirdr", "Fourth","Total ads")
-xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") %>%
-xtable2kable(include.rownames = FALSE) %>%
-kable_styling(full_width = F) %>%
-column_spec(1, width = "4em", italic = FALSE) %>%
-column_spec(2, width = "3em") %>%
-column_spec(3, width = "3em", latex_column_spec = "c") %>%
-column_spec(4, width = "3em", latex_column_spec = "c") %>%
-column_spec(5, width = "3em", latex_column_spec = "c") %>%
-column_spec(6, width = "2em", latex_column_spec = "c")
-```
-
-Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including the exact location of the dwelling (see Section [Anonymizing the data set](#anonymizing)).
-Table \ref{tab:variables} lists the main indoor variables with a short description and the mean value of each variable.
-The dwelling listings were enriched with a number of additional attributes from the Spanish cadastre [@Catastro].
-The cadastral information is described in Table \ref{tab:variables}, with the prefix CAD in the variable name.
-The cadastral features were assigned by applying the features of the nearest parcel to the coordinates.
-The year the dwelling was built (CONSTRUCTIONYEAR) given by the advertiser was revised since the year of construction is entered on the website by users, and it is therefore subject to errors and incomplete data (40\% missing data).
-To resolve this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included, assigning the cadastral construction year from the nearest cadastral parcel whenever the value was outstanding (date was after publication date or year of construction was before 1500) or when the value supplied by the advertiser was missing.
-
-Additionally, the distance of each dwelling to three urban points of interest was included in the sf object: distance to the city center, distance to the closest metro station, and distance to a major street (La Diagonal for Barcelona, La Castellana for Madrid, and Blasco Ibañez for Valencia).
-The last rows of Table \ref{tab:variables} show the mean values of these variables.
-
-```{r table-variables, echo = FALSE, message = FALSE}
-dummyvariables <- c(3,4,5,6,7,27,30,31,32,33,37,38,39)
-text_tbl <- data.frame(
-Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
-"Sort Description" = c("Asking price","Asking price per m^2 (euros)","Surface (m^2)",
-                       "Number of bedrooms","Number of bathrooms",
-                       "Construction year (advertiser)","Construction year (cadastre)",
-                       "Max build floor","Dwelling count in the building",
-                       "Cadastral quality. 0 Best-10 Worst",
-                       "Distance to city center","Distance to subway station","Distance to major street"),
-# Barcelona
-Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables]), na.rm = TRUE),
-# Madrid
-Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables]), na.rm = TRUE),
-# Valencia
-"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables]), na.rm = TRUE),
-check.names = FALSE)
-text_tbl[dim(text_tbl)[1],1]<-"DISTANCE_TO_(MAINSTREET)"
-xtable::xtable(text_tbl, caption = "List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities.
-See the help section in the \\textbf{idealista18} R package for details and formal definitions.
-Some variables have been excluded from this table to save space.
-Check the full list in the \\textbf{idealista18} package.\\label{tab:variables}") %>%
-xtable2kable(include.rownames = FALSE) %>%
-kable_styling(full_width = F, font_size=8) %>%
-column_spec(1, width = "13em", italic = FALSE) %>%
-column_spec(2, width = "14em") %>%
-column_spec(3, width = "2em", latex_column_spec = "c") %>%
-column_spec(4, width = "2em", latex_column_spec = "c") %>%
-column_spec(5, width = "2em", latex_column_spec = "c")
-```
-
-In addition to the variables listed in Table \ref{tab:variables}, the sf object includes a set of dummy variables with information about the basic characteristics of the dwelling.
-Table \ref{tab:Dummy-variables} shows the more relevant variables included in the sf object.
-
-```{r table-dummy-variables, echo = FALSE, message = FALSE}
-dummyvariables <- c(8,9,10,12,15,16,17,18,19,20,21,22,23,24,25,26,34,35,36)
-text_tbl <- data.frame(
-Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
-"Sort Description" = c("=1 if has terrace","=1 if has lift",
-                       "=1 if has air conditioning","=1 if has parking",
-                       "=1 if has north orientation","=1 if has south orientation",
-                       "=1 if has east orientation","=1 if has west orientation",
-                       "=1 if has boxroom","=1 if has wardrobe",
-                       "=1 if has swimmingpool","=1 if has doorman",
-                       "=1 if has garden","=1 if is duplex",
-                       "=1 if is studio","=1 is on the top floor",
-                       "=1 if is new contruction","=1 is second hand to be restored","=1 is second hand in good condition"),
-# Barcelona
-Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables])),
-# Madrid
-Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables])),
-# Valencia
-"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables])),
-check.names = FALSE)
-
-xtable::xtable(text_tbl, caption = "List of dummy variables, sort description, and ratios of dwellings with the specific characteristics.
-See the help section in the \\textbf{idealista18} R package for details and formal definitions.
-Some dummy variables have been excluded from this table to save space\\label{tab:Dummy-variables}") %>%
-xtable2kable(include.rownames = FALSE) %>%
-kable_styling(full_width = F, font_size = 8) %>%
-column_spec(1, width = "12em", italic = FALSE) %>%
-column_spec(2, width = "14em") %>%
-column_spec(3, width = "2em", latex_column_spec = "c") %>%
-column_spec(4, width = "2em", latex_column_spec = "c") %>%
-column_spec(5, width = "2em", latex_column_spec = "c")
-```
-
-## Neighboorhood polygons
-
-The second block of data included in the ‘idealista18’ R package is the spatial features of the three cities divided into neighborhoods.
-There is an sf object for each city with the name of the city and the suffix '_Polygons'.
-Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities.
-The neighborhoods are based on the official boundaries but slightly changed by Idealista\footnote{There are two criteria used to make this division.
-If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}.
-In practical terms, we can assume they are the same since the website combines areas when there are few ads for that area.
-In the case of Madrid, they combined four areas into two.
-
-```{r, echo = FALSE,message=FALSE,fig.width=5, fig.cap="\\label{fig:all-polygons}Quantile maps of the number of dwellings in each neighborhood.
-Boundary for Barcelona (Left), Madrid (Center), and Valencia (Right)."}
-library("ggplot2")
-library("dplyr")
-library("gridExtra")
-# Barcelona
-Barcelona_counts <- Barcelona_Polygons %>% 
-  mutate(counts = lengths(st_intersects(., Barcelona_Sale)))
-q <- quantile(Barcelona_counts$counts)
-Barcelona_Polygons$Quantile<- as.factor((Barcelona_counts$counts > q[2]) + (Barcelona_counts$counts > q[3]) +(Barcelona_counts$counts >= q[4]) + 1)
-plot4 <- ggplot(data = Barcelona_Polygons) +
-  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
-  theme_bw(base_size = 6) +
-  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
-        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
-# Madrid
-Madrid_counts <- Madrid_Polygons %>% 
-  mutate(counts = lengths(st_intersects(., Madrid_Sale)))
-q <- quantile(Madrid_counts$counts)
-Madrid_Polygons$Quantile<- as.factor((Madrid_counts$counts > q[2]) + (Madrid_counts$counts > q[3]) +(Madrid_counts$counts >= q[4]) + 1)
-plot5 <- ggplot(data = Madrid_Polygons) +
-  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
-  theme_bw(base_size = 6) +
-  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
-        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
-# Valencia
-Valencia_counts <- Valencia_Polygons %>% 
-  mutate(counts = lengths(st_intersects(., Valencia_Sale)))
-q <- quantile(Valencia_counts$counts)
-Valencia_Polygons$Quantile<- as.factor((Valencia_counts$counts > q[2]) + (Valencia_counts$counts > q[3]) +(Valencia_counts$counts >= q[4]) + 1)
-plot6 <- ggplot(data = Valencia_Polygons) +
-  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
-  ylim(39.405,39.52) +
-  theme_bw(base_size = 6) +
-  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
-        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
-grid.arrange(plot4, plot5, plot6, ncol = 3, nrow = 1)
-```
-
-There are a total of 69 neighborhoods in Barcelona, 135 in Madrid, and 73 in Valencia. The sf object includes a unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).
-
-## Points of Interest
-
-The last block of data included in the data package is a set of Points of Interest in each city as an object of the class list.
-The name of the list includes the name of the city with the suffix '_POIS'.
-These lists include three elements: (i) the coordinates of the city center, the central business district; (ii) a set of coordinates that define the main street of each city; and (iii) the coordinates of the metro stations.
-
-# Anonymizing the data set {#anonymizing}
-
-To comply with Spanish regulations, two variables were slightly modified to provide anonymity. A masking process was applied to asking prices and location (coordinates).
-
-In terms of the asking prices, the original values were obfuscated with the addition or subtraction of a random percentage of their original values, ranging from -2.5% to +2.5%.
-Since asking prices are usually multiples of 1,000, after the first price modification, the prices were aligned to multiples of 1,000.
-
-```{=tex}
-\begin{algorithm}[!ht]
- \KwData{all idealista listings}
- \KwResult{all idealista listings with masked coordinates}
- initialization\;
- \For{each listing L}{
-  take geographical location of L as $(X,Y)$
-  \Repeat{this stop condition}{
-    take a random angle $\alpha$ from 0 to 360 degrees
-    take a distance $R$ as a random value from 30 to 60 meters
-    determine a new point $(X',Y')$ calculated as a point located $R$ with the angle $\alpha$
-  }
-  set $(X',Y')$ as the new location for the listing L
- }
- \caption{Coordinate displacement process for anonymisation purposes}
- \label{algo:coordinates-displacement}
-\end{algorithm}
-```
-
-With respect to the location of the dwelling, a spatial masking process was implemented  to maintain the spatial properties of the original data set.
-The coordinates of each listing were displaced using a stochastic procedure.
-The listings were recorded using coordinates contained in maximum and minimum displacement circles, as shown in Figure \ref{fig:Anonymizing} (left).
-To preserve inclusion in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates remained in the original neighborhood of the listing.
-
-Algorithm \ref{algo:coordinates-displacement} iteratively displaces the coordinates of each listing with a minimum distance and a maximum distance with the restriction that the new coordinates do not fall into a different neighborhood.
-This ensures that neighborhood attributes are preserved.
-
-Figure \ref{fig:Anonymizing} (right) shows the histogram of the displacements in meters for all the listings in the city of Valencia.
-The average distance between the original and masked coordinates is 45 meters.
-
-```{r, echo = FALSE, out.width=c("29%","37%"), out.height="20%",fig.cap="\\label{fig:Anonymizing}(Left) Masking coordinates.
-Spatial range.
-(Right) Coordinate displacement in meters (Valencia)",fig.show='hold',fig.align='center'}
-knitr::include_graphics(c("EPB_files/points-moved-image.png","EPB_files/coordinates-valencia.png"))
-```
-
-# Conclusion
-
-This paper describes a data product of a geo-referenced micro-data set of Spain’s three largest cities.
-This is an excellent data product to help understand the complex mechanisms related to the housing market and housing prices.
-Researchers can apply hedonic models with spatial effects, identifying housing submarkets or machine learning techniques.
-The data product can also be used for educational proposes and teaching activities.
-
-# Declaration of Competing Interest
-
-The authors wish to thank Alessandro Galesi for his support in the paper revision and  Juan  Ramón  Selva  for  collecting  and  cleaning  the  spatial  data.
-This  work  has been partially funded by the Spanish Ministry of Economy and Competitiveness Grant PID2019-107800GB-100.
-
-# Acknowledgments
-
-The authors wish to thank Alessandro Galesi for their support in the paper revision and Juan Ramón Selva for collecting and cleaning the spatial data.
-This work has been partially funded by the Spanish Ministry of Economy and Competitiveness Grants PID2019-107800GB-100

### 2022-07-12 README.Rmd
Copied decision to `editorial-materials` in preparation to begin revising the paper
Updated the description and README

@@
 <!-- badges: start -->
+[![License: CC BY-NC-SA 4.0](https://img.shields.io/badge/license-CC BY--NC--SA 4.0-blue.svg)](https://creativecommons.org/licenses/by-nc-sa/4.0/deed.en_GB)
+
+[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
+
+[![GitHub Stars](https://img.shields.io/github/stars/paezha/edashop?style=social)](https://github.com/paezha/idealista18/stargazers)
+
+![GitHub search hit counter: R](https://img.shields.io/github/search/paezha/idealista18/real-estate)  
+![GitHub search hit counter: exploratory](https://img.shields.io/github/search/paezha/idealista18/spatial)
+![GitHub search hit counter: data](https://img.shields.io/github/search/paezha/idealista18/data)
+![GitHub search hit counter: analysis](https://img.shields.io/github/search/paezha/idealista18/Spain)
+![GitHub search hit counter: workshop](https://img.shields.io/github/search/paezha/idealista18/open-data-products)
+
+![GitHub issues](https://img.shields.io/github/issues/paezha/idealista18)
+![GitHub release](https://img.shields.io/github/release-date/paezha/idealista18)
+![GitHub commit activity](https://img.shields.io/github/commit-activity/y/paezha/idealista18)
+![GitHub last commit](https://img.shields.io/github/last-commit/paezha/idealista18)
 <!-- badges: end -->
 
-The package idealista18 includes real estate data for three major markets in Spain (Barcelona, Madrid, and Valencia) for the year 2018.
+{idealista18} is an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain.
+These data were originally published on the idealista.com real estate website.
+The data sets are for the three largest cities in Spain: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations), and Valencia (n = 33,622 observations), and include approximate coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics.
+The listings were enriched with official information from the Spanish cadastre (e.g., building material quality) plus other relevant geographical features, such as distance to urban points of interest.
+Along with the real estate listings, the data product also includes neighborhood boundaries for each city.
+The data sets are offered as a fully documented R package and are available for scientific and educational purposes, particularly for geo-spatial studies.
 
@@
 ```
-## Example
+# Data description
 
-This is a basic example which shows you how to solve a common problem:
+There are nine data objects in the package: three objects for each of the cities included.
+For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the R package.
+The following subsections describe each object.
+A full description of the data is available in the help section of the package.
 
-```{r example}
-library(idealista18)
-## basic example code
+```{r}
+library(dplyr) # A Grammar of Data Manipulation
+library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
+library(idealista18) # Idealista 2018 Data Package
+#library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
+library(sf) # Simple Features for R
+library(skimr) 
 ```
+
+## Dwelling listings
+
+The dwelling listing of each city includes a set of characteristics for each dwelling published on the idealista real estate website as an ad.
+The dwelling listing has been included in the ‘idealista18’ package as an sf object.
+The name of the sf object containing the dwelling listing includes the name of the city, followed by '_Sale' (e.g., Madrid_Sale) and includes a total of 42 variables.
+Each sf object includes the complete set of listings corresponding to the four quarters of the year 2018.
+Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter.
+The record counts for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622 for Valencia.
+Note that the same dwelling may be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter.
+The variable ASSETID, included in the sf objects, is the unique identifier of the dwelling.
+
+```{r}
+Madrid_Sale |>
+  sf::st_drop_geometry() |> 
+  skimr::skim()
+```
+
+```{r}
+Barcelona_Sale |>
+  sf::st_drop_geometry() |> 
+  skimr::skim()
+```
+
+```{r}
+Valencia_Sale |>
+  sf::st_drop_geometry() |> 
+  skimr::skim()
+```
+## Neighboorhood polygons
+
+The second block of data included in the ‘idealista18’ R package is the spatial features of the three cities divided into neighborhoods.
+There is an sf object for each city with the name of the city and the suffix '_Polygons'.
+Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities.
+The neighborhoods are based on the official boundaries but slightly changed by Idealista\footnote{There are two criteria used to make this division.
+If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}.
+In practical terms, we can assume they are the same since the website combines areas when there are few ads for that area.
+In the case of Madrid, they combined four areas into two.
+
+There are a total of 69 neighborhoods in Barcelona, 135 in Madrid, and 73 in Valencia. The sf object includes a unique identifier (LOCATIONID) and the neighborhood name (LOCATIONNAME).
+
+## Points of Interest
+
+The last block of data included in the data package is a set of Points of Interest in each city as an object of the class list.
+The name of the list includes the name of the city with the suffix '_POIS'.
+These lists include three elements: (i) the coordinates of the city center, the central business district; (ii) a set of coordinates that define the main street of each city; and (iii) the coordinates of the metro stations.

### 2022-07-18 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd
Added comments on license and expected impact

@@
 
+We really appreciate this suggestion.
+We have included the license under which the data is distributed in the main text of the introduction.
+We are also working to publish the dataset via the R package usign CRAN.
+
 > 2.	Please state your anticipated research impacts from the publication of this data set in the introduction.
@@
 The relocated and randomised data may be difficult to further enrich with other variables, the authors need to show how the 44 variables is enough for future users.
+
+Thanks for this comment. We have added discussed these concerns in the introduction.
 
@@
 
+We have added some examples of research topics that can be addressed with this dataset  in the introduction.
+
 # References

### 2022-07-18 EPB/EPB.Rmd
Apellido Pelayo
Added comments on license and expected impact

@@
 Moreover, the data set has been supplied directly by Idealista, and therefore is clean and free of download errors.
+However, in order to be able to share publicly the data we have slightly modified some variables by applying some random noise which will not bias the main results derived from its usage.
+
+This micro-data set is expected to be used as a benchmark dataset to test empirical methods in a reproducible fashion.
+Applied and theoretical researchers on Real Estate Mass Appraisal and Valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others.
+This can also be used to study the segmentation of housing submarkets and related topics such as the impact of suburban areas on house prices.
 The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest.
+In any case, the data might be easily extended by spatially joining other datasets that contain information at administrative, census tract, or postal code levels.
+
 The data set is distributed as an R package, named ‘idealista18’, which can be accessed from the Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
+The database idealista18 is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
+We also share the randomisation process applied to the original dataset in the aforementioned Github repository for full transparency on the data collection process.
 

### 2022-07-29 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd
Adds comment on microsite
Added a comment to remind AP to respond to comment 5 of reviewer report
Draft of comment to reviewer

@@
 We are also working to publish the dataset via the R package usign CRAN.
+In addition we have also created a [microsite](https://paezha.github.io/idealista18/) for the package that shows how to access the data.
 
@@
 
+We impose a key constraint in the relocation process that is keeping the listing in the same neighborhood.
+For the process clarity sake we released the anonymization process in the R package code on github.
+[microsite](https://paezha.github.io/idealista18/anonymization)
+
 >4.	More details on the data collection methods of Idealista would be useful in interpreting the data.
@@
 
+Idealista covers fairy well all segments in spanish market, individual and profesional adverstiser.
+Although listing prices reflects the reality of real estate markets from the offer persperctive, strong correlation of idealista and transaction prices can be established see [Banco de España](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3400031).
+
 >5.	It would be useful for the authors to evaluate the datasets in the light of the FAIR guiding principles for research data stewardship (findability, accessibility, interoperability, and reusability).
@@
 >6.	Given the asking price has resized the original asking price with a random percentage between -2.5% and +2.5%, please elaborate on how this influences quantitative research results and findings with this data.
+
+It should not affect much given it is random noise.
 

### 2022-07-29 anonymization/record_anonymization.Rmd
Draft of comment to reviewer

---
title: 'Listings anonymization (price and coordinates)'
output: pdf_document
---

# Introduction

This markdown file contains the code used for the anonymization of the package's listing records (due to legal requirement). It introduces a small random disturbance in the total price and exact position.

# Set parameters

\tiny
```{r Load libraries, echo=F}
knitr::opts_chunk$set(eval = F)
knitr::opts_chunk$set(size="small")
```
\normalsize

Set process parameters

\tiny
```{r Set process parameters}
dir_geom_path <- paste0(here::here(),"/data/wkt/")  
dir_path <- paste0(here::here(),"/data/assets/raw/")
output_path <- str_remove(dir_path,  "raw/")

#
# Process parameters (expected from a PARAM list)
# here we define some settings
#
COUNTRY = 'es'     # Spain 
TYPOLOGY = 'home'  # This type refers to flats/apartments
OPERATION = 'rent' # This refers to rental
YEAR = 2018
OUTPUT_FORMAT = 'csv.gz'   # Format used for the generated data set
CLEAN_CITYNAME = 'Madrid' # City name: Madrid, Barcelona or Valencio

knitr::opts_chunk$set(echo = TRUE)
```
\normalsize

We set the used CRS, input parameters are coded using and EPSG:4326 and we need to use a projected CRS, in this case we will use "+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs".
\tiny
```{r Set process CRS}
#
# Set the projections used
#
projcrs_src = '+proj=longlat +datum=WGS84 +no_defs'
projcrs_dest = '+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs'
```
\normalsize

# Load data

Load input data (with actual coordinates and prices).
\tiny
```{r Load polygons, message=F, warning=F}
#
# Set input file names
#

file_name = paste0(CLEAN_CITYNAME,'_polygons.',OUTPUT_FORMAT)

#
# Load file and project to make the spatial transformations
#

polygons.sf <- read_delim(paste0(dir_geom_path, file_name), 
                          delim = ";") |>
  st_as_sf(wkt = 'WKT', 
           crs = projcrs_src) |>
  st_transform(projcrs_dest)

#
# Plot geometry
#

plot(st_geometry(polygons.sf))
```
\normalsize

# Load and preprocess listing data

## Load data

Load data and remove some extreme values (taken from idealista history data).
\tiny
```{r Data load, message=F, warning=F}
#
# Variable boundaries come are defined as valid values 
#
max_unitprice = list('sale' = 10443.64, 'rent' = 47.3)
min_price = list('sale' = 5000, 'rent' = 100)
max_price = list('sale' = 4E6, 'rent' = 20E3)
range_area = c(20, 500)

file_name = paste0(COUNTRY,'_',TYPOLOGY,
                   '_',OPERATION,'_',CLEAN_CITYNAME,
                   '_',YEAR,'.',OUTPUT_FORMAT)

# Load individual records
assets.sf <- read_delim(paste0(dir_path, file_name),  delim = ";")

```
\normalsize

## Fix cadastral construction year

Given we take as right the year user fills in we need to fix it, if we see extreme values < 1500 or in the future we assign the median year in the area:
\tiny
```{r Fix construction year, message=F, warning=F}
#
# Impute construction year if it is out of bounds ]-inf, MIN_YEAR] or [YEAR, inf]
#

MIN_YEAR = 1500
imputed_years.sf <- assets.sf |> 
  dplyr::group_by(LOCATIONID, .drop=F) |>
  dplyr::filter(CADCONSTRUCTIONYEAR > MIN_YEAR & CADCONSTRUCTIONYEAR <= YEAR) |>
  dplyr::summarise(IMPUTED_YEAR=as.integer(median(CADCONSTRUCTIONYEAR), .groups=NULL))

assets.sf <- assets.sf |>
  dplyr::inner_join(imputed_years.sf, by='LOCATIONID') |>
  # 1st attempt: Impute with idealista construction year
  dplyr::mutate(CADCONSTRUCTIONYEAR = ifelse(CADCONSTRUCTIONYEAR < MIN_YEAR 
                                          | CADCONSTRUCTIONYEAR > YEAR 
                                          | is.na(CADCONSTRUCTIONYEAR), 
                                          CONSTRUCTIONYEAR, CADCONSTRUCTIONYEAR)) |>  
  dplyr::mutate(-IMPUTED_YEAR)
```
\normalsize
Remove duplicate records, we only take as repeated instance an asset located in the same place with the same price
\tiny
```{r Remove duplicates, message=F, warning=F}
assets.sf <- assets.sf |>
  dplyr::group_by(PERIOD,LATITUDE, LONGITUDE, PRICE, .drop=T) |>
  dplyr::filter(row_number() == 1) |>
  dplyr::ungroup()
```
\normalsize

We turn our regular dataframe to a spatial features data frame, we project the source coordinates.
\tiny
```{r Create spatial features, message=F, warning=F}
#
# Createsf and reproject
#

assets.sf <- assets.sf |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = projcrs_src) |>
  st_transform(projcrs_dest)

#
# Save coordinates in a separate dataframe for later use
#

assets.coordinates = as_tibble(st_coordinates(assets.sf)) |>
  dplyr::mutate(IDX = row_number())

#
# Remove geometry (turn sf into ordinary a dataframe)
#
st_geometry(assets.sf) <-  NULL
```
\normalsize

# Price obfuscation

We will modify slighlty the prices, as price are almost-continuous in a two-step process:

* Randomize price with +/- 2.5%
* Adversisers usually round up figures to multiples of 10 (rent) or 1000 euros (sale), we align prices to multiples of 1000 (sales) and 10 (rent)
\tiny
```{r Transform prices, message=F, warning=F, echo=F}
#
# changes randomly the price with this amount
#
price_random_pct = 2.5

#
# Set the rounding grain
#
round_grain = list('sale' =  1000.0, 'rent'= 10)

#
# Perform the price adjustment
#
assets.sf <- assets.sf |>
  dplyr::mutate(VAR = runif(nrow(.),
                            -price_random_pct/100, price_random_pct/100.0),
                #
                # Add random disturbance and realign
                #
                PRICE = round(((1+VAR) * PRICE)/round_grain[[OPERATION]])*round_grain[[OPERATION]],
                PRICE = round(PRICE),
                NEW_UNITPRICE = PRICE / CONSTRUCTEDAREA) |>
  select(-VAR)
```
\normalsize

# Relocation of coordinates

We move the source coordinates with a radius between 30 and 60 meters:

* Take a random angle 0 - 360 degrees
* Scale the unitary radius to a value between 30 and 60 meters

The process makes sure listings are not moved our their neighborhood boundary (for this matter we use the zoning schema used on www.idealista.com). 
\tiny
```{r Set coordinates and locations for the relocation process, message=F, warning=F}
#
# Set distance parameters
#

MIN = 30
MAX = 60
DIFF = MAX - MIN

#
# Set location codes for each record: join between polygon areas and listings' cooordinates
#

coordinates.sf <- assets.coordinates |>
    st_as_sf(coords = c("X", "Y"), crs = projcrs_dest) |>
    st_join(polygons.sf, left = T)

#
# Remove spatial features from the dataframe
#

st_geometry(coordinates.sf) <- NULL

#
# It should not necessary if polygons do not overlap, just in case they do we take the first location 
#  the listing is located within
#

coordinates.sf <- coordinates.sf |>
  dplyr::group_by(IDX, .drop=T) |>
  dplyr::summarize(LOCATIONID = min(LOCATIONID), groups = NULL) |>
  dplyr::arrange(IDX)

#
# We assign the idealista's location code (called LOCATIONID)
#

assets.coordinates$LOCATIONID = coordinates.sf$LOCATIONID
```
\normalsize

We move the coordinates, bearing in mind we must keep the item in the same polygon that it was, if not the case try get new coordinates again (to a maximum of 10 fold, in practice it needed 2 iteration maximum). 
\tiny
```{r Regenerate coordinates, message=F, warning=F}
#
# Set a maximum reprocessing iteration
# a reprocess takes care of all items that are moved out from their original idealista polygons (LOCATIONID)
#
MAX_ITERATIONS = 10

#
# Init the process
#
pending.coordinates = assets.coordinates
new.coordinates = NULL
iteration = 1

cat(paste0('Start process objective ', nrow(pending.coordinates), ' rows\n'))

while(nrow(pending.coordinates) > 0 | iteration > MAX_ITERATIONS){

    cat(paste0('\tPerforming relocation iteration ', iteration, 
             ' pending rows: ',nrow(pending.coordinates),'\n' ))

  #
  # Relocate coordinates has two steps:
  #
  #    1) Get a random angle alpha
  #    2) Calculate a distance between d in [MIN, MAX]
  #    3) Set new coordinates at (X + cos(alpha) * d, Y + sin(alpha) * d) 
  #
  candidate = pending.coordinates |>
      dplyr::mutate(ANGLE_RADIANS = runif(nrow(.), 0, 360) * pi / 180) |>
      dplyr::mutate(DX = cos(ANGLE_RADIANS) * (MIN + DIFF * runif(nrow(.), 0, 1)),
                    DY = sin(ANGLE_RADIANS) * (MIN + DIFF * runif(nrow(.), 0, 1))) |>
      dplyr::mutate(NEW_X = round(X + DX),
                    NEW_Y = round(Y + DY)) |>
      dplyr::select(-ANGLE_RADIANS, -DX, -DY)

  #
  # Append the new location to the data frame`
  #
  candidates.sf = candidate |>
      st_as_sf(coords = c("NEW_X", "NEW_Y"), crs = projcrs_dest) |>
      st_join(polygons.sf) |>
      dplyr::rename(LOCATIONID = LOCATIONID.x,
                    NEW_LOCATIONID = LOCATIONID.y)

  # Take locationid use the min
  candidates.sf <- candidates.sf |>
    dplyr::group_by(IDX, LOCATIONID, .drop = T) |>
    dplyr::summarize(NEW_LOCATIONID = min(NEW_LOCATIONID), .groups = NULL) |>
    dplyr::arrange(IDX)

  candidate$NEW_LOCATIONID = candidates.sf$LOCATIONID

  # Add created end redo the pending
  pending.coordinates <- candidate |>
    dplyr::filter(NEW_LOCATIONID != LOCATIONID & !is.na(LOCATIONID))

  done.coordinates  <- candidate |>
    dplyr::filter((NEW_LOCATIONID == LOCATIONID) | is.na(LOCATIONID))

  if (is.null(new.coordinates)){
    new.coordinates = done.coordinates
  } else {
    new.coordinates <- new.coordinates |>
      dplyr::bind_rows(done.coordinates)
  }
  cat(paste0('\tPerformed relocation iteration ', iteration, 
             ' total rows: ',nrow(done.coordinates),'\n-------\n' ))
  iteration = iteration + 1

}
new.coordinates <- new.coordinates |>
  mutate(MOVED_DISTANCE = sqrt((NEW_X - X)^2 + (NEW_Y - Y)^2))
rm(pending.coordinates, done.coordinates, candidates.sf, candidate)
```
\normalsize

Assign the new coordinates to the listings.
\tiny
```{r Reproject coordinates, message=F, warning=F, echo=F}
assets.sf <- assets.sf |>
  dplyr::mutate(LONGITUDE = new.coordinates$NEW_X,
                LATITUDE = new.coordinates$NEW_Y) |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = projcrs_dest) |>
  st_transform(projcrs_src) 
```
\normalsize
\tiny
```{r Assign final new coordinates}
#
# Save coordinates
#
assets.coordinates = as_tibble(st_coordinates(assets.sf))

#
# Remove geometry
#
st_geometry(assets.sf) <-  NULL

#
# Reasign coordinates once geom was removed
#
assets.sf <- assets.sf |>
  dplyr::mutate(LONGITUDE = assets.coordinates$X,
                LATITUDE = assets.coordinates$Y)

```
\normalsize

# Save results

Create final dataframe and save results
\tiny
```{r Save results, message=F, warning=F}
#
# Save results
#
assets.sf |>
  fwrite(file = paste0(output_path, iconv(file_name,from="UTF-8",to="ASCII//TRANSLIT")), 
         sep = ";",
         compress = "gzip")
```
\normalsize

### 2022-08-02 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd
Adds comments in the response to reviewer

@@
 
-Idealista covers fairy well all segments in spanish market, individual and profesional adverstiser.
-Although listing prices reflects the reality of real estate markets from the offer persperctive, strong correlation of idealista and transaction prices can be established see [Banco de España](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3400031).
+Idealista covers fairly well all segments in the spanish market, individual and professional advertisers.
+Although listing prices reflects the reality of real estate markets from the offer perspective, strong correlation between idealista and transaction prices can be established see [Banco de España](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3400031).
+
+The main real estate portals in Spain are idealista and Fotocasa, further away are Habitaclia and Milanuncios, the latter focusing almost exclusively on private individuals.
+In September 2021, according to data
+from Similarweb, in Spain, there were a total of 103 million page views on real estate portals, out of page views on real estate portals, of which the four main portals, idealista, Fotocasa, Habitaclia, and Pisos.com, in that order, accounted for 94% of the traffic.
+
+Another relevant feature of this sector is that it is highly concentrated, idealista being the leader by far with 58.6 million visits (57% of the total traffic)  versus its immediate competitor with 19.9 million visits (19.3 % of total traffic).
+In terms of the evolution over time of interest in the content of each portal, based on data from Google Trends, the leadership of idealista Trends, the leadership of idealista has been shared from 2004 to 2015, subsequently consolidating idealista as the leader by a wide margin.
 
@@
 
-It should not affect much given it is random noise.
+We are aware that performing data obfuscation in our dataset has a trade-off between data privacy and data utility.
+The data anonymization process maintains data utility and data properties for most empirical research as we are adding a white noise perturbation to the original prices and coordinates.
+This dataset is expected to be used to compare different valuation algorithms and in this sense, the random disturbances in the data should not cause major concerns.
+
+However, for some applications researchers should proceed with care with this dataset.
+We would like to warn researchers that the spatial perturbation we introduced in the data might impact the computation of travel times between the properties' location and Point of Interest coordinates.
+In this case, travel times wouldn't be completely accurate although they are likely to be good enough for most accessibility measure computations.
 

### 2022-08-02 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd
Corrects bad grammar in the second question
Adds comments about registradores

@@
 
-Thanks for this comment. We have added discussed these concerns in the introduction.
+Thanks for this comment. We have added a discussion about these concerns in the introduction.
 
@@
 Idealista covers fairly well all segments in the spanish market, individual and professional advertisers.
+This dataset involves information about listing prices.
+Actual sale prices information is not publicly available as it is information that can only be accessed by paying high fees to Colegio de Registradores.
 Although listing prices reflects the reality of real estate markets from the offer perspective, strong correlation between idealista and transaction prices can be established see [Banco de España](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3400031).

### 2022-08-12 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd

--- 2022-08-02 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd
@@
 
-We impose a key constraint in the relocation process that is keeping the listing in the same neighborhood.
+The possible advese effects due to mislocation assets after coordinate randomisation is mitigated by the imposition of a boundary constraint in the relocation process.
+This constraint keeps the listing in the same neighborhood it belonged before relocation.
 For the process clarity sake we released the anonymization process in the R package code on github.
@@
 
-Idealista covers fairly well all segments in the spanish market, individual and professional advertisers.
-Although listing prices reflects the reality of real estate markets from the offer perspective, strong correlation between idealista and transaction prices can be established see [Banco de España](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3400031).
+Idealista covers fairly well all segments in the spanish market, both individual and professional advertisers.
+This dataset comprehends information about listing prices, thus it contains the market situation from the offer.
+
+Nevertheless, actual sale prices information is not publicly available as it is information that can only be accessed by paying high fees to Colegio de Registradores.
+However listing prices reflect quite well the (transaction) reality of real estate markets using the offer perspective, they keep strong correlation between idealista and transaction prices can be established see [Banco de España](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3400031).
+Even though a great extent of correlation does exist between official transactional and asking records, both databases can be taken as complementary and are of great interest when studying asking-transaction price gaps or the relation between listing site demand variables (ie.
+ad contacts or ad views) and price gaps.
 
 The main real estate portals in Spain are idealista and Fotocasa, further away are Habitaclia and Milanuncios, the latter focusing almost exclusively on private individuals.
-In September 2021, according to data
-from Similarweb, in Spain, there were a total of 103 million page views on real estate portals, out of page views on real estate portals, of which the four main portals, idealista, Fotocasa, Habitaclia, and Pisos.com, in that order, accounted for 94% of the traffic.
+In September 2021, according to data from Similarweb (a site specialized in sites' traffic volume comparison), in Spain, there were a total of 103 million page views on real estate portals, out of page views on real estate portals, of which the four main portals, idealista, Fotocasa, Habitaclia, and Pisos.com, in that order, accounted for 94% of the traffic.
 
-Another relevant feature of this sector is that it is highly concentrated, idealista being the leader by far with 58.6 million visits (57% of the total traffic)  versus its immediate competitor with 19.9 million visits (19.3 % of total traffic).
+Another relevant feature of this sector is that it is highly concentrated, idealista being the leader by far with 58.6 monthly million visits (57% of the total traffic)  versus its immediate competitor with 19.9 million visits (19.3 % of total traffic).
 In terms of the evolution over time of interest in the content of each portal, based on data from Google Trends, the leadership of idealista Trends, the leadership of idealista has been shared from 2004 to 2015, subsequently consolidating idealista as the leader by a wide margin.
@@
 In this case, travel times wouldn't be completely accurate although they are likely to be good enough for most accessibility measure computations.
+We must say in any case that we decided to take these perturbation since they are limited when it comes to walk mode travel times.
+As the average displacement length is around 30 meters, this distance would take an adult, on average, about 21-40 seconds at a normal pace  [Travel times walk](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0190753).
 

### 2022-08-12 README.Rmd

@@
 ```
+
+Due to data protection requirements, coordinates and prices have been anonymized by introducing as slight degree of random noise, for more details about this process please find the process code in the folder named "anonyzation".
+
 ## Neighboorhood polygons

### 2022-08-12 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd
Updated license in repository and website and discussed FAIR guiding principles in paper

@@
 
-Idealista covers fairly well all segments in the spanish market, both individual and professional advertisers.
+Idealista covers fairly well all segments in the Spanish market, both individual and professional advertisers.
 This dataset comprehends information about listing prices, thus it contains the market situation from the offer.
@@
 >5.	It would be useful for the authors to evaluate the datasets in the light of the FAIR guiding principles for research data stewardship (findability, accessibility, interoperability, and reusability).
+
+We discuss our efforts to comply with FAIR guiding principles in the data description of the revised paper.
 

### 2022-08-12 EPB/EPB.Rmd
Updated license in repository and website and discussed FAIR guiding principles in paper

--- 2022-02-18 EPB/EPB.Rmd
@@
 
-We are also witnessing a growing interest in open data in geography and data science [@arribasl2021editorial; @arribas2021]  using reproducible or replicable  research  [@paez2021open].
+We are also witnessing a growing interest in open data in geography and data science [@arribasl2021editorial; @arribas2021] using reproducible or replicable  research  [@paez2021open; @brunsdon2021opening].
 But  to  work  openly in  science, it  is  necessary to  have  free software and open data.
-While great efforts have been made to make free software available to researchers (e.g., R or Python), not much data is currently out in the open.
-In the particular case of the real estate market, to our knowledge, there are few open micro-data sets of housing markets available [@Song2021].
+While great efforts have been made to make free software available to researchers (e.g., R or Python), data regarding real estate markets tend to be treated as confidential, and there are still few open micro-data sets of housing markets available [@Song2021].
 
-To overcome these limitations, this paper presents a sort description of an open micro-data set of geo-referenced dwelling listings.
+To contribute to the growing inventory of international micro-data sets of real estate markets, this paper introduces an open micro-data set of geo-referenced dwelling listings.
 The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain’s three largest cities.
@@
 Moreover, the data set has been supplied directly by Idealista, and therefore is clean and free of download errors.
+However, in order to be able to share publicly the data we have slightly modified some variables by applying some random noise which will not bias the main results derived from its usage.
+
+This micro-data set is expected to be used as a benchmark dataset to test empirical methods in a reproducible fashion.
+Applied and theoretical researchers on real estate mass appraisal and valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others.
+The data can also be used to study the segmentation of housing submarkets and related topics such as the impact of suburban areas on house prices.
 The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest.
-The data set is distributed as an R package, named ‘idealista18’, which can be accessed from the Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
+In any case, the data might be easily extended by spatially joining other datasets that contain information at administrative, census tract, or postal code levels.
+
+The data set is distributed as an R package, name {idealista18}, which can be accessed from a public Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
+The open data product {idealista18} is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
+For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.
 
@@
 
-The open data set ‘idealista18’ is an R package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia.
+The open data product {idealista18} is an R package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia.
 For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the R package.
@@
 A full description of the data is available in the help section of the package.
+We have strived to the best of our ability to comply with FAIR principles regarding research data [@Wilkinson2016fair]: upon publication, the dataset has a persistent digital object identifier; publication as a data article makes the data findable; the data and metadata are packaged together and protocols for help files in the R ecosystem mean that documentation is easily searchable; distribution as an R package means that only open software is needed to access the data; and a public repository documents all data processes followed to generate the distributed open data product.
 

### 2022-08-12 EPB/main_EPB.Rmd
Updated license in repository and website and discussed FAIR guiding principles in paper

@@
 title: A geo-referenced micro-data set of real estate listings for Spain’s three largest cities
-runninghead: Rey-Blanco \emph{et al}.
+runninghead: author \emph{et al}.
 author:
@@
 Moreover, the data set has been supplied directly by Idealista, and therefore is clean and free of download errors.
+However, in order to be able to share publicly the data we have slightly modified some variables by applying some random noise which will not bias the main results derived from its usage.
+
+This micro-data set is expected to be used as a benchmark dataset to test empirical methods in a reproducible fashion.
+Applied and theoretical researchers on real estate mass appraisal and valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others.
+The data can also be used to study the segmentation of housing submarkets and related topics such as the impact of suburban areas on house prices.
 The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest.
-The data set is distributed as an R package, named ‘idealista18’, which can be accessed from the Github repository.
+In any case, the data might be easily extended by spatially joining other datasets that contain information at administrative, census tract, or postal code levels.
+
+The data set is distributed as an R package, name {idealista18}, which can be accessed from a public Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
+The database idealista18 is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
+For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.
+
+<!-- @cohen2008spatial (508 observations) -->
+<!-- @lazrak2014market (19000 observations) -->
 
@@
 The data product can also be used for educational proposes and teaching activities.
+
+# Declaration of Competing Interest
+
+Author One and author Two are employed by Idealista.
+They have been granted permission to share the data presented in this article.
+None of the authors have financial interests or personal relationships which have, or could be perceived to have, influenced the work reported in this article.
+
+# Acknowledgments
+
+The authors wish to thank Alessandro Galesi for their support in the paper revision and Juan Ramón Selva for collecting and cleaning the spatial data.
+This work has been partially funded by the Spanish Ministry of Economy and Competitiveness Grants PID2019-107800GB-100

### 2022-08-12 README.Rmd
Updated description and licensing information

@@
 <!-- badges: start -->
-[![License: CC BY-NC-SA 4.0](https://img.shields.io/badge/license-CC BY--NC--SA 4.0-blue.svg)](https://creativecommons.org/licenses/by-nc-sa/4.0/deed.en_GB)
+[![License: CC BY-NC-SA 4.0](https://img.shields.io/badge/license-ODbLv1.0-blue.svg)](https://opendatacommons.org/licenses/odbl/1-0/)
 

### 2022-08-12 R/idealista18.R

@@ -10,14 +10,14 @@
 #'
 #' @docType package
 #' @name idealista18
-#' @details Escribir algo
+#' @details "Open data product with real estate listings from Idealista. The datasets are for year 2018 and three major cities in Spain: Barcelona, Madrid, and Valencia."
 #'
 #' @author
 #'   \tabular{ll}{
 #'   David Rey \tab \email{drey@@idealista.com} \cr
 #'   Pelayo Arbues \tab \email{pgarbues@@idealista.com} \cr
 #'   Fernando López  \tab \email{fernando.lopez@@upct.es} \cr
-#'   Antonio Páez \tab \email{paezha@@gmail.com} \cr
+#'   Antonio Páez \tab \email{paezha@@mcmaster.ca} \cr
 #'   }
 #'
 #' @references \url{https://github.com/paezha/Reproducible-Research-Workflow}

### 2022-08-12 editorial-materials/EPB-2022-0116_response_to_reviewer.Rmd
Revisions complete for resubmission to EPB

@@
 ---
-title: "Using machine learning to identify spatial market segments: A reproducible study of major Spanish markets - Response to referees"
-bibliography: bibliography.bib
+title: "A geo-referenced micro-data set of real estate listings for Spain’s three largest cities - Response to referees"
 output: pdf_document
@@
 We have added some examples of research topics that can be addressed with this dataset in the introduction.
-
-# References

### 2022-08-12 EPB/main_EPB.tex
Revisions complete for resubmission to EPB

@@
 
-\runninghead{Rey-Blanco \emph{et al}.}
+\runninghead{author \emph{et al}.}
 
@@
 data science \citep{arribasl2021editorial, arribas2021} using
-reproducible or replicable research \citep{paez2021open}. But to work
-openly in science, it is necessary to have free software and open data.
-While great efforts have been made to make free software available to
-researchers (e.g., R or Python), not much data is currently out in the
-open. In the particular case of the real estate market, to our
-knowledge, there are few open micro-data sets of housing markets
-available \citep{Song2021}.
+reproducible or replicable research
+\citep{paez2021open, brunsdon2021opening}. But to work openly in
+science, it is necessary to have free software and open data. While
+great efforts have been made to make free software available to
+researchers (e.g., R or Python), data regarding real estate markets tend
+to be treated as confidential, and there are still few open micro-data
+sets of housing markets available \citep{Song2021}.
 
-To overcome these limitations, this paper presents a sort description of
-an open micro-data set of geo-referenced dwelling listings. The data
-have been provided by the Idealista
+To contribute to the growing inventory of international micro-data sets
+of real estate markets, this paper introduces an open micro-data set of
+geo-referenced dwelling listings. The data have been provided by the
+Idealista
 company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal}
@@
 the data set has been supplied directly by Idealista, and therefore is
-clean and free of download errors. The listings have been enriched with
-official information from the Spanish cadastre along with other relevant
-geographical features, such as distance to urban points of interest. The
-data set is distributed as an R package, named `idealista18', which can
-be accessed from the Github repository.
+clean and free of download errors. However, in order to be able to share
+publicly the data we have slightly modified some variables by applying
+some random noise which will not bias the main results derived from its
+usage.
+
+This micro-data set is expected to be used as a benchmark dataset to
+test empirical methods in a reproducible fashion. Applied and
+theoretical researchers on real estate mass appraisal and valuation
+methods might use this dataset to canonically compare the performance of
+their proposed comparable and hedonic models, among others. The data can
+also be used to study the segmentation of housing submarkets and related
+topics such as the impact of suburban areas on house prices. The
+listings have been enriched with official information from the Spanish
+cadastre along with other relevant geographical features, such as
+distance to urban points of interest. In any case, the data might be
+easily extended by spatially joining other datasets that contain
+information at administrative, census tract, or postal code levels.
+
+The data set is distributed as an R package, name \{idealista18\}, which
+can be accessed from a public Github
+repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
+The database idealista18 is made available under the
+\href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
+For transparency, we also share the randomisation process applied to the
+original data in the aforementioned Github repository.
 
@@
 
-The open data set `idealista18' is an R package composed of nine
+The open data product \{idealista18\} is an R package composed of nine
 objects, three objects for each of the three main Spanish cities:
@@
 object. A full description of the data is available in the help section
-of the package.
+of the package. We have strived to the best of our ability to comply
+with FAIR principles regarding research data \citep{Wilkinson2016fair}:
+upon publication, the dataset has a persistent digital object
+identifier; publication as a data article makes the data findable; the
+data and metadata are packaged together and protocols for help files in
+the R ecosystem mean that documentation is easily searchable;
+distribution as an R package means that only open software is needed to
+access the data; and a public repository documents all data processes
+followed to generate the distributed open data product.
 
@@
 
+\hypertarget{declaration-of-competing-interest}{%
+\section{Declaration of Competing
+Interest}\label{declaration-of-competing-interest}}
+
+Author One and author Two are employed by Idealista. They have been
+granted permission to share the data presented in this article. None of
+the authors have financial interests or personal relationships which
+have, or could be perceived to have, influenced the work reported in
+this article.
+
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
+
+The authors wish to thank Alessandro Galesi for their support in the
+paper revision and Juan Ramón Selva for collecting and cleaning the
+spatial data. This work has been partially funded by the Spanish
+Ministry of Economy and Competitiveness Grants PID2019-107800GB-100
+
 \bibliographystyle{sageh}

### 2022-08-12 README.Rmd
Revisions complete for resubmission to EPB

@@
 <!-- badges: start -->
-[![License: CC BY-NC-SA 4.0](https://img.shields.io/badge/license-ODbLv1.0-blue.svg)](https://opendatacommons.org/licenses/odbl/1-0/)
+[![License: ODbLv1.0](https://img.shields.io/badge/license-ODbLv1.0-blue.svg)](https://opendatacommons.org/licenses/odbl/1-0/)
 

### 2023-05-16 editorial-materials/EPB-2022-0116.R1_response_to_reviewer.Rmd
Edits in preparation for resubmission and initializing renv

---
title: "A geo-referenced micro-data set of real estate listings for Spain’s three largest cities - Response to referees"
output: pdf_document
bibliography: bibEPB.bib
---

<!--
```{r install-data-package, include = FALSE}
# Run only once if needed to install data package `sobiEquity`
if (!require("idealista18", character.only = TRUE)) {
      devtools::install_github("https://github.com/paezha/idealista18")
  }
```

```{r setup,  cache=FALSE, warning = FALSE, include=FALSE, message = FALSE}
library(idealista18)
library(tidyverse)

set.seed(1234)
```
-->

Referee: 1

> Comments to the Author

> Dear Authors, Thank you for the opportunity to review this paper. I would appreciate it if you would highlight the changed text in the next submission. Otherwise it is difficult to identify which parts have changed. The unknown quality of this one year randomised data is the main issue for potential users and I repeat my suggestion that this data should be validated for the benefit potential readers and users. Below is a list of comments and questions which  need to be addressed.

> 1.Please clearly state the nature of the data in the abstract. It is worthwhile to mention the relocated and randomised dataset in the abstract. This house price dataset only has 189,923 records, which could be not really be described as a big dataset as you describe it in the abstract. Except the authors could provide an evidence/reference on why this is big dataset.

**Response:**

We appreciate your comment. There are a few publicly available datasets for hedonic price analysis, but they tend to be smaller. For example, @bonifaci2015real includes $n=1,042$ observations for Padua, in Italy; @delgiudice2018housing share a dataset with $n=576$ observations relating to rental prices in Naples, Italy; @solano2019dataset present a dataset with $n=1,623$ daily rental prices in Seville, Spain; and the dataset of @fuerst2020real includes $n=4,201$ property prices geocoded to the level of nine regions in England and Wales. The most similar in terms of geographical disaggregation and sample size is the dataset of @song2021hedonic which includes transactions for four cities in South Korea, namely Busan ($n=61,152$), Daegu ($n=32,363$), Daejeon ($n=21,114$), and Gwangju ($n=25,984$). In response to your comment we drop the terminology of "big data" and describe our dataset as "large" instead.

>2. Please check the typographical errors before your next submission. For example, the “thirdr” in your table1.

**Response:**

Thank you for highlighting this. We have carefully proof-read the document for typos and other errors.

>3. Please use part of your answer to my previous fourth question in the manuscript. Your answer is quite useful for your future users, especially those researchers who are not familiar with house price data in Spain. I quote your previous answer below:

> “idealista covers fairly well all segments in the Spanish market, both individual and professional advertisers. This dataset comprehends information about listing prices, thus it contains the market situation from the offer. Nevertheless, actual sale prices information is not publicly available as it is information that can only be accessed by paying high fees to Colegio de Registradores. However listing prices reflect quite well the (transaction) reality of real estate markets using the offer perspective, they keep strong correlation between idealista and transaction prices can be established see Banco de España. Even though a great extent of correlation does exist between official transactional and asking records, both databases can be taken as complementary and are of great interest when studying asking-transaction price gaps or the relation between listing site demand variables (ie. ad contacts or ad views) and price gaps. The main real estate portals in Spain are idealista and Fotocasa, further away are Habitaclia and Milanuncios, the latter focusing almost exclusively on private individuals. In September 2021, according to data from Similarweb (a site specialized in sites’ traffic volume comparison), in Spain, there were a total of 103 million page views on real estate portals, out of page views on real estate portals, of which the four main portals, idealista, Fotocasa, Habitaclia, and Pisos.com, in that order, accounted for 94% of the traffic. Another relevant feature of this sector is that it is highly concentrated, idealista being the leader by far with 58.6 monthly million visits (57% of the total traffic) versus its immediate competitor with 19.9 million visits (19.3 % of total traffic). In terms of the evolution over time of interest in the content of each portal, based on data from Google Trends, the leadership of idealista Trends, the leadership of idealist”

**Response:**

Thank you for this suggestion. We have adapted this answer to expand the text in the section "Data Description".

>4.For data description, it is worth showing solid evidence of the quality of this dataset. For example, what is the data coverage of this data compared to all the residential properties in the city?  Listing the advertised volume is not very helpful for users who wish to understand the quality.

Thank you. We appreciate your comment. To provide better context about the coverage of the {idealista18} dataset Table \ref{tab:transactions} now shows the number of listings with respect to the total residential stock in each city (in 2018). Globally, the number of listings to the total housing stock is around 7%. Additional information about the number of free-market home purchases can be obtained from the Spanish National Institute. Globally, the number of listingss 2,7 times the number of transactions.

To provide some context about the coverage of the {idealista18} dataset Table \ref{tab:transactions} shows the number of listings with respect to the total residential stock in each city in 2018. As seen in the table, the number of listings ranges between 6.1% of the total number of properties (in Madrid) and 8.1% (in Valencia). Information from Instituto Nacional de Estadistica^[\url{https://www.ine.es/jaxiT3/Tabla.htm?t=6150&L=1}] shows that the number of listings in the {idealista18} package correspond to 81.3\% of recorded real estate transactions in Barcelona, 80.8\% in Madrid, and 91.1\% in Valencia.

```{=tex}
\begin{table}[!ht]
    \centering
      \caption{Total properties and transactionst three Spanish cities. Year 2018} 
  \label{tab:transactions} 
    \begin{tabular}{lccccc}
        City & Total properties (P) & Total transactions (T) & Listings (L) & L/P & T/L \\
        \hline
        Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
        Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
        Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
        \hline
        Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
        \hline
        \multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
        \multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
        \multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
    \end{tabular}
\end{table}
```

>5. Given the asking price is a resized version of the original asking price with a random percentage between -2.5% and +2.5%, please elaborate on how this influences your quantitative research results and findings. Please provide critical evidence to explain how this randomised price will not influence academic research. For example, how does the house price variation at local level change after the price is randomised. To what degree will this influence a typical hedonic house price research approach? I am concerned about whether this randomised data could bias conclusions.

**Response:**

We appreciate the opportunity to explain this issue, as it may be of interest to academic researchers. Adding a small amount of random noise ($\pm 2.5$) to the prices has the effect of increasing the variance of estimates, but does not affect the mean. This is a well-known result in algebra of random variables.

Prices ($P$) in the dataset are calculated as the sum of two random variables, the raw price ($RP$) and some random noise ($\eta$) drawn from a uniform distribution as follows:
$$

$$

The mean of the raw price is:
$$
\mathrm{E}[RP] = \frac{1}{N}\sum_{i=1}^N R_i
$$

The mean of the white noise is zero by design:
$$
\mathrm{E}[\eta] = 0
$$

We know that the expectation of the sum of two random variables is:
$$
\mathrm{E}[P] = \mathrm{E}[RP + \eta] = \mathrm{E}[RP] + \mathrm{E}[\eta]
$$

\noindent and since $\mathrm{E}[\eta] = 0$ it follows that the prices are unbiased.

The variance of the sum of two random variables is:
$$
\mathrm{V}[P] = \mathrm{V}[RP + \eta] = \mathrm{V}[RP] + 2\mathrm{Cov}[RP,\eta] + \mathrm{V}[\eta]
$$

The variance of $\eta$ is:
$$
\mathrm{V}[\eta]=\frac{1}{12}\big(0.025 - (-0.25)\big)^2
$$

>6. In figure 1, is it possible to standardise this data rather than showing only the number of dwelling records in your house price data? I guess the city centre always has high advertised volumes but it also has a high density of dwellings. It is worthwhile to show values standardised by the number of dwellings.

**Response:** After this comment we included a new figure to display the percentage of listings relative to the total number of dwellings by district.

```{r, echo = FALSE,message=FALSE, fig.width=5, fig.cap="\\label{fig:districts}Percentage of listings relative to total number of dwellings by district. Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
library(idealista18)
library(dplyr)
library(ggplot2)
library(sf)
library(gridExtra)

data("properties_by_district")

# Barcelona
Barcelona_prop <- properties_by_district |>
  filter(CITY == "Barcelona") |>
  mutate(proportion = N/N_CADASTRE)

q <- quantile(Barcelona_prop$proportion)
pq <- round(q * 100, 2)

Barcelona_prop <- Barcelona_prop |>
  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))

plot4 <- ggplot(data = Barcelona_prop) +
  geom_sf(aes(fill = Quantile),
          color = "black",
          size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

# Madrid
Madrid_prop <- properties_by_district |>
  filter(CITY == "Madrid") |>
  mutate(proportion = N/N_CADASTRE)

q <- quantile(Madrid_prop$proportion)
pq <- round(q * 100, 2)

Madrid_prop <- Madrid_prop |>
  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))

plot5 <- ggplot(data = Madrid_prop) +
  geom_sf(aes(fill = Quantile),
          color = "black",
          size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

# Valencia
Valencia_prop <- properties_by_district |>
  filter(CITY == "Valencia") |>
  mutate(proportion = N/N_CADASTRE)

q <- quantile(Valencia_prop$proportion)
pq <- round(q * 100, 2)

Valencia_prop <- Valencia_prop |>
  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))

plot6 <- ggplot(data = Valencia_prop) +
  geom_sf(aes(fill = Quantile),
          color = "black",
          size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

# Arrange plots
grid.arrange(plot4, plot5, plot6, 
             ncol = 1, 
             nrow = 3)
```

---

*Dear Authors,* *Thank you for the opportunity to review this paper. I would appreciate it if you would highlight the changed text in the next submission. Otherwise it is difficult to identify which parts have changed. The unknown quality of this one year randomised data is the main issue for potential users and I repeat my suggestion that this data should be validated for the benefit potential readers and users. Below is a list of comments and questions which need to be addressed.*

> **Answer:** Thnak you.
> In the new version of the manuscrip, the changes are writen in red.

## First Comment

>Please clearly state the nature of the data in the abstract. It is worthwhile to mention the relocated and randomised dataset in the abstract. This house price dataset only has 189,923 records, which could be not really be described as a big dataset as you describe it in the abstract. Except the authors could provide an evidence/reference on why this is big dataset.

## Second Comment

2.  *Please check the typographical errors before your next submission. For example, the "thirdr" in your table1.*

> **Answer:** Thank you.
> A native English speaker has reviewed the document.

## Third Comment

3.  *Please use part of your answer to my previous fourth question in the manuscript. Your answer is quite useful for your future users, especially those researchers who are not familiar with house price data in Spain. I quote your previous answer below:*

*"dealista covers fairly well all segments in the Spanish market, both individual and professional advertisers. This dataset comprehends information about listing prices, thus it contains the market situation from the offer. Nevertheless, actual sale prices information is not publicly available as it is information that can only be accessed by paying high fees to Colegio de Registradores. However listing prices reflect quite well the (transaction) reality of real estate markets using the offer perspective, they keep strong correlation between idealista and transaction prices can be established see Banco de España. Even though a great extent of correlation does exist between official transactional and asking records, both databases can be taken as complementary and are of great interest when studying asking-transaction price gaps or the relation between listing site demand variables (ie. ad contacts or ad views) and price gaps. The main real estate portals in Spain are idealista and Fotocasa, further away are Habitaclia and Milanuncios, the latter focusing almost exclusively on private individuals. In September 2021, according to data from Similarweb (a site specialized in sites' traffic volume comparison), in Spain, there were a total of 103 million page views on real estate portals, out of page views on real estate portals, of which the four main portals, idealista, Fotocasa, Habitaclia, and Pisos.com, in that order, accounted for 94% of the traffic. Another relevant feature of this sector is that it is highly concentrated, idealista being the leader by far with 58.6 monthly million visits (57% of the total traffic) versus its immediate competitor with 19.9 million visits (19.3 % of total traffic). In terms of the evolution over time of interest in the content of each portal, based on data from Google Trends, the leadership of idealista Trends, the leadership of idealist"*

> **Answer:** Thank you. We include it.

## Fourth Comment

4.*For data description, it is worth showing solid evidence of the quality of this dataset. For example, what is the data coverage of this data compared to all the residential properties in the city? Listing the advertised volume is not very helpful for users who wish to understand the quality.*

> **Answer:** 
## Fifth Comment

5.  *Given the asking price is a resized version of the original asking price with a random percentage between -2.5% and +2.5%, please elaborate on how this influences your quantitative research results and findings.* **Please provide critical evidence to explain how this randomised price will not influence academic research.** *For example, how does the house price variation at local level change after the price is randomised. To what degree will this influence a typical hedonic house price research approach? I am concerned about whether this randomised data could bias conclusions.*

> **Answer:** Of course, the results of any econometric model will be differents if the original data are used.
> Unfortunally, the data protection laws in Spain, specifically the Organic Law on the Protection of Personal Data and the Guarantee of Digital Rights (LOPDGDD) and the General Data Protection Regulation (GDPR) of the European Union, establish that personal data, including information on the location of a home, is protected and must be treated confidentially.
> It is not permitted to disclose the specific location of a home without the explicit consent of the owner or person who has the right to dispose of it.

> The small changes in the prices will be, in mean, a small impact on the results.

## Sixth Comment

6.  *In figure 1, is it possible to standardise this data rather than showing only the number of dwelling records in your house price data? I guess the city centre always has high advertised volumes but it also has a high density of dwellings. It is worthwhile to show values standardised by the number of dwellings*

```{r, eval = FALSE, include = FALSE, message=FALSE, fig.width=5, }
library(readxl)
library(idealista18)
library(dplyr)
library(ggplot2)
library(sf)
library("gridExtra")
Data <- read_excel("Adds_by_Distrits_idealista18.xlsx")
Data$Por <- Data$N*100/Data$N_CADASTRE
Barcelona_Polygons$District <- as.numeric(substr(Barcelona_Polygons$LOCATIONID,23,24))
Barcelona_District <- Barcelona_Polygons %>% group_by(District) %>% summarise(District = unique(District))
Madrid_Polygons$District <- as.numeric(substr(Madrid_Polygons$LOCATIONID,23,24))
Madrid_District <- Madrid_Polygons %>% group_by(District) %>% summarise(District = unique(District))
Valencia_Polygons$District <- as.numeric(substr(Valencia_Polygons$LOCATIONID,23,24))
Valencia_District <- Valencia_Polygons %>% group_by(District) %>% summarise(District = unique(District))

Data.bar <- Data[Data$CITY=="Barcelona",]
Data.mad <- Data[Data$CITY=="Madrid",]
Data.val <- Data[Data$CITY=="Valencia",]

Barcelona_District <- merge(Barcelona_District,Data.bar,by="District")
q <- quantile(Barcelona_District$Por)
Barcelona_District$Quantile<- as.factor((Barcelona_District$Por > q[2]) + (Barcelona_District$Por > q[3]) +(Barcelona_District$Por >= q[4]) + 1)
levels(Barcelona_District$Quantile) <- c("(4.7%,5.7%]","(5.7,6.9]","(6.9,8.2]","(8.2,14.9]")
plot.d1 <- ggplot(data = Barcelona_District) +
  geom_sf(aes(fill = Quantile),size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_discrete("") +
    scale_fill_manual(name = "",values=c("#D4E7F1","#6DAFD3","#2470B2","darkblue")) +
  theme_bw(base_size = 6) +
  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
        legend.key.size = unit(.2, 'cm'),text = element_text(size=4))

Madrid_District <- merge(Madrid_District,Data.mad,by="District")
q <- quantile(Madrid_District$Por)
Madrid_District$Quantile<- as.factor((Madrid_District$Por > q[2]) + (Madrid_District$Por > q[3]) +(Madrid_District$Por >= q[4]) + 1)
plot.d2 <- ggplot(data = Madrid_District) +
  geom_sf(aes(fill = Quantile),size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(name = "",values=c("#D4E7F1","#6DAFD3","#2470B2","darkblue")) +
  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
        legend.key.size = unit(.2, 'cm'),text = element_text(size=4))

Valencia_District <- merge(Valencia_District,Data.val,by="District")
q <- quantile(Valencia_District$Por)
#levels(B$Quartile) <- c("Q1: (1200,4050]","Q2: (4050,5450]","Q3: (5450,8290]","Q4: (8290,830000]")
Valencia_District$Quantile<- as.factor((Valencia_District$Por > q[2]) + (Valencia_District$Por > q[3]) +(Valencia_District$Por >= q[4]) + 1)
plot.d3 <- ggplot(data = Valencia_District) +
  geom_sf(aes(fill = Quantile),size=.2) +
  theme_bw(base_size = 6) +
  ylim(39.405,39.52) +
  scale_fill_manual(name = "",values=c("#D4E7F1","#6DAFD3","#2470B2","darkblue"))+
  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
        legend.key.size = unit(.2, 'cm'),text = element_text(size=4))
grid.arrange(plot.d1, plot.d2, plot.d3, ncol = 3, nrow = 1)
```

### 2023-05-16 EPB/main_EPB.Rmd
Edits in preparation for resubmission and initializing renv

@@
 ---
-title: A geo-referenced micro-data set of real estate listings for Spain’s three largest cities
+title: A geo-referenced micro-data set of real estate listings for Spain's three largest cities
 runninghead: author \emph{et al}.
@@
 - name: 
-abstract: "This article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain.
+abstract: "This article presents an open data product with large geo-referenced micro-data sets of 2018 real estate listings in Spain.
 These data were originally published on the idealista.com real estate website.
@@
 
-keywords: Housing market; idealista.com; geo-referenced data; point-level data; open data; Spain
+keywords: Housing prices; hedonic price analysis; idealista.com; geo-referenced data; point-level data; open data; Spain
 classoption:
@@
 
+```{r load-packages, include=FALSE}
+library(dplyr)
+library(ggplot2)
+library(glue)
+library(gridExtra)
+library(idealista18)
+library(kableExtra)
+library(xtable)
+library(sf)
+```
+
+```{r load-data, include=FALSE}
+# Listings
+data("Barcelona_Sale")
+data("Madrid_Sale")
+data("Valencia_Sale")
+
+# Polygons
+data("Barcelona_Polygons")
+data("Madrid_Polygons")
+data("Valencia_Polygons")
+
+# Points of interest
+data("Barcelona_POIS")
+data("Madrid_POIS")
+data("Valencia_POIS")
+
+# Districts
+data("properties_by_district")
+```
+
 # Introduction
@@
 Interest in the characteristics of the housing market and housing prices has been a growing area of research in recent decades, generating a vast amount of theoretical and empirical literature.
-Including the spatial component to analyze the real estate market and incorporating geographic variables has significantly improved the understanding of this market.
-But to really understand the characteristics of the housing market, it is essential to have information/data at the point level.
-Therefore, it is becoming common for spatial analysis of urban environments to be developed with geo- referenced micro-data sets [@lopez2015].
-However, the availability of this type of open data at the point level is limited, and not many data sets contain latitude/longitude coordinates for each dwelling.
-In some cases, researchers have had to resort to web scraping processes to obtain the large volumes of information that permit robust analyses [@gupta2022take; @arbia2020spatial; @Li2019; @lopez2015].
-These  web  scraping  processes  can  include  missing  data, download errors, duplicate records, etc.
-Furthermore, the authors of this research do not generally share the data sets.
+Including the spatial component to analyze the real estate market and incorporating geographic variables has significantly improved the understanding of this market: to this end, it is essential to have information/data at the point level.
+Therefore, it is becoming common for spatial analysis of urban environments to be use micro-data, geo-referenced as points [@lopez2015].
+In some cases, researchers have had to resort to webscraping processes to obtain data for research [e.g., @gupta2022take; @arbia2020spatial; @Li2019; @lopez2015].
+Alas, webscraping is a chancy process prone to download errors, missing  data, duplicated records, etc.
+Furthermore, researchers do not always share their webscrapped datasets, which limits reproducibility of their research.
+As we witness a growing interest in openness and reproducibility in geographical data science [@arribasl2021editorial; @paez2021open; @brunsdon2021opening], it  becomes increasingly urgent to have open data products to support research [@arribas2021].
 
-We are also witnessing a growing interest in open data in geography and data science [@arribasl2021editorial; @arribas2021] using reproducible or replicable  research  [@paez2021open; @brunsdon2021opening].
-But  to  work  openly in  science, it  is  necessary to  have  free software and open data.
-While great efforts have been made to make free software available to researchers (e.g., R or Python), data regarding real estate markets tend to be treated as confidential, and there are still few open micro-data sets of housing markets available [@Song2021].
+Some researchers have already responded to this need for publicly available, geo-referenced datasets to support open, reproducible research.
+A few such datasets are now available to support the analysis of real estate markets, but they are sometimes geo-referenced at the level of large geographical zones, such as @fuerst2020real, an open dataset that includes $n=4,201$ property prices geocoded to the level of nine regions in England and Wales.
+Other datasets are geo-coded as points, including @bonifaci2015real, which includes $n=1,042$ observations for Padua, in Italy; @delgiudice2018housing, who share a dataset with $n=576$ observations relating to rental prices in Naples, Italy; and @solano2019dataset present a dataset with $n=1,623$ daily rental prices in Seville, Spain.
 
 To contribute to the growing inventory of international micro-data sets of real estate markets, this paper introduces an open micro-data set of geo-referenced dwelling listings.
-The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain’s three largest cities.
-To date, this data product is the biggest open geo-referenced micro-data set of the housing market in Spain.
-Moreover, the data set has been supplied directly by Idealista, and therefore is clean and free of download errors.
+The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain's three largest cities.
+To date, this data product is one of the largest open geo-referenced micro-data set of the housing market in the world.
+The most similar in terms of geographical disaggregation and sample size is the dataset of @song2021hedonic which includes transactions for four cities in South Korea, namely Busan ($n=61,152$), Daegu ($n=32,363$), Daejeon ($n=21,114$), and Gwangju ($n=25,984$).
+It is certainly the largest in Spain.
+
+The data set has been supplied directly by Idealista, and therefore is clean and free of download errors.
 However, in order to be able to share publicly the data we have slightly modified some variables by applying some random noise which will not bias the main results derived from its usage.
 
-This micro-data set is expected to be used as a benchmark dataset to test empirical methods in a reproducible fashion.
+This micro-data set can be used to benchmark new analytical methods in a reproducible fashion.
 Applied and theoretical researchers on real estate mass appraisal and valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others.
@@
 
-The data set is distributed as an R package, name {idealista18}, which can be accessed from a public Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
-The database idealista18 is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
+The data set is distributed as an `R` package, named {idealista18}, which can be accessed from a public Github repository\footnote{\url{https://github.com/paezha/idealista18}}.
+The open data product is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
 For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.
@@
 
-The open data product {idealista18} is an R package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia.
-For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the R package.
+The open data product {idealista18} is an `R` package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia.
+For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the package.
+
+The data provider (idealista.com) is a leading real estate portal in Spain, on par with Fotocasa.
+Smaller portals include Habitaclia and Milanuncios, the latter focusing almost exclusively on individual (i.e., non-professional) advertisers.
+In September 2021, according to data from Similarweb (a site specialized in sites' traffic volume comparison), there were a total of 103 million page views on real estate portals in Spain.
+Of this, 94% of traffic was concentrated on four portals, of which idealista was the most important, followed by Fotocasa, Habitaclia, and Pisos.com.
+Traffic is also highly concentrated, and idealista.com was the the leader by far with 58.6 monthly million visits (57% of the total traffic) compared to its closest competitor with 19.9 million visits (19.3 % of total traffic).
+
+As a result of its share of advertisments, idealista.com covers fairly well all segments of the Spanish market, including both individual and professional advertisers.
+This dataset includes information about listing prices, and therefore represents the market situation from the perspective of asking prices.
+This is a necessary compromise in the present case, since actual sale prices are not publicly available, and the information can only be accessed by paying high fees to Colegio de Registradores.
+However listing prices reflect quite well the (transaction) reality of real estate markets, and correlations between idealista listings and transaction prices can be established see [Banco de España](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3400031).
+Official transactions and asking prices can be taken as complementary and are of great interest when studying asking-transaction price gaps or the relation between listing site demand variables (i.e., ad contacts or ad views) and price gaps.
+
+To provide some context about the coverage of the {idealista18} dataset Table \ref{tab:transactions} shows the number of listings with respect to the total residential stock in each city in 2018.
+As seen in the table, the number of listings ranges between 6.1% of the total number of properties (in Madrid) and 8.1% (in Valencia).
+Information from Instituto Nacional de Estadistica^[\url{https://www.ine.es/jaxiT3/Tabla.htm?t=6150&L=1}] shows that the number of listings in the {idealista18} package correspond to 81.3\% of recorded real estate transactions in Barcelona, 80.8\% in Madrid, and 91.1\% in Valencia.
+
+```{=tex}
+\begin{table}[!ht]
+    \centering
+      \caption{Total properties and transactionst three Spanish cities. Year 2018} 
+  \label{tab:transactions} 
+    \begin{tabular}{lccccc}
+        City & Total properties (P) & Total transactions (T) & Listings (L) & L/P & T/L \\
+        \hline
+        Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
+        Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
+        Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
+        \hline
+        Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
+        \hline
+        \multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
+        \multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
+        \multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
+    \end{tabular}
+\end{table}
+```
+
 The following subsections describe each object.
 A full description of the data is available in the help section of the package.
-We have strived to the best of our ability to comply with FAIR principles regarding research data [@Wilkinson2016fair]: upon publication, the dataset has a persistent digital object identifier; publication as a data article makes the data findable; the data and metadata are packaged together and protocols for help files in the R ecosystem mean that documentation is easily searchable; distribution as an R package means that only open software is needed to access the data; and a public repository documents all data processes followed to generate the distributed open data product.
+We have tried to the best of our ability to comply with FAIR principles regarding research data [@Wilkinson2016fair]: upon publication, the dataset has a persistent digital object identifier; publication as a data article makes the data findable; the data and metadata are packaged together and protocols for help files in the R ecosystem mean that documentation is easily searchable; distribution as an R package means that only open software is needed to access the data; and a public repository documents all data processes followed to generate the distributed open data product.
 
@@
 The dwelling listing of each city includes a set of characteristics for each dwelling published on the idealista real estate website as an ad.
-The dwelling listing has been included in the ‘idealista18’ package as an sf object [@Pebesma].
-The name of the sf object containing the dwelling listing includes the name of the city, followed by '_Sale' (e.g., Madrid_Sale) and includes a total of 42 variables.
-Each sf object includes the complete set of listings corresponding to the four quarters of the year 2018.
+The dwelling listing has been included in the {idealista18} package as a simple features (sf) object [@Pebesma] with point geometry in latitude and longitude.
+The name of each sf object with the list of dwellings is the name of the city followed by '_Sale' (e.g., Madrid_Sale).
+Each data object includes a total of 42 variables and the complete set of listings corresponding to the four quarters of 2018.
 Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter.
@@
 ```{r number-ads, echo = FALSE, message = FALSE}
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
 H <- t(cbind(table(Barcelona_Sale$PERIOD),table(Madrid_Sale$PERIOD),table(Valencia_Sale$PERIOD)))
@@
 check.names = FALSE)
-colnames(text_tbl) <- c("City\\Quarter", "First", "Second ", "Thirdr", "Fourth","Total ads")
-xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") %>%
-xtable2kable(include.rownames = FALSE) %>%
-kable_styling(full_width = F) %>%
-column_spec(1, width = "4em", italic = FALSE) %>%
-column_spec(2, width = "3em") %>%
-column_spec(3, width = "3em", latex_column_spec = "c") %>%
-column_spec(4, width = "3em", latex_column_spec = "c") %>%
-column_spec(5, width = "3em", latex_column_spec = "c") %>%
+colnames(text_tbl) <- c("City\\Quarter", "First", "Second ", "Third", "Fourth","Total ads")
+xtable::xtable(text_tbl, digits=0, caption = "Number of dwelling  listing ads for each city and quarter. \\label{tab:number-ads}") |>
+xtable2kable(include.rownames = FALSE) |>
+kable_styling(full_width = F) |>
+column_spec(1, width = "4em", italic = FALSE) |>
+column_spec(2, width = "3em") |>
+column_spec(3, width = "3em", latex_column_spec = "c") |>
+column_spec(4, width = "3em", latex_column_spec = "c") |>
+column_spec(5, width = "3em", latex_column_spec = "c") |>
 column_spec(6, width = "2em", latex_column_spec = "c")
@@
 
-Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including the exact location of the dwelling (see Section [Anonymizing the data set](#anonymizing)).
+Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including the location of the dwelling.
+The exact location has been masked, as described in Section [Anonymizing the data set](#anonymizing)).
 Table \ref{tab:variables} lists the main indoor variables with a short description and the mean value of each variable.
@@
 
-Additionally, the distance of each dwelling to three urban points of interest was included in the sf object: distance to the city center, distance to the closest metro station, and distance to a major street (La Diagonal for Barcelona, La Castellana for Madrid, and Blasco Ibañez for Valencia).
+Additionally, the distance of each dwelling to several points of interest was included in the sf object: distance to the city center, distance to the closest metro station, and distance to a major street (La Diagonal for Barcelona, La Castellana for Madrid, and Blasco Ibañez for Valencia).
 The last rows of Table \ref{tab:variables} show the mean values of these variables.
@@
 Some variables have been excluded from this table to save space.
-Check the full list in the \\textbf{idealista18} package.\\label{tab:variables}") %>%
-xtable2kable(include.rownames = FALSE) %>%
-kable_styling(full_width = F, font_size=8) %>%
-column_spec(1, width = "13em", italic = FALSE) %>%
-column_spec(2, width = "14em") %>%
-column_spec(3, width = "2em", latex_column_spec = "c") %>%
-column_spec(4, width = "2em", latex_column_spec = "c") %>%
+Check the full list in the \\textbf{idealista18} package.\\label{tab:variables}") |>
+xtable2kable(include.rownames = FALSE) |>
+kable_styling(full_width = F, font_size=8) |>
+column_spec(1, width = "13em", italic = FALSE) |>
+column_spec(2, width = "14em") |>
+column_spec(3, width = "2em", latex_column_spec = "c") |>
+column_spec(4, width = "2em", latex_column_spec = "c") |>
 column_spec(5, width = "2em", latex_column_spec = "c")
@@
 See the help section in the \\textbf{idealista18} R package for details and formal definitions.
-Some dummy variables have been excluded from this table to save space\\label{tab:Dummy-variables}") %>%
-xtable2kable(include.rownames = FALSE) %>%
-kable_styling(full_width = F, font_size = 8) %>%
-column_spec(1, width = "12em", italic = FALSE) %>%
-column_spec(2, width = "14em") %>%
-column_spec(3, width = "2em", latex_column_spec = "c") %>%
-column_spec(4, width = "2em", latex_column_spec = "c") %>%
+Some dummy variables have been excluded from this table to save space\\label{tab:Dummy-variables}") |>
+xtable2kable(include.rownames = FALSE) |>
+kable_styling(full_width = F, font_size = 8) |>
+column_spec(1, width = "12em", italic = FALSE) |>
+column_spec(2, width = "14em") |>
+column_spec(3, width = "2em", latex_column_spec = "c") |>
+column_spec(4, width = "2em", latex_column_spec = "c") |>
 column_spec(5, width = "2em", latex_column_spec = "c")
@@
 
-The second block of data included in the ‘idealista18’ R package is the spatial features of the three cities divided into neighborhoods.
+The second data object included in {idealista18} is the neighborhoods in the three cities as polygons.
 There is an sf object for each city with the name of the city and the suffix '_Polygons'.
@@
 
-```{r, echo = FALSE,message=FALSE,fig.width=5, fig.cap="\\label{fig:all-polygons}Quantile maps of the number of dwellings in each neighborhood.
-Boundary for Barcelona (Left), Madrid (Center), and Valencia (Right)."}
-library("ggplot2")
-library("dplyr")
-library("gridExtra")
+```{r, echo = FALSE,message=FALSE, fig.width=5, fig.cap="\\label{fig:all-polygons}Number of listings in each neighborhood. Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
+
 # Barcelona
-Barcelona_counts <- Barcelona_Polygons %>% 
-  mutate(counts = lengths(st_intersects(., Barcelona_Sale)))
+Barcelona_counts <- Barcelona_Polygons |> 
+  mutate(counts = lengths(st_intersects(Barcelona_Polygons, Barcelona_Sale)))
 q <- quantile(Barcelona_counts$counts)
-Barcelona_Polygons$Quantile<- as.factor((Barcelona_counts$counts > q[2]) + (Barcelona_counts$counts > q[3]) +(Barcelona_counts$counts >= q[4]) + 1)
-plot4 <- ggplot(data = Barcelona_Polygons) +
-  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
+
+Barcelona_counts <- Barcelona_counts |>
+  mutate(Quantile = case_when(counts <= q[2] ~ glue("[{q[1]}-{q[2]})"),
+                              counts > q[2] & counts <= q[3] ~ glue("[{q[2]}-{q[3]})"),
+                              counts > q[3] & counts <= q[4] ~ glue("[{q[3]}-{q[4]})"),
+                              counts > q[4] ~ glue("[{q[4]}-{q[5]})")))
+
+barcelona_listings <- ggplot(data = Barcelona_counts) +
+  geom_sf(aes(fill = Quantile),
+          color = "black",size=.2) +
   theme_bw(base_size = 6) +
   scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
-        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
+  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm"),
+        legend.position = "bottom",
+        legend.title = element_blank(),
+        legend.key.size = unit(0.2, 'cm'),
+        text = element_text(size=4))
+
 # Madrid
-Madrid_counts <- Madrid_Polygons %>% 
-  mutate(counts = lengths(st_intersects(., Madrid_Sale)))
+Madrid_counts <- Madrid_Polygons |> 
+  mutate(counts = lengths(st_intersects(Madrid_Polygons, Madrid_Sale)))
 q <- quantile(Madrid_counts$counts)
-Madrid_Polygons$Quantile<- as.factor((Madrid_counts$counts > q[2]) + (Madrid_counts$counts > q[3]) +(Madrid_counts$counts >= q[4]) + 1)
-plot5 <- ggplot(data = Madrid_Polygons) +
-  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
+
+Madrid_counts <- Madrid_counts |>
+  mutate(Quantile = case_when(counts <= q[2] ~ glue("[{q[1]}-{q[2]})"),
+                              counts > q[2] & counts <= q[3] ~ glue("[{q[2]}-{q[3]})"),
+                              counts > q[3] & counts <= q[4] ~ glue("[{q[3]}-{q[4]})"),
+                              counts > q[4] ~ glue("[{q[4]}-{q[5]})")))
+
+madrid_listings <- ggplot(data = Madrid_counts) +
+  geom_sf(aes(fill = Quantile),
+          color = "black",size=.2) +
   theme_bw(base_size = 6) +
   scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
-        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
+  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm"),
+        legend.position = "bottom",
+        legend.title = element_blank(),
+        legend.key.size = unit(0.2, 'cm'),
+        text = element_text(size=4))
+
 # Valencia
-Valencia_counts <- Valencia_Polygons %>% 
-  mutate(counts = lengths(st_intersects(., Valencia_Sale)))
+Valencia_counts <- Valencia_Polygons |> 
+  mutate(counts = lengths(st_intersects(Valencia_Polygons, Valencia_Sale)))
 q <- quantile(Valencia_counts$counts)
-Valencia_Polygons$Quantile<- as.factor((Valencia_counts$counts > q[2]) + (Valencia_counts$counts > q[3]) +(Valencia_counts$counts >= q[4]) + 1)
-plot6 <- ggplot(data = Valencia_Polygons) +
-  geom_sf(aes(fill = Quantile),color = "black",size=.2) +
-  ylim(39.405,39.52) +
+
+Valencia_counts <- Valencia_counts |>
+  mutate(Quantile = case_when(counts <= q[2] ~ glue("[{q[1]}-{q[2]})"),
+                              counts > q[2] & counts <= q[3] ~ glue("[{q[2]}-{q[3]})"),
+                              counts > q[3] & counts <= q[4] ~ glue("[{q[3]}-{q[4]})"),
+                              counts > q[4] ~ glue("[{q[4]}-{q[5]})")))
+
+valencia_listings <- ggplot(data = Valencia_counts) +
+  geom_sf(aes(fill = Quantile),
+          color = "black",size=.2) +
   theme_bw(base_size = 6) +
   scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
-        legend.key.size = unit(0.2, 'cm'),text = element_text(size=4))
-grid.arrange(plot4, plot5, plot6, ncol = 3, nrow = 1)
+  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm"),
+        legend.position = "bottom",
+        legend.title = element_blank(),
+        legend.key.size = unit(0.2, 'cm'),
+        text = element_text(size=4))
+
+# Arrange plots
+grid.arrange(barcelona_listings, madrid_listings, valencia_listings,
+             ncol = 1,
+             nrow = 3)
 ```
@@
 
+The total number of dwellings is available from the Spanish cadastre aggregated by district (districts are groups of neighborhoods).
+Figure \ref{fig:districts} shows the percentage of listed dwellings relative to the total number of dwelling by district in each of the three cities.
+This gives a sense of how active residential real estate markets were in different parts of each city in 2018.
+
+```{r, echo = FALSE,message=FALSE, fig.width=5, fig.cap="\\label{fig:districts}Percentage of listings relative to total number of dwellings by district.
+Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
+# Barcelona
+Barcelona_prop <- properties_by_district |>
+  filter(CITY == "Barcelona") |>
+  mutate(proportion = N/N_CADASTRE)
+
+q <- quantile(Barcelona_prop$proportion)
+pq <- round(q * 100, 2)
+
+Barcelona_prop <- Barcelona_prop |>
+  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
+                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
+                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
+                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))
+
+plot4 <- ggplot(data = Barcelona_prop) +
+  geom_sf(aes(fill = Quantile),
+          color = "black",
+          size=.2) +
+  theme_bw(base_size = 6) +
+  scale_fill_manual(values=c("#FFFEDE",
+                             "#FFDFA2",
+                             "#FFA93F",
+                             "#D5610D"))+
+  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
+                         "cm"),
+        legend.position = "bottom",
+        legend.title = element_blank(),
+        legend.key.size = unit(0.2, 'cm'),
+        text = element_text(size=4))
+
+# Madrid
+Madrid_prop <- properties_by_district |>
+  filter(CITY == "Madrid") |>
+  mutate(proportion = N/N_CADASTRE)
+
+q <- quantile(Madrid_prop$proportion)
+pq <- round(q * 100, 2)
+
+Madrid_prop <- Madrid_prop |>
+  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
+                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
+                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
+                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))
+
+plot5 <- ggplot(data = Madrid_prop) +
+  geom_sf(aes(fill = Quantile),
+          color = "black",
+          size=.2) +
+  theme_bw(base_size = 6) +
+  scale_fill_manual(values=c("#FFFEDE",
+                             "#FFDFA2",
+                             "#FFA93F",
+                             "#D5610D"))+
+  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
+                         "cm"),
+        legend.position = "bottom",
+        legend.title = element_blank(),
+        legend.key.size = unit(0.2, 'cm'),
+        text = element_text(size=4))
+
+# Valencia
+Valencia_prop <- properties_by_district |>
+  filter(CITY == "Valencia") |>
+  mutate(proportion = N/N_CADASTRE)
+
+q <- quantile(Valencia_prop$proportion)
+pq <- round(q * 100, 2)
+
+Valencia_prop <- Valencia_prop |>
+  mutate(Quantile = case_when(proportion <= q[2] ~ glue("[{pq[1]}%-{pq[2]}%)"),
+                              proportion > q[2] & proportion <= q[3] ~ glue("[{pq[2]}%-{pq[3]}%)"),
+                              proportion > q[3] & proportion <= q[4] ~ glue("[{pq[3]}%-{pq[4]}%)"),
+                              proportion > q[4] ~ glue("[{pq[4]}%-{pq[5]}%)")))
+
+plot6 <- ggplot(data = Valencia_prop) +
+  geom_sf(aes(fill = Quantile),
+          color = "black",
+          size=.2) +
+  theme_bw(base_size = 6) +
+  scale_fill_manual(values=c("#FFFEDE",
+                             "#FFDFA2",
+                             "#FFA93F",
+                             "#D5610D"))+
+  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
+                         "cm"),
+        legend.position = "bottom",
+        legend.title = element_blank(),
+        legend.key.size = unit(0.2, 'cm'),
+        text = element_text(size=4))
+
+# Arrange plots
+grid.arrange(plot4, plot5, plot6, 
+             ncol = 1, 
+             nrow = 3)
+```
+
 ## Points of Interest
 
-The last block of data included in the data package is a set of Points of Interest in each city as an object of the class list.
+The last data object included in the package is a set of Points of Interest in each city as an object of the class list.
 The name of the list includes the name of the city with the suffix '_POIS'.
@@
 
-In terms of the asking prices, the original values were obfuscated with the addition or subtraction of a random percentage of their original values, ranging from -2.5% to +2.5%.
+In terms of the asking prices, the original values were obfuscated with the addition or subtraction of a random percentage of their original values, ranging from $-2.5%$ to $+2.5%$.
 Since asking prices are usually multiples of 1,000, after the first price modification, the prices were aligned to multiples of 1,000.
@@
 
-This paper describes a data product of a geo-referenced micro-data set of Spain’s three largest cities.
+This paper describes a data product of a geo-referenced micro-data set of Spain's three largest cities.
 This is an excellent data product to help understand the complex mechanisms related to the housing market and housing prices.
-Researchers can apply hedonic models with spatial effects, identifying housing submarkets or machine learning techniques.
+Researchers can apply hedonic models with spatial effects, identifying housing submarkets or machine learning techniques [e.g., @rey2023using].
 The data product can also be used for educational proposes and teaching activities.

### 2023-05-16 EPB/main_EPB.tex
Edits in preparation for resubmission and initializing renv

@@
 \begin{abstract}
-This article shares an open data product with big geo-referenced
+This article presents an open data product with large geo-referenced
 micro-data sets of 2018 real estate listings in Spain. These data were
@@
 
-\keywords{Housing market; idealista.com; geo-referenced data;
-point-level data; open data; Spain}
+\keywords{Housing prices; hedonic price analysis; idealista.com;
+geo-referenced data; point-level data; open data; Spain}
 
@@
 component to analyze the real estate market and incorporating geographic
-variables has significantly improved the understanding of this market.
-But to really understand the characteristics of the housing market, it
-is essential to have information/data at the point level. Therefore, it
-is becoming common for spatial analysis of urban environments to be
-developed with geo- referenced micro-data sets \citep{lopez2015}.
-However, the availability of this type of open data at the point level
-is limited, and not many data sets contain latitude/longitude
-coordinates for each dwelling. In some cases, researchers have had to
-resort to web scraping processes to obtain the large volumes of
-information that permit robust analyses
-\citep{gupta2022take, arbia2020spatial, Li2019, lopez2015}. These web
-scraping processes can include missing data, download errors, duplicate
-records, etc. Furthermore, the authors of this research do not generally
-share the data sets.
+variables has significantly improved the understanding of this market:
+to this end, it is essential to have information/data at the point
+level. Therefore, it is becoming common for spatial analysis of urban
+environments to be use micro-data, geo-referenced as points
+\citep{lopez2015}. In some cases, researchers have had to resort to
+webscraping processes to obtain data for research
+\citep[e.g.,][]{gupta2022take, arbia2020spatial, Li2019, lopez2015}.
+Alas, webscraping is a chancy process prone to download errors, missing
+data, duplicated records, etc. Furthermore, researchers do not always
+share their webscrapped datasets, which limits reproducibility of their
+research. As we witness a growing interest in openness and
+reproducibility in geographical data science
+\citep{arribasl2021editorial, paez2021open, brunsdon2021opening}, it
+becomes increasingly urgent to have open data products to support
+research \citep{arribas2021}.
 
-We are also witnessing a growing interest in open data in geography and
-data science \citep{arribasl2021editorial, arribas2021} using
-reproducible or replicable research
-\citep{paez2021open, brunsdon2021opening}. But to work openly in
-science, it is necessary to have free software and open data. While
-great efforts have been made to make free software available to
-researchers (e.g., R or Python), data regarding real estate markets tend
-to be treated as confidential, and there are still few open micro-data
-sets of housing markets available \citep{Song2021}.
+Some researchers have already responded to this need for publicly
+available, geo-referenced datasets to support open, reproducible
+research. A few such datasets are now available to support the analysis
+of real estate markets, but they are sometimes geo-referenced at the
+level of large geographical zones, such as \citet{fuerst2020real}, an
+open dataset that includes \(n=4,201\) property prices geocoded to the
+level of nine regions in England and Wales. Other datasets are geo-coded
+as points, including \citet{bonifaci2015real}, which includes
+\(n=1,042\) observations for Padua, in Italy;
+\citet{delgiudice2018housing}, who share a dataset with \(n=576\)
+observations relating to rental prices in Naples, Italy; and
+\citet{solano2019dataset} present a dataset with \(n=1,623\) daily
+rental prices in Seville, Spain.
 
@@
 and contain information about 189,923 dwellings located in Spain's three
-largest cities. To date, this data product is the biggest open
-geo-referenced micro-data set of the housing market in Spain. Moreover,
-the data set has been supplied directly by Idealista, and therefore is
+largest cities. To date, this data product is one of the largest open
+geo-referenced micro-data set of the housing market in the world. The
+most similar in terms of geographical disaggregation and sample size is
+the dataset of \citet{song2021hedonic} which includes transactions for
+four cities in South Korea, namely Busan (\(n=61,152\)), Daegu
+(\(n=32,363\)), Daejeon (\(n=21,114\)), and Gwangju (\(n=25,984\)). It
+is certainly the largest in Spain.
+
+The data set has been supplied directly by Idealista, and therefore is
 clean and free of download errors. However, in order to be able to share
@@
 
-This micro-data set is expected to be used as a benchmark dataset to
-test empirical methods in a reproducible fashion. Applied and
-theoretical researchers on real estate mass appraisal and valuation
-methods might use this dataset to canonically compare the performance of
-their proposed comparable and hedonic models, among others. The data can
-also be used to study the segmentation of housing submarkets and related
-topics such as the impact of suburban areas on house prices. The
-listings have been enriched with official information from the Spanish
-cadastre along with other relevant geographical features, such as
-distance to urban points of interest. In any case, the data might be
-easily extended by spatially joining other datasets that contain
-information at administrative, census tract, or postal code levels.
+This micro-data set can be used to benchmark new analytical methods in a
+reproducible fashion. Applied and theoretical researchers on real estate
+mass appraisal and valuation methods might use this dataset to
+canonically compare the performance of their proposed comparable and
+hedonic models, among others. The data can also be used to study the
+segmentation of housing submarkets and related topics such as the impact
+of suburban areas on house prices. The listings have been enriched with
+official information from the Spanish cadastre along with other relevant
+geographical features, such as distance to urban points of interest. In
+any case, the data might be easily extended by spatially joining other
+datasets that contain information at administrative, census tract, or
+postal code levels.
 
-The data set is distributed as an R package, name \{idealista18\}, which
-can be accessed from a public Github
-repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
-The database idealista18 is made available under the
+The data set is distributed as an \texttt{R} package, named
+\{idealista18\}, which can be accessed from a public Github
+repository\footnote{\url{https://github.com/paezha/idealista18}}. The
+open data product is made available under the
 \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
@@
 
-The open data product \{idealista18\} is an R package composed of nine
-objects, three objects for each of the three main Spanish cities:
-Barcelona, Madrid, and Valencia. For each city, dwelling listings,
-neighborhood polygons, and a set of points of interest have been
-included in the R package. The following subsections describe each
-object. A full description of the data is available in the help section
-of the package. We have strived to the best of our ability to comply
-with FAIR principles regarding research data \citep{Wilkinson2016fair}:
-upon publication, the dataset has a persistent digital object
-identifier; publication as a data article makes the data findable; the
-data and metadata are packaged together and protocols for help files in
-the R ecosystem mean that documentation is easily searchable;
-distribution as an R package means that only open software is needed to
-access the data; and a public repository documents all data processes
-followed to generate the distributed open data product.
+The open data product \{idealista18\} is an \texttt{R} package composed
+of nine objects, three objects for each of the three main Spanish
+cities: Barcelona, Madrid, and Valencia. For each city, dwelling
+listings, neighborhood polygons, and a set of points of interest have
+been included in the package.
+
+The data provider (idealista.com) is a leading real estate portal in
+Spain, on par with Fotocasa. Smaller portals include Habitaclia and
+Milanuncios, the latter focusing almost exclusively on individual (i.e.,
+non-professional) advertisers. In September 2021, according to data from
+Similarweb (a site specialized in sites' traffic volume comparison),
+there were a total of 103 million page views on real estate portals in
+Spain. Of this, 94\% of traffic was concentrated on four portals, of
+which idealista was the most important, followed by Fotocasa,
+Habitaclia, and Pisos.com. Traffic is also highly concentrated, and
+idealista.com was the the leader by far with 58.6 monthly million visits
+(57\% of the total traffic) compared to its closest competitor with 19.9
+million visits (19.3 \% of total traffic).
+
+As a result of its share of advertisments, idealista.com covers fairly
+well all segments of the Spanish market, including both individual and
+professional advertisers. This dataset includes information about
+listing prices, and therefore represents the market situation from the
+perspective of asking prices. This is a necessary compromise in the
+present case, since actual sale prices are not publicly available, and
+the information can only be accessed by paying high fees to Colegio de
+Registradores. However listing prices reflect quite well the
+(transaction) reality of real estate markets, and correlations between
+idealista listings and transaction prices can be established see
+\href{https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3400031}{Banco
+de España}. Official transactions and asking prices can be taken as
+complementary and are of great interest when studying asking-transaction
+price gaps or the relation between listing site demand variables (i.e.,
+ad contacts or ad views) and price gaps.
+
+To provide some context about the coverage of the \{idealista18\}
+dataset Table \ref{tab:transactions} shows the number of listings with
+respect to the total residential stock in each city in 2018. As seen in
+the table, the number of listings ranges between 6.1\% of the total
+number of properties (in Madrid) and 8.1\% (in Valencia). Information
+from Instituto Nacional de Estadistica\footnote{\url{https://www.ine.es/jaxiT3/Tabla.htm?t=6150&L=1}}
+shows that the number of listings in the \{idealista18\} package
+correspond to 81.3\% of recorded real estate transactions in Barcelona,
+80.8\% in Madrid, and 91.1\% in Valencia.
+
+\begin{table}[!ht]
+    \centering
+      \caption{Total properties and transactionst three Spanish cities. Year 2018} 
+  \label{tab:transactions} 
+    \begin{tabular}{lccccc}
+        City & Total properties (P) & Total transactions (T) & Listings (L) & L/P & T/L \\
+        \hline
+        Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
+        Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
+        Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
+        \hline
+        Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
+        \hline
+        \multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
+        \multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
+        \multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
+    \end{tabular}
+\end{table}
+
+The following subsections describe each object. A full description of
+the data is available in the help section of the package. We have tried
+to the best of our ability to comply with FAIR principles regarding
+research data \citep{Wilkinson2016fair}: upon publication, the dataset
+has a persistent digital object identifier; publication as a data
+article makes the data findable; the data and metadata are packaged
+together and protocols for help files in the R ecosystem mean that
+documentation is easily searchable; distribution as an R package means
+that only open software is needed to access the data; and a public
+repository documents all data processes followed to generate the
+distributed open data product.
 
@@
 each dwelling published on the idealista real estate website as an ad.
-The dwelling listing has been included in the `idealista18' package as
-an sf object \citep{Pebesma}. The name of the sf object containing the
-dwelling listing includes the name of the city, followed by '\_Sale'
-(e.g., Madrid\_Sale) and includes a total of 42 variables. Each sf
-object includes the complete set of listings corresponding to the four
-quarters of the year 2018. Table \ref{tab:number-ads} shows the number
-of dwelling listing ads included in the data set for each city and
-quarter. The record counts for each city are: 94,815 listings for
-Madrid, 61,486 for Barcelona, and 33,622 for Valencia. Note that the
-same dwelling may be found in more than one period when a property
-listed for sale in one quarter was sold in a subsequent quarter. The
-variable ASSETID, included in the sf objects, is the unique identifier
-of the dwelling.
+The dwelling listing has been included in the \{idealista18\} package as
+a simple features (sf) object \citep{Pebesma} with point geometry in
+latitude and longitude. The name of each sf object with the list of
+dwellings is the name of the city followed by '\_Sale' (e.g.,
+Madrid\_Sale). Each data object includes a total of 42 variables and the
+complete set of listings corresponding to the four quarters of 2018.
+Table \ref{tab:number-ads} shows the number of dwelling listing ads
+included in the data set for each city and quarter. The record counts
+for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and
+33,622 for Valencia. Note that the same dwelling may be found in more
+than one period when a property listed for sale in one quarter was sold
+in a subsequent quarter. The variable ASSETID, included in the sf
+objects, is the unique identifier of the dwelling.
 
@@
   \hline
-City$\backslash$Quarter & First & Second  & Thirdr & Fourth & Total ads \\ 
+City$\backslash$Quarter & First & Second  & Third & Fourth & Total ads \\ 
   \hline
@@
 (e.g., price, surface area, number of bedrooms, basic features, etc.),
-including the exact location of the dwelling (see Section
+including the location of the dwelling. The exact location has been
+masked, as described in Section
 \protect\hyperlink{anonymizing}{Anonymizing the data set}). Table
@@
 
-Additionally, the distance of each dwelling to three urban points of
+Additionally, the distance of each dwelling to several points of
 interest was included in the sf object: distance to the city center,
@@
 
-The second block of data included in the `idealista18' R package is the
-spatial features of the three cities divided into neighborhoods. There
-is an sf object for each city with the name of the city and the suffix
-'\_Polygons'. Figure \ref{fig:all-polygons} shows the quantile maps of
-the number of dwellings in the listing for the different neighborhoods
-in the three cities. The neighborhoods are based on the official
-boundaries but slightly changed by
+The second data object included in \{idealista18\} is the neighborhoods
+in the three cities as polygons. There is an sf object for each city
+with the name of the city and the suffix '\_Polygons'. Figure
+\ref{fig:all-polygons} shows the quantile maps of the number of
+dwellings in the listing for the different neighborhoods in the three
+cities. The neighborhoods are based on the official boundaries but
+slightly changed by
 Idealista\footnote{There are two criteria used to make this division.
@@
 
+The total number of dwellings is available from the Spanish cadastre
+aggregated by district (districts are groups of neighborhoods). Figure
+\ref{fig:districts} shows the percentage of listed dwellings relative to
+the total number of dwelling by district in each of the three cities.
+This gives a sense of how active residential real estate markets were in
+different parts of each city in 2018.
+
+\begin{figure}
+\centering
+\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-2-1.pdf}
+\caption{\label{fig:districts}Quantile maps of the percentage of
+listings relative to total number of dwellings by district. Boundary for
+Barcelona (Left), Madrid (Center), and Valencia (Right).}
+\end{figure}
+
 \hypertarget{points-of-interest}{%
@@
 
-The last block of data included in the data package is a set of Points
-of Interest in each city as an object of the class list. The name of the
+The last data object included in the package is a set of Points of
+Interest in each city as an object of the class list. The name of the
 list includes the name of the city with the suffix '\_POIS'. These lists
@@
 
-\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)}\label{fig:unnamed-chunk-2}
+\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)}\label{fig:unnamed-chunk-3}
 \end{figure}
@@
 housing prices. Researchers can apply hedonic models with spatial
-effects, identifying housing submarkets or machine learning techniques.
-The data product can also be used for educational proposes and teaching
-activities.
+effects, identifying housing submarkets or machine learning techniques
+\citep[e.g.,][]{rey2023using}. The data product can also be used for
+educational proposes and teaching activities.
 

### 2023-05-16 R/idealista18.R

@@ -23,6 +23,31 @@
 #' @references \url{https://github.com/paezha/Reproducible-Research-Workflow}
 NULL
 
+#' Total number of properties by district in 2018.
+#'
+#' A dataset containing information about the number of properties by district in Barcelona, Madrid, and Valencia during 2018 according to the cadastre. Districts are collections of neighborhoods.
+#'
+#' @format An object of class `sf` (inherits from `data.frame`) with 50 rows and 5 variables:
+#' \describe{
+#'   \item{District}{Number of district}
+#'   \item{N_CADASTRE}{Date YYYYMM, indicates the quarter when the ad was extracted we used YYYY03 for the 1st Q, YYYY06 the 2nd, YYYY09 for the 3rd and YYYY12 for the 4th}
+#'   \item{N}{Asking price for the ad at idealista expressed in euros}
+#'   \item{CITY}{Price in euros per square meter}
+#'   \item{geometry}{geometry of simple features in latitude and longitude}
+#'   }
+#'
+#' @docType data
+#' @keywords datasets
+#' @name properties_by_district
+#' @usage data(properties_by_district)
+#' @source Idealista (https://www.idealista.com/)
+"properties_by_district"
+#' @examples
+#'  library(idealista18)
+#'  data(properties_by_district)
+#'  summary(properties_by_district$N)
+"properties_by_district"
+
 #' Sale prices of residential property in Barcelona in 2018.
 #'
 #' A dataset containing information about 84,280 real estate listings on idealista
@@ -68,7 +93,7 @@ NULL
 #'   \item{BUILTTYPEID_3}{Dummy value for flat condition: 1 second hand in good condition 0 otherwise (source: advertiser))}
 #'   \item{DISTANCE_TO_CITY_CENTER}{Distance to center of city in km}
 #'   \item{DISTANCE_TO_DIAGONAL}{Distance to Avinguda Diagonal in km; Diagonal is a major street that cuts across the city diagonally to the street grid}
-#'   \item{geometry}{geometry of simple features}
+#'   \item{geometry}{geometry of simple features in latitude and longitude}
 #'   }
 #'
 #' @docType data
@@ -93,7 +118,7 @@ NULL
 #'   \item{ASSETID}{Unique identifier of the asset}
[…]

### 2023-05-16 data-raw/DATASET.R

@@ -1,0 +1,103 @@
+## code to prepare `DATASET` dataset goes here
+
+library(dplyr)
+library(ggplot2)
+library(here)
+library(readxl)
+library(sf)
+
+# Load datasets
+
+load(paste0(here(), "/data-raw/Barcelona_Polygons.RData"))
+load(paste0(here(), "/data-raw/Madrid_Polygons.RData"))
+load(paste0(here(), "/data-raw/Valencia_Polygons.RData"))
+
+load(paste0(here(), "/data-raw/Barcelona_Sale.RData"))
+load(paste0(here(), "/data-raw/Madrid_Sale.RData"))
+load(paste0(here(), "/data-raw/Valencia_Sale.RData"))
+
+load(paste0(here(), "/data-raw/Barcelona_POIS.RData"))
+load(paste0(here(), "/data-raw/Madrid_POIS.RData"))
+load(paste0(here(), "/data-raw/Valencia_POIS.RData"))
+
+# Rename geometry columns
+Barcelona_Polygons <- Barcelona_Polygons |>
+  rename(geometry = WKT)
+
+Madrid_Polygons <- Madrid_Polygons |>
+  rename(geometry = WKT)
+
+Valencia_Polygons <- Valencia_Polygons |>
+  rename(geometry = WKT)
+
+# Read data on properties by district
+properties_by_district <- readxl::read_xlsx(paste0(here(), "/data-raw/Ads-by-District.xlsx"))
+
+# Summarize districts, which are groups of geometries
+Barcelona_Districts <- Barcelona_Polygons |>
+  mutate(District = as.numeric(substr(LOCATIONID,23,24))) |>
+  group_by(District) |>
+  summarise(District = unique(District)) |>
+  mutate(CITY = "Barcelona")
+
+Madrid_Districts <- Madrid_Polygons |>
+  mutate(District = as.numeric(substr(LOCATIONID,23,24))) |>
+  group_by(District) |>
+  summarise(District = unique(District)) |>
+  mutate(CITY = "Madrid")
+
+Valencia_Districts <- Valencia_Polygons |>
+  mutate(District = as.numeric(substr(LOCATIONID,23,24))) |>
+  group_by(District) |>
+  summarise(District = unique(District)) |>
+  mutate(CITY = "Valencia")
+
+
+Districts <- rbind(Barcelona_Districts,
+                   Madrid_Districts,
+                   Valencia_Districts)
+
[…]

### 2023-05-17 editorial-materials/EPB-2022-0116.R1_response_to_reviewer.Rmd
Further work in preparation for resubmission

@@
 
-> Comments to the Author
+>Comments to the Author
 
-> Dear Authors, Thank you for the opportunity to review this paper.
+>Dear Authors, Thank you for the opportunity to review this paper.
 I would appreciate it if you would highlight the changed text in the next submission.
@@
 
-> 1.Please clearly state the nature of the data in the abstract.
+Thank you for your constructive comments. Below we describe how we improved the paper in response.
+
+---
+
+>1.
+Please clearly state the nature of the data in the abstract.
 It is worthwhile to mention the relocated and randomised dataset in the abstract.
@@
 
-We appreciate your comment.
+We appreciate this comment.
 There are a few publicly available datasets for hedonic price analysis, but they tend to be smaller.
@@
 In response to your comment we drop the terminology of "big data" and describe our dataset as "large" instead.
+
+---
 
@@
 
+---
+
 >3.
@@
 
-> “idealista covers fairly well all segments in the Spanish market, both individual and professional advertisers.
+>“idealista covers fairly well all segments in the Spanish market, both individual and professional advertisers.
 This dataset comprehends information about listing prices, thus it contains the market situation from the offer.
@@
 
->4.For data description, it is worth showing solid evidence of the quality of this dataset.
+---
+
+>4.
+For data description, it is worth showing solid evidence of the quality of this dataset.
 For example, what is the data coverage of this data compared to all the residential properties in the city?
@@
 
-Thank you.
-We appreciate your comment.
-To provide better context about the coverage of the {idealista18} dataset Table \ref{tab:transactions} now shows the number of listings with respect to the total residential stock in each city (in 2018).
-Globally, the number of listings to the total housing stock is around 7%.
-Additional information about the number of free-market home purchases can be obtained from the Spanish National Institute.
-Globally, the number of listingss 2,7 times the number of transactions.
+**Response:**
 
-To provide some context about the coverage of the {idealista18} dataset Table \ref{tab:transactions} shows the number of listings with respect to the total residential stock in each city in 2018.
+To better place in context the coverage of the {idealista18} dataset, Table \ref{tab:transactions} now shows the number of listings with respect to the total residential stock in each city (in 2018).
 As seen in the table, the number of listings ranges between 6.1% of the total number of properties (in Madrid) and 8.1% (in Valencia).
-Information from Instituto Nacional de Estadistica^[\url{https://www.ine.es/jaxiT3/Tabla.htm?t=6150&L=1}] shows that the number of listings in the {idealista18} package correspond to 81.3\% of recorded real estate transactions in Barcelona, 80.8\% in Madrid, and 91.1\% in Valencia.
+Information from Instituto Nacional de Estadistica^[\url{https://www.ine.es/jaxiT3/Tabla.htm?t=6150&L=1}] for 2018 also shows that the number of transactions in 2018 are equivalent to 81.3\% of idealista listings in Barcelona, 80.8\% in Madrid, and 91.1\% in Valencia.
+While it is not possible to link actual transactions to idealista listings, this gives a quantitative sense of the coverage of the data package.
 
@@
 
+---
+
 >5.
@@
 
-We appreciate the opportunity to explain this issue, as it may be of interest to academic researchers.
-Adding a small amount of random noise ($\pm 2.5$) to the prices has the effect of increasing the variance of estimates, but does not affect the mean.
-This is a well-known result in algebra of random variables.
+We appreciate the opportunity to delve more deeply into this issue, as it may be of interest to users of the data package.
+The short version of this story is that adding a small amount of random noise to the prices ($\pm 2.5\%$) does not introduce bias but does increase the variance of the masked variable.
+This can be shown with some standard algebra of random variables.
 
-Prices ($P$) in the dataset are calculated as the sum of two random variables, the raw price ($RP$) and some random noise ($\eta$) drawn from a uniform distribution as follows:
+Now the longer story.
+
+The masked prices ($P$) in the dataset are calculated as the sum of two random variables, the original raw prices ($RP$) and a random masking variable ($\eta$).
+The masking variable is generated as follows:
+$$
+\eta = RP\cdot\epsilon
 $$
 
+The masking variable uses values drawn at random from the uniform distribution with parameters $b = 0.025$ and $a = -0.025$:
+$$
+f(\epsilon) = 
+\begin{cases}
+\frac{1}{b - a} & \text{for } a\le \epsilon \le b\\
+0 & \text{otherwise}
+\end{cases}
 $$
 
-The mean of the raw price is:
+The effect of $\epsilon$ is to produce a random percentage of the raw price ($\pm 2.5\%$). The expectation of $\epsilon$ given these parameters is:
 $$
-\mathrm{E}[RP] = \frac{1}{N}\sum_{i=1}^N R_i
+E[\epsilon] = \frac{a+b}{2} = \frac{-0.025 + 0.025}{2} = 0
 $$
 
-The mean of the white noise is zero by design:
+\noindent and the variance is:
 $$
-\mathrm{E}[\eta] = 0
+\mathrm{V}[\epsilon]=\frac{(b - a)^2}{12} = \frac{25}{120000}
 $$
 
-We know that the expectation of the sum of two random variables is:
+The masked prices $P$ are the raw prices plus the masking variable:
 $$
-\mathrm{E}[P] = \mathrm{E}[RP + \eta] = \mathrm{E}[RP] + \mathrm{E}[\eta]
+P = RP + RP\cdot\epsilon = RP + \eta
 $$
 
-\noindent and since $\mathrm{E}[\eta] = 0$ it follows that the prices are unbiased.
-
-The variance of the sum of two random variables is:
+Since the mean of $\epsilon$ is zero, and $RP$ and $\epsilon$ are independent, it turns out that the mean of the masking variable is zero:
 $$
-\mathrm{V}[P] = \mathrm{V}[RP + \eta] = \mathrm{V}[RP] + 2\mathrm{Cov}[RP,\eta] + \mathrm{V}[\eta]
+\mathrm{E}[\eta] = \mathrm{E}[RP\cdot\epsilon] = \mathrm{E}[RP] \cdot \mathrm{E}[\epsilon] = 0
 $$
 
-The variance of $\eta$ is:
+It follows that the masked prices are unbiased since their expectation is identical to the expectation of the raw prices:
 $$
-\mathrm{V}[\eta]=\frac{1}{12}\big(0.025 - (-0.25)\big)^2
+\mathrm{E}[P] = \mathrm{E}[RP] + \mathrm{E}[\eta] = \mathrm{E}[RP]
 $$
+
+Now let us examine the covariance. Again, since $RP$ and $\epsilon$ are independent, the variance of the masking variable $\eta$ is:
+$$
+\mathrm{V}[\eta] = \mathrm{V}[RP\cdot\epsilon] = \mathrm{V}[RP]\cdot\mathrm{V}[\epsilon] + \mathrm{V}[RP]\cdot(\mathrm{E}[\epsilon])^2 + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2  
+$$
+
+Considering that $\mathrm{E}[\epsilon] = 0$, then:
+$$
+\mathrm{V}[\eta] = \mathrm{V}[RP]\cdot\mathrm{V}[\epsilon] + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2 = \mathrm{V}[\epsilon]\cdot\big(\mathrm{V}[RP] + (\mathrm{E}[RP])^2\big) = \frac{25}{120000}\big(\mathrm{V}[RP] + (\mathrm{E}[RP])^2\big)
+$$
+
+Accordingly, the variance of the masked prices is:
+$$
+\mathrm{V}[P] = \mathrm{V}[RP + \eta] = \mathrm{V}[RP] + \mathrm{V}[\eta] + 2\mathrm{Cov}[RP,\eta] 
+$$
+
+If we recall that $\mathrm{E}[\eta] = 0$ and $\mathrm{E}[\epsilon] = 0$, the covariance of $RP$ and $\eta$ can be shown to be zero:
+$$
+\mathrm{Cov}[RP,\eta] = \mathrm{E}[RP\cdot \eta] - \mathrm{E}[RP]\cdot \mathrm{E}[\eta] = \mathrm{E}[RP\cdot \eta] = \mathrm{E}[RP^2\cdot \epsilon] = \mathrm{E}[RP^2]\cdot \mathrm{E}[\epsilon] = 0
+$$
+
+From the above we can conclude that the variance of the masked prices is:
+$$
+\mathrm{V}[P] = \mathrm{V}[RP] + \mathrm{V}[\eta] = \mathrm{V}[RP] + \frac{25}{120000}\big(\mathrm{V}[RP] + (\mathrm{E}[RP])^2\big) = \frac{120025}{120000}\cdot\mathrm{V}[RP] + \frac{25}{120000}\cdot(\mathrm{E}[RP])^2
+$$
+
+\noindent which is $0.0208%$ greater than the variance of the raw prices, plus $0.0208%$ of the square of the mean of the raw prices.
+
+What does this mean for users of the data? The masked prices are an unbiased version of the raw prices, but they have a somewhat larger variance than the original (unreported) prices.
+
+---
 
@@
 
-**Response:** After this comment we included a new figure to display the percentage of listings relative to the total number of dwellings by district.
+**Response:**
 
-```{r, echo = FALSE,message=FALSE, fig.width=5, fig.cap="\\label{fig:districts}Percentage of listings relative to total number of dwellings by district.
+After this comment we included a new figure in the paper to display the percentage of listings relative to the total number of dwellings by district (see Figure \ref{fig:districts} below).
+
+```{r, echo = FALSE,message=FALSE, fig.height=8, fig.cap="\\label{fig:districts}Percentage of listings relative to total number of dwellings by district.
 Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
-library(idealista18)
 library(dplyr)
 library(ggplot2)
+library(glue)
+library(gridExtra)
+library(idealista18)
 library(sf)
-library(gridExtra)
 
@@
   geom_sf(aes(fill = Quantile),
-          color = "black",
-          size=.2) +
+          color = NA) +
+  geom_sf(data = Barcelona_Polygons,
+          fill = NA) +
   theme_bw(base_size = 6) +
@@
         legend.position = "bottom",
-        legend.title = element_blank(),
-        legend.key.size = unit(0.2, 'cm'),
-        text = element_text(size=4))
+        legend.title = element_blank())
 
@@
   geom_sf(aes(fill = Quantile),
-          color = "black",
+          color = NA,
           size=.2) +
+    geom_sf(data = Madrid_Polygons,
+          fill = NA) +
   theme_bw(base_size = 6) +
@@
   geom_sf(aes(fill = Quantile),
-          color = "black",
+          color = NA,
           size=.2) +
+    geom_sf(data = Valencia_Polygons,
+          fill = NA) +
   theme_bw(base_size = 6) +
@@
 
-*Dear Authors,* *Thank you for the opportunity to review this paper.
-I would appreciate it if you would highlight the changed text in the next submission.
-Otherwise it is difficult to identify which parts have changed.
-The unknown quality of this one year randomised data is the main issue for potential users and I repeat my suggestion that this data should be validated for the benefit potential readers and users.
-Below is a list of comments and questions which need to be addressed.*
+Thank you again for your generous feedback.
+We hope that our responses will prove satisfactory and you will agree that the paper makes a worthwhile contribution to growing catalog of datasets to support research on real estate markets.
 
-> **Answer:** Thnak you.
-> In the new version of the manuscrip, the changes are writen in red.
+\newpage
 
-## First Comment
-
->Please clearly state the nature of the data in the abstract.
-It is worthwhile to mention the relocated and randomised dataset in the abstract.
-This house price dataset only has 189,923 records, which could be not really be described as a big dataset as you describe it in the abstract.
-Except the authors could provide an evidence/reference on why this is big dataset.
-
-## Second Comment
-
-2.  *Please check the typographical errors before your next submission. For example, the "thirdr" in your table1.*
-
-> **Answer:** Thank you.
-> A native English speaker has reviewed the document.
-
-## Third Comment
-
-3.
-*Please use part of your answer to my previous fourth question in the manuscript.
-Your answer is quite useful for your future users, especially those researchers who are not familiar with house price data in Spain.
-I quote your previous answer below:*
-
-*"dealista covers fairly well all segments in the Spanish market, both individual and professional advertisers.
-This dataset comprehends information about listing prices, thus it contains the market situation from the offer.
-Nevertheless, actual sale prices information is not publicly available as it is information that can only be accessed by paying high fees to Colegio de Registradores.
-However listing prices reflect quite well the (transaction) reality of real estate markets using the offer perspective, they keep strong correlation between idealista and transaction prices can be established see Banco de España.
-Even though a great extent of correlation does exist between official transactional and asking records, both databases can be taken as complementary and are of great interest when studying asking-transaction price gaps or the relation between listing site demand variables (ie.
-ad contacts or ad views) and price gaps.
-The main real estate portals in Spain are idealista and Fotocasa, further away are Habitaclia and Milanuncios, the latter focusing almost exclusively on private individuals.
-In September 2021, according to data from Similarweb (a site specialized in sites' traffic volume comparison), in Spain, there were a total of 103 million page views on real estate portals, out of page views on real estate portals, of which the four main portals, idealista, Fotocasa, Habitaclia, and Pisos.com, in that order, accounted for 94% of the traffic.
-Another relevant feature of this sector is that it is highly concentrated, idealista being the leader by far with 58.6 monthly million visits (57% of the total traffic) versus its immediate competitor with 19.9 million visits (19.3 % of total traffic).
-In terms of the evolution over time of interest in the content of each portal, based on data from Google Trends, the leadership of idealista Trends, the leadership of idealist"*
-
-> **Answer:** Thank you. We include it.
-
-## Fourth Comment
-
-4.*For data description, it is worth showing solid evidence of the quality of this dataset.
-For example, what is the data coverage of this data compared to all the residential properties in the city?
-Listing the advertised volume is not very helpful for users who wish to understand the quality.*
-
-> **Answer:** 
-## Fifth Comment
-
-5.
-*Given the asking price is a resized version of the original asking price with a random percentage between -2.5% and +2.5%, please elaborate on how this influences your quantitative research results and findings.* **Please provide critical evidence to explain how this randomised price will not influence academic research.** *For example, how does the house price variation at local level change after the price is randomised.
-To what degree will this influence a typical hedonic house price research approach?
-I am concerned about whether this randomised data could bias conclusions.*
-
-> **Answer:** Of course, the results of any econometric model will be differents if the original data are used.
-> Unfortunally, the data protection laws in Spain, specifically the Organic Law on the Protection of Personal Data and the Guarantee of Digital Rights (LOPDGDD) and the General Data Protection Regulation (GDPR) of the European Union, establish that personal data, including information on the location of a home, is protected and must be treated confidentially.
-> It is not permitted to disclose the specific location of a home without the explicit consent of the owner or person who has the right to dispose of it.
-
-> The small changes in the prices will be, in mean, a small impact on the results.
-
-## Sixth Comment
-
-6.
-*In figure 1, is it possible to standardise this data rather than showing only the number of dwelling records in your house price data?
-I guess the city centre always has high advertised volumes but it also has a high density of dwellings.
-It is worthwhile to show values standardised by the number of dwellings*
-
-```{r, eval = FALSE, include = FALSE, message=FALSE, fig.width=5, }
-library(readxl)
-library(idealista18)
-library(dplyr)
-library(ggplot2)
-library(sf)
-library("gridExtra")
-Data <- read_excel("Adds_by_Distrits_idealista18.xlsx")
-Data$Por <- Data$N*100/Data$N_CADASTRE
-Barcelona_Polygons$District <- as.numeric(substr(Barcelona_Polygons$LOCATIONID,23,24))
-Barcelona_District <- Barcelona_Polygons %>% group_by(District) %>% summarise(District = unique(District))
-Madrid_Polygons$District <- as.numeric(substr(Madrid_Polygons$LOCATIONID,23,24))
-Madrid_District <- Madrid_Polygons %>% group_by(District) %>% summarise(District = unique(District))
-Valencia_Polygons$District <- as.numeric(substr(Valencia_Polygons$LOCATIONID,23,24))
-Valencia_District <- Valencia_Polygons %>% group_by(District) %>% summarise(District = unique(District))
-
-Data.bar <- Data[Data$CITY=="Barcelona",]
-Data.mad <- Data[Data$CITY=="Madrid",]
-Data.val <- Data[Data$CITY=="Valencia",]
-
-Barcelona_District <- merge(Barcelona_District,Data.bar,by="District")
-q <- quantile(Barcelona_District$Por)
-Barcelona_District$Quantile<- as.factor((Barcelona_District$Por > q[2]) + (Barcelona_District$Por > q[3]) +(Barcelona_District$Por >= q[4]) + 1)
-levels(Barcelona_District$Quantile) <- c("(4.7%,5.7%]","(5.7,6.9]","(6.9,8.2]","(8.2,14.9]")
-plot.d1 <- ggplot(data = Barcelona_District) +
-  geom_sf(aes(fill = Quantile),size=.2) +
-  theme_bw(base_size = 6) +
-  scale_fill_discrete("") +
-    scale_fill_manual(name = "",values=c("#D4E7F1","#6DAFD3","#2470B2","darkblue")) +
-  theme_bw(base_size = 6) +
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
-        legend.key.size = unit(.2, 'cm'),text = element_text(size=4))
-
-Madrid_District <- merge(Madrid_District,Data.mad,by="District")
-q <- quantile(Madrid_District$Por)
-Madrid_District$Quantile<- as.factor((Madrid_District$Por > q[2]) + (Madrid_District$Por > q[3]) +(Madrid_District$Por >= q[4]) + 1)
-plot.d2 <- ggplot(data = Madrid_District) +
-  geom_sf(aes(fill = Quantile),size=.2) +
-  theme_bw(base_size = 6) +
-  scale_fill_manual(name = "",values=c("#D4E7F1","#6DAFD3","#2470B2","darkblue")) +
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
-        legend.key.size = unit(.2, 'cm'),text = element_text(size=4))
-
-Valencia_District <- merge(Valencia_District,Data.val,by="District")
-q <- quantile(Valencia_District$Por)
-#levels(B$Quartile) <- c("Q1: (1200,4050]","Q2: (4050,5450]","Q3: (5450,8290]","Q4: (8290,830000]")
-Valencia_District$Quantile<- as.factor((Valencia_District$Por > q[2]) + (Valencia_District$Por > q[3]) +(Valencia_District$Por >= q[4]) + 1)
-plot.d3 <- ggplot(data = Valencia_District) +
-  geom_sf(aes(fill = Quantile),size=.2) +
-  theme_bw(base_size = 6) +
-  ylim(39.405,39.52) +
-  scale_fill_manual(name = "",values=c("#D4E7F1","#6DAFD3","#2470B2","darkblue"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1),"cm"),legend.position = "bottom",
-        legend.key.size = unit(.2, 'cm'),text = element_text(size=4))
-grid.arrange(plot.d1, plot.d2, plot.d3, ncol = 3, nrow = 1)
-```
+## References

### 2023-05-17 EPB/main_EPB.Rmd
Further work in preparation for resubmission

@@
 The data set has been supplied directly by Idealista, and therefore is clean and free of download errors.
-However, in order to be able to share publicly the data we have slightly modified some variables by applying some random noise which will not bias the main results derived from its usage.
+However, to comply with data legislation, to share the data we have slightly modified the prices by applying a small amount of random noise which will not bias the main results derived from its usage.
 
@@
 The open data product is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
-For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.
+For transparency, we also share the randomization process applied to the original data in the aforementioned Github repository.
 
@@
 
-The open data product {idealista18} is an `R` package composed of nine objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia.
+The open data product {idealista18} is an `R` package composed of ten objects, three objects for each of the three main Spanish cities: Barcelona, Madrid, and Valencia.
 For each city, dwelling listings, neighborhood polygons, and a set of points of interest have been included in the package.
+There is in addition a data object with the number of dwellings by district (a collection of neighborhoods), according to the Spanish cadastre.
 
-The data provider (idealista.com) is a leading real estate portal in Spain, on par with Fotocasa.
-Smaller portals include Habitaclia and Milanuncios, the latter focusing almost exclusively on individual (i.e., non-professional) advertisers.
+The data provider (idealista.com) is a leading real estate portal in Spain, on par with its nearest competitor Fotocasa.
+Smaller listing portals include Habitaclia and Milanuncios, the latter focusing almost exclusively on individual (i.e., non-professional) advertisers.
 In September 2021, according to data from Similarweb (a site specialized in sites' traffic volume comparison), there were a total of 103 million page views on real estate portals in Spain.

### 2023-05-17 EPB/main_EPB.tex
Further work in preparation for resubmission

@@
 The data set has been supplied directly by Idealista, and therefore is
-clean and free of download errors. However, in order to be able to share
-publicly the data we have slightly modified some variables by applying
-some random noise which will not bias the main results derived from its
-usage.
+clean and free of download errors. However, to comply with data
+legislation, to share the data we have slightly modified the prices by
+applying a small amount of random noise which will not bias the main
+results derived from its usage.
 
@@
 \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
-For transparency, we also share the randomisation process applied to the
+For transparency, we also share the randomization process applied to the
 original data in the aforementioned Github repository.
@@
 The open data product \{idealista18\} is an \texttt{R} package composed
-of nine objects, three objects for each of the three main Spanish
-cities: Barcelona, Madrid, and Valencia. For each city, dwelling
-listings, neighborhood polygons, and a set of points of interest have
-been included in the package.
+of ten objects, three objects for each of the three main Spanish cities:
+Barcelona, Madrid, and Valencia. For each city, dwelling listings,
+neighborhood polygons, and a set of points of interest have been
+included in the package. There is in addition a data object with the
+number of dwellings by district (a collection of neighborhoods),
+according to the Spanish cadastre.
 
 The data provider (idealista.com) is a leading real estate portal in
-Spain, on par with Fotocasa. Smaller portals include Habitaclia and
-Milanuncios, the latter focusing almost exclusively on individual (i.e.,
-non-professional) advertisers. In September 2021, according to data from
-Similarweb (a site specialized in sites' traffic volume comparison),
-there were a total of 103 million page views on real estate portals in
-Spain. Of this, 94\% of traffic was concentrated on four portals, of
-which idealista was the most important, followed by Fotocasa,
-Habitaclia, and Pisos.com. Traffic is also highly concentrated, and
-idealista.com was the the leader by far with 58.6 monthly million visits
-(57\% of the total traffic) compared to its closest competitor with 19.9
-million visits (19.3 \% of total traffic).
+Spain, on par with its nearest competitor Fotocasa. Smaller listing
+portals include Habitaclia and Milanuncios, the latter focusing almost
+exclusively on individual (i.e., non-professional) advertisers. In
+September 2021, according to data from Similarweb (a site specialized in
+sites' traffic volume comparison), there were a total of 103 million
+page views on real estate portals in Spain. Of this, 94\% of traffic was
+concentrated on four portals, of which idealista was the most important,
+followed by Fotocasa, Habitaclia, and Pisos.com. Traffic is also highly
+concentrated, and idealista.com was the the leader by far with 58.6
+monthly million visits (57\% of the total traffic) compared to its
+closest competitor with 19.9 million visits (19.3 \% of total traffic).
 
@@
 \includegraphics{main_EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:all-polygons}Quantile maps of the number of
-dwellings in each neighborhood. Boundary for Barcelona (Left), Madrid
-(Center), and Valencia (Right).}
+\caption{\label{fig:all-polygons}Number of listings in each
+neighborhood. Boundary for Barcelona (Top), Madrid (Center), and
+Valencia (Bottom).}
 \end{figure}
@@
 \includegraphics{main_EPB_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{\label{fig:districts}Quantile maps of the percentage of
-listings relative to total number of dwellings by district. Boundary for
-Barcelona (Left), Madrid (Center), and Valencia (Right).}
+\caption{\label{fig:districts}Percentage of listings relative to total
+number of dwellings by district. Boundary for Barcelona (Top), Madrid
+(Center), and Valencia (Bottom).}
 \end{figure}
@@
 the addition or subtraction of a random percentage of their original
-values, ranging from -2.5\% to +2.5\%. Since asking prices are usually
-multiples of 1,000, after the first price modification, the prices were
-aligned to multiples of 1,000.
+values, ranging from \(-2.5%
+\) to \(+2.5%
+\). Since asking prices are usually multiples of 1,000, after the first
+price modification, the prices were aligned to multiples of 1,000.
 

### 2023-05-19 editorial-materials/EPB-2022-0116.R1_response_to_reviewer.Rmd
Resubmission to EPB

@@
 bibliography: bibEPB.bib
+link-citations: yes
+linkcolor: blue
 ---
+
+```{r load-packages, include=FALSE}
+library(dplyr)
+library(ggplot2)
+library(glue)
+library(gridExtra)
+library(idealista18)
+library(kableExtra)
+library(sf)
+```
 
@@
 \begin{tabular}{lccccc}
-        City & Total properties (P) & Total transactions (T) & Listings (L) & L/P & T/L \\
+City & Total properties (TP) & Total transactions (TT) & Listings (L) & L/TP & TT/L \\
 \hline
@@
 
-The masked prices ($P$) in the dataset are calculated as the sum of two random variables, the original raw prices ($RP$) and a random masking variable ($\eta$).
-The masking variable is generated as follows:
+The masked prices are given by:
 $$
-\eta = RP\cdot\epsilon
+P = RP\cdot \epsilon
 $$
 
-The masking variable uses values drawn at random from the uniform distribution with parameters $b = 0.025$ and $a = -0.025$:
+\noindent where $\epsilon$ is drawn from the uniform distribution with parameters $a=0.975$ and $b=1.025$:
 $$
@@
 
-The effect of $\epsilon$ is to produce a random percentage of the raw price ($\pm 2.5\%$). The expectation of $\epsilon$ given these parameters is:
+The expectation of $\epsilon$ given these parameters is:
 $$
-E[\epsilon] = \frac{a+b}{2} = \frac{-0.025 + 0.025}{2} = 0
+E[\epsilon] = \frac{a+b}{2} = \frac{0.975 + 1.025}{2} = 1
 $$
 
-\noindent and the variance is:
+\noindent and the variance of $\epsilon$ is:
 $$
-\mathrm{V}[\epsilon]=\frac{(b - a)^2}{12} = \frac{25}{120000}
+\mathrm{V}[\epsilon]=\frac{(b - a)^2}{12} = \frac{1}{4800}
 $$
 
-The masked prices $P$ are the raw prices plus the masking variable:
+Therefore, the expectation of the masked prices is:
 $$
-P = RP + RP\cdot\epsilon = RP + \eta
+\mathrm{E}[P] = \mathrm{E}[RP\cdot \epsilon] = \mathrm{E}[RP]\cdot \mathrm{E}[\epsilon] = \mathrm{E}[RP]
 $$
 
-Since the mean of $\epsilon$ is zero, and $RP$ and $\epsilon$ are independent, it turns out that the mean of the masking variable is zero:
+In other words, the masked prices $P$ are an unbiased version of the raw prices $RP$.
+
+Considering that $RP$ and $\epsilon$ are independent, the variance of the masked prices is as follows:
 $$
-\mathrm{E}[\eta] = \mathrm{E}[RP\cdot\epsilon] = \mathrm{E}[RP] \cdot \mathrm{E}[\epsilon] = 0
+\mathrm{V}[P] = \mathrm{V}[RP\cdot \epsilon] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP]\cdot (\mathrm{E}[\epsilon])^2 + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
 $$
 
-It follows that the masked prices are unbiased since their expectation is identical to the expectation of the raw prices:
+Since $\mathrm{E}[\epsilon] = 1$, we have that:
 $$
-\mathrm{E}[P] = \mathrm{E}[RP] + \mathrm{E}[\eta] = \mathrm{E}[RP]
+\mathrm{V}[P] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP] + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2 = \mathrm{V}[RP]\cdot (1 + \mathrm{V}[\epsilon]) + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
 $$
 
-Now let us examine the covariance. Again, since $RP$ and $\epsilon$ are independent, the variance of the masking variable $\eta$ is:
+Solving for $\mathrm{V}[RP]$, and replacing the expectation of the raw prices by its unbiased version ($\mathrm{E}[P]$), yields the variance of the raw prices:
 $$
-\mathrm{V}[\eta] = \mathrm{V}[RP\cdot\epsilon] = \mathrm{V}[RP]\cdot\mathrm{V}[\epsilon] + \mathrm{V}[RP]\cdot(\mathrm{E}[\epsilon])^2 + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2  
+\mathrm{V}[RP] = \frac{\mathrm{V}[P] - \frac{1}{4800}\cdot (\mathrm{E}[P])^2}{1 + \frac{1}{4800}}
 $$
 
-Considering that $\mathrm{E}[\epsilon] = 0$, then:
-$$
-\mathrm{V}[\eta] = \mathrm{V}[RP]\cdot\mathrm{V}[\epsilon] + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2 = \mathrm{V}[\epsilon]\cdot\big(\mathrm{V}[RP] + (\mathrm{E}[RP])^2\big) = \frac{25}{120000}\big(\mathrm{V}[RP] + (\mathrm{E}[RP])^2\big)
-$$
+Table \ref{tab:variance} reports the mean and the standard deviation of the prices in the package, and the standard deviation of the raw prices calculated using the formula above.
+The last column of the table can be read as an inflation factor.
+We see that the variance of the masked prices is inflated with respect to the variance of the original values by less than $1\%$ in all cases examined.
+Users can use the formula above to calculate the inflation of the variance if they use subsamples other than those shown, to assess the potential impacts of the masking (e.g., when computing intervals of confidence).
 
-Accordingly, the variance of the masked prices is:
-$$
-\mathrm{V}[P] = \mathrm{V}[RP + \eta] = \mathrm{V}[RP] + \mathrm{V}[\eta] + 2\mathrm{Cov}[RP,\eta] 
-$$
+```{r echo = FALSE}
+data("Barcelona_Sale")
+data("Madrid_Sale")
+data("Valencia_Sale")
+```
 
-If we recall that $\mathrm{E}[\eta] = 0$ and $\mathrm{E}[\epsilon] = 0$, the covariance of $RP$ and $\eta$ can be shown to be zero:
-$$
-\mathrm{Cov}[RP,\eta] = \mathrm{E}[RP\cdot \eta] - \mathrm{E}[RP]\cdot \mathrm{E}[\eta] = \mathrm{E}[RP\cdot \eta] = \mathrm{E}[RP^2\cdot \epsilon] = \mathrm{E}[RP^2]\cdot \mathrm{E}[\epsilon] = 0
-$$
+```{r echo=FALSE}
+# Barcelona
+Barcelona_Price <- Barcelona_Sale |>
+  select(PRICE, PERIOD) |>
+  st_drop_geometry()
 
-From the above we can conclude that the variance of the masked prices is:
-$$
-\mathrm{V}[P] = \mathrm{V}[RP] + \mathrm{V}[\eta] = \mathrm{V}[RP] + \frac{25}{120000}\big(\mathrm{V}[RP] + (\mathrm{E}[RP])^2\big) = \frac{120025}{120000}\cdot\mathrm{V}[RP] + \frac{25}{120000}\cdot(\mathrm{E}[RP])^2
-$$
+BQ <- Barcelona_Price |>
+  group_by(PERIOD) |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "BARCELONA")
 
-\noindent which is $0.0208%$ greater than the variance of the raw prices, plus $0.0208%$ of the square of the mean of the raw prices.
+BT <- Barcelona_Price |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(PERIOD = "2018",
+         sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "BARCELONA")
 
-What does this mean for users of the data? The masked prices are an unbiased version of the raw prices, but they have a somewhat larger variance than the original (unreported) prices.
+# Madrid
+Madrid_Price <- Madrid_Sale |>
+  select(PRICE, PERIOD) |>
+  st_drop_geometry()
+
+MQ <- Madrid_Price |>
+  group_by(PERIOD) |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "MADRID")
+
+MT <- Madrid_Price |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(PERIOD = "2018",
+         sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "MADRID")
+
+# Valencia
+Valencia_Price <- Valencia_Sale |>
+  select(PRICE, PERIOD) |>
+  st_drop_geometry()
+
+VQ <- Valencia_Price |>
+  group_by(PERIOD) |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "VALENCIA")
+
+VT <- Valencia_Price |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(PERIOD = "2018",
+         sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "VALENCIA")
+
+Q <- rbind(BQ,
+           BT,
+           MQ,
+           MT,
+           VQ,
+           VT)
+
+Q |>
+  mutate(PERIOD = case_when(PERIOD == "201803" ~ "Q1",
+                            PERIOD == "201806" ~ "Q2",
+                            PERIOD == "201809" ~ "Q3",
+                            PERIOD == "201812" ~ "Q4",
+                            PERIOD == "2018" ~ "2018"),
+         sd = round(sqrt(variance), 2),
+         sd_raw = round(sqrt(var_raw), 2),
+         inflation = round(inflation, 5),
+         across(c(mean, sd, sd_raw, inflation), ~ prettyNum(.x, big.mark = ",")))  |>
+  select(CITY, PERIOD, mean, sd, sd_raw, inflation) |>
+  kableExtra::kable(booktabs = TRUE,
+                    caption = "\\label{tab:variance}Inflation of the variance of masked prices with respect to raw prices",
+                    col.names = c("City", "Period", "mean(P)", "sd(P)", "sd(RP)", "sd(P)/sd(RP)")) |>
+  collapse_rows(columns = 1) |>
+  footnote(general = c("P: masked prices; ", "RP: raw prices; ", "sd: standard deviation (square root of the variance)"),
+           threeparttable = T)
+```
 
@@
 Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
-library(dplyr)
-library(ggplot2)
-library(glue)
-library(gridExtra)
-library(idealista18)
-library(sf)
-
 data("properties_by_district")

### 2023-05-19 EPB/main_EPB.Rmd
Resubmission to EPB

@@
 Along with the real estate listings, the data product also includes neighborhood boundaries for each city.
-The data product is offered as a fully documented R package and is available for scientific and educational purposes, particularly for geo-spatial studies"
+The data product is offered as a fully documented `R` package and is available for scientific and educational purposes, particularly for geo-spatial studies"
+
+link-citations: yes
+linkcolor: blue
 
@@
 Therefore, it is becoming common for spatial analysis of urban environments to be use micro-data, geo-referenced as points [@lopez2015].
-In some cases, researchers have had to resort to webscraping processes to obtain data for research [e.g., @gupta2022take; @arbia2020spatial; @Li2019; @lopez2015].
+In some cases, researchers have had to resort to webscraping processes to obtain data for research [e.g., @Li2019; @lopez2015].
 Alas, webscraping is a chancy process prone to download errors, missing  data, duplicated records, etc.
@@
 
-To contribute to the growing inventory of international micro-data sets of real estate markets, this paper introduces an open micro-data set of geo-referenced dwelling listings.
+To contribute to the growing inventory of international micro-data sets of real estate markets, this paper introduces an open data product with geo-referenced dwelling listings, called {idealista18}.
 The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain's three largest cities.
@@
 The most similar in terms of geographical disaggregation and sample size is the dataset of @song2021hedonic which includes transactions for four cities in South Korea, namely Busan ($n=61,152$), Daegu ($n=32,363$), Daejeon ($n=21,114$), and Gwangju ($n=25,984$).
-It is certainly the largest in Spain.
+{idealista18} is certainly the largest open data product of its kind in Spain.
 
-The data set has been supplied directly by Idealista, and therefore is clean and free of download errors.
-However, to comply with data legislation, to share the data we have slightly modified the prices by applying a small amount of random noise which will not bias the main results derived from its usage.
+The data set has been supplied directly by Idealista, and therefore is clean and free from download errors.
+However, to comply with data legislation, we have masked the prices by applying a small amount of random noise that will not bias the main results derived from its usage.
 
-This micro-data set can be used to benchmark new analytical methods in a reproducible fashion.
+This micro-data set can be used to benchmark new methods in a reproducible fashion  [e.g., @rey2023using].
 Applied and theoretical researchers on real estate mass appraisal and valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others.
@@
 
-The data set is distributed as an `R` package, named {idealista18}, which can be accessed from a public Github repository\footnote{\url{https://github.com/paezha/idealista18}}.
+The data set is distributed as an `R` package and it can be accessed from a public Github repository\footnote{\url{https://github.com/paezha/idealista18}}.
 The open data product is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
-For transparency, we also share the randomization process applied to the original data in the aforementioned Github repository.
+For transparency, we also share the masking process applied to the original data in the aforementioned Github repository.
 
@@
 
-The following subsections describe each object.
-A full description of the data is available in the help section of the package.
-We have tried to the best of our ability to comply with FAIR principles regarding research data [@Wilkinson2016fair]: upon publication, the dataset has a persistent digital object identifier; publication as a data article makes the data findable; the data and metadata are packaged together and protocols for help files in the R ecosystem mean that documentation is easily searchable; distribution as an R package means that only open software is needed to access the data; and a public repository documents all data processes followed to generate the distributed open data product.
+The following subsections describe the data objects.
+A full description of the data is also available in the help section of the package.
+We have tried to the best of our ability to comply with FAIR principles regarding research data [@Wilkinson2016fair]: upon publication, the dataset has a persistent digital object identifier; publication as a data article makes the data findable; the data and metadata are packaged together and protocols for help files in the `R` ecosystem mean that documentation is easily searchable; distribution as an `R` package means that only open software is needed to access the data; and a public repository documents all data processes followed to generate the distributed open data product.
 
@@
 
-The dwelling listing of each city includes a set of characteristics for each dwelling published on the idealista real estate website as an ad.
-The dwelling listing has been included in the {idealista18} package as a simple features (sf) object [@Pebesma] with point geometry in latitude and longitude.
-The name of each sf object with the list of dwellings is the name of the city followed by '_Sale' (e.g., Madrid_Sale).
-Each data object includes a total of 42 variables and the complete set of listings corresponding to the four quarters of 2018.
+The listing for each city includes a set of characteristics for each dwelling published on the idealista real estate website.
+The listing included in the {idealista18} package are simple features (sf) objects [@Pebesma] with point geometry in latitude and longitude.
+The name of each sf object with the list of dwellings is the name of the city followed by '_Sale' (e.g., `Madrid_Sale`).
+Each data object includes a total of 42 variables and the complete set of listings corresponding to the four quarters of 2018 (Q1 through Q4).
 Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter.
@@
 
-Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including the location of the dwelling.
-The exact location has been masked, as described in Section [Anonymizing the data set](#anonymizing)).
+Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including an approximated location of the dwelling (the exact location has been masked, as described in Section [Masking the prices](#anonymizing)).
 Table \ref{tab:variables} lists the main indoor variables with a short description and the mean value of each variable.
@@
 xtable::xtable(text_tbl, caption = "List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities.
-See the help section in the \\textbf{idealista18} R package for details and formal definitions.
+See the help section in the \\{idealista18\\} package for details and formal definitions.
 Some variables have been excluded from this table to save space.
-Check the full list in the \\textbf{idealista18} package.\\label{tab:variables}") |>
+Check the full list in the package.\\label{tab:variables}") |>
 xtable2kable(include.rownames = FALSE) |>
@@
 There is an sf object for each city with the name of the city and the suffix '_Polygons'.
-Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities.
+The left column of Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities.
 The neighborhoods are based on the official boundaries but slightly changed by Idealista\footnote{There are two criteria used to make this division.
@@
 
-```{r, echo = FALSE,message=FALSE, fig.width=5, fig.cap="\\label{fig:all-polygons}Number of listings in each neighborhood. Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
+```{r listings-neighborhood, include = FALSE}
 
@@
 
-barcelona_listings <- ggplot(data = Barcelona_counts) +
+plot1 <- ggplot(data = Barcelona_counts) +
   geom_sf(aes(fill = Quantile),
           color = "black",size=.2) +
+  labs(caption = "Barcelona neighborhoods: listings") +
   theme_bw(base_size = 6) +
   scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm"),
+  theme(plot.margin=unit(c(0.1, 0.5, 0.1, 0.5), "cm"),
         legend.position = "bottom",
@@
 
-madrid_listings <- ggplot(data = Madrid_counts) +
+plot2 <- ggplot(data = Madrid_counts) +
   geom_sf(aes(fill = Quantile),
           color = "black",size=.2) +
+  labs(caption = "Madrid neighborhoods: listings") +
   theme_bw(base_size = 6) +
   scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm"),
+  theme(plot.margin=unit(c(0.1, 0.5, 0.1, 0.5), "cm"),
         legend.position = "bottom",
@@
 
-valencia_listings <- ggplot(data = Valencia_counts) +
+plot3 <- ggplot(data = Valencia_counts) +
   geom_sf(aes(fill = Quantile),
           color = "black",size=.2) +
+  labs(caption = "Valencia neighborhoods: listings") +
+  ylim(39.405,39.52) +
   theme_bw(base_size = 6) +
   scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
-  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm"),
+  theme(plot.margin=unit(c(0.1, 0.5, 0.1, 0.5), "cm"),
         legend.position = "bottom",
@@
         text = element_text(size=4))
-
-# Arrange plots
-grid.arrange(barcelona_listings, madrid_listings, valencia_listings,
-             ncol = 1,
-             nrow = 3)
 ```
@@
 The total number of dwellings is available from the Spanish cadastre aggregated by district (districts are groups of neighborhoods).
-Figure \ref{fig:districts} shows the percentage of listed dwellings relative to the total number of dwelling by district in each of the three cities.
+The right column in Figure \ref{fig:all-polygons} shows the percentage of listed dwellings relative to the total number of dwelling by district in each of the three cities.
 This gives a sense of how active residential real estate markets were in different parts of each city in 2018.
 
-```{r, echo = FALSE,message=FALSE, fig.width=5, fig.cap="\\label{fig:districts}Percentage of listings relative to total number of dwellings by district.
-Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
+```{r listings-to-properties, include = FALSE}
 # Barcelona
@@
           fill = NA) +
+  labs(caption = "Barcelona districts: listings/total dwellings") +
   theme_bw(base_size = 6) +
@@
                              "#D5610D"))+
-  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
+  theme(plot.margin=unit(c(0.1, 0.5, 0.1, 0.5),
                          "cm"),
@@
           fill = NA) +
+  labs(caption = "Madrid districts: listings/total dwellings") +
   theme_bw(base_size = 6) +
@@
           fill = NA) +
+  labs(caption = "Valencia districts: listings/total dwellings") +
+  ylim(39.405,39.52) +
   theme_bw(base_size = 6) +
@@
         text = element_text(size=4))
+```
 
-# Arrange plots
-grid.arrange(plot4, plot5, plot6, 
-             ncol = 1, 
-             nrow = 3)
+```{r figure-listings, echo=FALSE, fig.height=6.5, message=FALSE, fig.height=6.5, fig.cap="\\label{fig:all-polygons}Listings by neighborhood (left column) and percentage of listings relative to total dwellings (right column).
+Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
+
+library(patchwork)
+(plot1 + plot4)/(plot2 + plot5)/(plot3 + plot6)
 ```
@@
 
-# Anonymizing the data set {#anonymizing}
+# Masking the prices {#anonymizing}
 
-To comply with Spanish regulations, two variables were slightly modified to provide anonymity. A masking process was applied to asking prices and location (coordinates).
+To comply with Spanish regulations, three variables were slightly modified to provide a measure of anonymity. A masking process was applied to asking prices and location (coordinates).
 
@@
 
+To understand the implications of this masking process, we use some standard results from algebra of random variables. The masked prices $P$ are given by:
+$$
+P = RP\cdot \epsilon
+$$
+
+\noindent where $RP$ are the original (raw) prices, and $\epsilon$ is a random variable drawn from the uniform distribution with parameters $a=0.975$ and $b=1.025$:
+$$
+f(\epsilon) = 
+\begin{cases}
+\frac{1}{b - a} & \text{for } a\le \epsilon \le b\\
+0 & \text{otherwise}
+\end{cases}
+$$
+
+The expectation of $\epsilon$ given these parameters is:
+$$
+E[\epsilon] = \frac{a+b}{2} = \frac{0.975 + 1.025}{2} = 1
+$$
+
+\noindent and the variance of $\epsilon$ is:
+$$
+\mathrm{V}[\epsilon]=\frac{(b - a)^2}{12} = \frac{1}{4800}
+$$
+
+Therefore, the expectation of the masked prices is:
+$$
+\mathrm{E}[P] = \mathrm{E}[RP\cdot \epsilon] = \mathrm{E}[RP]\cdot \mathrm{E}[\epsilon] = \mathrm{E}[RP]
+$$
+
+In other words, the masked prices $P$ are an unbiased version of the raw prices $RP$.
+
+Considering that $RP$ and $\epsilon$ are independent, the variance of the masked prices is as follows:
+$$
+\mathrm{V}[P] = \mathrm{V}[RP\cdot \epsilon] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP]\cdot (\mathrm{E}[\epsilon])^2 + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
+$$
+
+Since $\mathrm{E}[\epsilon] = 1$, we have that:
+$$
+\mathrm{V}[P] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP] + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2 = \mathrm{V}[RP]\cdot (1 + \mathrm{V}[\epsilon]) + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
+$$
+
+Solving for $\mathrm{V}[RP]$, and replacing the expectation of the raw prices by its unbiased version ($\mathrm{E}[P]$), yields the variance of the raw prices:
+$$
+\mathrm{V}[RP] = \frac{\mathrm{V}[P] - \frac{1}{4800}\cdot (\mathrm{E}[P])^2}{1 + \frac{1}{4800}}
+$$
+
+Table \ref{tab:variance} reports the mean and the standard deviation (i.e., the square root of the variance) of the prices in the package, and the standard deviation  (again, the square root of the variance) of the raw prices calculated using the formula above.
+This is done for each quarter and for the full year.
+The last column of the table can be read as an inflation factor.
+We see that the variance of the masked prices is inflated with respect to the variance of the original values by less than $1\%$ in all cases examined.
+Users can use the formula above to calculate the inflation of the variance if they use subsamples other than those shown, to assess the potential impacts of the masking (e.g., when computing intervals of confidence).
+
+```{r variance-inflation-calculations, echo=FALSE}
+# Barcelona
+Barcelona_Price <- Barcelona_Sale |>
+  select(PRICE, PERIOD) |>
+  st_drop_geometry()
+
+BQ <- Barcelona_Price |>
+  group_by(PERIOD) |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "BARCELONA")
+
+BT <- Barcelona_Price |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(PERIOD = "2018",
+         sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "BARCELONA")
+
+# Madrid
+Madrid_Price <- Madrid_Sale |>
+  select(PRICE, PERIOD) |>
+  st_drop_geometry()
+
+MQ <- Madrid_Price |>
+  group_by(PERIOD) |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "MADRID")
+
+MT <- Madrid_Price |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(PERIOD = "2018",
+         sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "MADRID")
+
+# Valencia
+Valencia_Price <- Valencia_Sale |>
+  select(PRICE, PERIOD) |>
+  st_drop_geometry()
+
+VQ <- Valencia_Price |>
+  group_by(PERIOD) |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "VALENCIA")
+
+VT <- Valencia_Price |>
+  summarize(mean = mean(PRICE),
+            variance = var(PRICE),
+            .groups = "drop") |>
+  mutate(PERIOD = "2018",
+         sq_mean = mean^2,
+         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
+         inflation = variance/var_raw,
+         CITY = "VALENCIA")
+
+Q <- rbind(BQ,
+           BT,
+           MQ,
+           MT,
+           VQ,
+           VT)
+
+Q |>
+  mutate(PERIOD = case_when(PERIOD == "201803" ~ "Q1",
+                            PERIOD == "201806" ~ "Q2",
+                            PERIOD == "201809" ~ "Q3",
+                            PERIOD == "201812" ~ "Q4",
+                            PERIOD == "2018" ~ "2018"),
+         sd = round(sqrt(variance), 2),
+         sd_raw = round(sqrt(var_raw), 2),
+         inflation = round(inflation, 5),
+         across(c(mean, sd, sd_raw, inflation), ~ prettyNum(.x, big.mark = ",")))  |>
+  select(CITY, PERIOD, mean, sd, sd_raw, inflation) |>
+  kableExtra::kable(booktabs = TRUE,
+                    caption = "\\label{tab:variance}Inflation of the variance of masked prices with respect to raw prices",
+                    col.names = c("City", "Period", "mean(P)", "sd(P)", "sd(RP)", "sd(P)/sd(RP)")) |>
+  kable_styling(font_size = 9) |>
+  collapse_rows(columns = 1) |>
+  footnote(general = c("P: masked prices; ", "RP: raw prices; ", "sd: standard deviation (square root of the variance)"),
+           threeparttable = T)
+```
+
+With respect to the location of the dwelling, a spatial masking process was implemented  to maintain the spatial properties of the original data set.
+The coordinates of each listing were displaced using a stochastic procedure.
+The listings were recorded using coordinates contained in maximum and minimum displacement circles, as shown in Figure \ref{fig:Anonymizing} (left).
+To preserve inclusion in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates remained in the original neighborhood of the listing.
 ```{=tex}
@@
  }
- \caption{Coordinate displacement process for anonymisation purposes}
+ \caption{Coordinate displacement process for masking purposes}
  \label{algo:coordinates-displacement}
@@
 ```
-
-With respect to the location of the dwelling, a spatial masking process was implemented  to maintain the spatial properties of the original data set.
-The coordinates of each listing were displaced using a stochastic procedure.
-The listings were recorded using coordinates contained in maximum and minimum displacement circles, as shown in Figure \ref{fig:Anonymizing} (left).
-To preserve inclusion in a neighborhood, the spatial masking procedure was constrained to ensure that the masked coordinates remained in the original neighborhood of the listing.
 
@@
 
-```{r, echo = FALSE, out.width=c("29%","37%"), out.height="20%",fig.cap="\\label{fig:Anonymizing}(Left) Masking coordinates.
+```{r figure-masking-coordinates, echo = FALSE, out.width=c("29%","37%"), out.height="20%",fig.cap="\\label{fig:Anonymizing}(Left) Masking coordinates.
 Spatial range.
@@
 This paper describes a data product of a geo-referenced micro-data set of Spain's three largest cities.
-This is an excellent data product to help understand the complex mechanisms related to the housing market and housing prices.
-Researchers can apply hedonic models with spatial effects, identifying housing submarkets or machine learning techniques [e.g., @rey2023using].
+This data product can be of value to support research into the mechanisms of housing market and housing prices.
+Researchers can apply hedonic models with spatial effects, identify housing submarkets, or experiment with machine learning techniques.
 The data product can also be used for educational proposes and teaching activities.
+To the best of our knowledge, this is the largest, publicly available data set of its type that is also analysis ready and fully documented.
 
@@
 
-Author One and author Two are employed by Idealista.
+Authors One and Two are employed by Idealista.
 They have been granted permission to share the data presented in this article.
@@
 The authors wish to thank Alessandro Galesi for their support in the paper revision and Juan Ramón Selva for collecting and cleaning the spatial data.
-This work has been partially funded by the Spanish Ministry of Economy and Competitiveness Grants PID2019-107800GB-100
+This work has been partially funded by the Spanish Ministry of Economy and Competitiveness Grants PID2019-107800GB-100, but it was not funded by any of Canada's research councils.

### 2023-05-19 EPB/main_EPB.tex
Resubmission to EPB

@@
 neighborhood boundaries for each city. The data product is offered as a
-fully documented R package and is available for scientific and
+fully documented \texttt{R} package and is available for scientific and
 educational purposes, particularly for geo-spatial studies
@@
 webscraping processes to obtain data for research
-\citep[e.g.,][]{gupta2022take, arbia2020spatial, Li2019, lopez2015}.
-Alas, webscraping is a chancy process prone to download errors, missing
-data, duplicated records, etc. Furthermore, researchers do not always
-share their webscrapped datasets, which limits reproducibility of their
-research. As we witness a growing interest in openness and
-reproducibility in geographical data science
+\citep[e.g.,][]{Li2019, lopez2015}. Alas, webscraping is a chancy
+process prone to download errors, missing data, duplicated records, etc.
+Furthermore, researchers do not always share their webscrapped datasets,
+which limits reproducibility of their research. As we witness a growing
+interest in openness and reproducibility in geographical data science
 \citep{arribasl2021editorial, paez2021open, brunsdon2021opening}, it
@@
 To contribute to the growing inventory of international micro-data sets
-of real estate markets, this paper introduces an open micro-data set of
-geo-referenced dwelling listings. The data have been provided by the
-Idealista
+of real estate markets, this paper introduces an open data product with
+geo-referenced dwelling listings, called \{idealista18\}. The data have
+been provided by the Idealista
 company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal}
@@
 four cities in South Korea, namely Busan (\(n=61,152\)), Daegu
-(\(n=32,363\)), Daejeon (\(n=21,114\)), and Gwangju (\(n=25,984\)). It
-is certainly the largest in Spain.
+(\(n=32,363\)), Daejeon (\(n=21,114\)), and Gwangju (\(n=25,984\)).
+\{idealista18\} is certainly the largest open data product of its kind
+in Spain.
 
 The data set has been supplied directly by Idealista, and therefore is
-clean and free of download errors. However, to comply with data
-legislation, to share the data we have slightly modified the prices by
-applying a small amount of random noise which will not bias the main
-results derived from its usage.
+clean and free from download errors. However, to comply with data
+legislation, we have masked the prices by applying a small amount of
+random noise that will not bias the main results derived from its usage.
 
-This micro-data set can be used to benchmark new analytical methods in a
-reproducible fashion. Applied and theoretical researchers on real estate
-mass appraisal and valuation methods might use this dataset to
-canonically compare the performance of their proposed comparable and
-hedonic models, among others. The data can also be used to study the
-segmentation of housing submarkets and related topics such as the impact
-of suburban areas on house prices. The listings have been enriched with
-official information from the Spanish cadastre along with other relevant
-geographical features, such as distance to urban points of interest. In
-any case, the data might be easily extended by spatially joining other
-datasets that contain information at administrative, census tract, or
-postal code levels.
+This micro-data set can be used to benchmark new methods in a
+reproducible fashion \citep[e.g.,][]{rey2023using}. Applied and
+theoretical researchers on real estate mass appraisal and valuation
+methods might use this dataset to canonically compare the performance of
+their proposed comparable and hedonic models, among others. The data can
+also be used to study the segmentation of housing submarkets and related
+topics such as the impact of suburban areas on house prices. The
+listings have been enriched with official information from the Spanish
+cadastre along with other relevant geographical features, such as
+distance to urban points of interest. In any case, the data might be
+easily extended by spatially joining other datasets that contain
+information at administrative, census tract, or postal code levels.
 
-The data set is distributed as an \texttt{R} package, named
-\{idealista18\}, which can be accessed from a public Github
+The data set is distributed as an \texttt{R} package and it can be
+accessed from a public Github
 repository\footnote{\url{https://github.com/paezha/idealista18}}. The
@@
 \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
-For transparency, we also share the randomization process applied to the
+For transparency, we also share the masking process applied to the
 original data in the aforementioned Github repository.
@@
 
-The following subsections describe each object. A full description of
-the data is available in the help section of the package. We have tried
-to the best of our ability to comply with FAIR principles regarding
-research data \citep{Wilkinson2016fair}: upon publication, the dataset
-has a persistent digital object identifier; publication as a data
-article makes the data findable; the data and metadata are packaged
-together and protocols for help files in the R ecosystem mean that
-documentation is easily searchable; distribution as an R package means
-that only open software is needed to access the data; and a public
-repository documents all data processes followed to generate the
-distributed open data product.
+The following subsections describe the data objects. A full description
+of the data is also available in the help section of the package. We
+have tried to the best of our ability to comply with FAIR principles
+regarding research data \citep{Wilkinson2016fair}: upon publication, the
+dataset has a persistent digital object identifier; publication as a
+data article makes the data findable; the data and metadata are packaged
+together and protocols for help files in the \texttt{R} ecosystem mean
+that documentation is easily searchable; distribution as an \texttt{R}
+package means that only open software is needed to access the data; and
+a public repository documents all data processes followed to generate
+the distributed open data product.
 
@@
 
-The dwelling listing of each city includes a set of characteristics for
-each dwelling published on the idealista real estate website as an ad.
-The dwelling listing has been included in the \{idealista18\} package as
-a simple features (sf) object \citep{Pebesma} with point geometry in
-latitude and longitude. The name of each sf object with the list of
-dwellings is the name of the city followed by '\_Sale' (e.g.,
-Madrid\_Sale). Each data object includes a total of 42 variables and the
-complete set of listings corresponding to the four quarters of 2018.
-Table \ref{tab:number-ads} shows the number of dwelling listing ads
-included in the data set for each city and quarter. The record counts
-for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and
-33,622 for Valencia. Note that the same dwelling may be found in more
-than one period when a property listed for sale in one quarter was sold
-in a subsequent quarter. The variable ASSETID, included in the sf
-objects, is the unique identifier of the dwelling.
+The listing for each city includes a set of characteristics for each
+dwelling published on the idealista real estate website. The listing
+included in the \{idealista18\} package are simple features (sf) objects
+\citep{Pebesma} with point geometry in latitude and longitude. The name
+of each sf object with the list of dwellings is the name of the city
+followed by '\_Sale' (e.g., \texttt{Madrid\_Sale}). Each data object
+includes a total of 42 variables and the complete set of listings
+corresponding to the four quarters of 2018 (Q1 through Q4). Table
+\ref{tab:number-ads} shows the number of dwelling listing ads included
+in the data set for each city and quarter. The record counts for each
+city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622
+for Valencia. Note that the same dwelling may be found in more than one
+period when a property listed for sale in one quarter was sold in a
+subsequent quarter. The variable ASSETID, included in the sf objects, is
+the unique identifier of the dwelling.
 
@@
 (e.g., price, surface area, number of bedrooms, basic features, etc.),
-including the location of the dwelling. The exact location has been
-masked, as described in Section
-\protect\hyperlink{anonymizing}{Anonymizing the data set}). Table
+including an approximated location of the dwelling (the exact location
+has been masked, as described in Section
+\protect\hyperlink{anonymizing}{Masking the prices}). Table
 \ref{tab:variables} lists the main indoor variables with a short
@@
 \caption{List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities.
-See the help section in the \textbf{idealista18} R package for details and formal definitions.
+See the help section in the \{idealista18\} package for details and formal definitions.
 Some variables have been excluded from this table to save space.
-Check the full list in the \textbf{idealista18} package.\label{tab:variables}}
+Check the full list in the package.\label{tab:variables}}
 \end{table}
@@
 in the three cities as polygons. There is an sf object for each city
-with the name of the city and the suffix '\_Polygons'. Figure
-\ref{fig:all-polygons} shows the quantile maps of the number of
-dwellings in the listing for the different neighborhoods in the three
+with the name of the city and the suffix '\_Polygons'. The left column
+of Figure \ref{fig:all-polygons} shows the quantile maps of the number
+of dwellings in the listing for the different neighborhoods in the three
 cities. The neighborhoods are based on the official boundaries but
@@
 
-\begin{figure}
-\centering
-\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
-\caption{\label{fig:all-polygons}Number of listings in each
-neighborhood. Boundary for Barcelona (Top), Madrid (Center), and
-Valencia (Bottom).}
-\end{figure}
-
 There are a total of 69 neighborhoods in Barcelona, 135 in Madrid, and
@@
 The total number of dwellings is available from the Spanish cadastre
-aggregated by district (districts are groups of neighborhoods). Figure
-\ref{fig:districts} shows the percentage of listed dwellings relative to
-the total number of dwelling by district in each of the three cities.
-This gives a sense of how active residential real estate markets were in
-different parts of each city in 2018.
+aggregated by district (districts are groups of neighborhoods). The
+right column in Figure \ref{fig:all-polygons} shows the percentage of
+listed dwellings relative to the total number of dwelling by district in
+each of the three cities. This gives a sense of how active residential
+real estate markets were in different parts of each city in 2018.
 
@@
 \centering
-\includegraphics{main_EPB_files/figure-latex/unnamed-chunk-2-1.pdf}
-\caption{\label{fig:districts}Percentage of listings relative to total
-number of dwellings by district. Boundary for Barcelona (Top), Madrid
-(Center), and Valencia (Bottom).}
+\includegraphics{main_EPB_files/figure-latex/figure-listings-1.pdf}
+\caption{\label{fig:all-polygons}Listings by neighborhood (left column)
+and percentage of listings relative to total dwellings (right column).
+Barcelona (Top), Madrid (Center), and Valencia (Bottom).}
 \end{figure}
@@
 \hypertarget{anonymizing}{%
-\section{Anonymizing the data set}\label{anonymizing}}
+\section{Masking the prices}\label{anonymizing}}
 
-To comply with Spanish regulations, two variables were slightly modified
-to provide anonymity. A masking process was applied to asking prices and
-location (coordinates).
+To comply with Spanish regulations, three variables were slightly
+modified to provide a measure of anonymity. A masking process was
+applied to asking prices and location (coordinates).
 
@@
 price modification, the prices were aligned to multiples of 1,000.
+
+To understand the implications of this masking process, we use some
+standard results from algebra of random variables. The masked prices
+\(P\) are given by: \[
+P = RP\cdot \epsilon
+\]
+
+\noindent where \(RP\) are the original (raw) prices, and \(\epsilon\)
+is a random variable drawn from the uniform distribution with parameters
+\(a=0.975\) and \(b=1.025\): \[
+f(\epsilon) = 
+\begin{cases}
+\frac{1}{b - a} & \text{for } a\le \epsilon \le b\\
+0 & \text{otherwise}
+\end{cases}
+\]
+
+The expectation of \(\epsilon\) given these parameters is: \[
+E[\epsilon] = \frac{a+b}{2} = \frac{0.975 + 1.025}{2} = 1
+\]
+
+\noindent and the variance of \(\epsilon\) is: \[
+\mathrm{V}[\epsilon]=\frac{(b - a)^2}{12} = \frac{1}{4800}
+\]
+
+Therefore, the expectation of the masked prices is: \[
+\mathrm{E}[P] = \mathrm{E}[RP\cdot \epsilon] = \mathrm{E}[RP]\cdot \mathrm{E}[\epsilon] = \mathrm{E}[RP]
+\]
+
+In other words, the masked prices \(P\) are an unbiased version of the
+raw prices \(RP\).
+
+Considering that \(RP\) and \(\epsilon\) are independent, the variance
+of the masked prices is as follows: \[
+\mathrm{V}[P] = \mathrm{V}[RP\cdot \epsilon] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP]\cdot (\mathrm{E}[\epsilon])^2 + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
+\]
+
+Since \(\mathrm{E}[\epsilon] = 1\), we have that: \[
+\mathrm{V}[P] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP] + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2 = \mathrm{V}[RP]\cdot (1 + \mathrm{V}[\epsilon]) + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
+\]
+
+Solving for \(\mathrm{V}[RP]\), and replacing the expectation of the raw
+prices by its unbiased version (\(\mathrm{E}[P]\)), yields the variance
+of the raw prices: \[
+\mathrm{V}[RP] = \frac{\mathrm{V}[P] - \frac{1}{4800}\cdot (\mathrm{E}[P])^2}{1 + \frac{1}{4800}}
+\]
+
+Table \ref{tab:variance} reports the mean and the standard deviation
+(i.e., the square root of the variance) of the prices in the package,
+and the standard deviation (again, the square root of the variance) of
+the raw prices calculated using the formula above. This is done for each
+quarter and for the full year. The last column of the table can be read
+as an inflation factor. We see that the variance of the masked prices is
+inflated with respect to the variance of the original values by less
+than \(1\%\) in all cases examined. Users can use the formula above to
+calculate the inflation of the variance if they use subsamples other
+than those shown, to assess the potential impacts of the masking (e.g.,
+when computing intervals of confidence).
+
+\begin{table}
+
+\caption{\label{tab:variance-inflation-calculations}\label{tab:variance}Inflation of the variance of masked prices with respect to raw prices}
+\centering
+\fontsize{9}{11}\selectfont
+\begin{threeparttable}
+\begin{tabular}[t]{llllll}
+\toprule
+City & Period & mean(P) & sd(P) & sd(RP) & sd(P)/sd(RP)\\
+\midrule
+ & Q1 & 405,166.8 & 308,623.8 & 308,536.2 & 1.00057\\
+\cmidrule{2-6}
+ & Q2 & 388,053.5 & 252,520.5 & 252,432.1 & 1.0007\\
+\cmidrule{2-6}
+ & Q3 & 382,692.5 & 268,880.9 & 268,796.1 & 1.00063\\
+\cmidrule{2-6}
+ & Q4 & 398,157.8 & 275,459.3 & 275,370.6 & 1.00064\\
+\cmidrule{2-6}
+\multirow{-5}{*}{\raggedright\arraybackslash BARCELONA} & 2018 & 395,770.6 & 281,554.8 & 281,467.5 & 1.00062\\
+\cmidrule{1-6}
+ & Q1 & 404,960.8 & 447,935.3 & 447,850.5 & 1.00038\\
+\cmidrule{2-6}
+ & Q2 & 367,527.5 & 383,093.9 & 383,017.2 & 1.0004\\
+\cmidrule{2-6}
+ & Q3 & 365,467.2 & 359,118.3 & 359,042.2 & 1.00042\\
+\cmidrule{2-6}
+ & Q4 & 410,952.7 & 428,852.7 & 428,767 & 1.0004\\
+\cmidrule{2-6}
+\multirow{-5}{*}{\raggedright\arraybackslash MADRID} & 2018 & 396,110.1 & 417,074.4 & 416,991.8 & 1.0004\\
+\cmidrule{1-6}
+ & Q1 & 204,836.5 & 187,957.4 & 187,914.6 & 1.00046\\
+\cmidrule{2-6}
+ & Q2 & 176,661 & 141,002.4 & 140,964.6 & 1.00054\\
+\cmidrule{2-6}
+ & Q3 & 188,189.4 & 167,146 & 167,106.6 & 1.00047\\
+\cmidrule{2-6}
+ & Q4 & 208,523.5 & 183,452.7 & 183,409 & 1.00048\\
+\cmidrule{2-6}
+\multirow{-5}{*}{\raggedright\arraybackslash VALENCIA} & 2018 & 199,678.3 & 177,156 & 177,114.1 & 1.00047\\
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item P: masked prices; 
+\item RP: raw prices; 
+\item sd: standard deviation (square root of the variance)
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+With respect to the location of the dwelling, a spatial masking process
+was implemented to maintain the spatial properties of the original data
+set. The coordinates of each listing were displaced using a stochastic
+procedure. The listings were recorded using coordinates contained in
+maximum and minimum displacement circles, as shown in Figure
+\ref{fig:Anonymizing} (left). To preserve inclusion in a neighborhood,
+the spatial masking procedure was constrained to ensure that the masked
+coordinates remained in the original neighborhood of the listing.
 
@@
  }
- \caption{Coordinate displacement process for anonymisation purposes}
+ \caption{Coordinate displacement process for masking purposes}
  \label{algo:coordinates-displacement}
 \end{algorithm}
-
-With respect to the location of the dwelling, a spatial masking process
-was implemented to maintain the spatial properties of the original data
-set. The coordinates of each listing were displaced using a stochastic
-procedure. The listings were recorded using coordinates contained in
-maximum and minimum displacement circles, as shown in Figure
-\ref{fig:Anonymizing} (left). To preserve inclusion in a neighborhood,
-the spatial masking procedure was constrained to ensure that the masked
-coordinates remained in the original neighborhood of the listing.
 
@@
 
-\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)}\label{fig:unnamed-chunk-3}
+\caption{\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)}\label{fig:figure-masking-coordinates}
 \end{figure}
@@
 This paper describes a data product of a geo-referenced micro-data set
-of Spain's three largest cities. This is an excellent data product to
-help understand the complex mechanisms related to the housing market and
-housing prices. Researchers can apply hedonic models with spatial
-effects, identifying housing submarkets or machine learning techniques
-\citep[e.g.,][]{rey2023using}. The data product can also be used for
-educational proposes and teaching activities.
+of Spain's three largest cities. This data product can be of value to
+support research into the mechanisms of housing market and housing
+prices. Researchers can apply hedonic models with spatial effects,
+identify housing submarkets, or experiment with machine learning
+techniques. The data product can also be used for educational proposes
+and teaching activities. To the best of our knowledge, this is the
+largest, publicly available data set of its type that is also analysis
+ready and fully documented.
 
@@
 
-Author One and author Two are employed by Idealista. They have been
-granted permission to share the data presented in this article. None of
-the authors have financial interests or personal relationships which
-have, or could be perceived to have, influenced the work reported in
-this article.
+Authors One and Two are employed by Idealista. They have been granted
+permission to share the data presented in this article. None of the
+authors have financial interests or personal relationships which have,
+or could be perceived to have, influenced the work reported in this
+article.
 
@@
 spatial data. This work has been partially funded by the Spanish
-Ministry of Economy and Competitiveness Grants PID2019-107800GB-100
+Ministry of Economy and Competitiveness Grants PID2019-107800GB-100, but
+it was not funded by any of Canada's research councils.
 

### 2023-05-19 EPB/main_EPB_v0.Rmd
Resubmission to EPB

--- 2022-08-12 EPB/EPB.Rmd
@@
 title: A geo-referenced micro-data set of real estate listings for Spain’s three largest cities
-runninghead: Rey-Blanco \emph{et al}.
+runninghead: author \emph{et al}.
 author:
-- name: D. Rey-Blanco
-  num: 1
-- name: P. Arbues
-  num: 1
-- name: F. López
-  num: 2
-- name: A. Páez*
-  num: 3
-address:
-- num: 1
-  org: Idealista, Plaza de las Cortes 5, 28014 Madrid, Spain
-- num: 2
-  org: Facultad de CC de la Empresa, C/ Real, 3. 30201 Cartagena, Murcia, Spain
-- num: 3
-  org: School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada
-corrauth: "Antonio Páez, School of Earth, Environment and Society, McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada."
-email: paezha@mcmaster.ca
+- name: 
 abstract: "This article shares an open data product with big geo-referenced micro-data sets of 2018 real estate listings in Spain.
@@
 The data set is distributed as an R package, name {idealista18}, which can be accessed from a public Github repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
-The open data product {idealista18} is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
+The database idealista18 is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
 For transparency, we also share the randomisation process applied to the original data in the aforementioned Github repository.

### 2023-05-19 EPB/main_EPB_v0.tex
Resubmission to EPB

--- 2022-02-18 EPB/EPB.tex
@@
 
-\runninghead{Rey-Blanco \emph{et al}.}
+\runninghead{author \emph{et al}.}
 
-\author{D. Rey-Blanco\affilnum{1}, P. González-Arbues\affilnum{1}, F.
-López\affilnum{2}, A. Páez*\affilnum{3}}
+\author{\affilnum{}}
 
-\affiliation{\affilnum{1}{Idealista, Plaza de las Cortes 5, 28014
-Madrid, Spain}\\\affilnum{2}{Facultad de CC de la Empresa, C/ Real, 3.
-30201 Cartagena, Murcia, Spain}\\\affilnum{3}{School of Earth,
-Environment and Society, McMaster University, 1280 Main St W, Hamilton,
-Ontario L8S 4K1 Canada}}
-
-\corrauth{Antonio Páez, School of Earth, Environment and Society,
-McMaster University, 1280 Main St W, Hamilton, Ontario L8S 4K1 Canada.}
-
-\email{\href{mailto:paezha@mcmaster.ca}{\nolinkurl{paezha@mcmaster.ca}}}
+\affiliation{}
 
@@
 the data set has been supplied directly by Idealista, and therefore is
-clean and free of download errors. The listings have been enriched with
-official information from the Spanish cadastre along with other relevant
-geographical features, such as distance to urban points of interest. The
-data set is distributed as an R package, named `idealista18', which can
-be accessed from the Github
+clean and free of download errors. However, in order to be able to share
+publicly the data we have slightly modified some variables by applying
+some random noise which will not bias the main results derived from its
+usage.
+
+This micro-data set is expected to be used as a benchmark dataset to
+test empirical methods in a reproducible fashion. Applied and
+theoretical researchers on real estate mass appraisal and valuation
+methods might use this dataset to canonically compare the performance of
+their proposed comparable and hedonic models, among others. The data can
+also be used to study the segmentation of housing submarkets and related
+topics such as the impact of suburban areas on house prices. The
+listings have been enriched with official information from the Spanish
+cadastre along with other relevant geographical features, such as
+distance to urban points of interest. In any case, the data might be
+easily extended by spatially joining other datasets that contain
+information at administrative, census tract, or postal code levels.
+
+The data set is distributed as an R package, name \{idealista18\}, which
+can be accessed from a public Github
 repository\footnote{The direct URL to the GitHub repository is \url{https://github.com/paezha/idealista18}}.
+The database idealista18 is made available under the
+\href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}.
+For transparency, we also share the randomisation process applied to the
+original data in the aforementioned Github repository.
 
@@
 \centering
-\includegraphics{EPB_files/figure-latex/unnamed-chunk-1-1.pdf}
+\includegraphics{main_EPB_v0_files/figure-latex/unnamed-chunk-1-1.pdf}
 \caption{\label{fig:all-polygons}Quantile maps of the number of

### 2023-10-18 editorial-materials/EPB-2022-0116.R1_response_to_reviewer.Rmd
Revisions to cover letter requested by Editorial Assistant

@@
 -->
+
+To the Editor:
+
+We have responded to the comments of the reviewers in a thorough and conscientious manner. This has made the paper longer.
+
+At this point, the paper has been stuck in limbo for months with the Editorial Assistant, who initially requested that we cut the length of the paper, but did not indicate just _what_ to cut.
+In our opinion, the paper needs to be seen by an Editor to decide whether the paper _needs_ to be cut, and if so, _what_ can be left out without compromising the changes made to satisfy the requests of the reviewers.
+
+More generally, to avoid this situation in the future with Data Papers, the editors need to:
+
+1. Instruct reviewers as to the expectations for criticism. If the papers cannot be longer, they should not ask questions that require long responses; and/or
+1. Instruct the Editorial Assistant that papers _will be_ longer after revisions; and/or
+1. Provide clear guidelines to the authors to ensure that they know what is essential to revise so that they do not spend time and effort responding to critiques that the Editors may find superfluous.
+
+Below are our detailed responses to the comments from the Reviewers.
 

### 2024-03-12 EPB/main_EPB.Rmd
Prepared files for final submission to production with EPB

@@
 ```{r load-packages, include=FALSE}
-library(dplyr)
-library(ggplot2)
-library(glue)
-library(gridExtra)
-library(idealista18)
-library(kableExtra)
-library(xtable)
-library(sf)
+library(dplyr) # A Grammar of Data Manipulation
+library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
+library(glue) # Interpreted String Literals
+library(gridExtra) # Miscellaneous Functions for "Grid" Graphics
+library(idealista18) # Idealista 2018 Data Package
+library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
+library(patchwork) # The Composer of Plots
+library(sf) # Simple Features for R
+library(xtable) # Export Tables to LaTeX or HTML
 ```
@@
 
-This micro-data set can be used to benchmark new methods in a reproducible fashion  [e.g., @rey2023using].
+This micro-data set can be used to benchmark new methods in a reproducible fashion  [e.g., @rey2024using].
 Applied and theoretical researchers on real estate mass appraisal and valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others.
@@
         text = element_text(size=4))
+
+fig_01 <- ((plot1 + plot4)/(plot2 + plot5)/(plot3 + plot6))
+
+ggsave(fig_01,
+       filename = "main_EPB_figures/fig-01.jpeg",
+       height = 6.5,
+       units = "in")
 ```
@@
 
-library(patchwork)
-(plot1 + plot4)/(plot2 + plot5)/(plot3 + plot6)
+fig_01
 ```
@@
 (Right) Coordinate displacement in meters (Valencia)",fig.show='hold',fig.align='center'}
-knitr::include_graphics(c("EPB_files/points-moved-image.png","EPB_files/coordinates-valencia.png"))
+knitr::include_graphics(c("main_EPB_figures/points-moved-image.jpeg","main_EPB_figures/coordinates-valencia.jpeg"))
 ```

### 2024-03-12 EPB/main_EPB.tex
Prepared files for final submission to production with EPB

@@
 This micro-data set can be used to benchmark new methods in a
-reproducible fashion \citep[e.g.,][]{rey2023using}. Applied and
+reproducible fashion \citep[e.g.,][]{rey2024using}. Applied and
 theoretical researchers on real estate mass appraisal and valuation
@@
 
-{\centering \includegraphics[width=0.29\linewidth,height=0.2\textheight]{EPB_files/points-moved-image} \includegraphics[width=0.37\linewidth,height=0.2\textheight]{EPB_files/coordinates-valencia}
+{\centering \includegraphics[width=0.29\linewidth,height=0.2\textheight]{main_EPB_figures/points-moved-image} \includegraphics[width=0.37\linewidth,height=0.2\textheight]{main_EPB_figures/coordinates-valencia}
 

### 2024-04-04 README.Rmd
Updated with bibliographic information

@@
 The data sets are offered as a fully documented R package and are available for scientific and educational purposes, particularly for geo-spatial studies.
+
+## Paper
+
+**A geo-referenced micro-data set of real estate listings for Spain’s three largest cities**
+
+David Rey-Blanco and Pelayo Arbues (idealista)  
+Fernando Lopez (Universidad Politecnica de Cartagena)  
+Antonio Paez (McMaster University)
+
+Published in *Environment and Planning B: Urban Analytics and City Science* (https://doi.org/10.1177/23998083241242844)
 
